### 2014-12-11 writeup/paper.tex
add paper skeleton

\documentclass[12pt,preprint]{aastex}

% has to be before amssymb it seems
\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
            filecolor=linkcolor,urlcolor=linkcolor}

\usepackage{url}
\usepackage{algorithmic,algorithm}
\usepackage{amssymb,amsmath}

\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{arXiv:#1}}

\usepackage{listings}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
\lstset{language=Python,
        basicstyle=\footnotesize\ttfamily,
        showspaces=false,
        showstringspaces=false,
        tabsize=2,
        breaklines=false,
        breakatwhitespace=true,
        identifierstyle=\ttfamily,
        keywordstyle=\bfseries\color[rgb]{0.133,0.545,0.133},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
    }

\newcommand{\project}[1]{{\sffamily #1}}
\newcommand{\Python}{\project{Python}}
\newcommand{\numpy}{\project{numpy}}
\newcommand{\Ubuntu}{\project{Ubuntu}}
\newcommand{\github}{\project{GitHub}}
\newcommand{\pip}{\project{pip}}
\newcommand{\acor}{\project{acor}}
\newcommand{\thisplain}{emcee}
\newcommand{\this}{\project{\thisplain}}
\newcommand{\paper}{document}
\newcommand{\license}{MIT License}

\newcommand{\foreign}[1]{\emph{#1}}
\newcommand{\etal}{\foreign{et\,al.}}
\newcommand{\etc}{\foreign{etc.}}

\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\fig}[1]{\Fig{#1}}
\newcommand{\figlabel}[1]{\label{fig:#1}}
\newcommand{\Tab}[1]{Table~\ref{tab:#1}}
\newcommand{\tab}[1]{\Tab{#1}}
\newcommand{\tablabel}[1]{\label{tab:#1}}
\newcommand{\Eq}[1]{Equation~(\ref{eq:#1})}
\newcommand{\eq}[1]{\Eq{#1}}
\newcommand{\eqlabel}[1]{\label{eq:#1}}
\newcommand{\Sect}[1]{Section~\ref{sect:#1}}
\newcommand{\sect}[1]{\Sect{#1}}
\newcommand{\App}[1]{Appendix~\ref{sect:#1}}
\newcommand{\app}[1]{\App{#1}}
\newcommand{\sectlabel}[1]{\label{sect:#1}}
\newcommand{\Algo}[1]{Algorithm~\ref{algo:#1}}
\newcommand{\algo}[1]{\Algo{#1}}
\newcommand{\algolabel}[1]{\label{algo:#1}}

% math symbols
\newcommand{\dd}{\mathrm{d}}
\newcommand{\like}{\mathscr{L}}
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\paramvector}[1]{\bvec{#1}}
\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
\newcommand{\ensemble}{S}
\newcommand{\colorens}[1]{\ensemble^{(#1)}}
\newcommand{\red}{\colorens{0}}
\newcommand{\blue}{\colorens{1}}
\renewcommand{\vector}[1]{#1}
\renewcommand{\matrix}[1]{#1}
\newcommand{\pr}[1]{\ensuremath{p(#1)}}
\newcommand{\af}{\ensuremath{a_f}}
\newcommand{\expect}[1]{\left<#1\right>}

% model parameters
\newcommand{\model}{\ensuremath{\vector{\Theta}}}
\newcommand{\data}{\ensuremath{\vector{D}}}
\newcommand{\nuisance}{\ensuremath{\vector{\alpha}}}
\newcommand{\link}{\ensuremath{X}}

% units
\newcommand{\unit}[1]{\mathrm{#1}}

\begin{document}

\title{Multiband Periodograms of Astronomical Sources}

\newcommand{\escience}{1}
\newcommand{\uwastro}{2}
\author{Jacob T. VanderPlas\altaffilmark{\escience}}
\author{{\v Z}eljko Ivezi{\'c}\altaffilmark{\uwastro}}
\altaffiltext{\escience}{eScience Institute, University of Washington}
\altaffiltext{\uwastro}{Department of Astronomy, University of Washington}

\begin{abstract}
  This paper introduces a general method for least-squares spectral fitting of multi-band periodic data.
\end{abstract}

\keywords{
    methods: data analysis ---
    methods: numerical ---
    methods: statistical
}

\section{Introduction}

\begin{itemize}
  \item Lomb Scargle periodograms \& their use in Astronomy
  \item Other methods of period finding (Supersmoother, ARMA-type models, etc.)
  \item Need for multi-band periodograms
\end{itemize}

\section{Multiterm Periodograms: Background}

\begin{itemize}
  \item Relationship between Fourier Periodogram \& Least Squares fit (cite Jaynes)
  \item Matrix form of linear models
  \item Generalization of single-term to multi-term models
  \item Show expression for Power; refer to appendix for computational details
\end{itemize}

\section{Moving to Multiple Bands}

\begin{itemize}
  \item Write-out the linear model
  \item Demonstrate the matrix form of this
  \item Use light regularization
  \item Correspondance of the single-term multiband result to the weighted sum of single-band Lomb-Scargle results
\end{itemize}

\section{Examples}
\subsection{Simulated Example}

\begin{itemize}
  \item Show power spectrum for a single-band RR-Lyrae
  \item Reduce observations, show multiple power spectra for multi-band RR-Lyrae
  \item Show single power spectrum for multi-band RR-Lyrae
\end{itemize}

\subsection{Stripe 82}

\begin{itemize}
  \item Compute periods using majority method
  \item Compute periods using unified method
  \item Compare the results
\end{itemize}

\section{Further work}

\begin{itemize}
  \item issues with window function corrections
  \item issues with physicality of model
\end{itemize}

\section{Appendices}

\appendix

\section{The Lomb-Scargle Periodogram and Least-Squares Spectral Fitting}

We start with time-series data

\begin{equation}
  D = \{t_i,y_i,\sigma_i\}_{i=1}^N
\end{equation}

without loss of generality, we will assume throughout that the $y_i$ values are centered; that is, the weighted mean

\begin{equation}
  \label{eq:centered-data}
  \frac{\sum_{i=1}^N w_i y_i}{\sum_{i=1}^N w_i} = 0
\end{equation}

where the weights are $w_i \equiv \sigma_i^{-2}$. We will seek a single sinusoid model which fits this data; that is, our model is

\begin{equation}
M(t~|~A,\omega,\phi) = A \sin(\omega t + \phi)
\end{equation}

to make this a linear problem,

\begin{equation}
M(t~|~\omega,\theta) = \theta_0 \sin(\omega t) + \theta_1\cos(\omega t)
\end{equation}

which can be related to the above form by $\theta_0 = A\cos\phi$ and $\theta_1 = A\sin\phi$.
The $\chi^2$ for this model is given by

\begin{equation}
\chi^2(D, \omega, \theta) = \sum_i\frac{[y_i - M(t_i~|~\omega,\theta)]^2}{2\sigma_i^2}
\end{equation}

This can be expressed more compactly by defining the following matrices:

\begin{equation}
X_\omega = \left[
\begin{array}{cc}
\sin(\omega t_1) & \cos(\omega t_1)\\
\sin(\omega t_2) & \cos(\omega t_2)\\
\vdots & \vdots \\
\sin(\omega t_N) & \cos(\omega t_N)\\
\end{array}
\right];~~
y = \left[
\begin{array}{c}
y_1 \\
y_2\\
\vdots \\
y_N\\
\end{array}
\right];~~
\Sigma = \left[
\begin{array}{ccccc}
\sigma_1^2 & 0 & 0 & \cdots & 0\\
0 & \sigma_2^2 & 0 & \cdots & 0\\
0 & 0 & \sigma_3^2 & \cdots & 0\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
0 & 0 & 0 & \cdots & \sigma_N^2
\end{array}
\right]
\end{equation}

Then the model is given by $X_\omega\theta$ and the $\chi^2$ can be written
compactly as:

\begin{equation}
  \chi^2(\omega, \theta) = (y - X_\omega\theta)^T\Sigma^{-1}(y - X_\omega\theta)
\end{equation}

Minimizing this cost funciton with respect to $\theta$ gives the maximum likelihood parameters:

\begin{equation}
\hat{\theta}_\omega = (X_\omega^T\Sigma^{-1}X_\omega)^{-1}X_\omega^T\Sigma^{-1}y
\end{equation}

We'll simplify this slightly by defining the quantities $\tilde{X}_\omega = \Sigma^{-1/2}X_\omega$ and $\tilde{y} = \Sigma^{-1/2}y$, so that the maximum likelihood parameters can be written

\begin{equation}
  \hat{\theta}_\omega = (\tilde{X}_\omega^T\tilde{X}_\omega)^{-1}\tilde{X}_\omega^T\tilde{y}
\end{equation}

Now the value of $\chi^2$ at maximum likelihood can be expressed:

\begin{equation}
  \chi^2(\omega, \hat{\theta}_\omega) = \tilde{y}^T\tilde{y} - \tilde{y}^T\tilde{X}_\omega(\tilde{X}_\omega^T\tilde{X}_\omega)^{-1}\tilde{X}_\omega^T \tilde{y}
\end{equation}

The form of this equation suggests defining a reference $\chi^2$ value which is given by a best-fit constant model to the data: because of the constraint given in Eqn.~\ref{eq:centered-data}, this reference is simply written $\chi^2_0 = \tilde{y}^T\tilde{y}$. We can now follow Lomb [ref] \& Scargle [ref] and write the normalized periodogram:

\begin{eqnarray}
  \label{eq:power}
  P(\omega) &\equiv& 1 -\frac{\chi^2(\omega, \hat{\theta}_\omega)}{\chi^2_0(\omega)}\\ &=& \frac{\tilde{y}^T\tilde{X}_\omega(\tilde{X}_\omega^T\tilde{X}_\omega)^{-1}\tilde{X}_\omega^T \tilde{y}}{\tilde{y}^T\tilde{y}}
\end{eqnarray}

Typically, the Lomb-Scargle periodogram is expressed in terms of various weighted sums of sines and cosines of the data; this expression is equivalent, but has the advantage that it can be trivially generalized to an arbitrary linear model through adding columns to the $X$ matrix, and can handle arbitrary correlated measurement errors by adding appropriate off-diagonal terms to the $\Sigma$ matrix.

\section{Higher-order Models}

As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of Zechmeister [cite]. This adjusts the classic Lomb-Scargle algorithm by using a model with a floating mean, which can be more accurate for certain observing cadences:

\begin{equation}
  M(t~|~\omega, \theta) = \theta_0 + \theta_1\sin\omega t + \theta_2\cos\omega t
\end{equation}

The normalized power under this model can be computed by simply adding a column of ones to the $X$ matrix.

In a similar vein, the power for a truncated Fourier series model of any order can be constructed by extending the $X$ matrix with appropriate columns; for example, for a 2-term truncated Fourier fit with a floating mean we can write

\begin{equation}
X_\omega = \left[
\begin{array}{ccccc}
1 & \sin(\omega t_1) & \cos(\omega t_1) & \sin(2\omega t_1) & \cos(2\omega t_1)\\
1 & \sin(\omega t_2) & \cos(\omega t_2) & \sin(2\omega t_2) & \cos(2\omega t_2)\\
1 & \sin(\omega t_3) & \cos(\omega t_3) & \sin(2\omega t_3) & \cos(2\omega t_3)\\
\vdots & \vdots & \vdots & \vdots & \vdots \\
1 & \sin(\omega t_N) & \cos(\omega t_N) & \sin(2\omega t_N) & \cos(2\omega t_N)\\
\end{array}
\right]
\end{equation}

Plugging this $X$ matrix into Eqn.~\ref{eq:power} will give us the normalized periodogram associated with the two-term truncated Fourier model.

\section{Regularized Models}

(Show how regularization works its way through the equations \& derive expression for power).

\end{document}

### 2014-12-11 multiband_LS/data.py
change data directory

@@ -20,7 +20,7 @@ def set_data_directory(data_directory):
 
 DATA_DIRECTORY = ''
 set_data_directory(os.path.abspath(os.path.join(os.path.dirname(__file__),
-                                                'data')))
+                                                '_data_downloads')))
 
 SESAR_RRLYRAE_URL = 'http://www.astro.washington.edu/users/bsesar/S82_RRLyr/'
 

### 2014-12-11 writeup/paper.tex
add bibliography stuff to paper

@@
 \begin{itemize}
-  \item Relationship between Fourier Periodogram \& Least Squares fit (cite Jaynes)
+  \item Relationship between Fourier Periodogram \& Least Squares fit \citep{Jaynes96}
   \item Matrix form of linear models
@@
 
-\section{Appendices}
+\bibliographystyle{apj}
+\bibliography{paper}
 
@@
 
-As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of Zechmeister [cite].
+As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of \citep{Zechmeister09}.
 This adjusts the classic Lomb-Scargle algorithm by using a model with a floating mean, which can be more accurate for certain observing cadences:

### 2014-12-12 writeup/paper.tex
add some example figures

@@
 \usepackage{amssymb,amsmath}
+\usepackage{graphicx}
+\graphicspath{{figures/}}
 
@@
 \subsection{Simulated Example}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig01.pdf}
+  \caption{
+    An illustration of the typical approach to multiband periodograms,
+    in which each band is fit individually. The data consists of 60 coeval
+    u-g-r-i-z observations spread over 180 nights, and is based on an
+    RR Lyrae template from \citet{Sesar2010}. With this much data in each
+    band, individual periodograms can be constructed and compared to find the
+    true period $P=0.622$ days.
+  }
+  \label{fig:01}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig02.pdf}
+  \caption{
+    Another realization of the data from \fig{01}, with only a single-band
+    observation each night (i.e. 12 observations per band, spread over 180
+    days). In this case, there is not enough data in each band to recover
+    periodicity information. By applying the multiband approach which utilizes
+    all the data at once (lower right), we again see a significant spike of
+    power at the true period $P=0.622$ days.    
+  } 
+  \label{fig:02}
+\end{figure}
 

### 2014-12-12 writeup/figures/fig_rrlyrae_sim.py

@@ -1,0 +1,122 @@
+"""
+Here we plot a typical approach to multi-band Lomb-Scargle: treating each band
+separately, and taking a majority vote between the bands.
+"""
+
+import sys
+import os
+sys.path.append(os.path.abspath('../..'))
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Use seaborn settings for plot styles
+import seaborn; seaborn.set()
+
+from multiband_LS.generated import RRLyraeObject
+from multiband_LS.lomb_scargle import LombScargleAstroML, LombScargleMultiband
+
+
+# Choose a Sesar 2010 object to base our fits on
+lcid = 1019544
+rrlyrae = RRLyraeObject(lcid, random_state=0)
+
+# Generate data in a 6-month observing season
+Nobs = 60
+dy = 0.07
+rng = np.random.RandomState(0)
+
+nights = np.arange(180)
+rng.shuffle(nights)
+nights = nights[:Nobs]
+
+t = 57000 + nights + 0.2 * rng.rand(Nobs)
+mags = np.array([rrlyrae.generated(band, t, err=dy, corrected=False)
+                 for band in 'ugriz'])
+
+#----------------------------------------------------------------------
+# First figure:
+# Compute the lomb-scargle periodogram in each band
+
+periods = np.linspace(0.2, 0.9, 10000)
+omegas = 2 * np.pi / periods
+P = [LombScargleAstroML().fit(t, m, dy).power(omegas) for m in mags]
+
+fig, ax = plt.subplots(1, 2, figsize=(10, 4))
+fig.subplots_adjust(left=0.07, right=0.95, wspace=0.1, bottom=0.15)
+
+for i, band in enumerate('ugriz'):
+    ax[0].errorbar((t / rrlyrae.period) % 1, mags[i], dy,
+                   fmt='.', label=band)
+ax[0].set_ylim(18, 14.5)
+ax[0].legend(loc='upper left', fontsize=12, ncol=3)
+ax[0].set_title('Folded Data (P={0:.3f} days)'.format(rrlyrae.period))
+ax[0].set_xlabel('phase')
+ax[0].set_ylabel('magnitude')
+
+for i, (Pi, band) in enumerate(zip(P, 'ugriz')):
+    offset = 4 - i
+    ax[1].plot(periods, Pi + offset, lw=1)
[…]

### 2014-12-12 writeup/paper.tex
initial draft of introduction & background

--- 2014-12-11 writeup/paper.tex
@@
 % has to be before amssymb it seems
-\usepackage{color,hyperref}
-\definecolor{linkcolor}{rgb}{0,0,0.5}
-\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
-            filecolor=linkcolor,urlcolor=linkcolor}
+%\usepackage{color,hyperref}
+%\definecolor{linkcolor}{rgb}{0,0,0.5}
+%\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
+%            filecolor=linkcolor,urlcolor=linkcolor}
 
+\usepackage{color}
 \usepackage{url}
@@
 \usepackage{amssymb,amsmath}
+\usepackage{graphicx}
+\graphicspath{{figures/}}
 
-\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{arXiv:#1}}
+%\usepackage{listings}
+%\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
+%\lstset{language=Python,
+%        basicstyle=\footnotesize\ttfamily,
+%        showspaces=false,
+%        showstringspaces=false,
+%        tabsize=2,
+%        breaklines=false,
+%        breakatwhitespace=true,
+%        identifierstyle=\ttfamily,
+%        keywordstyle=\bfseries\color[rgb]{0.133,0.545,0.133},
+%        commentstyle=\color[rgb]{0.133,0.545,0.133},
+%        stringstyle=\color[rgb]{0.627,0.126,0.941},
+%    }
 
-\usepackage{listings}
-\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
-\lstset{language=Python,
-        basicstyle=\footnotesize\ttfamily,
-        showspaces=false,
-        showstringspaces=false,
-        tabsize=2,
-        breaklines=false,
-        breakatwhitespace=true,
-        identifierstyle=\ttfamily,
-        keywordstyle=\bfseries\color[rgb]{0.133,0.545,0.133},
-        commentstyle=\color[rgb]{0.133,0.545,0.133},
-        stringstyle=\color[rgb]{0.627,0.126,0.941},
-    }
+\newcommand{\foreign}[1]{{\it #1}}
 
-\newcommand{\project}[1]{{\sffamily #1}}
-\newcommand{\Python}{\project{Python}}
-\newcommand{\numpy}{\project{numpy}}
-\newcommand{\Ubuntu}{\project{Ubuntu}}
-\newcommand{\github}{\project{GitHub}}
-\newcommand{\pip}{\project{pip}}
-\newcommand{\acor}{\project{acor}}
-\newcommand{\thisplain}{emcee}
-\newcommand{\this}{\project{\thisplain}}
-\newcommand{\paper}{document}
-\newcommand{\license}{MIT License}
-
-\newcommand{\foreign}[1]{\emph{#1}}
+\newcommand{\adhoc}{\foreign{ad hoc}}
 \newcommand{\etal}{\foreign{et\,al.}}
@@
 \newcommand{\figlabel}[1]{\label{fig:#1}}
-\newcommand{\Tab}[1]{Table~\ref{tab:#1}}
-\newcommand{\tab}[1]{\Tab{#1}}
-\newcommand{\tablabel}[1]{\label{tab:#1}}
 \newcommand{\Eq}[1]{Equation~(\ref{eq:#1})}
@@
 \newcommand{\sectlabel}[1]{\label{sect:#1}}
-\newcommand{\Algo}[1]{Algorithm~\ref{algo:#1}}
-\newcommand{\algo}[1]{\Algo{#1}}
-\newcommand{\algolabel}[1]{\label{algo:#1}}
-
-% math symbols
-\newcommand{\dd}{\mathrm{d}}
-\newcommand{\like}{\mathscr{L}}
-\newcommand{\bvec}[1]{\boldsymbol{#1}}
-\newcommand{\paramvector}[1]{\bvec{#1}}
-\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
-\newcommand{\ensemble}{S}
-\newcommand{\colorens}[1]{\ensemble^{(#1)}}
-\newcommand{\red}{\colorens{0}}
-\newcommand{\blue}{\colorens{1}}
-\renewcommand{\vector}[1]{#1}
-\renewcommand{\matrix}[1]{#1}
-\newcommand{\pr}[1]{\ensuremath{p(#1)}}
-\newcommand{\af}{\ensuremath{a_f}}
-\newcommand{\expect}[1]{\left<#1\right>}
-
-% model parameters
-\newcommand{\model}{\ensuremath{\vector{\Theta}}}
-\newcommand{\data}{\ensuremath{\vector{D}}}
-\newcommand{\nuisance}{\ensuremath{\vector{\alpha}}}
-\newcommand{\link}{\ensuremath{X}}
-
-% units
-\newcommand{\unit}[1]{\mathrm{#1}}
 
@@
 
-\begin{itemize}
-  \item Lomb Scargle periodograms \& their use in Astronomy
-  \item Other methods of period finding (Supersmoother, ARMA-type models, etc.)
-  \item Need for multi-band periodograms
-\end{itemize}
+The detection and quantification of periodicity in time-varying signals is an important area of data analysis within modern astronomical surveys.
+For evenly-spaced data, the {\it Schuster periodogram}, introduced in 1905, gives a quantitative measure of the periodicity of data as a function of the angular frequency $\omega$.
+For data $\{d_k\}_{k=1}^N$ measured at equal intervals $t_k = t_0 + k\Delta t$, the Schuster periodogram, which measures the spectral power as a function of the angular frequency, is given by
+\begin{equation}
+  \eqlabel{Schuster}
+  C(\omega) = \frac{1}{N}\left| \sum_{k=1}^N d_k e^{i\omega t_k} \right|^2,
+\end{equation}
+and can be computed very efficiently using the Fast Fourier Transform.
 
-\section{Multiterm Periodograms: Background}
+Because astronomical observing cadences are rarely so uniform, many have looked at extending the concept of the periodogram.
+Most famously, \citep{Lomb76} and \citep{Scargle82} extended earlier work to define the {\it normalized periodogram}:
+\begin{equation}
+  \eqlabel{LombScargle}
+  P_N(\omega) = \frac{1}{2\sigma^2}\left[
+    \frac{\left[\sum_k(d_k - \bar{d})\cos\omega(t_k - \tau)\right]^2}
+    {\sum_k \cos^2\omega(t_k - \tau)}
+    +
+    \frac{\left[\sum_k(d_k - \bar{d})\sin\omega(t_k - \tau)\right]^2}
+    {\sum_k \sin^2\omega(t_k - \tau)}
+\right],
+\end{equation}
+where $\bar{d}$ is the mean and $\sigma^2$ is the variance of the data $\{d_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[See][for an in-depth discussion]{NumRec}.
+\citep{Lomb76} showed that this offset has a more important effect: namely, it makes $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
 
-\begin{itemize}
-  \item Relationship between Fourier Periodogram \& Least Squares fit \citep{Jaynes96}
-  \item Matrix form of linear models
-  \item Generalization of single-term to multi-term models
-  \item Show expression for Power; refer to appendix for computational details
-\end{itemize}
+\begin{equation}
+  \eqlabel{SingleModel}
+  d(t) = A\sin(\omega t + \phi).
+\end{equation}
+
+This deep connection between spectral power and least squares fitting methods was solidified by \citep{Jaynes87}, who demonstrated that the normalized periodogram of Lomb and Scargle is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
+Building on this result, \citep{Bretthorst88} explored the extension of these methods to more complicated models with multiple frequency terms, nonstationary frequencies, and other more sophisticated models within a Bayesian framework.
+
+(Mention supersmoother and ARMA-based methods?)
+
+A weakness with the above methods is that they require homogeneous measurements -- for astronomy data, this means that successive measurements must be taken through a single band.
+This has not been a problem for past surveys, as measurements are generally taken through a single photometric filter (e.g.
+LINEAR, ??), or in all bands at each observation (e.g.
+SDSS, ??).
+In such cases, past studies have generally relied on \adhoc{} methods such as a majority vote among multiple single-band estimates of the periodogram.
+
+For future multicolor surveys such as LSST, this \adhoc{} approach will not be sufficient.
+In order to take advantage of the full available data, it would be desirable to have a single estimate of the periodogram which accounts for all observed data in a manner which is not dependent on the underlying spectrum of the object.
+We propose such a method below.
+
+\section{Background: Matrix Formalism for Periodograms}
+
+In this section we will give a quick quantitative introduction to the least squares fitting formulation of the normalized periodogram of \eq{LombScargle}. We'll denote our $N$ observed data points as
+\begin{equation}
+  D = \{t_k, y_k, \sigma_k\}_{k=1}^N
+\end{equation}
+where $t_k$ is the time of observation, $y_k$ is the observed value (typically a magnitude), and $\sigma_k$ describes the Gaussian errors on each value.
+Without loss of generality we will assume that the data $y_k$ are centered such that the measurements satisfy
+\begin{equation}
+  \eqlabel{ycentered}
+  \frac{\sum_k w_ky_k}{\sum_k w_k} = 0
+\end{equation}
+where the weights are $w_k = \sigma_k^{-2}$.
+
+\subsection{Stationary Sinusoid Model}
+The Lomb-Scargle periodogram of \eq{LombScargle} can be derived from the normalized $\chi^2$ of the best-fit single-term stationary sinusoidal model given in \eq{SingleModel}.
+To make the problem linear, we can re-express the model in terms of the parameter vector $\theta = [A\cos\phi, A\sin\phi]$ so that our model is
+\begin{equation}
+  \eqlabel{simplemodel}
+  M(t|\omega,\theta) = \theta_1\sin(\omega t) + \theta_2\cos(\omega t).
+\end{equation}
+We can find the maximum likelihood estimate of the parameters $\theta$ by minimizing the $\chi^2$ of the model, which is given by
+\begin{equation}
+  \chi^2(\omega) = \sum_k \frac{(y_k - M(t_k|\omega,\theta))^2}{2\sigma_k^2}.
+\end{equation}
+For the single-term Fourier model, it can be shown \citep[See, e.g.][]{ICVG2014} that
+\begin{equation}
+  \eqlabel{chi2_PN}
+  \chi_{min}^2(\omega) = \chi^2_0[1 - P_N(\omega)]
+\end{equation}
+where $P_N(\omega)$ is the normalized periodogram given in \eq{LombScargle} and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$
+
+These quantities can be expressed more compactly by defining the following matrices:
+
+\begin{equation}
+X_\omega = \left[
+\begin{array}{cc}
+\sin(\omega t_1) & \cos(\omega t_1)\\
+\sin(\omega t_2) & \cos(\omega t_2)\\
+\vdots & \vdots \\
+\sin(\omega t_N) & \cos(\omega t_N)\\
+\end{array}
+\right];~~
+y = \left[
+\begin{array}{c}
+y_1 \\
+y_2\\
+\vdots \\
+y_N\\
+\end{array}
+\right];~~
+\Sigma = \left[
+\begin{array}{cccc}
+\sigma_1^2 & 0 &  \cdots & 0\\
+0 & \sigma_2^2 &  \cdots & 0\\
+\vdots & \vdots &  \ddots & \vdots\\
+0 & 0 &  \cdots & \sigma_N^2
+\end{array}
+\right]
+\end{equation}
+
+With these definitions, the model in \eq{simplemodel} is a simple linear product: $M(t|\omega,\theta) = X_\omega\theta$ and the model and reference $\chi^2$ can be written
+
+\begin{eqnarray}
+  \chi^2(\omega) &=& (y - X_\omega\theta)^T\Sigma^{-1}(y - X_\omega\theta)\\
+  \chi^2_0 &=& y^T \Sigma^{-1} y
+\end{eqnarray}
+
+the normalized periodogram can be computed by finding the minimum of $\chi^2(\omega)$ via standard methods and plugging the result into \eq{chi2_PN} to find
+
+\begin{equation}
+  \eqlabel{LombScargle2}
+  P_N(\omega) = \frac{y^T\Sigma^{-1}X_\omega(X_\omega^T\Sigma^{-1}X_\omega)^{-1}X_\omega^T\Sigma^{-1}y}{y^T\Sigma^{-1}y}
+\end{equation}
+
+For $\Sigma \propto I$, this expression is equivalent to \eq{LombScargle}.
+
+While \eq{LombScargle} can be computed more efficiently than \eq{LombScargle2}, the matrix-based formulation has several advantages
+\begin{enumerate}
+  \item It is trivially extended for heteroscedastic and/or uncorrelated noise in the data $y_k$ by appropriately modifying the noise matrix $\Sigma$
+  \item It is trivially extended to more sophisticated models by appropriately modifying the design matrix $X_\omega$.
+  \item It is trivially extended to L2-regularized likelihoods by adding an appropriate diagonal term to $X_\omega^T\Sigma^{-1}X_\omega$ before inverting.
+\end{enumerate}
+
+\subsection{Lomb-Scargle with Floating Mean}
+
+As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of \citep{Zechmeister09}.
+This adjusts the classic Lomb-Scargle algorithm by using a model with a floating mean, which can be more accurate for certain observing cadences:
+
+\begin{equation}
+  M(t~|~\omega, \theta) = \theta_0 + \theta_1\sin\omega t + \theta_2\cos\omega t
+\end{equation}
+
+While \citep{Zechmeister09} details a formalism to express this {\it generalized periodogram} model similarly to \eq{LombScargle}, in the matrix formalism all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
+
+\subsection{Truncated Fourier Models}
+In a similar vein, the power for a truncated Fourier series model of any order can be constructed by extending the $X$ matrix with appropriate columns; for example, for a 2-term truncated Fourier fit with a floating mean we can write
+
+\begin{equation}
+X_\omega = \left[
+\begin{array}{ccccc}
+1 & \sin(\omega t_1) & \cos(\omega t_1) & \sin(2\omega t_1) & \cos(2\omega t_1)\\
+1 & \sin(\omega t_2) & \cos(\omega t_2) & \sin(2\omega t_2) & \cos(2\omega t_2)\\
+1 & \sin(\omega t_3) & \cos(\omega t_3) & \sin(2\omega t_3) & \cos(2\omega t_3)\\
+\vdots & \vdots & \vdots & \vdots & \vdots \\
+1 & \sin(\omega t_N) & \cos(\omega t_N) & \sin(2\omega t_N) & \cos(2\omega t_N)\\
+\end{array}
+\right]
+\end{equation}
+
+Plugging this $X$ matrix into Eqn.~\ref{eq:power} will give us the normalized periodogram associated with the two-term truncated Fourier model.
+
+\subsection{Regularized Models}
+
+Show how regularized models fit into this.
 
@@
 \subsection{Simulated Example}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig01.pdf}
+  \caption{
+    An illustration of the typical approach to multiband periodograms,
+    in which each band is fit individually. The data consists of 60 coeval
+    {\it ugriz} observations spread over 180 nights, and is based on an
+    RR Lyrae template from \citet{Sesar2010}. With this much data in each
+    band, individual periodograms can be constructed and compared to find the
+    true period $P=0.622$ days. Note also the presence of aliasing typical
+    in Lomb-Scargle periodograms, located at beat frequencies between the
+    true Period and the 1-day observing cadence.
+  }
+  \figlabel{01}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig02.pdf}
+  \caption{
+    Another realization of the data from \fig{01}, with only a single filter
+    observed each night (i.e. 12 observations per band, spread over 180
+    days). In this case, single-band data is not sufficient to recover the
+    Lomb-Scargle peaks, and the classic band-by-band approach fails.
+    By utilizing all the data at once with the multiband approach,
+    we recover a significant spike of power at the period $P=0.622$ days.    
+  } 
+  \figlabel{02}
+\end{figure}
 
@@
 
-\appendix
-
-\section{The Lomb-Scargle Periodogram and Least-Squares Spectral Fitting}
-
-We start with time-series data
-
-\begin{equation}
-  D = \{t_i,y_i,\sigma_i\}_{i=1}^N
-\end{equation}
-
-without loss of generality, we will assume throughout that the $y_i$ values are centered; that is, the weighted mean
-
-\begin{equation}
-  \label{eq:centered-data}
-  \frac{\sum_{i=1}^N w_i y_i}{\sum_{i=1}^N w_i} = 0
-\end{equation}
-
-where the weights are $w_i \equiv \sigma_i^{-2}$. We will seek a single sinusoid model which fits this data; that is, our model is
-
-\begin{equation}
-M(t~|~A,\omega,\phi) = A \sin(\omega t + \phi)
-\end{equation}
-
-to make this a linear problem,
-
-\begin{equation}
-M(t~|~\omega,\theta) = \theta_0 \sin(\omega t) + \theta_1\cos(\omega t)
-\end{equation}
-
-which can be related to the above form by $\theta_0 = A\cos\phi$ and $\theta_1 = A\sin\phi$.
-The $\chi^2$ for this model is given by
-
-\begin{equation}
-\chi^2(D, \omega, \theta) = \sum_i\frac{[y_i - M(t_i~|~\omega,\theta)]^2}{2\sigma_i^2}
-\end{equation}
-
-This can be expressed more compactly by defining the following matrices:
-
-\begin{equation}
-X_\omega = \left[
-\begin{array}{cc}
-\sin(\omega t_1) & \cos(\omega t_1)\\
-\sin(\omega t_2) & \cos(\omega t_2)\\
-\vdots & \vdots \\
-\sin(\omega t_N) & \cos(\omega t_N)\\
-\end{array}
-\right];~~
-y = \left[
-\begin{array}{c}
-y_1 \\
-y_2\\
-\vdots \\
-y_N\\
-\end{array}
-\right];~~
-\Sigma = \left[
-\begin{array}{ccccc}
-\sigma_1^2 & 0 & 0 & \cdots & 0\\
-0 & \sigma_2^2 & 0 & \cdots & 0\\
-0 & 0 & \sigma_3^2 & \cdots & 0\\
-\vdots & \vdots & \vdots & \ddots & \vdots\\
-0 & 0 & 0 & \cdots & \sigma_N^2
-\end{array}
-\right]
-\end{equation}
-
-Then the model is given by $X_\omega\theta$ and the $\chi^2$ can be written
-compactly as:
-
-\begin{equation}
-  \chi^2(\omega, \theta) = (y - X_\omega\theta)^T\Sigma^{-1}(y - X_\omega\theta)
-\end{equation}
-
-Minimizing this cost funciton with respect to $\theta$ gives the maximum likelihood parameters:
-
-\begin{equation}
-\hat{\theta}_\omega = (X_\omega^T\Sigma^{-1}X_\omega)^{-1}X_\omega^T\Sigma^{-1}y
-\end{equation}
-
-We'll simplify this slightly by defining the quantities $\tilde{X}_\omega = \Sigma^{-1/2}X_\omega$ and $\tilde{y} = \Sigma^{-1/2}y$, so that the maximum likelihood parameters can be written
-
-\begin{equation}
-  \hat{\theta}_\omega = (\tilde{X}_\omega^T\tilde{X}_\omega)^{-1}\tilde{X}_\omega^T\tilde{y}
-\end{equation}
-
-Now the value of $\chi^2$ at maximum likelihood can be expressed:
-
-\begin{equation}
-  \chi^2(\omega, \hat{\theta}_\omega) = \tilde{y}^T\tilde{y} - \tilde{y}^T\tilde{X}_\omega(\tilde{X}_\omega^T\tilde{X}_\omega)^{-1}\tilde{X}_\omega^T \tilde{y}
-\end{equation}
-
-The form of this equation suggests defining a reference $\chi^2$ value which is given by a best-fit constant model to the data: because of the constraint given in Eqn.~\ref{eq:centered-data}, this reference is simply written $\chi^2_0 = \tilde{y}^T\tilde{y}$.
-We can now follow Lomb [ref] \& Scargle [ref] and write the normalized periodogram:
-
-\begin{eqnarray}
-  \label{eq:power}
-  P(\omega) &\equiv& 1 -\frac{\chi^2(\omega, \hat{\theta}_\omega)}{\chi^2_0(\omega)}\\ &=& \frac{\tilde{y}^T\tilde{X}_\omega(\tilde{X}_\omega^T\tilde{X}_\omega)^{-1}\tilde{X}_\omega^T \tilde{y}}{\tilde{y}^T\tilde{y}}
-\end{eqnarray}
-
-Typically, the Lomb-Scargle periodogram is expressed in terms of various weighted sums of sines and cosines of the data; this expression is equivalent, but has the advantage that it can be trivially generalized to an arbitrary linear model through adding columns to the $X$ matrix, and can handle arbitrary correlated measurement errors by adding appropriate off-diagonal terms to the $\Sigma$ matrix.
-
-\section{Higher-order Models}
-
-As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of \citep{Zechmeister09}.
-This adjusts the classic Lomb-Scargle algorithm by using a model with a floating mean, which can be more accurate for certain observing cadences:
-
-\begin{equation}
-  M(t~|~\omega, \theta) = \theta_0 + \theta_1\sin\omega t + \theta_2\cos\omega t
-\end{equation}
-
-The normalized power under this model can be computed by simply adding a column of ones to the $X$ matrix.
-
-In a similar vein, the power for a truncated Fourier series model of any order can be constructed by extending the $X$ matrix with appropriate columns; for example, for a 2-term truncated Fourier fit with a floating mean we can write
-
-\begin{equation}
-X_\omega = \left[
-\begin{array}{ccccc}
-1 & \sin(\omega t_1) & \cos(\omega t_1) & \sin(2\omega t_1) & \cos(2\omega t_1)\\
-1 & \sin(\omega t_2) & \cos(\omega t_2) & \sin(2\omega t_2) & \cos(2\omega t_2)\\
-1 & \sin(\omega t_3) & \cos(\omega t_3) & \sin(2\omega t_3) & \cos(2\omega t_3)\\
-\vdots & \vdots & \vdots & \vdots & \vdots \\
-1 & \sin(\omega t_N) & \cos(\omega t_N) & \sin(2\omega t_N) & \cos(2\omega t_N)\\
-\end{array}
-\right]
-\end{equation}
-
-Plugging this $X$ matrix into Eqn.~\ref{eq:power} will give us the normalized periodogram associated with the two-term truncated Fourier model.
-
-\section{Regularized Models}
-
-(Show how regularization works its way through the equations \& derive expression for power).
+%\appendix
 

### 2014-12-12 writeup/figures/fig_rrlyrae_sim.py

@@ -38,7 +38,7 @@ mags = np.array([rrlyrae.generated(band, t, err=dy, corrected=False)
 # First figure:
 # Compute the lomb-scargle periodogram in each band
 
-periods = np.linspace(0.2, 0.9, 10000)
+periods = np.linspace(0.2, 0.9, 1000)
 omegas = 2 * np.pi / periods
 P = [LombScargleAstroML().fit(t, m, dy).power(omegas) for m in mags]
 
@@ -57,7 +57,7 @@ ax[0].set_ylabel('magnitude')
 for i, (Pi, band) in enumerate(zip(P, 'ugriz')):
     offset = 4 - i
     ax[1].plot(periods, Pi + offset, lw=1)
-    ax[1].text(0.89, 0.98 + offset, band, fontsize=14, ha='right', va='top')
+    ax[1].text(0.89, 0.7 + offset, band, fontsize=14, ha='right', va='top')
 ax[1].set_title('Lomb-Scargle in Each Band')
 ax[1].set_ylim(0, 5)
 ax[1].yaxis.set_major_formatter(plt.NullFormatter())

### 2014-12-12 writeup/paper.tex
add more writeup

@@
 %            filecolor=linkcolor,urlcolor=linkcolor}
+%\usepackage{amssymb,amsmath}
 
@@
 \usepackage{url}
-\usepackage{algorithmic,algorithm}
-\usepackage{amssymb,amsmath}
 \usepackage{graphicx}
@@
 
+\newcommand{\todo}[1]{{\color{red} [TODO: #1]}}
 \newcommand{\foreign}[1]{{\it #1}}
@@
 \begin{abstract}
-  This paper introduces a general method for least-squares spectral fitting of multi-band periodic data.
+  This paper introduces the {\it multiband periodogram} a general extension of the well-known Lomb-Scargle method for detecting periodic signals in time-domain data.
 \end{abstract}
@@
 
-(Mention supersmoother and ARMA-based methods?)
+\todo{Add a paragraph about Supersmoother \citep{Reimann94}, CARMA \citep{Kelly14}, and other methods}
 
@@
 \begin{equation}
-  \eqlabel{chi2_PN}
+  \eqlabel{chi2PN}
   \chi_{min}^2(\omega) = \chi^2_0[1 - P_N(\omega)]
@@
 
-the normalized periodogram can be computed by finding the minimum of $\chi^2(\omega)$ via standard methods and plugging the result into \eq{chi2_PN} to find
-
+the normalized periodogram can be computed by finding the minimum of $\chi^2(\omega)$ via standard methods and plugging the result into \eq{chi2PN} to find
 \begin{equation}
   \eqlabel{LombScargle2}
-  P_N(\omega) = \frac{y^T\Sigma^{-1}X_\omega(X_\omega^T\Sigma^{-1}X_\omega)^{-1}X_\omega^T\Sigma^{-1}y}{y^T\Sigma^{-1}y}
+  P_N(\omega) = \frac{y^T\Sigma^{-1}X_\omega~[X_\omega^T\Sigma^{-1}X_\omega]^{-1}~X_\omega^T\Sigma^{-1}y}{y^T\Sigma^{-1}y}.
 \end{equation}
+This expression is equivalent to \eq{LombScargle} in the case that $\Sigma \propto I$.
 
-For $\Sigma \propto I$, this expression is equivalent to \eq{LombScargle}.
-
-While \eq{LombScargle} can be computed more efficiently than \eq{LombScargle2}, the matrix-based formulation has several advantages
+While the expression in \eq{LombScargle} can be computed more efficiently than that in \eq{LombScargle2}, the matrix-based formulation has several advantages:
 \begin{enumerate}
@@
 \end{enumerate}
+In the remainder of this section, we'll explore a few of these modifications.
 
 \subsection{Lomb-Scargle with Floating Mean}
-
-As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of \citep{Zechmeister09}.
+As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of \citet{Zechmeister09}.
 This adjusts the classic Lomb-Scargle algorithm by using a model with a floating mean, which can be more accurate for certain observing cadences:
@@
 
-While \citep{Zechmeister09} details a formalism to express this {\it generalized periodogram} model similarly to \eq{LombScargle}, in the matrix formalism all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
+While \citet{Zechmeister09} details a formalism to express this {\it generalized periodogram} model similarly to \eq{LombScargle}, in the matrix formalism all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
+
+\todo{example of this in action?}
 
 \subsection{Truncated Fourier Models}
-In a similar vein, the power for a truncated Fourier series model of any order can be constructed by extending the $X$ matrix with appropriate columns; for example, for a 2-term truncated Fourier fit with a floating mean we can write
+\sectlabel{multiterm}
+The classic Lomb-Scargle approach is equivalent to fitting a single-term stationary sinusoidal model to the data.
+A natural extension is to use a multiple-term sinusoidal model.
+With $M$ terms, there are $2M + 1$ free parameters, and the model is given by:
 
 \begin{equation}
-X_\omega = \left[
+  M(t|\omega,\theta) = \theta_0 + \sum_{n=1}^M \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]
+\end{equation}
+
+Because this remains a linear model, it can be easily accommodated into the matrix formalism above.
+For example, an $M = 2$-term model can be constructed by building a design matrix $X_\omega$ with $2M + 1 = 5$ columns:
+
+\begin{equation}
+X_\omega^{(2)} = \left[
 \begin{array}{ccccc}
@@
 \end{equation}
+Using $X_\omega^{(2)}$ in \eq{LombScargle2} will give the two-term periodogram. In general, we can fit an $M$-term periodogram by simply adding terms to the fit.
 
-Plugging this $X$ matrix into Eqn.~\ref{eq:power} will give us the normalized periodogram associated with the two-term truncated Fourier model.
+\todo{Show a truncated Fourier model?}
 
 \subsection{Regularized Models}
+\sectlabel{regularization}
+The previous sections raise the question: how complicated a model should we use?
+If we added many terms to the truncated Fourier series, we might fit the data perfectly, but eventually we would {\it overfit} the data: that is, our model is fitting the noise rather than the signal.
+This can be avoided by explicitly truncating the series, but another approach is to use a {\it regularization} term to enforce a less complicated model.
 
-Show how regularized models fit into this.
+A regularization term is an explicit penalty on the magnitude of the model parameters $\theta$, and can take any number of forms.
+For computational simplicity here we'll use an {\it L2 regularization} (also known as Tikhonov Regularization), which is a quadratic penalty term in the model parameters.
+This is mathematically equivalent to the Bayesian approach of using a Gaussian prior on the model parameters.
+
+We'll encode our regularization in the matrix $\Lambda = {\rm diag}([\lambda_1, \lambda_2 \cdots \lambda_M])$ for a model with $M$ parameters, and our expression for $\chi^2$ becomes
+\begin{equation}
+  \eqlabel{chi2reg}
+  \chi_\Lambda^2(\omega) = (y - X_\omega\theta)^T\Sigma^{-1}(y - X_\omega\theta) + \theta^T\Lambda\theta
+\end{equation}
+Minimizing this $\chi^2$, solving for $\theta$, and plugging into the expression for $P_N$ gives us the regularized counterpart of \eq{LombScargle2}:
+\begin{equation}
+  \eqlabel{LombScargleReg}
+  P_N(\omega) = \frac{y^T\Sigma^{-1}X_\omega~[X_\omega^T\Sigma^{-1}X_\omega + \Lambda]^{-1}~X_\omega^T\Sigma^{-1}y}{y^T\Sigma^{-1}y}.
+\end{equation}
+Note that the effect of this regularization term is to add a diagonal term to the matrix inversion in the numerator, which also has the effect of correcting ill-posed models.
+This feature of the regularization will become important below.
+
+\todo{Show a regularized model?}
 
@@
 
-\begin{itemize}
-  \item Write-out the linear model
-  \item Demonstrate the matrix form of this
-  \item Use light regularization
-  \item Correspondance of the single-term multiband result to the weighted sum of single-band Lomb-Scargle results
-\end{itemize}
+To compute a periodogram for multi-band data, we'll take advantage of many of the nice features of the matrix form of the normalized periodogram, which allows us to specify an arbitrary linear model covering our data.
+We will construct a multi-term Fourier model with the following features:
+\begin{enumerate}
+  \item An $M_{base}$-term truncated Fourier fit which models a latent parameter, which we'll call the ``overall variability''.
+  \item An $M_{band}$-term truncated Fourier fit which models the residual of each band from the overall variability.
+\end{enumerate}
+The total number of parameters for $N_{filt}$ filters is then $M = (2M_{base} + 1) + N_{filt}(2M_{band} + 1)$. The model looks as follows, for a band $b$:
+\begin{eqnarray}
+  M(t|\omega,\theta,b) = &&\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]\\ 
+  &+& \theta^{(b)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(b)}_{2n - 1}\sin(n\omega t) + \theta^{(b)}_{2n}\cos(n\omega t)\right].
+\end{eqnarray}
+The important feature is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(b)}$ are fit within each band.
+
+We can build the matrix form of this model by building a sparse $X_{\omega}$ matrix with $M$ columns.
+Each row corresponds to a single observation through a single band; columns corresponding to the base model and the observation band will have nonzero entries; all other columns will be filled with zeros.
+For example, the $X_\omega$ matrix with $N_{base}=1$ and $N_{band}=0$ (that is, a single sinusoid model with independent offsets to each band) will look as follows:
+
+\begin{equation}
+X_\omega^{(1,0)} = \left[
+\begin{array}{cccccccc}
+1 & \sin(\omega t_1) & \cos(\omega t_1) & 1 & 0 & 0 & 0 & 0\\
+1 & \sin(\omega t_2) & \cos(\omega t_2) & 0 & 1 & 0 & 0 & 0\\
+1 & \sin(\omega t_3) & \cos(\omega t_3) & 1 & 0 & 0 & 0 & 0\\
+1 & \sin(\omega t_4) & \cos(\omega t_4) & 0 & 0 & 1 & 0 & 0\\
+\vdots & \vdots & \vdots & & & \vdots & &\\
+1 & \sin(\omega t_N) & \cos(\omega t_N) & 0 & 0 & 0 & 0 & 1\\
+\end{array}
+\right]
+\end{equation}
+Here the structure of the final five columns indicates the band observed at the given time: the first row is a $u$-band measurement, the second is a $g$-band, the third is a $u$-band, etc.
+
+From reasoning about the free parameters, or by examining the above matrix it is easy to see that some of the parameters will be degenerate (i.e.
+$X_\omega$ is low-rank).
+Intuitively, this is due to the fact that if we add an overall offset to the base model, this can be perfectly accounted for by subtracting that same offset from each of the band columns.
+In order to proceed, then, we'll either have to correct this model deficiency, or use a regularization term on one of the offending parameters.
+We'll choose the latter here, and regularize all the band columns to drive the common behavior into the base model.
+For the above matrix,
+this regularization will look like
+\begin{equation}
+  \Lambda = {\rm diag}([0, 0, 0, \epsilon, \epsilon, \epsilon, \epsilon, \epsilon])
+\end{equation}
+where $\epsilon$ is some small fraction of the trace of the matrix $[X_\omega^T\Sigma^{-1}X_\omega]$.
+With this regularization in place, the model is well-posed and \eq{LombScargleReg} can be used to straightforwardly compute the power.
+We'll see an example of this in action in \sect{Simulated}.
+
+\todo{Talk about the $N_{base} = 0$, $N_{band} = 1$ case.
+This is equivalent to fitting a floating-mean Lomb-Scargle independently to each band, and constructing a weighted sum of each individual periodogram.}
 
 \section{Examples}
+Here we'll show several examples of the multiband periodogram method in action.
+
 \subsection{Simulated Example}
+\sectlabel{Simulated}
 
@@
     in Lomb-Scargle periodograms, located at beat frequencies between the
-    true Period and the 1-day observing cadence.
+    true period and the 1-day observing cadence (see footnote \ref{foot1}).
   }
-  \figlabel{01}
+  \figlabel{adhoc_example}
 \end{figure}
+
+\fig{adhoc_example} shows a simulated RR Lyrae light curve based on a template from \citet{Sesar2010}.
+The observations take place on 60 random nights over a 6-month period, and each night all five bands ({\it u,g,r,i,z}) are recorded.
+The left panel of \fig{adhoc_example} shows the observations folded as a function of phase.
+Using the typical approach from the literature prior to this paper, we individually compute the Lomb-Scargle periodogram within each band: the results are shown in the right panel.
+Here the data is well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band.\footnote{\label{foot1}
+  An aliasing effect is readily apparent in the left panel of \fig{adhoc_example}: these additional peaks are due to beat frequencies between the true period $P$ and the observing cadence of $\sim 1$ day.
+That is, for nightly observations, we'd expect excess power to be found at frequencies $P / (1 + nP)$ for a positive or negative integer $n$.
+For the current work, we'll ignore this (important) piece of the problem; for a thorough discussion of aliasing within astronomical time series, see \citet{Roberts87}.}
 
@@
   \caption{
-    Another realization of the data from \fig{01}, with only a single filter
-    observed each night (i.e. 12 observations per band, spread over 180
+    Another realization of the data from \fig{adhoc_example}, with only a single
+    band observed each night (i.e. 12 observations per band, spread over 180
     days). In this case, single-band data is not sufficient to recover the
     Lomb-Scargle peaks, and the classic band-by-band approach fails.
-    By utilizing all the data at once with the multiband approach,
-    we recover a significant spike of power at the period $P=0.622$ days.    
+    By utilizing all the data at once with the multiband approach
+    (with $N_{base}=1,~N_{band}=0$),
+    we faithfully reconstruct the periodogram and recover
+    the dominant power at period $P=0.622$ days.
   } 
-  \figlabel{02}
+  \figlabel{multi_example}
 \end{figure}
 
-\begin{itemize}
-  \item Show power spectrum for a single-band RR-Lyrae
-  \item Reduce observations, show multiple power spectra for multi-band RR-Lyrae
-  \item Show single power spectrum for multi-band RR-Lyrae
-\end{itemize}
+\fig{multi_example} shows the same 60 nights of data, except that each night only a {\it single} band observation is recorded.
+The left panel again shows the observations as a function of phase, and the right panel shows the Lomb-Scargle fits to the data.
+Now, with only 12 observations for each indiviudal band, there is not enough data to accurately determine the period band-by-band.
+With the multiband approach (here computed for $N_{base}=1$ and $N_{band}=0$), as seen in the lower-right panel of \fig{multi_example}, the period is cleanly recovered.
 
@@
 \begin{itemize}
-  \item issues with window function corrections
+  \item issues with aliasing \& window function corrections
   \item issues with physicality of model

### 2014-12-12 writeup/figures/fig_rrlyrae_sim.py

@@ -23,14 +23,14 @@ rrlyrae = RRLyraeObject(lcid, random_state=0)
 
 # Generate data in a 6-month observing season
 Nobs = 60
-dy = 0.07
 rng = np.random.RandomState(0)
 
 nights = np.arange(180)
 rng.shuffle(nights)
 nights = nights[:Nobs]
 
-t = 57000 + nights + 0.2 * rng.rand(Nobs)
+t = 57000 + nights + 0.05 * rng.randn(Nobs)
+dy = 0.06 + 0.01 * rng.randn(Nobs)
 mags = np.array([rrlyrae.generated(band, t, err=dy, corrected=False)
                  for band in 'ugriz'])
 
@@ -77,7 +77,7 @@ mags = mags[np.arange(Nobs) % 5, np.arange(Nobs)]
     
 masks = [(filts == band) for band in 'ugriz']
 
-P = [LombScargleAstroML().fit(t[mask], mags[mask], dy).power(omegas)
+P = [LombScargleAstroML().fit(t[mask], mags[mask], dy[mask]).power(omegas)
      for mask in masks]
     
 LS_multi = LombScargleMultiband(Nterms_base=1, Nterms_band=0)
@@ -93,7 +93,7 @@ ax = [fig.add_subplot(gs[:, 0]),
       fig.add_subplot(gs[3, 1])]
     
 for band, mask in zip('ugriz', masks):
-    ax[0].errorbar((t[mask] / rrlyrae.period) % 1, mags[mask], dy,
+    ax[0].errorbar((t[mask] / rrlyrae.period) % 1, mags[mask], dy[mask],
                    fmt='.', label=band)
 ax[0].set_ylim(18, 14.5)
 ax[0].legend(loc='upper left', fontsize=12, ncol=3)
@@ -118,5 +118,16 @@ ax[2].yaxis.set_major_formatter(plt.NullFormatter())
 ax[2].set_xlabel('Period (days)')
 
 fig.savefig('fig02.pdf')
+
+#----------------------------------------------------------------------
+# Write the results to file
+import pandas as pd
+df = pd.DataFrame({'t': t})
+df['band'] = filts
+df['mag'] = mags
+df['dmag'] = dy
+df = df.sort('t')
+
+df.to_csv('fig01data.csv', sep=' ', float_format='%.4f', index=False)
 
 plt.show()

### 2014-12-15 writeup/paper.tex
add some introductory figures

@@
 
-Because astronomical observing cadences are rarely so uniform, many have looked at extending the concept of the periodogram.
+Because astronomical observing cadences are rarely so uniform, many have looked at extending the ideas behind the periodogram to work with unevenly-sampled data.
 Most famously, \citep{Lomb76} and \citep{Scargle82} extended earlier work to define the {\it normalized periodogram}:
@@
 
-This deep connection between spectral power and least squares fitting methods was solidified by \citep{Jaynes87}, who demonstrated that the normalized periodogram of Lomb and Scargle is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
+This long-recognized connection between spectral power and least squares fitting methods was solidified by \citep{Jaynes87}, who demonstrated that the normalized periodogram of Lomb and Scargle is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
 Building on this result, \citep{Bretthorst88} explored the extension of these methods to more complicated models with multiple frequency terms, nonstationary frequencies, and other more sophisticated models within a Bayesian framework.
@@
 \subsection{Stationary Sinusoid Model}
-The Lomb-Scargle periodogram of \eq{LombScargle} can be derived from the normalized $\chi^2$ of the best-fit single-term stationary sinusoidal model given in \eq{SingleModel}.
+
+The normalized periodogram of \eq{LombScargle} can be derived from the normalized $\chi^2$ of the best-fit single-term stationary sinusoidal model given in \eq{SingleModel}.
 To make the problem linear, we can re-express the model in terms of the parameter vector $\theta = [A\cos\phi, A\sin\phi]$ so that our model is
@@
   \eqlabel{simplemodel}
-  M(t|\omega,\theta) = \theta_1\sin(\omega t) + \theta_2\cos(\omega t).
+  y(t|\omega,\theta) = \theta_1\sin(\omega t) + \theta_2\cos(\omega t).
 \end{equation}
@@
 \begin{equation}
-  \chi^2(\omega) = \sum_k \frac{(y_k - M(t_k|\omega,\theta))^2}{2\sigma_k^2}.
+  \chi^2(\omega) = \sum_k \frac{[y_k - y(t_k|\omega,\theta)]^2}{2\sigma_k^2}.
 \end{equation}
@@
 where $P_N(\omega)$ is the normalized periodogram given in \eq{LombScargle} and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$
+
+\subsubsection{Matrix Formalism}
 
@@
 
-With these definitions, the model in \eq{simplemodel} is a simple linear product: $M(t|\omega,\theta) = X_\omega\theta$ and the model and reference $\chi^2$ can be written
+With these definitions, the model in \eq{simplemodel} is a simple linear product: $y(t|\omega,\theta) = X_\omega\theta$ and the model and reference $\chi^2$ can be written
 
@@
 \end{eqnarray}
-
-the normalized periodogram can be computed by finding the minimum of $\chi^2(\omega)$ via standard methods and plugging the result into \eq{chi2PN} to find
+The normalized periodogram can be computed by finding the minimum of $\chi^2(\omega)$ via standard methods and plugging the result into \eq{chi2PN} to find
 \begin{equation}
@@
 
-While the expression in \eq{LombScargle} can be computed more efficiently than that in \eq{LombScargle2}, the matrix-based formulation has several advantages:
+\fig{basic_example} shows an example periodogram for a simulated RR Lyrae light curve based on a template from \citet{Sesar2010}.
+The particular RR Lyrae star has a period of 0.622 days, and the observations take place on 60 random nights over a 6-month period.
+The left panel of the figure shows the input data, and the right panels show the result of a simple Lomb-Scargle periodogram on the data.
+
+\subsubsection{Simple Period Finding}
+\sectlabel{simple_period}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig01.pdf}
+  \caption{
+    An illustration of the basic periodogram and its relationship to the single-term sinusoid model.
+The left panel shows the input data, while the right panels show the fit derived from the data.
+The top-right panel shows the periodogram, showing a distinct peak at the true period of 0.622 days, and the bottom-right panel shows the data as a function of phase as defined by this period.
+Note in the periodogram the presence of the typical aliasing effect, with power located at beat frequencies between the true period and the 1-day observing cadence (see \sect{simple_period} for further discussion).
+  }
+  \figlabel{basic_example}
+\end{figure}
+
+In the upper right, we see the normalized periodogram as a function of period.
+While the power does peak at the true period of 0.622 days, an aliasing effect is apparent.
+These additional peaks are due to beat frequencies between the true period $P$ and the observing cadence of $\sim 1$ day.
+That is, for nightly observations, we'd expect excess power to be found at frequencies $P / (1 + nP)$ for a positive or negative integer $n$.
+This explains the prominent peak at $0.383$.
+For the current work, we'll ignore this (important) piece of the period-finding problem; for a thorough discussion of aliasing within astronomical time series, see \citet{Roberts87}.
+
+The lower-right panel shows the maximum likelihood interpretation of this periodogram: it is a measure of the normalized $\chi^2$ for a single-term sinusoidal model.
+Here we visualize the data from the left panel, but folded as a function of phase with the best-fit single-term model over-plot.
+Here its apparent that the single-term model is biased: RR Lyrae are much more complicated than a simple sinusoid.
+Nevertheless, the simplistic sinusoidal model does well to zero-in
+
+\subsubsection{Discussion}
+We have shown two forms of the classic normalized periodogram: \eq{LombScargle} and \eq{LombScargle2}.
+Though the two expressions are equivalent, they each have distinct advantages.
+The expression in \eq{LombScargle} avoids the explicit construction of a matrix, and thus can be computed very efficiently.
+In particular, through some computational tricks involving the FFT, expressions of the form of \eq{LombScargle} can be evaluated for $N$ frequencies in $\mathcal{O}[\log{N}]$ time \citep{Press89}.
+
+The matrix-based formulation of \eq{LombScargle2} sacrifices some potential computational efficiency in order to attain several advantages:
 \begin{enumerate}
-  \item It is trivially extended for heteroscedastic and/or uncorrelated noise in the data $y_k$ by appropriately modifying the noise matrix $\Sigma$
-  \item It is trivially extended to more sophisticated models by appropriately modifying the design matrix $X_\omega$.
-  \item It is trivially extended to L2-regularized likelihoods by adding an appropriate diagonal term to $X_\omega^T\Sigma^{-1}X_\omega$ before inverting.
+  \item It is trivially extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the noise matrix $\Sigma$
+  \item It is trivially extended to more sophisticated linear models by appropriately modifying the design matrix $X_\omega$.
+  \item It is trivially extended to L2-regularized likelihoods by adding an appropriate diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$.
 \end{enumerate}
@@
 
-\subsection{Lomb-Scargle with Floating Mean}
+\subsection{Stationary Sinusoid with Floating Mean}
+\sectlabel{floating_mean}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig02.pdf}
+  \caption{
+    An illustration of the effect of the floating mean model for biased data.
+    The data consist of 80 observations drawn from a sinusoidal model; all observations with magnitude fainter than 16 are masked, as shown in the left panel.
+The standard and floating-mean periodograms are computed from the unmasked data; these are shown in the right panel.
+Because of this biased observing pattern, the mean of the observed data is not a good predictor of the true mean, and the standard model fails to recover the true period of 0.622 days, while the floating mean model still finds the correct period.
+  }
+  \figlabel{floating_mean}
+\end{figure}
+
 As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of \citet{Zechmeister09}.
@@
 \begin{equation}
-  M(t~|~\omega, \theta) = \theta_0 + \theta_1\sin\omega t + \theta_2\cos\omega t
+  y(t~|~\omega, \theta) = \theta_0 + \theta_1\sin\omega t + \theta_2\cos\omega t
 \end{equation}
 
-While \citet{Zechmeister09} details a formalism to express this {\it generalized periodogram} model similarly to \eq{LombScargle}, in the matrix formalism all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
+While \citet{Zechmeister09} details a formalism to express this floating-mean model (which they call a ``generalized periodogram'') similarly to \eq{LombScargle}, in the matrix formalism all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
 
-\todo{example of this in action?}
+For well-sampled data, there is little difference between a standard periodogram and a floating-mean periodogram.
+Where this becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases: an example of this situation is shown in \fig{floating_mean}.
+The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias.
+Because of this bias, the mean of the observed data to be a poor predictor of the true mean, which makes it so that the standard approach fails to find a suitable fit to the data.
+The floating-mean approach is able to adjust for this, resulting in a periodogram which detects the input period of 0.622 days.
 
@@
 \sectlabel{multiterm}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig03.pdf}
+  \caption{
+    An illustration of a periodogram based on the truncated Fourier model. The data are the same as those in \fig{basic_example}. 
+  }
+  \figlabel{multiterm_example}
+\end{figure}
+
 The classic Lomb-Scargle approach is equivalent to fitting a single-term stationary sinusoidal model to the data.
@@
 \begin{equation}
-  M(t|\omega,\theta) = \theta_0 + \sum_{n=1}^M \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]
+  y(t|\omega,\theta) = \theta_0 + \sum_{n=1}^M \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]
 \end{equation}
@@
 \end{equation}
-Using $X_\omega^{(2)}$ in \eq{LombScargle2} will give the two-term periodogram. In general, we can fit an $M$-term periodogram by simply adding terms to the fit.
+Using $X_\omega^{(2)}$ in \eq{LombScargle2} will result in a two-term periodogram.
+In general, we can fit an $M$-term periodogram by simply adding terms to the fit.
+\fig{multiterm_example} shows a few examples of this multiterm Fourier approach as applied to a simulated RR Lyrae light curve constructed from templates derived in \citet{Sesar2010}.
+There are several important features of this figure that give us insight into the subtleties of this type of fit:
 
-\todo{Show a truncated Fourier model?}
+First, we see in the right panel that all three models show a significant spike at the true period of $P_0 = 0.622$.
+The higher-order models, however, also show a a spike of poer at $P_1 = 2 P_0$: the reason for this is that the fundamental model of the single-term model at $P_0$ is the first harmonic of the 2-term model at $2P_0$, so the higher-order models contain this single-term result.
+
+Second, notice that as the number of terms is increased, the general ``background'' level of the periodogram increases.
+This is due to the fact that the periodogram is directly related to the $\chi^2$ of the fit.
+A more complicated model will have the flexibility to conform to the data at {\it all frequencies}, and thus the observed ``power'' is stronger at all frequencies.
 
@@
 \end{equation}
-Note that the effect of this regularization term is to add a diagonal term to the matrix inversion in the numerator, which also has the effect of correcting ill-posed models.
+Note that the effect of this regularization term is to add a diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$, which has the additional feature that it corrects ill-posed models.
 This feature of the regularization will become important below.
@@
 \begin{eqnarray}
-  M(t|\omega,\theta,b) = &&\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]\\ 
+  y(t|\omega,\theta,b) = &&\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]\\ 
   &+& \theta^{(b)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(b)}_{2n - 1}\sin(n\omega t) + \theta^{(b)}_{2n}\cos(n\omega t)\right].
@@
   \centering
-  \includegraphics[width=\textwidth]{fig01.pdf}
+  \includegraphics[width=\textwidth]{fig04.pdf}
   \caption{
@@
     band, individual periodograms can be constructed and compared to find the
-    true period $P=0.622$ days. Note also the presence of aliasing typical
-    in Lomb-Scargle periodograms, located at beat frequencies between the
-    true period and the 1-day observing cadence (see footnote \ref{foot1}).
+    true period $P=0.622$ days
   }
@@
   \centering
-  \includegraphics[width=\textwidth]{fig02.pdf}
+  \includegraphics[width=\textwidth]{fig05.pdf}
   \caption{
@@
 
-\section{Further work}
+\section{Discussion and Conclusion}
 
+\subsection{Further Study}
 \begin{itemize}

### 2014-12-15 writeup/figures/fig_basic_example.py

@@ -1,0 +1,85 @@
+"""
+Here we plot a typical approach to the basic periodogram
+"""
+
+import sys
+import os
+sys.path.append(os.path.abspath('../..'))
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Use seaborn settings for plot styles
+import seaborn; seaborn.set()
+
+from multiband_LS.generated import RRLyraeObject
+from multiband_LS.lomb_scargle import LombScargle
+
+
+# Choose a Sesar 2010 object to base our fits on
+lcid = 1019544
+rrlyrae = RRLyraeObject(lcid, random_state=0)
+
+# Generate data in a 6-month observing season
+Nobs = 60
+rng = np.random.RandomState(0)
+
+nights = np.arange(180)
+rng.shuffle(nights)
+nights = nights[:Nobs]
+
+t = 57000 + nights + 0.05 * rng.randn(Nobs)
+dmag = 0.06 + 0.01 * rng.randn(Nobs)
+mag = rrlyrae.generated('r', t, err=dmag, corrected=False)
+omega_best = 2 * np.pi / rrlyrae.period
+
+periods = np.linspace(0.2, 1.4, 1000)
+omegas = 2 * np.pi / periods
+
+phase = (t / rrlyrae.period) % 1
+phasefit = np.linspace(0, 1, 1000)
+tfit = rrlyrae.period * phasefit
+
+fig = plt.figure(figsize=(10, 4))
+gs = plt.GridSpec(2, 2, left=0.07, right=0.95,
+                  wspace=0.15, hspace=0.7,
+                  bottom=0.15)
+ax = [fig.add_subplot(gs[:, 0]),
+      fig.add_subplot(gs[1, 1]),
+      fig.add_subplot(gs[0, 1])]
+
+# Plot the data
+ax[0].errorbar(t, mag, dmag, fmt='o', color='#333333')
+ax[1].errorbar(phase, mag, dmag, fmt='.', color='#888888')
+
+# Fit and plot the model
+model = LombScargle().fit(t, mag, dmag)
+
+phase = (t / model.best_period) % 1
+phasefit = np.linspace(0, 1, 1000)
+tfit = model.best_period * phasefit
+P = model.periodogram(omegas)
+        
+lines = ax[2].plot(periods, model.periodogram(omegas), lw=1)
[…]

### 2014-12-15 writeup/figures/fig_floating_mean.py

@@ -1,0 +1,81 @@
+"""
+Illustration of a floating-mean periodogram
+"""
+
+import sys
+import os
+sys.path.append(os.path.abspath('../..'))
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Use seaborn settings for plot styles
+import seaborn; seaborn.set()
+
+from multiband_LS.generated import RRLyraeObject
+from multiband_LS.lomb_scargle import LombScargle
+
+
+# Choose a Sesar 2010 object to base our fits on
+lcid = 1019544
+rrlyrae = RRLyraeObject(lcid, random_state=0)
+
+# Generate data in a 6-month observing season
+Nobs = 60
+rng = np.random.RandomState(0)
+
+nights = np.arange(180)
+rng.shuffle(nights)
+nights = nights[:Nobs]
+
+t = 57000 + nights + 0.1 * rng.randn(Nobs)
+dmag = 0.06 + 0.01 * rng.randn(Nobs)
+mag = (16 + 0.5 * np.sin(2 * np.pi * t / rrlyrae.period - 0.4)
+       + dmag * rng.randn(Nobs))
+omega_best = 2 * np.pi / rrlyrae.period
+phase = (t / rrlyrae.period) % 1
+
+mask = mag < 16.05
+
+phasefit = np.linspace(0, 1, 1000)
+tfit = rrlyrae.period * phasefit
+
+periods = np.linspace(0.2, 1.1, 1000)
+omegas = 2 * np.pi / periods
+
+fig, ax = plt.subplots(1, 2, figsize=(10, 4))
+fig.subplots_adjust(left=0.07, right=0.95, wspace=0.1, bottom=0.15)
+
+ax[0].errorbar(phase[mask], mag[mask], dmag[mask], fmt='o',
+               color='#666666')
+ax[0].errorbar(phase[~mask], mag[~mask], dmag[~mask], fmt='o',
+               color='#CCCCCC')
+ax[0].invert_yaxis()
+
+for fit_offset in [False, True]:
+    model = LombScargle(fit_offset=fit_offset).fit(t[mask],
+                                                   mag[mask], dmag[mask])
+    P = model.periodogram(omegas)
+    if fit_offset:
+        label = 'floating mean'
+    else:
+        label = 'standard'
+    ax[0].plot(phasefit, model.predict(tfit, omega=2 * np.pi / rrlyrae.period),
[…]

### 2014-12-15 writeup/figures/fig_multiband_sim.py

@@ -1,6 +1,6 @@
 """
-Here we plot a typical approach to multi-band Lomb-Scargle: treating each band
-separately, and taking a majority vote between the bands.
+Here we plot a typical approach to the multi-band periodogram: treating each
+band separately, and taking a majority vote between the bands.
 """
 
 import sys
@@ -40,7 +40,7 @@ mags = np.array([rrlyrae.generated(band, t, err=dy, corrected=False)
 
 periods = np.linspace(0.2, 0.9, 1000)
 omegas = 2 * np.pi / periods
-P = [LombScargleAstroML().fit(t, m, dy).power(omegas) for m in mags]
+P = [LombScargleAstroML().fit(t, m, dy).periodogram(omegas) for m in mags]
 
 fig, ax = plt.subplots(1, 2, figsize=(10, 4))
 fig.subplots_adjust(left=0.07, right=0.95, wspace=0.1, bottom=0.15)
@@ -58,13 +58,13 @@ for i, (Pi, band) in enumerate(zip(P, 'ugriz')):
     offset = 4 - i
     ax[1].plot(periods, Pi + offset, lw=1)
     ax[1].text(0.89, 0.7 + offset, band, fontsize=14, ha='right', va='top')
-ax[1].set_title('Lomb-Scargle in Each Band')
+ax[1].set_title('Periodogram in Each Band')
 ax[1].set_ylim(0, 5)
 ax[1].yaxis.set_major_formatter(plt.NullFormatter())
 ax[1].set_xlabel('Period (days)')
-ax[1].set_ylabel('Lomb-Scargle Power + offset')
+ax[1].set_ylabel('power + offset')
 
-fig.savefig('fig01.pdf')
+fig.savefig('fig04.pdf')
 
 #----------------------------------------------------------------------
 # Second figure:
@@ -77,12 +77,13 @@ mags = mags[np.arange(Nobs) % 5, np.arange(Nobs)]
     
 masks = [(filts == band) for band in 'ugriz']
 
-P = [LombScargleAstroML().fit(t[mask], mags[mask], dy[mask]).power(omegas)
+P = [LombScargleAstroML().fit(t[mask], mags[mask],
+                              dy[mask]).periodogram(omegas)
      for mask in masks]
     
 LS_multi = LombScargleMultiband(Nterms_base=1, Nterms_band=0)
[…]

### 2014-12-15 writeup/figures/fig_multiterm_example.py

@@ -1,0 +1,83 @@
+"""
+Here we plot a typical approach to multi-band Lomb-Scargle: treating each band
+separately, and taking a majority vote between the bands.
+"""
+
+import sys
+import os
+sys.path.append(os.path.abspath('../..'))
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Use seaborn settings for plot styles
+import seaborn; seaborn.set()
+
+from multiband_LS.generated import RRLyraeObject
+from multiband_LS.lomb_scargle import LombScargle
+
+
+
+# Choose a Sesar 2010 object to base our fits on
+lcid = 1019544
+rrlyrae = RRLyraeObject(lcid, random_state=0)
+
+# Generate data in a 6-month observing season
+Nobs = 60
+rng = np.random.RandomState(0)
+
+nights = np.arange(180)
+rng.shuffle(nights)
+nights = nights[:Nobs]
+
+t = 57000 + nights + 0.05 * rng.randn(Nobs)
+dmag = 0.06 + 0.01 * rng.randn(Nobs)
+mag = rrlyrae.generated('r', t, err=dmag, corrected=False)
+omega_best = 2 * np.pi / rrlyrae.period
+
+periods = np.linspace(0.2, 1.4, 1000)
+omegas = 2 * np.pi / periods
+
+phase = (t / rrlyrae.period) % 1
+phasefit = np.linspace(0, 1, 1000)
+tfit = rrlyrae.period * phasefit
+
+fig, ax = plt.subplots(1, 2, figsize=(10, 4))
+fig.subplots_adjust(left=0.07, right=0.95, wspace=0.1, bottom=0.15)
+
+# Plot the data
+ax[0].errorbar(phase, mag, dmag, fmt='o', color='#AAAAAA')
+
+# Plot the fits
+models = [1, 2, 6]
+
+for i, Nterms in enumerate(models):
+    model = LombScargle(Nterms=Nterms).fit(t, mag, dmag)
+    P = model.periodogram(omegas)
+
+    label = "{0} terms".format(Nterms)
+    if Nterms == 1:
+        label = label[:-1]
+        
+    offset = len(models) - 1 - i
+    lines = ax[1].plot(periods, offset + model.periodogram(omegas), lw=1)
+    ax[1].text(0.22, offset + 0.9, label, ha='left', va='top')
+    ax[0].plot(phasefit, model.predict(tfit, omega=omega_best),
[…]

### 2014-12-15 writeup/paper.tex
update discussion of weighting single fits

@@
 
-\section{Moving to Multiple Bands}
+\section{A Multiple-Band Model}
 
@@
 
-\todo{Talk about the $N_{base} = 0$, $N_{band} = 1$ case.
-This is equivalent to fitting a floating-mean Lomb-Scargle independently to each band, and constructing a weighted sum of each individual periodogram.}
+The final remaining piece is our assumption that the data are centered (i.e.
+\eq{ycentered}).
+This is required so that the simple form of the reference $\chi^2_0$ remains valid.
+For the multiband model, this assumption requires that the data satisfy \eq{ycentered} {\it within each band}.
+
+\subsection{Relationship of Multiband and Single-band approaches}
+There is a special case of the multi-band approach when $N_{base}=0$ and $N_{band}=1$.
+In this case, the $X$ matrix can be arranged to be block-diagonal, and the computation can be split into an independent single-band computation for each of the $K$ bands.
+For band $k$, denote the single-band periodogram as
+\begin{equation}
+  P_N^{(k)}(\omega) = 1 - \frac{\chi^2_{min, k}(\omega)}{\chi^2_{0,k}}
+\end{equation}
+The full multiband periodogram is given by
+\begin{equation}
+  P_N(\omega) = 1 - \frac{\sum_{k=1}^K\chi^2_{min, k}(\omega)}{\sum_{k=1}^K\chi^2_{0,k}}
+\end{equation}
+and it can easily be shown that the $P_N$ can be constructed as a weighted sum of $P_N^{(k)}$:
+\begin{equation}
+  P_N(\omega) = \frac{\sum_{k=1}^K\left[\chi^2_{0,k}P_N^{k}\right]}{\sum_{k=1}^K\chi^2_{0,k}}
+\end{equation}
+where, recall $\chi^2_{0,k}$ is simply the $\chi^2$ of the data in each band about the mean.
+This tells us how we can construct a particular case of the multi-term periodogram from a weighted sum of periodograms in each band.
+Here the weights reflect both how many measurements there are in each band, and how poorly those measurements are fit by a simple constant reference model.
 
@@
 
+\todo{Compare different options of $N_{base}$ and $N_{band}$?}
+
 \subsection{Stripe 82}

### 2014-12-16 writeup/paper.tex
minor paper fixes

@@
 \begin{abstract}
-  This paper introduces the {\it multiband periodogram} a general extension of the well-known Lomb-Scargle method for detecting periodic signals in time-domain data.
+  This paper introduces the {\it multiband periodogram} a general extension of the standard Lomb-Scargle approach to detecting periodic signals in time-domain data.
 \end{abstract}
@@
   \item It is trivially extended to more sophisticated linear models by appropriately modifying the design matrix $X_\omega$.
-  \item It is trivially extended to L2-regularized likelihoods by adding an appropriate diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$.
+  \item It is trivially extended to Tikhonov/L2-regularized likelihoods by adding an appropriate diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$.
 \end{enumerate}
-In the remainder of this section, we'll explore a few of these modifications.
+In the remainder of this section, we'll explore a few of these modifications and how they affect the periodogram and resulting model fits.
 
@@
     An illustration of the effect of the floating mean model for biased data.
-    The data consist of 80 observations drawn from a sinusoidal model; all observations with magnitude fainter than 16 are masked, as shown in the left panel.
-The standard and floating-mean periodograms are computed from the unmasked data; these are shown in the right panel.
+    The data consist of 80 observations drawn from a sinusoidal model.
+To mimick a potentially damaging selection effect, all observations with magnitude fainter than 16 are removed (indicated by the light-gray points).
+The standard and floating-mean periodograms are computed from the remaining data; these fits are shown over the data in the right panel.
 Because of this biased observing pattern, the mean of the observed data is not a good predictor of the true mean, and the standard model fails to recover the true period of 0.622 days, while the floating mean model still finds the correct period.
@@
 As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of \citet{Zechmeister09}.
-This adjusts the classic Lomb-Scargle algorithm by using a model with a floating mean, which can be more accurate for certain observing cadences:
+This adjusts the classic normalized periodogram by allowing the mean of the model to be fit alongside the amplitudes:
 
@@
 
+This model can be more accurate than the standard periodogram for certain observing cadences and selection functions.
 While \citet{Zechmeister09} details a formalism to express this floating-mean model (which they call a ``generalized periodogram'') similarly to \eq{LombScargle}, in the matrix formalism all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
@@
 
-The classic Lomb-Scargle approach is equivalent to fitting a single-term stationary sinusoidal model to the data.
+As mentioned above, the standard periodogram is equivalent to fitting a single-term stationary sinusoidal model to the data.
 A natural extension is to use a multiple-term sinusoidal model.
@@
 \end{enumerate}
-The total number of parameters for $N_{filt}$ filters is then $M = (2M_{base} + 1) + N_{filt}(2M_{band} + 1)$. The model looks as follows, for a band $b$:
+The total number of parameters for $K$ filters is then $M_K = (2M_{base} + 1) + K(2M_{band} + 1)$. The model looks as follows, for a band $b$:
 \begin{eqnarray}
-  y(t|\omega,\theta,b) = &&\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]\\ 
-  &+& \theta^{(b)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(b)}_{2n - 1}\sin(n\omega t) + \theta^{(b)}_{2n}\cos(n\omega t)\right].
+  y(t|\omega,\theta,b) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\\ 
+  &\theta^{(b)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(b)}_{2n - 1}\sin(n\omega t) + \theta^{(b)}_{2n}\cos(n\omega t)\right].&
 \end{eqnarray}
@@
 1 & \sin(\omega t_2) & \cos(\omega t_2) & 0 & 1 & 0 & 0 & 0\\
-1 & \sin(\omega t_3) & \cos(\omega t_3) & 1 & 0 & 0 & 0 & 0\\
+1 & \sin(\omega t_3) & \cos(\omega t_3) & 0 & 0 & 0 & 1 & 0\\
 1 & \sin(\omega t_4) & \cos(\omega t_4) & 0 & 0 & 1 & 0 & 0\\
 \vdots & \vdots & \vdots & & & \vdots & &\\
-1 & \sin(\omega t_N) & \cos(\omega t_N) & 0 & 0 & 0 & 0 & 1\\
+1 & \sin(\omega t_N) & \cos(\omega t_N) & 0 & 1 & 0 & 0 & 0\\
 \end{array}
@@
 \end{equation}
-Here the structure of the final five columns indicates the band observed at the given time: the first row is a $u$-band measurement, the second is a $g$-band, the third is a $u$-band, etc.
+Here the nonzero entries of the final five columns indicate the band observed at the given time: the first row is a $u$-band measurement, the second is a $g$-band, the third is a $i$-band, etc.
 
@@
 The left panel of \fig{adhoc_example} shows the observations folded as a function of phase.
-Using the typical approach from the literature prior to this paper, we individually compute the Lomb-Scargle periodogram within each band: the results are shown in the right panel.
+Using the typical approach from the literature prior to this paper, we individually compute the standard normalized periodogram within each band: the results are shown in the right panel.
 Here the data is well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band.\footnote{\label{foot1}
@@
     days). In this case, single-band data is not sufficient to recover the
-    Lomb-Scargle peaks, and the classic band-by-band approach fails.
+    periodogram peaks, and the classic band-by-band approach fails.
     By utilizing all the data at once with the multiband approach
@@
 \fig{multi_example} shows the same 60 nights of data, except that each night only a {\it single} band observation is recorded.
-The left panel again shows the observations as a function of phase, and the right panel shows the Lomb-Scargle fits to the data.
+The left panel again shows the observations as a function of phase, and the right panel shows the periodograms derived from the data.
 Now, with only 12 observations for each indiviudal band, there is not enough data to accurately determine the period band-by-band.

### 2014-12-16 multiband_LS/interactive.py
update figure 5

@@ -97,4 +97,4 @@ def interact_multifit():
         ax3.set_xlabel('r')
 
     return interact(plot_multifit, object_index=[0, len(lcids) - 1],
-                    Nterms_base=[1, 10], Nterms_band=[0, 10])
+                    Nterms_base=[0, 10], Nterms_band=[0, 10])

### 2014-12-16 writeup/figures/compute_all_periods.py

@@ -1,0 +1,24 @@
+import sys, os; sys.path.append(os.path.abspath('../..'))
+
+from multiband_LS import LombScargleMultiband
+from multiband_LS.memoize import CacheResults
+
+from multiband_LS.data import fetch_light_curves
+rrlyrae = fetch_light_curves()
+
+def find_best_period(lcid, rrlyrae, Nterms_base=1, Nterms_band=0):
+    t, y, dy, filts = rrlyrae.get_lightcurve(lcid, return_1d=True)
+    LS = LombScargleMultiband(Nterms_base=Nterms_base,
+                              Nterms_band=Nterms_band,
+                              period_range=(0.2, 1.1))
+    LS.fit(t, y, dy, filts)
+    return LS.best_period
+
+
+cache = CacheResults('results_1_0', verbose=True)
+periods_1_0 = cache.call_iter(find_best_period, rrlyrae.ids,
+                              args=(rrlyrae, 1, 0))
+
+cache = CacheResults('results_0_1', verbose=True)
+periods_0_1 = cache.call_iter(find_best_period, rrlyrae.ids,
+                              args=(rrlyrae, 0, 1))

### 2014-12-16 writeup/figures/fig_multiband_sim.py

@@ -81,18 +81,14 @@ masks = [(filts == band) for band in 'ugriz']
 P = [LombScargleAstroML().fit(t[mask], mags[mask],
                               dy[mask]).periodogram(omegas)
      for mask in masks]
-    
-LS_multi = LombScargleMultiband(Nterms_base=1, Nterms_band=0)
-LS_multi.fit(t, mags, dy, filts)
-P_multi = LS_multi.periodogram(omegas)
 
 fig = plt.figure(figsize=(10, 4))
-gs = plt.GridSpec(4, 2,
+gs = plt.GridSpec(5, 2,
                   left=0.07, right=0.95, bottom=0.15,
-                  wspace=0.1, hspace=0.5)
+                  wspace=0.1, hspace=0.6)
 ax = [fig.add_subplot(gs[:, 0]),
-      fig.add_subplot(gs[:3, 1]),
-      fig.add_subplot(gs[3, 1])]
+      fig.add_subplot(gs[:-2, 1]),
+      fig.add_subplot(gs[-2:, 1])]
     
 for band, mask in zip('ugriz', masks):
     ax[0].errorbar((t[mask] / rrlyrae.period) % 1, mags[mask], dy[mask],
@@ -106,16 +102,25 @@ ax[0].set_ylabel('magnitude')
 for i, (Pi, band) in enumerate(zip(P, 'ugriz')):
     offset = 4 - i
     ax[1].plot(periods, Pi + offset, lw=1)
-    ax[1].text(0.89, 1 + offset, band, fontsize=14, ha='right', va='top')
+    ax[1].text(0.89, 1 + offset, band,
+               fontsize=10, ha='right', va='top')
 ax[1].set_title('Standard Periodogram in Each Band')
 ax[1].yaxis.set_major_formatter(plt.NullFormatter())
 ax[1].xaxis.set_major_formatter(plt.NullFormatter())
-ax[1].set_ylabel('power + offset')
+ax[1].set_ylabel('power + offset' + 30 * ' ')
     
-ax[2].plot(periods, P_multi, lw=1, color='gray')
+for (i, Nbase, Nband) in [(0, 1, 0), (1, 0, 1)]:
+    LS_multi = LombScargleMultiband(Nterms_base=Nbase, Nterms_band=Nband)
+    LS_multi.fit(t, mags, dy, filts)
+    P_multi = LS_multi.periodogram(omegas)
[…]

### 2014-12-17 writeup/paper.tex
take a pass at the paper

@@
 
-\title{Multiband Periodograms of Astronomical Sources}
+\title{Periodograms for Multi-band Astronomical Time Series}
 
@@
 
-The detection and quantification of periodicity in time-varying signals is an important area of data analysis within modern astronomical surveys.
+The detection and quantification of periodicity in time-varying signals is an important area of data analysis within modern time-domain astronomical surveys.
 For evenly-spaced data, the {\it Schuster periodogram}, introduced in 1905, gives a quantitative measure of the periodicity of data as a function of the angular frequency $\omega$.
-For data $\{d_k\}_{k=1}^N$ measured at equal intervals $t_k = t_0 + k\Delta t$, the Schuster periodogram, which measures the spectral power as a function of the angular frequency, is given by
+For data $\{y_k\}_{k=1}^N$ measured at equal intervals $t_k = t_0 + k\Delta t$, the Schuster periodogram, which measures the spectral power as a function of the angular frequency, is given by
 \begin{equation}
   \eqlabel{Schuster}
-  C(\omega) = \frac{1}{N}\left| \sum_{k=1}^N d_k e^{i\omega t_k} \right|^2,
+  C(\omega) = \frac{1}{N}\left| \sum_{k=1}^N y_k e^{i\omega t_k} \right|^2,
 \end{equation}
@@
 Because astronomical observing cadences are rarely so uniform, many have looked at extending the ideas behind the periodogram to work with unevenly-sampled data.
-Most famously, \citep{Lomb76} and \citep{Scargle82} extended earlier work to define the {\it normalized periodogram}:
+Most famously, \citet{Lomb76} and \citet{Scargle82} extended earlier work to define the {\it normalized periodogram}:
 \begin{equation}
@@
   P_N(\omega) = \frac{1}{2\sigma^2}\left[
-    \frac{\left[\sum_k(d_k - \bar{d})\cos\omega(t_k - \tau)\right]^2}
+    \frac{\left[\sum_k(y_k - \bar{d})\cos\omega(t_k - \tau)\right]^2}
     {\sum_k \cos^2\omega(t_k - \tau)}
     +
-    \frac{\left[\sum_k(d_k - \bar{d})\sin\omega(t_k - \tau)\right]^2}
+    \frac{\left[\sum_k(y_k - \bar{d})\sin\omega(t_k - \tau)\right]^2}
     {\sum_k \sin^2\omega(t_k - \tau)}
@@
 \end{equation}
-where $\bar{d}$ is the mean and $\sigma^2$ is the variance of the data $\{d_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[See][for an in-depth discussion]{NumRec}.
-\citep{Lomb76} showed that this offset has a more important effect: namely, it makes $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
+where $\bar{d}$ is the mean and $\sigma^2$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[See][for an in-depth discussion]{NumRec}.
+\citet{Lomb76} showed that this offset has a deeper effect: namely, it makes $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
 
@@
 
-This long-recognized connection between spectral power and least squares fitting methods was solidified by \citep{Jaynes87}, who demonstrated that the normalized periodogram of Lomb and Scargle is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
-Building on this result, \citep{Bretthorst88} explored the extension of these methods to more complicated models with multiple frequency terms, nonstationary frequencies, and other more sophisticated models within a Bayesian framework.
+This long-recognized connection between spectral power and least squares fitting methods was solidified by \citet{Jaynes87}, who demonstrated that the normalized periodogram of Lomb and Scargle is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
+Building on this result, \citet{Bretthorst88} explored the extension of these methods to more complicated models with multiple frequency terms, nonstationary frequencies, and other more sophisticated models within a Bayesian framework.
 
@@
 where $t_k$ is the time of observation, $y_k$ is the observed value (typically a magnitude), and $\sigma_k$ describes the Gaussian errors on each value.
-Without loss of generality we will assume that the data $y_k$ are centered such that the measurements satisfy
+Without loss of generality we will assume that the data $y_k$ are centered such that the measurements within each band satisfy
 \begin{equation}
@@
 \end{equation}
-where $P_N(\omega)$ is the normalized periodogram given in \eq{LombScargle} and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$
+where $P_N(\omega)$ is the normalized periodogram given in \eq{LombScargle} and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$.
 
 \subsubsection{Matrix Formalism}
-
-These quantities can be expressed more compactly by defining the following matrices:
-
+The expressions related to the stationary sinusoid model can be expressed more compactly by defining the following matrices:
 \begin{equation}
@@
 \end{equation}
-
-With these definitions, the model in \eq{simplemodel} is a simple linear product: $y(t|\omega,\theta) = X_\omega\theta$ and the model and reference $\chi^2$ can be written
+With these definitions, the model in \eq{simplemodel} can be expressed as a simple linear product: $y(t|\omega,\theta) = X_\omega\theta$, and the model and reference $\chi^2$ can be written
 
@@
 \end{eqnarray}
-The normalized periodogram can be computed by finding the minimum of $\chi^2(\omega)$ via standard methods and plugging the result into \eq{chi2PN} to find
+The expression for the normalized periodogram can be computed by finding the value of $\theta$ which minizes $\chi^2(\omega)$ via standard methods and plugging the result into \eq{chi2PN} to find
 \begin{equation}
@@
 \end{equation}
-This expression is equivalent to \eq{LombScargle} in the case that $\Sigma \propto I$.
-
-\fig{basic_example} shows an example periodogram for a simulated RR Lyrae light curve based on a template from \citet{Sesar2010}.
-The particular RR Lyrae star has a period of 0.622 days, and the observations take place on 60 random nights over a 6-month period.
-The left panel of the figure shows the input data, and the right panels show the result of a simple Lomb-Scargle periodogram on the data.
+Take note that this expression is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto \sigma^2 I$.
 
@@
 
-In the upper right, we see the normalized periodogram as a function of period.
-While the power does peak at the true period of 0.622 days, an aliasing effect is apparent.
-These additional peaks are due to beat frequencies between the true period $P$ and the observing cadence of $\sim 1$ day.
-That is, for nightly observations, we'd expect excess power to be found at frequencies $P / (1 + nP)$ for a positive or negative integer $n$.
-This explains the prominent peak at $0.383$.
-For the current work, we'll ignore this (important) piece of the period-finding problem; for a thorough discussion of aliasing within astronomical time series, see \citet{Roberts87}.
+As an example of the standard periodogram in action, we perform a simple single-band harmonic analysis of simulated $r$-band observations of an RR Lyrae light curve, based on empirical templates derived in \citet{Sesar2010} (\fig{basic_example}).
+The observations are of a star with a period of 0.622 days, and take place on 60 random nights over a 6-month period, as seen in the left panel.
 
-The lower-right panel shows the maximum likelihood interpretation of this periodogram: it is a measure of the normalized $\chi^2$ for a single-term sinusoidal model.
+The upper-right panel shows the normalized periodogram for this source as a function of period.
+While the power does peak at the true period of 0.622 days, an aliasing effect is readily apparent near $P=0.38$.
+This additional peak is due to beat frequency between the true period $P$ and the observing cadence of $\sim 1$ day.
+This beat frequency is the first in a large sequence: for nightly observations, we'd expect to find excess power at periods That is, for nightly observations, we'd expect excess power to be found at frequencies $P_n = P / (1 + nP)$ days, for any positive or negative integer $n$.
+The strong alias in \fig{basic_example} corresponds to the $n=1$ beat period $P_n=0.383$.
+Though it is possible to carefully correct for such aliasing via the estimated window function \citep[e.g.][]{Roberts87}, we'll ignore this detail in the current work.
+
+The lower-right panel of \fig{basic_example} shows the maximum likelihood interpretation of this periodogram: it is a measure of the normalized $\chi^2$ for a single-term sinusoidal model.
 Here we visualize the data from the left panel, but folded as a function of phase with the best-fit single-term model over-plot.
-Here its apparent that the single-term model is biased: RR Lyrae are much more complicated than a simple sinusoid.
-Nevertheless, the simplistic sinusoidal model does well to zero-in
+Here its apparent that the single-term model is biased: RR Lyrae light curves are, in general, much more complicated than a simple sinusoid.
+Nevertheless, the simplistic sinusoidal model does well is able to recover the correct frequency to a high degree of accuracy (roughly related to the width of the peak) and significance (roughly related to the height of the peak).
+For a more complete introduction to and discussion of the single-term normalized periodogram, refer to, e.g.
+\citet{Bretthorst88} or \citet{ICVG2014}.
 
-\subsubsection{Discussion}
+\subsection{Extending the Periodogram}
 We have shown two forms of the classic normalized periodogram: \eq{LombScargle} and \eq{LombScargle2}.
@@
 The expression in \eq{LombScargle} avoids the explicit construction of a matrix, and thus can be computed very efficiently.
-In particular, through some computational tricks involving the FFT, expressions of the form of \eq{LombScargle} can be evaluated for $N$ frequencies in $\mathcal{O}[\log{N}]$ time \citep{Press89}.
+Furthermore, through some computational tricks taking advantage of the Fast Fourier Transform, expressions of the form of \eq{LombScargle} can be evaluated exactly for $N$ frequencies in $\mathcal{O}[\log{N}]$ time \citep{Press89}.
 
-The matrix-based formulation of \eq{LombScargle2} sacrifices some potential computational efficiency in order to attain several advantages:
+The matrix-based formulation of \eq{LombScargle2}, though slower than the Fourier-derived formulation, is a more general expression and allows several advantages:
 \begin{enumerate}
-  \item It is trivially extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the noise matrix $\Sigma$
+  \item It is trivially extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the noise covariance matrix $\Sigma$
   \item It is trivially extended to more sophisticated linear models by appropriately modifying the design matrix $X_\omega$.
-  \item It is trivially extended to Tikhonov/L2-regularized likelihoods by adding an appropriate diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$.
+  \item It is trivially extended to include Tikhonov/L2-regularization terms by adding an appropriate diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$.
 \end{enumerate}
@@
 
-\subsection{Stationary Sinusoid with Floating Mean}
+\subsubsection{Stationary Sinusoid with Floating Mean}
 \sectlabel{floating_mean}
@@
 This adjusts the classic normalized periodogram by allowing the mean of the model to be fit alongside the amplitudes:
-
 \begin{equation}
@@
 \end{equation}
+This model can be more accurate than the standard periodogram for certain observing cadences and selection functions.
+\citet{Zechmeister09} detail the modifications required to the harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
+In the matrix formalism, the modification is much more straightforward: all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
 
-This model can be more accurate than the standard periodogram for certain observing cadences and selection functions.
-While \citet{Zechmeister09} details a formalism to express this floating-mean model (which they call a ``generalized periodogram'') similarly to \eq{LombScargle}, in the matrix formalism all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
+For well-sampled data, there is usually very little difference between a standard periodogram and a floating-mean periodogram.
+Where this becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases of the light curve: a toy example demonstrating this situation is shown in \fig{floating_mean}.
+The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias (right panel).
+Because of this selection effect, the mean of the observed data are a poor predictor of the true mean, causing the standard fixed-mean method to poorly fit the data.
+The floating-mean approach is able to automatically adjust for this bias, resulting in a periodogram which readily detects the input period of 0.622 days (lower-right panel).
 
-For well-sampled data, there is little difference between a standard periodogram and a floating-mean periodogram.
-Where this becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases: an example of this situation is shown in \fig{floating_mean}.
-The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias.
-Because of this bias, the mean of the observed data to be a poor predictor of the true mean, which makes it so that the standard approach fails to find a suitable fit to the data.
-The floating-mean approach is able to adjust for this, resulting in a periodogram which detects the input period of 0.622 days.
-
-\subsection{Truncated Fourier Models}
+\subsubsection{Truncated Fourier Models}
 \sectlabel{multiterm}
@@
   \caption{
-    An illustration of a periodogram based on the truncated Fourier model. The data are the same as those in \fig{basic_example}. 
+    An illustration of a periodogram based on the truncated Fourier model.
+    The data are the same as those in \fig{basic_example}. Note that the
+    higher-order models will show periodogram peaks at multiples of the true
+    fundamental frequency $P_0$: this is because for integer $n$ less than
+    the number of Fourier terms in the model, $P_0$ is a higher harmonic of
+    the model at $P=nP_0$.
   }
@@
 As mentioned above, the standard periodogram is equivalent to fitting a single-term stationary sinusoidal model to the data.
-A natural extension is to use a multiple-term sinusoidal model.
-With $M$ terms, there are $2M + 1$ free parameters, and the model is given by:
-
+A natural extension is to instead use a multiple-term sinusoidal model, with frequencies at integer multiples of the fundamental frequency.
+With $M$ terms, there are $2M + 1$ free parameters, and the model is given by
 \begin{equation}
-  y(t|\omega,\theta) = \theta_0 + \sum_{n=1}^M \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]
+  y(t|\omega,\theta) = \theta_0 + \sum_{n=1}^M \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right].
 \end{equation}
-
 Because this remains a linear model, it can be easily accommodated into the matrix formalism above.
-For example, an $M = 2$-term model can be constructed by building a design matrix $X_\omega$ with $2M + 1 = 5$ columns:
-
+For example, an $M = 2$-term floating-mean model can be constructed by building a design matrix $X_\omega$ with $2M + 1 = 5$ columns:
 \begin{equation}
@@
 \end{equation}
-Using $X_\omega^{(2)}$ in \eq{LombScargle2} will result in a two-term periodogram.
-In general, we can fit an $M$-term periodogram by simply adding terms to the fit.
-\fig{multiterm_example} shows a few examples of this multiterm Fourier approach as applied to a simulated RR Lyrae light curve constructed from templates derived in \citet{Sesar2010}.
-There are several important features of this figure that give us insight into the subtleties of this type of fit:
+Computing the power via \eq{LombScargle2} using $X_\omega^{(2)}$ will give the two-term periodogram.
+As $M$ grows, the size of the design matrix grows, but the periodogram can be computed in the same manner.
+\fig{multiterm_example} shows a few examples of this multiterm Fourier approach as applied to the simulated RR Lyrae light curve from \fig{basic_example}.
+There are several important features of this figure that give us insight into the subtleties of this type of multiterm fit.
 
 First, we see in the right panel that all three models show a significant spike at the true period of $P_0 = 0.622$.
-The higher-order models, however, also show a a spike of poer at $P_1 = 2 P_0$: the reason for this is that the fundamental model of the single-term model at $P_0$ is the first harmonic of the 2-term model at $2P_0$, so the higher-order models contain this single-term result.
+The higher-order models, however, also show a a spike in power at $P_1 = 2 P_0$: the reason for this is that for and $M>1$-term model, the period $P_0$ is the first harmonic of a model with fundamental frequency $2P_0$, and the higher-order models contain the single-period result.
 
 Second, notice that as the number of terms is increased, the general ``background'' level of the periodogram increases.
-This is due to the fact that the periodogram is directly related to the $\chi^2$ of the fit.
-A more complicated model will have the flexibility to conform to the data at {\it all frequencies}, and thus the observed ``power'' is stronger at all frequencies.
+This is due to the fact that the periodogram is directly related to the $\chi^2$ of the fit at each frequency.
+Because of the increased model complexity of a higher-order model, it is able to better fit the data at all periods, not just the true period.
+Thus the observed ``power'' of a higher-order model will be everywhere higher than the power of a lower-order model.
 
-\subsection{Regularized Models}
+\subsubsection{Regularized Models}
 \sectlabel{regularization}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig04.pdf}
+  \caption{
+    The effect of regularization on a high-order model. The data is the same as
+    those in \fig{basic_example}. We fit a 20-term truncated Fourier model to
+    the data, with and without a regularization term. Without regularization,
+    the model oscillates widely to fit the noise in the data. The
+    regularization term effectively damps the higher-order Fourier modes and
+    removes this oscillating behavior, leading to a model which is much more
+    robust to noise in the data.
+  }
+  \figlabel{regularized_example}
+\end{figure}
+
 The previous sections raise the question: how complicated a model should we use?
-If we added many terms to the truncated Fourier series, we might fit the data perfectly, but eventually we would {\it overfit} the data: that is, our model is fitting the noise rather than the signal.
-This can be avoided by explicitly truncating the series, but another approach is to use a {\it regularization} term to enforce a less complicated model.
+It is easily shown that as we add more terms to the fit, the model will come closer and closer to the observed data.
+When the number of model parameters equals the number of data points, we can expect to fit the data {\it perfectly} (though in most cases numerical inaccuracies in the fit prevent a truly perfect fit).
+For very high-order models, the result become very sensitive to the noise in the data, and we say we have {\it over-fit} the data.
+This can be addressed by explicitly truncating the series, but we can also use a {\it regularization} term to mathematically ensure a less complicated model.
 
-A regularization term is an explicit penalty on the magnitude of the model parameters $\theta$, and can take any number of forms.
-For computational simplicity here we'll use an {\it L2 regularization} (also known as Tikhonov Regularization), which is a quadratic penalty term in the model parameters.
-This is mathematically equivalent to the Bayesian approach of using a Gaussian prior on the model parameters.
+A regularization term is an explicit penalty on the magnitude of the model parameters $\theta$, and can take a number of forms.
+For computational simplicity here we'll use an {\it L2 regularization} (also known as Tikhonov Regularization), which is a quadratic penalty term in the model parameters added to the $\chi^2$.
+Mathematically, this is equivalent to the Bayesian approach of using a Gaussian prior on the model parameters.
 
@@
 \end{equation}
-Note that the effect of this regularization term is to add a diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$, which has the additional feature that it corrects ill-posed models.
+Notice that the effect of this regularization term is to add a diagonal penalty term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$, which has the additional feature that it can correct ill-posed models where the normal matrix is not invertible.
 This feature of the regularization will become important below.
 
-\todo{Show a regularized model?}
+In \fig{regularized_example}, we compare a regularized and unregularized 20-term truncated Fourier model on our simulated RR Lyrae light curve.
+We use $\lambda = 0$ on the offset term, and make $\lambda_j$ progressively larger for each harmonic component.
+The result of the regularized model is less over-fitting to the input data, and a more distinct peaks in the associated periodogram.
 
 \section{A Multiple-Band Model}
-
 To compute a periodogram for multi-band data, we'll take advantage of many of the nice features of the matrix form of the normalized periodogram, which allows us to specify an arbitrary linear model covering our data.
@@
   \item An $M_{base}$-term truncated Fourier fit which models a latent parameter, which we'll call the ``overall variability''.
-  \item An $M_{band}$-term truncated Fourier fit which models the residual of each band from the overall variability.
+  \item An $M_{band}$-term truncated Fourier fit which models the residual of each band from this overall variability.
 \end{enumerate}
-The total number of parameters for $K$ filters is then $M_K = (2M_{base} + 1) + K(2M_{band} + 1)$. The model looks as follows, for a band $b$:
+The total number of parameters for $K$ filters is then $M_K = (2M_{base} + 1) + K(2M_{band} + 1)$. As a result, for each band $b$ we have the following model of the observed magnitudes:
 \begin{eqnarray}
-  y(t|\omega,\theta,b) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\\ 
+  y_b(t|\omega,\theta) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\\ 
   &\theta^{(b)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(b)}_{2n - 1}\sin(n\omega t) + \theta^{(b)}_{2n}\cos(n\omega t)\right].&
 \end{eqnarray}
-The important feature is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(b)}$ are fit within each band.
+The important feature of this model is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(b)}$ are determined individually.
 
-We can build the matrix form of this model by building a sparse $X_{\omega}$ matrix with $M$ columns.
-Each row corresponds to a single observation through a single band; columns corresponding to the base model and the observation band will have nonzero entries; all other columns will be filled with zeros.
-For example, the $X_\omega$ matrix with $N_{base}=1$ and $N_{band}=0$ (that is, a single sinusoid model with independent offsets to each band) will look as follows:
-
+We can construct the normalized periodogram for this model by building a sparse design matrix with $M_K$ columns.
+Each row corresponds to a single observation through a single band.
+Columns corresponding to the base model and the matching observation band will have nonzero entries; all other columns will be filled with zeros.
+For example, the $N_{base}=1$ and $N_{band}=0$ model corresponds to one with a simple single-term periodic base frequency, and an independent constant offset term in each band.
+The associated design matrix will look as follows:
 \begin{equation}
@@
 \end{equation}
-Here the nonzero entries of the final five columns indicate the band observed at the given time: the first row is a $u$-band measurement, the second is a $g$-band, the third is a $i$-band, etc.
+Here the nonzero entries of the final five columns indicate the $ugriz$ band of the given observation: the first row is a $u$-band measurement, the second is a $g$-band, the third is a $i$-band, etc.
 
-From reasoning about the free parameters, or by examining the above matrix it is easy to see that some of the parameters will be degenerate (i.e.
-$X_\omega$ is low-rank).
+From reasoning about the free parameters, or by examining the above matrix it is easy to see that some of the parameters will be degenerate: i.e.
+$X_\omega$ is low-rank, and the problem is ill-posed.
 Intuitively, this is due to the fact that if we add an overall offset to the base model, this can be perfectly accounted for by subtracting that same offset from each of the band columns.
-In order to proceed, then, we'll either have to correct this model deficiency, or use a regularization term on one of the offending parameters.
-We'll choose the latter here, and regularize all the band columns to drive the common behavior into the base model.
-For the above matrix,
-this regularization will look like
+Mathematically, the result of this is that the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$ will be non-invertible, and thus the periodogram is ill-defined.
+In order to proceed, then, we'll either have to use a different model, or use a cleverly-constructed regularization term on one of the offending parameters.
+
+We'll choose the latter here, and regularize all the band columns while leaving the base columns un-regularized: for the above matrix, this regularization will look like
 \begin{equation}
-  \Lambda = {\rm diag}([0, 0, 0, \epsilon, \epsilon, \epsilon, \epsilon, \epsilon])
+  \Lambda^{(1,0)} = {\rm diag}([0, 0, 0, \epsilon, \epsilon, \epsilon, \epsilon, \epsilon])
 \end{equation}
-where $\epsilon$ is some small fraction of the trace of the matrix $[X_\omega^T\Sigma^{-1}X_\omega]$.
+where $\epsilon$ is some small fraction of the trace of the normal matrix $[X_\omega^T\Sigma^{-1}X_\omega]$.
+The logic of this choice of regularization is that any component of the model which is common to all bands will be accounted for in the base terms: only components of the fit that differ in each band will affect the band terms.
+Setting $\epsilon$ to some small fraction of the trace ensures that the regularization effect will be small.
 With this regularization in place, the model is well-posed and \eq{LombScargleReg} can be used to straightforwardly compute the power.
@@
 
-The final remaining piece is our assumption that the data are centered (i.e.
-\eq{ycentered}).
+The final remaining piece to mention is our assumption in \eq{ycentered} that the data are centered.
 This is required so that the simple form of the reference $\chi^2_0$ remains valid.
-For the multiband model, this assumption requires that the data satisfy \eq{ycentered} {\it within each band}.
+For the multiband model, this assumption requires that the data satisfy \eq{ycentered} {\it within each band}: equivalently, we could lift this assumption and compute the reference $\chi^2_0$ of the multiband model with an independent floating mean within each band; the results will be identical.
 
 \subsection{Relationship of Multiband and Single-band approaches}
+\sectlabel{relationship}
 There is a special case of the multi-band approach when $N_{base}=0$ and $N_{band}=1$.
-In this case, the $X$ matrix can be arranged to be block-diagonal, and the computation can be split into an independent single-band computation for each of the $K$ bands.
-For band $k$, denote the single-band periodogram as
+In this case, the base offset can be removed completely and the $X$ matrix can be arranged to be block-diagonal.
+A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here the independent components are a standard single-band generalized periodogram within each of the $K$ bands.
+
+For band $k$, we'll denote the single-band normalized periodogram as
 \begin{equation}
@@
 \begin{equation}
-  P_N(\omega) = 1 - \frac{\sum_{k=1}^K\chi^2_{min, k}(\omega)}{\sum_{k=1}^K\chi^2_{0,k}}
+  P_N^{(0,1)}(\omega) = 1 - \frac{\sum_{k=1}^K\chi^2_{min, k}(\omega)}{\sum_{k=1}^K\chi^2_{0,k}}
 \end{equation}
@@
 \begin{equation}
-  P_N(\omega) = \frac{\sum_{k=1}^K\left[\chi^2_{0,k}P_N^{k}\right]}{\sum_{k=1}^K\chi^2_{0,k}}
+  P_N^{(0,1)}(\omega) = \frac{\sum_{k=1}^K\left[\chi^2_{0,k}P_N^{k}\right]}{\sum_{k=1}^K\chi^2_{0,k}}.
 \end{equation}
-where, recall $\chi^2_{0,k}$ is simply the $\chi^2$ of the data in each band about the mean.
 This tells us how we can construct a particular case of the multi-term periodogram from a weighted sum of periodograms in each band.
-Here the weights reflect both how many measurements there are in each band, and how poorly those measurements are fit by a simple constant reference model.
+Here the weights reflect both how many measurements there are in each band, as well as how much those measurements deviate from a simple constant reference model.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{fig04.pdf}
+  \includegraphics[width=\textwidth]{fig05.pdf}
   \caption{
@@
 
-\fig{adhoc_example} shows a simulated RR Lyrae light curve based on a template from \citet{Sesar2010}.
-The observations take place on 60 random nights over a 6-month period, and each night all five bands ({\it u,g,r,i,z}) are recorded.
-The left panel of \fig{adhoc_example} shows the observations folded as a function of phase.
-Using the typical approach from the literature prior to this paper, we individually compute the standard normalized periodogram within each band: the results are shown in the right panel.
-Here the data is well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band.\footnote{\label{foot1}
-  An aliasing effect is readily apparent in the left panel of \fig{adhoc_example}: these additional peaks are due to beat frequencies between the true period $P$ and the observing cadence of $\sim 1$ day.
-That is, for nightly observations, we'd expect excess power to be found at frequencies $P / (1 + nP)$ for a positive or negative integer $n$.
-For the current work, we'll ignore this (important) piece of the problem; for a thorough discussion of aliasing within astronomical time series, see \citet{Roberts87}.}
-
 \begin{figure}
   \centering
-  \includegraphics[width=\textwidth]{fig05.pdf}
+  \includegraphics[width=\textwidth]{fig06.pdf}
   \caption{
@@
   } 
-  \figlabel{multi_example}
+  \figlabel{multiband_sim}
 \end{figure}
 
-\fig{multi_example} shows the same 60 nights of data, except that each night only a {\it single} band observation is recorded.
+\todo{We need the bottom-right panel to be bigger! It's our money result! How do we arrange the plot to do this? Also: label it new style and old style.}
+
+\fig{adhoc_example} shows a simulated multiband version of the RR Lyrae light curve from \fig{basic_example}.
+The observations take place on 60 random nights over a 6-month period, and each night all five bands ({\it u,g,r,i,z}) are recorded.
+The left panel of \fig{adhoc_example} shows the observations folded as a function of phase.
+Using the typical approach from the literature prior to this paper, we individually compute the standard normalized periodogram within each band: the results are shown in the right panel.
+Here the data is well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band, up to the aliasing effect discussed in \sect{simple_period}.
+
+\fig{multiband_sim} shows the same 60 nights of data, except that each night only a {\it single} band observation is recorded.
 The left panel again shows the observations as a function of phase, and the right panel shows the periodograms derived from the data.
-Now, with only 12 observations for each indiviudal band, there is not enough data to accurately determine the period band-by-band.
-With the multiband approach (here computed for $N_{base}=1$ and $N_{band}=0$), as seen in the lower-right panel of \fig{multi_example}, the period is cleanly recovered.
+Now, with only 12 observations for each indiviudal band, there is not enough data to accurately determine the period within each single band.
+With the multiband approach (here computed for $(N_{base},N_{band})=(1,0)$ and $(0,1)$), seen in the lower-right panel of \fig{multiband_sim}, the period is cleanly recovered.
 
-\todo{Compare different options of $N_{base}$ and $N_{band}$?}
+Here we can clearly see the difference between the $(0,1)$ model, which we'll call {\it old-style}, and the $(1,0)$ model, which we'll call {\it new-style}.
+The old-style model, as discussed in \sectlabel{relationship}, is equivalent to a weighted sum of single-band periodograms for each individual band.
+The model thus has effectively 15 free parameters, and the phase and amplitude of each band varies individually.
+As with the truncated Fourier series, this added complexity leads to higher power at off-frequencies.
+The new-style model, on the other hand, is effectively a 7-parameter model where a single oscillation phase and amplitude is used to describe all of the data.
+The only flexibility allowed in individual bands is the constant offset.
+This lower-parameter model produces a periodogram which nicely recovers the input frequency, and has a background level qualitatively similar to that of the single-band observations in \fig{adhoc_example}.
 

### 2014-12-17 writeup/figures/compute_all_periods.py

@@ -15,10 +15,12 @@ def find_best_period(lcid, rrlyrae, Nterms_base=1, Nterms_band=0):
     return LS.best_period
 
 
+lcids = list(rrlyrae.ids)
+
 cache = CacheResults('results_1_0', verbose=True)
-periods_1_0 = cache.call_iter(find_best_period, rrlyrae.ids,
+periods_1_0 = cache.call_iter(find_best_period, lcids,
                               args=(rrlyrae, 1, 0))
 
 cache = CacheResults('results_0_1', verbose=True)
-periods_0_1 = cache.call_iter(find_best_period, rrlyrae.ids,
+periods_0_1 = cache.call_iter(find_best_period, lcids,
                               args=(rrlyrae, 0, 1))

### 2014-12-17 writeup/figures/fig_floating_mean.py

@@ -43,8 +43,11 @@ tfit = rrlyrae.period * phasefit
 periods = np.linspace(0.2, 1.1, 1000)
 omegas = 2 * np.pi / periods
 
-fig, ax = plt.subplots(1, 2, figsize=(10, 4))
-fig.subplots_adjust(left=0.07, right=0.95, wspace=0.1, bottom=0.15)
+fig = plt.figure(figsize=(10, 4))
+gs = plt.GridSpec(2, 2, left=0.07, right=0.95, wspace=0.15, bottom=0.15)
+ax = [fig.add_subplot(gs[:, 0]),
+      fig.add_subplot(gs[0, 1]),
+      fig.add_subplot(gs[1, 1])]
 
 ax[0].errorbar(phase[mask], mag[mask], dmag[mask], fmt='o',
                color='#666666')
@@ -53,6 +56,7 @@ ax[0].errorbar(phase[~mask], mag[~mask], dmag[~mask], fmt='o',
 ax[0].invert_yaxis()
 
 for fit_offset in [False, True]:
+    i = int(fit_offset)
     model = LombScargle(fit_offset=fit_offset).fit(t[mask],
                                                    mag[mask], dmag[mask])
     P = model.periodogram(omegas)
@@ -60,22 +64,22 @@ for fit_offset in [False, True]:
         label = 'floating mean'
     else:
         label = 'standard'
-    ax[0].plot(phasefit, model.predict(tfit, omega=2 * np.pi / rrlyrae.period),
-               label=label)
-    ax[1].plot(periods, P + 1 - fit_offset, lw=1)
-    ax[1].text(1.08, 0.5 + 1 - fit_offset, label, ha='right')
+    lines = ax[0].plot(phasefit,
+                       model.predict(tfit, omega=2 * np.pi / rrlyrae.period),
+                       label=label)
+    ax[1 + i].plot(periods, P, lw=1, c=lines[0].get_color())
+    ax[1 + i].set_title('{0} Periodogram'.format(label.title()))
+    ax[1 + i].set_ylabel('power')
+    ax[1 + i].set_ylim(0, 1)
+    ax[1 + i].set_xlim(0.2, 1.1)
 
 ax[0].legend(loc='upper left')
 ax[0].set_xlabel('phase')
 ax[0].set_ylabel('magnitude')
 ax[0].set_title('Phased data (period=0.622 days)')
 
[…]

### 2014-12-17 writeup/figures/fig_multiterm_example.py

@@ -41,8 +41,14 @@ phase = (t / rrlyrae.period) % 1
 phasefit = np.linspace(0, 1, 1000)
 tfit = rrlyrae.period * phasefit
 
-fig, ax = plt.subplots(1, 2, figsize=(10, 4))
-fig.subplots_adjust(left=0.07, right=0.95, wspace=0.1, bottom=0.15)
+fig = plt.figure(figsize=(10, 4))
+gs = plt.GridSpec(3, 2, left=0.07, right=0.95, bottom=0.15,
+                  wspace=0.15, hspace=0.3)
+
+ax = [fig.add_subplot(gs[:, 0]),
+      fig.add_subplot(gs[0, 1]),
+      fig.add_subplot(gs[1, 1]),
+      fig.add_subplot(gs[2, 1])]
 
 # Plot the data
 ax[0].errorbar(phase, mag, dmag, fmt='o', color='#AAAAAA')
@@ -58,11 +64,18 @@ for i, Nterms in enumerate(models):
     if Nterms == 1:
         label = label[:-1]
         
-    offset = len(models) - 1 - i
-    lines = ax[1].plot(periods, offset + model.periodogram(omegas), lw=1)
-    ax[1].text(0.22, offset + 0.9, label, ha='left', va='top')
-    ax[0].plot(phasefit, model.predict(tfit, omega=omega_best),
-               c=lines[0].get_color(), label=label)
+    lines = ax[0].plot(phasefit, model.predict(tfit, omega=omega_best),
+                       label=label)
+    ax[1 + i].plot(periods, model.periodogram(omegas),
+                   c=lines[0].get_color(), lw=1)
+    ax[1 + i].set_title("{0}-term Periodogram".format(Nterms))
+    ax[1 + i].set_xlim(0.2, 1.4)
+    ax[1 + i].set_ylim(0, 1)
+
+ax[2].set_ylabel('power')
+ax[3].set_xlabel('period (days)')
+for i in [1, 2]:
+    ax[i].xaxis.set_major_formatter(plt.NullFormatter())
 
 ax[0].set_xlabel('phase')
 ax[0].set_ylabel('magnitude')
@@ -70,13 +83,6 @@ ax[0].invert_yaxis()
 ax[0].legend(loc='upper left')
 ax[0].set_title('Folded Data (P={0:.3f} days)'.format(rrlyrae.period))
 
-ax[1].set_title('Periodogram for each model')
[…]

### 2014-12-17 writeup/figures/fig_regularization_example.py

@@ -41,7 +41,7 @@ phasefit = np.linspace(0, 1, 1000)
 tfit = rrlyrae.period * phasefit
 
 fig = plt.figure(figsize=(10, 4))
-gs = plt.GridSpec(2, 2, left=0.07, right=0.95, wspace=0.1, bottom=0.15)
+gs = plt.GridSpec(2, 2, left=0.07, right=0.95, wspace=0.15, bottom=0.15)
 ax = [fig.add_subplot(gs[:, 0]),
       fig.add_subplot(gs[0, 1]),
       fig.add_subplot(gs[1, 1])]
@@ -54,7 +54,7 @@ ylim = ax[0].get_ylim()
 Nterms = 20
 sigma_r_inv =  np.vstack([np.arange(Nterms + 1),
                           np.arange(Nterms + 1)]).T.ravel()[1:] ** 2
-models = [None, 0.5 * sigma_r_inv ** 2]
+models = [0.5 * sigma_r_inv ** 2, None]
 
 for i, reg in enumerate(models):
     model = LombScargle(Nterms=Nterms, regularization=reg,
@@ -76,7 +76,7 @@ for i, reg in enumerate(models):
     ax[1 + i].set_ylabel('power')
     ax[1 + i].set_xlim(0.2, 1.4)
     ax[1 + i].set_ylim(0, 1)
-    ax[1 + i].yaxis.set_major_formatter(plt.NullFormatter())
+    #ax[1 + i].yaxis.set_major_formatter(plt.NullFormatter())
 
 ax[0].set_xlabel('phase')
 ax[0].set_ylabel('magnitude')

### 2014-12-17 writeup/figures/compute_all_periods.py
add compare periods script

@@ -1,26 +1,0 @@
-import sys, os; sys.path.append(os.path.abspath('../..'))
-
-from multiband_LS import LombScargleMultiband
-from multiband_LS.memoize import CacheResults
-
-from multiband_LS.data import fetch_light_curves
-rrlyrae = fetch_light_curves()
-
-def find_best_period(lcid, rrlyrae, Nterms_base=1, Nterms_band=0):
-    t, y, dy, filts = rrlyrae.get_lightcurve(lcid, return_1d=True)
-    LS = LombScargleMultiband(Nterms_base=Nterms_base,
-                              Nterms_band=Nterms_band,
-                              period_range=(0.2, 1.1))
-    LS.fit(t, y, dy, filts)
-    return LS.best_period
-
-
-lcids = list(rrlyrae.ids)
-
-cache = CacheResults('results_1_0', verbose=True)
-periods_1_0 = cache.call_iter(find_best_period, lcids,
-                              args=(rrlyrae, 1, 0))
-
-cache = CacheResults('results_0_1', verbose=True)
-periods_0_1 = cache.call_iter(find_best_period, lcids,
-                              args=(rrlyrae, 0, 1))

### 2014-12-17 writeup/figures/fig_compare_periods.py

@@ -1,0 +1,105 @@
+import os, sys; sys.path.append(os.path.abspath('../..'))
+
+import numpy as np
+from scipy.stats import mode
+import matplotlib.pyplot as plt
+import seaborn; seaborn.set()
+
+from multiband_LS import LombScargleAstroML, LombScargleMultiband
+from multiband_LS.memoize import CacheResults
+from multiband_LS.data import fetch_light_curves
+rrlyrae = fetch_light_curves()
+
+
+def find_best_period_adhoc(lcid):
+    t, y, dy, filts = rrlyrae.get_lightcurve(lcid, return_1d=True)
+    models = [LombScargleAstroML(period_range=(0.2, 1.2)).fit(t[filts==filt],
+                                                              y[filts==filt],
+                                                              dy[filts==filt])
+              for filt in 'ugriz']
+    return np.array([model.best_period for model in models])
+
+
+def compute_all_periods_adhoc():
+    lcids = list(rrlyrae.ids)
+    cache_dir = 'results_adhoc'
+    cache = CacheResults(cache_dir, verbose=True)
+    results = cache.call_iter(find_best_period_adhoc, lcids)
+    return np.array(results)
+
+
+def find_best_period(lcid, rrlyrae, Nterms_base=1, Nterms_band=0):
+    t, y, dy, filts = rrlyrae.get_lightcurve(lcid, return_1d=True)
+    LS = LombScargleMultiband(Nterms_base=Nterms_base,
+                              Nterms_band=Nterms_band,
+                              period_range=(0.2, 1.2))
+    LS.fit(t, y, dy, filts)
+    return LS.best_period
+
+
+def compute_all_periods(Nterms_base, Nterms_band):
+    lcids = list(rrlyrae.ids)
+    cache_dir = 'results_{0}_{1}'.format(Nterms_base, Nterms_band)
+    cache = CacheResults(cache_dir, verbose=True)
+    return cache.call_iter(find_best_period, lcids,
+                           args=(rrlyrae, Nterms_base, Nterms_band))
+
+
[…]

### 2014-12-22 writeup/paper.tex
ZI edits

@@
 \begin{abstract}
-  This paper introduces the {\it multiband periodogram} a general extension of the standard Lomb-Scargle approach to detecting periodic signals in time-domain data.
+This paper introduces the {\it multiband periodogram}, a general extension of the standard 
+Lomb-Scargle approach for detecting periodic signals in time-domain data. In addition to 
+advantanges of the Lomb-Scargle method such as treatment of non-uniform sampling and
+heteroscedastic errors, the multiband periodogram significantly improves period finding for randomly 
+sampled multi-band light curves (e.g., Pan-STARRS, DES and LSST). The light curves in 
+each band are modeled as arbitrary truncated Fourier series, with the period and phase 
+shared across all bands. The key aspect is the use of Tikhonov regularization which drives
+most of the variability into the so-called base model common to all bands, while 
+fits for individual bands describe residuals relative to the base model and typically require
+lower-order Fourier series. This decrease in the required number of fit parameters is the
+main reason for improved performance. We use simulated light curves 
+and randomly subsampled SDSS Stripe 82 data to demonstrate the superiority of this 
+method compared to other methods from the literature. The python implementation of 
+this method  is made publicly available as part of the astroML distribution. 
 \end{abstract}
@@
 \section{Introduction}
+
+Many types of variable stars show periodic flux variability \citep{EM2008}. Periodic variable stars are important 
+both for testing models of stellar evolution and for using such stars as distance indicators (e.g., Cepheids 
+and RR Lyrae stars). One of the first and main goals of the analysis is to detect variability and to estimate the 
+period and its uncertainty. A number of parametric and non-parametric methods have been proposed to 
+estimate the period of an astronomical time series \citep[e.g.,][and references therein]{Graham13}.
+
+The most popular non-parametric method is the phase dispersion minimization (PDM) introduced by \cite{PDM1978}. 
+Dispersion per bin is computed for binned phased light curves evaluated for a grid of trial periods. The best
+period minimizes the dispersion per bin.  A similar and related non-parametric method that has been recently 
+gaining popularity is the Supersmoother routine \citep{Reimann94}. It uses a running mean or running linear 
+regression on the data to fit the observations as a function of phase to a range of periods. The best period 
+minimizes a figure-of-merit, adopted as weighted sum of absolute residuals around the running mean. 
+Neither the Supersmoother algorithm nor the PDM method require a priori knowledge of the light curve shape. 
+We note that neither method produces posterior probability distribution for the period but rather a single point 
+estimate.
+
+The most popular parametric method is the Lomb-Scargle periodogram (discussed in detail in \S~\ref{sec:periodograms}.
+The Lomb-Scargle periodogram is related to the $\chi^2$ for a least-square fit of a single sinusoid to data
+and can treat non-uniformly sampled time series with heteroscedastic measurement uncertainties. 
+The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency and basis functions at different
+frequencies are not orthogonal. As a result, the periodogram has many local maxima and thus in practice the global 
+maximum of the periodogram is found by grid search \citep[for details see, e.g.][]{ICVG2014}.
+A more general parameteric method based on the use of continuous-time autoregressive moving average (CARMA) model
+was recently introduced by \citep{Kelly14}. CARMA models can also treat non-uniformly sampled time series with 
+heteroscedastic measurement uncertainties, and can handle complex variability patterns.
+
+A weakness of the above methods is that they require homogeneous measurements -- for astronomy data, this means 
+that successive measurements must be taken through a single photometric bandpass (filter). This has not been a major
+problem for past surveys because measurements are generally taken through a single photometric filter 
+\citep [e.g. LINEAR,][]{LINEAR1}, or nearly-simultaneously in all bands at each observation \citep [e.g. SDSS,][]{Sesar2010}.
+For the case of simultaneously taken multi-band measurements, \cite{Suveges12} utilized the principal component
+method to optimaly extract the best period. Their method is essentially a multi-band generalization of the well-known
+two-band Welch-Stetson variability index \citep{Stetson1996}. Unfortunately, when data in each band are taken at
+different times, the  principal component approach in not applicable. In such cases, past studies have generally relied 
+on \adhoc{} methods such as a majority vote among multiple single-band estimates of the 
+periodogram \citep[e.g.,][]{Oluseyi12}.
+
+For surveys that obtain multi-band data one band at a time, such as Pan-STARRS \citep{Kaiser2010} and DES \citep{Flaugher08},
+and for future multicolor surveys such as LSST \citep{Ivezic08LSST}, this \adhoc{} approach is not optimal. In order to take 
+advantage of the full information content in available data, it would be desirable to have a single estimate of the periodogram 
+which accounts for all observed data in a manner which is not dependent on the underlying spectrum of the object. 
+We propose such a method in this paper.
+
+The proposed method is essentially a generalization of the Lomb-Scargle method to 
+multi-band case. The light curves in  each band are modeled as arbitrary truncated Fourier series, 
+with the period and phase shared across all bands. The key aspect is the use of Tikhonov regularization 
+(discussed in detail in \S \ref{regularization}) which drives most of the variability into the so-called base 
+model common to all bands, while fits for individual bands describe residuals relative to the base model 
+and typically require lower-order Fourier series. This decrease in the required number of fit parameters is the
+main reason for improved performance.
+
+In \S 2, we provide a brief overview of periodograms, including the matrix-based formulation used in this work. 
+XXX Finish this paragraph after the paper is nearly completed...
+
+\section{Brief Overview of Periodograms \label{sec:periodograms}}
 
@@
   \eqlabel{LombScargle}
-  P_N(\omega) = \frac{1}{2\sigma^2}\left[
-    \frac{\left[\sum_k(y_k - \bar{d})\cos\omega(t_k - \tau)\right]^2}
+  P_N(\omega) = \frac{1}{2\,V}\left[
+    \frac{\left[\sum_k(y_k - \bar{y})\cos\omega(t_k - \tau)\right]^2}
     {\sum_k \cos^2\omega(t_k - \tau)}
     +
-    \frac{\left[\sum_k(y_k - \bar{d})\sin\omega(t_k - \tau)\right]^2}
+    \frac{\left[\sum_k(y_k - \bar{y})\sin\omega(t_k - \tau)\right]^2}
     {\sum_k \sin^2\omega(t_k - \tau)}
@@
 \end{equation}
-where $\bar{d}$ is the mean and $\sigma^2$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[See][for an in-depth discussion]{NumRec}.
+where $\bar{y}$ is the mean and $V$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
 \citet{Lomb76} showed that this offset has a deeper effect: namely, it makes $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
@@
 
-\todo{Add a paragraph about Supersmoother \citep{Reimann94}, CARMA \citep{Kelly14}, and other methods}
-
-A weakness with the above methods is that they require homogeneous measurements -- for astronomy data, this means that successive measurements must be taken through a single band.
-This has not been a problem for past surveys, as measurements are generally taken through a single photometric filter (e.g.
-LINEAR, ??), or in all bands at each observation (e.g.
-SDSS, ??).
-In such cases, past studies have generally relied on \adhoc{} methods such as a majority vote among multiple single-band estimates of the periodogram.
-
-For future multicolor surveys such as LSST, this \adhoc{} approach will not be sufficient.
-In order to take advantage of the full available data, it would be desirable to have a single estimate of the periodogram which accounts for all observed data in a manner which is not dependent on the underlying spectrum of the object.
-We propose such a method below.
-
 \section{Background: Matrix Formalism for Periodograms}
 
-In this section we will give a quick quantitative introduction to the least squares fitting formulation of the normalized periodogram of \eq{LombScargle}. We'll denote our $N$ observed data points as
+In this section we present a quick quantitative introduction to the least squares fitting formulation of the normalized periodogram of \eq{LombScargle}. We denote $N$ observed data points as
 \begin{equation}
@@
 \end{equation}
-We can find the maximum likelihood estimate of the parameters $\theta$ by minimizing the $\chi^2$ of the model, which is given by
+For a given $\omega$, we can find the maximum likelihood estimate of the parameters $\theta$ by minimizing the $\chi^2$ of the model, which is given by
 \begin{equation}
-  \chi^2(\omega) = \sum_k \frac{[y_k - y(t_k|\omega,\theta)]^2}{2\sigma_k^2}.
+  \chi^2(\omega) = \sum_k \frac{[y_k - y(t_k|\omega,\theta)]^2}{\sigma_k^2}.
 \end{equation}
@@
 \end{equation}
-Take note that this expression is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto \sigma^2 I$.
+We note that this expression is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto V I$.
 
@@
   \item It is trivially extended to more sophisticated linear models by appropriately modifying the design matrix $X_\omega$.
-  \item It is trivially extended to include Tikhonov/L2-regularization terms by adding an appropriate diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$.
+  \item It is trivially extended to include Tikhonov/L2-regularization (see \S \ref{regularization} for more details)  terms by 
+               adding an appropriate diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$.
 \end{enumerate}
@@
 
-\subsubsection{Regularized Models}
-\sectlabel{regularization}
+\subsubsection{Regularized Models \label{regularization}}
 
@@
 
-We'll encode our regularization in the matrix $\Lambda = {\rm diag}([\lambda_1, \lambda_2 \cdots \lambda_M])$ for a model with $M$ parameters, and our expression for $\chi^2$ becomes
+We encode our regularization in the matrix $\Lambda = {\rm diag}([\lambda_1, \lambda_2 \cdots \lambda_M])$ for a model with $M$ parameters, and our expression for ``regularized'' $\chi^2$ becomes
 \begin{equation}

### 2015-01-12 periodogram/__init__.py
add multiband model refactors

@@ -1,2 +1,6 @@
+"""General Periodic Modeling for Astronomical Time Series"""
+
 from .lomb_scargle import LombScargle, LombScargleAstroML
+from .lomb_scargle_multiband import (LombScargleMultiband,
+                                     LombScargleMultibandFast)
 from .supersmoother import SuperSmoother

### 2015-01-12 periodogram/lomb_scargle_multiband.py

@@ -1,0 +1,221 @@
+from __future__ import division, print_function
+
+import numpy as np
+
+from .modeler import PeriodicModeler
+from.lomb_scargle import LombScargle, LombScargleAstroML
+
+
+class LombScargleMultiband(LombScargle):
+    """Multiband Periodogram Implementation
+
+    This implements the generalized multi-band periodogram described in
+    VanderPlas & Ivezic 2015.
+
+    Parameters
+    ----------
+    optimizer : PeriodicOptimizer instance
+        Optimizer to use to find the best period. If not specified, the
+        LinearScanOptimizer will be used.
+    Nterms_base : integer (default = 1)
+        number of frequency terms to use for the base model common to all bands
+    Nterms_band : integer (default = 1)
+        number of frequency terms to use for the residuals between the base
+        model and each individual band
+    reg_base : float or None (default = None)
+        amount of regularization to use on the base model parameters
+    reg_band : float or None (default = 1E-6)
+        amount of regularization to use on the band model parameters
+    regularize_by_trace : bool (default = True)
+        if True, then regularization is expressed in units of the trace of
+        the normal matrix
+    center_data : boolean (default = True)
+        if True, then center the y data prior to the fit
+
+    See Also
+    --------
+    LombScargle
+    LombScargleAstroML
+    LombScargleMultibandFast
+    """
+    def __init__(self, optimizer=None, Nterms_base=1, Nterms_band=1,
+                 reg_base=None, reg_band=1E-6, regularize_by_trace=True,
+                 center_data=True):
+        self.Nterms_base = Nterms_base
+        self.Nterms_band = Nterms_band
+        self.reg_base = reg_base
+        self.reg_band = reg_band
+        self.regularize_by_trace = regularize_by_trace
+        self.center_data = center_data
+        PeriodicModeler.__init__(self, optimizer)
+
+    def _fit(self, t, y, dy, filts):
+        self.unique_filts_ = np.unique(filts)
+        self.ymean_ = self._compute_ymean()
+
+        masks = [(filts == filt) for filt in self.unique_filts_]
+        self.ymean_by_filt_ = [LombScargle._compute_ymean(self,
+                                                          y=y[mask],
+                                                          dy=dy[mask])
+                               for mask in masks]
+
+        self.yw_ = self._construct_y(weighted=True)
[…]

### 2015-01-12 writeup/paper.tex
some text/fig updates

@@
 
+\newcommand{\apriori}{\foreign{a priori}}
 \newcommand{\adhoc}{\foreign{ad hoc}}
@@
     methods: data analysis ---
-    methods: numerical ---
     methods: statistical
@@
 minimizes a figure-of-merit, adopted as weighted sum of absolute residuals around the running mean. 
-Neither the Supersmoother algorithm nor the PDM method require a priori knowledge of the light curve shape. 
+Neither the Supersmoother algorithm nor the PDM method require \apriori{} knowledge of the light curve shape. 
 We note that neither method produces posterior probability distribution for the period but rather a single point 
@@
   \item It is trivially extended to more sophisticated linear models by appropriately modifying the design matrix $X_\omega$.
-  \item It is trivially extended to include Tikhonov/L2-regularization (see \S \ref{regularization} for more details)  terms by 
-               adding an appropriate diagonal term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$.
+  \item It is trivially extended to include Tikhonov/L2-regularization terms (see \S \ref{regularization} for more details)  by adding an appropriate diagonal term to the {\it normal matrix} $X_\omega^T\Sigma^{-1}X_\omega$.
 \end{enumerate}
@@
 The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias (right panel).
-Because of this selection effect, the mean of the observed data are a poor predictor of the true mean, causing the standard fixed-mean method to poorly fit the data.
+Because of this observational bias, the mean of the observed data are a poor predictor of the true mean, causing the standard fixed-mean method to poorly fit the data.
 The floating-mean approach is able to automatically adjust for this bias, resulting in a periodogram which readily detects the input period of 0.622 days (lower-right panel).
@@
 A natural extension is to instead use a multiple-term sinusoidal model, with frequencies at integer multiples of the fundamental frequency.
-With $M$ terms, there are $2M + 1$ free parameters, and the model is given by
+With $M$ Fourier terms, there are $2M + 1$ free parameters, and the model is given by
 \begin{equation}
@@
 Second, notice that as the number of terms is increased, the general ``background'' level of the periodogram increases.
-This is due to the fact that the periodogram is directly related to the $\chi^2$ of the fit at each frequency.
-Because of the increased model complexity of a higher-order model, it is able to better fit the data at all periods, not just the true period.
-Thus the observed ``power'' of a higher-order model will be everywhere higher than the power of a lower-order model.
+This is due to the fact that the periodogram power is inversely related to the $\chi^2$ of the fit at each frequency.
+A more flexible higher-order model can better fit the data at all periods, not just the true period.
+Thus the observed power of a higher-order Fourier model will be everywhere higher than the power of a lower-order Fourier model.
 
@@
 \end{equation}
-Minimizing this $\chi^2$, solving for $\theta$, and plugging into the expression for $P_N$ gives us the regularized counterpart of \eq{LombScargle2}:
+Minimizing this regularized $\chi^2$, solving for $\theta$, and plugging into the expression for $P_N$ gives us the regularized counterpart of \eq{LombScargle2}:
 \begin{equation}
@@
 \end{equation}
-Notice that the effect of this regularization term is to add a diagonal penalty term to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$, which has the additional feature that it can correct ill-posed models where the normal matrix is not invertible.
+Notice that the effect of this regularization term is to add a diagonal penalty to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$, which has the additional feature that it can correct ill-posed models where the normal matrix is not invertible.
 This feature of the regularization will become important below.
@@
 In \fig{regularized_example}, we compare a regularized and unregularized 20-term truncated Fourier model on our simulated RR Lyrae light curve.
-We use $\lambda = 0$ on the offset term, and make $\lambda_j$ progressively larger for each harmonic component.
+We use $\lambda = 0$ on the offset term, and make the penalty $\lambda_j$ progressively larger for each harmonic component.
 The result of the regularized model is less over-fitting to the input data, and a more distinct peaks in the associated periodogram.
@@
   \item An $M_{base}$-term truncated Fourier fit which models a latent parameter, which we'll call the ``overall variability''.
-  \item An $M_{band}$-term truncated Fourier fit which models the residual of each band from this overall variability.
+  \item A set of $M_{band}$-term truncated Fourier fits, each of which models the residual of a single band from this overall variability.
 \end{enumerate}
@@
 
-From reasoning about the free parameters, or by examining the above matrix it is easy to see that some of the parameters will be degenerate: i.e.
-$X_\omega$ is low-rank, and the problem is ill-posed.
+On examination of the above matrix, it's clear that the columns are not linearly independent (i.e.
+$X_\omega$ is low-rank), and thus the parameters of the best-fit model will be degenerate.
 Intuitively, this is due to the fact that if we add an overall offset to the base model, this can be perfectly accounted for by subtracting that same offset from each of the band columns.
@@
 
-We'll choose the latter here, and regularize all the band columns while leaving the base columns un-regularized: for the above matrix, this regularization will look like
+We'll choose the latter here, and regularize all the band columns while leaving the base columns un-regularized: for the above $X_\omega$ matrix, this regularization will look like
 \begin{equation}
@@
 \begin{equation}
-  P_N^{(0,1)}(\omega) = \frac{\sum_{k=1}^K\left[\chi^2_{0,k}P_N^{k}\right]}{\sum_{k=1}^K\chi^2_{0,k}}.
+  P_N^{(0,1)}(\omega) = \frac{\sum_{k=1}^K\left[\chi^2_{0,k}P_N^{(k)}\right]}{\sum_{k=1}^K\chi^2_{0,k}}.
 \end{equation}
@@
 Here we can clearly see the difference between the $(0,1)$ model, which we'll call {\it old-style}, and the $(1,0)$ model, which we'll call {\it new-style}.
-The old-style model, as discussed in \sectlabel{relationship}, is equivalent to a weighted sum of single-band periodograms for each individual band.
+The old-style model, as discussed in \sect{relationship}, is equivalent to a weighted sum of single-band periodograms for each individual band.
 The model thus has effectively 15 free parameters, and the phase and amplitude of each band varies individually.
 As with the truncated Fourier series, this added complexity leads to higher power at off-frequencies.
-The new-style model, on the other hand, is effectively a 7-parameter model where a single oscillation phase and amplitude is used to describe all of the data.
+The new-style model, on the other hand, is effectively a 7-parameter model where a single oscillation phase and amplitude is used to describe all the variability in the data.
 The only flexibility allowed in individual bands is the constant offset.

### 2015-01-12 multiband_LS/memoize.py

@@ -11,9 +11,12 @@ class CacheResults(object):
         self.cache_dir = os.path.abspath(cache_dir)
         self.verbose = verbose
 
-    def key_to_file(self, f, key):
+    def key_to_file(self, funcname, key):
         return os.path.join(self.cache_dir,
-                            "_{0}_{1}.npy".format(f.__name__, key))
+                            "_{0}_{1}.npy".format(funcname, key))
+
+    def __call__(self, *args, **kwargs):
+        return self.call(*args, **kwargs)
 
     def call(self, f, key, overwrite=False, args=None, kwargs=None):
         if args is None:
@@ -24,7 +27,7 @@ class CacheResults(object):
         if not os.path.exists(self.cache_dir):
             os.makedirs(self.cache_dir)
 
-        cache_file = self.key_to_file(f, key)
+        cache_file = self.key_to_file(f.__name__, key)
         if os.path.exists(cache_file) and not overwrite:
             result = np.load(cache_file)
         else:
@@ -41,3 +44,23 @@ class CacheResults(object):
     def call_iter(self, f, keys, overwrite=False, args=None, kwargs=None):
         return np.array([self.call(f, key, overwrite, args, kwargs)
                          for key in keys])
+
+    def call_iter_parallel(self, f, keys, overwrite=False, args=None,
+                           kwargs=None, client_args=None, client_kwds=None):
+        from IPython import parallel
+        if client_args is None:
+            client_args = ()
+        if client_kwds is None:
+            client_kwds = {}
+
+        try:
+            client = parallel.Client(*client_args, **client_kwds)
+        except FileNotFoundError:
+            raise ValueError("Cannot start the cluster. Did you run this?\n"
+                             " $ ipcluster start")
+
+        lbv = client.load_balanced_view()
+
+        results = lbv.map_sync(self, keys, overwrite=overwrite,
+                               args=args, kwargs=kwargs)
+        return np.array(list(results))

### 2015-01-12 writeup/figures/fig_compare_periods.py

@@ -1,3 +1,5 @@
+from __future__ import print_function, division
+
 import os, sys; sys.path.append(os.path.abspath('../..'))
 
 import numpy as np
@@ -5,10 +7,19 @@ from scipy.stats import mode
 import matplotlib.pyplot as plt
 import seaborn; seaborn.set()
 
-from multiband_LS import LombScargleAstroML, LombScargleMultiband
+from multiband_LS import (LombScargleAstroML, LombScargleMultiband,
+                          SuperSmoother)
 from multiband_LS.memoize import CacheResults
 from multiband_LS.data import fetch_light_curves
 rrlyrae = fetch_light_curves()
+
+
+def approx_mode(a, axis=0, tol=1E-3):
+    a_trunc = a // tol
+    vals, counts = mode(a_trunc, axis)
+    mask = (a_trunc == vals)
+    # mean of each row
+    return np.sum(a * mask, axis) / np.sum(mask, axis)
 
 
 def find_best_period_adhoc(lcid):
@@ -45,16 +56,41 @@ def compute_all_periods(Nterms_base, Nterms_band):
                            args=(rrlyrae, Nterms_base, Nterms_band))
 
 
+def find_best_period_supersmoother(lcid, rrlyrae):
+    t, y, dy = rrlyrae.get_lightcurve(lcid)
+    res = []
+
+    for i in range(t.shape[1]):
+        ti, yi, dyi = t[:, i], y[:, i], dy[:, i]
+        mask = (np.isnan(ti) | np.isnan(yi) | np.isnan(dyi))
+        ti, yi, dyi = ti[~mask], yi[~mask], dyi[~mask]
+    
+        model = SuperSmoother(period_range=(0.2, 1.2))
+        model.fit(ti, yi, dyi)
+        res.append(model.best_period)
+    return res
+
+
+def compute_all_periods_supersmoother():
+    lcids = list(rrlyrae.ids)
+    cache = CacheResults('results_supersmoother', verbose=True)
+    results = cache.call_iter(find_best_period_supersmoother, lcids,
+                              args=(rrlyrae,))
+    return np.array(results)
+
+
 new_periods = compute_all_periods(1, 0)
 old_periods = compute_all_periods(0, 1)
 sesar_periods = np.array([rrlyrae.get_metadata(lcid)['P']
[…]

### 2015-01-12 writeup/figures/fig_floating_mean.py

@@ -40,7 +40,7 @@ mask = mag < 16.05
 phasefit = np.linspace(0, 1, 1000)
 tfit = rrlyrae.period * phasefit
 
-periods = np.linspace(0.2, 1.1, 1000)
+periods = np.linspace(0.2, 1.4, 1000)
 omegas = 2 * np.pi / periods
 
 fig = plt.figure(figsize=(10, 4))
@@ -71,7 +71,7 @@ for fit_offset in [False, True]:
     ax[1 + i].set_title('{0} Periodogram'.format(label.title()))
     ax[1 + i].set_ylabel('power')
     ax[1 + i].set_ylim(0, 1)
-    ax[1 + i].set_xlim(0.2, 1.1)
+    ax[1 + i].set_xlim(0.2, 1.4)
 
 ax[0].legend(loc='upper left')
 ax[0].set_xlabel('phase')

### 2015-01-12 multiband_LS/base.py
add base class

@@ -1,0 +1,195 @@
+from __future__ import division, print_function
+
+import sys
+
+import numpy as np
+
+
+class PeriodicModeler(object):
+    """Base class for periodic modeling"""
+    def __init__(self, *args, **kwargs):
+        raise NotImplementedError("PeriodicModeler")
+
+    def fit(self, t, y, dy, filts=None):
+        raise NotImplementedError()
+
+    def score(self, period):
+        raise NotImplementedError()
+
+    def best_params(self, omega=None):
+        """Compute the maximum likelihood model parameters at frequency omega
+
+        Parameters
+        ----------
+        omega : float (optional)
+            The angular frequency at which to compute the best parameters.
+            if not specified, it will be computed via self.find_best_omega().
+
+        Returns
+        -------
+        theta : np.ndarray
+            The array of model parameters for the best-fit model at omega
+        """
+        if not hasattr(self, 'fit_data_'):
+            raise ValueError("Must call fit() before periodogram/best_params")
+        if omega is None:
+            omega = self.best_omega
+        return self._best_params(omega)
+
+    def _best_params(self, omega):
+        raise NotImplementedError()
+
+    def predict(self, t, filts=None, omega=None):
+        """Compute the best-fit model at ``t`` for a given frequency omega
+
+        Parameters
+        ----------
+        t : float or array_like
+            times at which to predict
+        filts : array_like (optional)
+            the array specifying the filter/bandpass for each observation. This
+            is used only in multiband periodograms.
+        omega : float (optional)
+            The angular frequency at which to compute the best parameters.
+            if not specified, it will be computed via self.find_best_omega().
+
+        Returns
+        -------
+        y : np.ndarray
+            predicted model values at times t
+        """
+        if not hasattr(self, 'fit_data_'):
+            raise ValueError("Must call fit() before predict")
+        if omega is None:
+            omega = self.best_omega
+        return self._predict(t=t, omega=omega, filts=filts)
+
+    def _predict(self, t, omega, filts=None):
+        raise NotImplementedError()
+
+    def __call__(self, omegas):
+        return self.periodogram(omegas)
+
+    @property
+    def best_omega(self):
+        if not hasattr(self, '_best_omega'):
+            self._best_omega = self.find_best_omega()
+        return self._best_omega
+
+    @property
+    def best_period(self):
+        return 2 * np.pi / self.best_omega
+
+    def find_best_omega(self, omega_min=None, omega_max=None,
+                        Nzooms=10, verbose=1):
[…]

### 2015-01-12 multiband_LS/__init__.py
initial move of refactored code

@@ -1,7 +1,6 @@
-"""
-multiband_LS: tools for finding periods in multiband time series
-"""
+"""General Periodic Modeling for Astronomical Time Series"""
 
-from .lomb_scargle import (LombScargle, LombScargleAstroML,
-                           LombScargleMultiband, LombScargleMultibandFast)
+from .lomb_scargle import LombScargle, LombScargleAstroML
+from .lomb_scargle_multiband import (LombScargleMultiband,
+                                     LombScargleMultibandFast)
 from .supersmoother import SuperSmoother

### 2015-01-12 multiband_LS/decorators.py

@@ -1,41 +1,0 @@
-# hash which appropriately hashes numpy arrays
-from sklearn.externals.joblib import hash
-import functools
-
-
-def cache_results(arg, filename=None, overwrite=False):
-    """Generator for a decorator which allows cacheing results"""
-    if type(arg) == str:
-        if filename is None:
-            raise ValueError("unrecognized input!")
-        return functools.partial(cache_results, filename=arg, overwrite=False)
-    elif filename is None:
-        raise ValueError("must include filename")
-
-    func = arg
-
-    if not callable(func):
-        raise ValueError("cache_results must be applied to a function")
-
-    def new_func(*args, **kwargs):
-        arg_hash = hash([args, kwargs])
-
-        if os.path.exists(filename):
-            cached = np.load(filename)
-            if cached['arg_hash'].flat[0] == arg_hash:
-                nretvals = cached['nreturns']
-                if cached['nretvals'] > 1:
-                    return tuple(cached['retvals'])
-                else:
-                    return cached['retvals']
-            elif not overwrite:
-                raise ValueError("Cache file exists and input does not match. "
-                                 "Set overwrite=True to force overwrite")
-
-            # otherwise, do the computation
-            retvals = func(*args, **kwargs)
-                
-            if type(retvals) is tuple:
-                nretvals = len(retvals)
-            else:
-                nretvals = 1

### 2015-01-12 multiband_LS/lomb_scargle.py

@@ -1,12 +1,8 @@
 from __future__ import division, print_function
-__all__ = ['LombScargle', 'LombScargleAstroML',
-           'LombScargleMultiband', 'LombScargleMultibandFast']
-
-import warnings
 
 import numpy as np
 
-from .base import PeriodicModeler
+from .modeler import PeriodicModeler
 
 
 class LombScargle(PeriodicModeler):
@@ -19,6 +15,9 @@ class LombScargle(PeriodicModeler):
 
     Parameters
     ----------
+    optimizer : PeriodicOptimizer instance
+        Optimizer to use to find the best period. If not specified, the
+        LinearScanOptimizer will be used.
     center_data : boolean (default = True)
         If True, then compute the weighted mean of the input data and subtract
         before fitting the model.
@@ -40,10 +39,12 @@ class LombScargle(PeriodicModeler):
     >>> omega = 10
     >>> y = np.sin(omega * t) + dy * rng.randn(100)
     >>> ls = LombScargle().fit(t, y, dy)
-    >>> ls.best_omega
-    10.000707659055472
+    >>> ls.best_period
+    0.62827393156409295
+    >>> ls.score(ls.best_period)
+    array(0.9815178000850804)
     >>> ls.predict([0, 0.5])
-    array([-0.01203899, -0.92698316])
+    array([-0.01213809, -0.92700951])
 
     See Also
     --------
@@ -51,15 +52,14 @@ class LombScargle(PeriodicModeler):
     LombScargleMultiband
     LombScargleMultibandFast
     """
-    def __init__(self, center_data=True, fit_offset=True, Nterms=1,
-                 regularization=None, regularize_by_trace=True,
-                 period_range=(0.2, 1.1)):
+    def __init__(self, optimizer=None, center_data=True, fit_offset=True,
+                 Nterms=1, regularization=None, regularize_by_trace=True):
         self.center_data = center_data
         self.fit_offset = fit_offset
         self.Nterms = int(Nterms)
         self.regularization = regularization
         self.regularize_by_trace = regularize_by_trace
-        self.period_range = period_range
+        PeriodicModeler.__init__(self, optimizer)
 
         if not self.center_data and not self.fit_offset:
             warnings.warn("Not centering data or fitting offset can lead "
@@ -73,8 +73,8 @@ class LombScargle(PeriodicModeler):
 
     def _construct_X(self, omega, weighted=True, **kwargs):
[…]

### 2015-01-12 multiband_LS/period_search.py

@@ -1,52 +1,0 @@
-from __future__ import division, print_function
-
-import numpy as np
-from astroML.time_series import lomb_scargle as lomb_scargle_astroML
-
-
-def _single_curve(t, y, dy, i):
-    t = t[:, i]
-    y = y[:, i]
-    dy = dy[:, i]
-
-    good = ~(np.isnan(t) | np.isnan(y) | np.isnan(dy))
-
-    return (t[good], y[good], dy[good])    
-
-
-def period_search(t, y, dy, pmin=0.2, pmax=1.0, N=50000,
-                  LS_func=None, LS_kwargs=None):
-    """Search periods band-by-band"""
-
-    if LS_func is None:
-        LS_func = lomb_scargle_astroML
-    if LS_kwargs is None:
-        LS_kwargs = {}
-
-    # do a coarse-grained zoom for the first pass
-    data = [_single_curve(t, y, dy, i) for i in range(5)]
-    period = np.linspace(pmin, pmax, N)
-    dp = period[1] - period[0]
-
-    p_best = []
-    for i, filt in enumerate('ugriz'):
-        # shift period for greater coverage
-        omega = 2 * np.pi / (period + 0.2 * i * dp)
-        t, y, dy = data[i]
-        P = LS_func(t, y, dy, omega, **LS_kwargs)
-        p_best.append(period[np.argmax(P)])
-
-    # now zoom-in around the best periods
-    p_best = np.unique(0.001 * (np.array(p_best) // 0.001))
-
-    period = np.concatenate([np.linspace(p - 0.05, p + 0.05, N)
-                             for p in p_best])
-    omega = 2 * np.pi / period
-    
-    p_best = []
-    for i, filt in enumerate('ugriz'):
-        t, y, dy = data[i]
-        P = LS_func(t, y, dy, omega, **LS_kwargs)
-        p_best.append(period[np.argmax(P)])
-
-    return np.array(p_best)

### 2015-01-12 multiband_LS/supersmoother.py

@@ -1,99 +1,81 @@
-from __future__ import division, print_function
+from __future__ import print_function, division
+
+"""
+Supersmoother code for periodic modeling
+"""
+import numpy as np
+
 
 try:
     import supersmoother as ssm
 except ImportError:
-    raise ImportError("Package supersmoother required "
-                      "(``pip install supersmoother``)")
+    raise ImportError("Package supersmoother is required. "
+                      "Use ``pip install supersmoother`` to install")
 
-import numpy as np
 
-from .base import PeriodicModeler
+from .modeler import PeriodicModeler
 
 
 class SuperSmoother(PeriodicModeler):
-    """SuperSmoother periodogram implementation
+    """Periodogram based on Friedman's SuperSmoother.
+
+    Parameters
+    ----------
+    optimizer : PeriodicOptimizer instance
+        Optimizer to use to find the best period. If not specified, the
+        LinearScanOptimizer will be used.
+
+    Examples
+    --------
+    >>> rng = np.random.RandomState(0)
+    >>> t = 100 * rng.rand(100)
+    >>> dy = 0.1
+    >>> omega = 10
+    >>> y = np.sin(omega * t) + dy * rng.randn(100)
+    >>> ssm = SuperSmoother().fit(t, y, dy)
+    >>> ssm.best_period
+    0.62826749832108475
+    >>> ssm.score(ls.best_period)
+    array(0.9951882158877049)
+    >>> ssm.predict([0, 0.5])
+    array([ 0.06759746, -0.90006247])
 
     See Also
     --------
     LombScargle
     """
-    def __init__(self, period_range=(0.2, 1.1)):
-        self.period_range = period_range
+    def __init__(self, optimizer=None):
+        PeriodicModeler.__init__(self, optimizer)
 
-    def fit(self, t, y, dy=1.0, filts=None):
-        """Fit the Supersmoother model to the data.
+    def _fit(self, t, y, dy, filts):
+        if filts is not None:
+            raise NotImplementedError("``filts`` keyword is not supported")
 
-        Parameters
-        ----------
-        t : array_like, one-dimensional
-            sequence of observation times
-        y : array_like, one-dimensional
-            sequence of observed values
-        dy : float or array_like
-            errors on observed values
-        """
-        t, y, dy = np.broadcast_arrays(t, y, dy)
-        self.fit_data_ = dict(t=t, y=y, dy=dy)
-
[…]

### 2015-01-12 multiband_LS/tests/test_lomb_scargle.py

@@ -1,133 +1,0 @@
-import numpy as np
-from numpy.testing import assert_allclose, assert_
-from nose import SkipTest
-
-from ..lomb_scargle import (LombScargle, LombScargleAstroML,
-                            LombScargleMultiband, LombScargleMultibandFast)
-
-
-def _generate_data(N=100, omega=10, theta=[10, 2, 3], dy=1, rseed=0):
-    """Generate some data for testing"""
-    rng = np.random.RandomState(rseed)
-    t = 20 * (2 * np.pi / omega) * rng.rand(N)
-    y = theta[0] + theta[1] * np.sin(omega * t) + theta[2] * np.cos(omega * t)
-    dy = dy * (0.5 + rng.rand(N))
-    y += dy * rng.randn(N)
-    return t, y, dy
-
-
-def test_lomb_scargle(N=100, omega=10):
-    """Test whether the standard and generalized lomb-scargle
-    give close to the same results for non-centered data"""
-    t, y, dy = _generate_data(N, omega)
-    omegas = np.linspace(1, omega + 1, 100)
-
-    P1 = LombScargle(fit_offset=True).fit(t, y, dy).periodogram(omegas)
-    P2 = LombScargle(fit_offset=False).fit(t, y, dy).periodogram(omegas)
-
-    rms = np.sqrt(np.mean((P1 - P2) ** 2))
-    assert_(rms < 0.005)
-
-
-def test_dy_scalar(N=100, omega=10):
-    """Test whether the standard and generalized lomb-scargle
-    give close to the same results for non-centered data"""
-    t, y, dy = _generate_data(N, omega)
-
-    # Make dy array all the same
-    dy[:] = dy.mean()
-    omegas = np.linspace(1, omega + 1, 100)
-
-    def compare(cls):
-        P1 = cls().fit(t, y, dy).periodogram(omegas)
-        P2 = cls().fit(t, y, dy[0]).periodogram(omegas)
-
-        assert_allclose(P1, P2)
-
-    for cls in [LombScargle, LombScargleAstroML]:
-        yield compare, cls
-
-
-def test_lomb_scargle_multiband(N=100, omega=10):
-    """Test that results are the same with/without filter labels"""
-    t, y, dy = _generate_data(N, omega)
-    omegas = np.linspace(1, omega + 1, 100)
-    model = LombScargle(center_data=False, fit_offset=True)
-    P_singleband = model.fit(t, y, dy).periodogram(omegas)
-
-    filts = np.ones(N)
[…]

### 2015-01-12 periodogram/__init__.py

@@ -1,6 +1,0 @@
-"""General Periodic Modeling for Astronomical Time Series"""
-
-from .lomb_scargle import LombScargle, LombScargleAstroML
-from .lomb_scargle_multiband import (LombScargleMultiband,
-                                     LombScargleMultibandFast)
-from .supersmoother import SuperSmoother

### 2015-01-12 periodogram/lomb_scargle.py

@@ -1,237 +1,0 @@
-from __future__ import division, print_function
-
-import numpy as np
-
-from .modeler import PeriodicModeler
-
-
-class LombScargle(PeriodicModeler):
-    """Lomb-Scargle Periodogram Implementation
-
-    This is a generalized periodogram implementation using the matrix formalism
-    outlined in VanderPlas & Ivezic 2015. It allows computation of periodograms
-    and best-fit models for both the classic normalized periodogram and
-    truncated Fourier series generalizations.
-
-    Parameters
-    ----------
-    optimizer : PeriodicOptimizer instance
-        Optimizer to use to find the best period. If not specified, the
-        LinearScanOptimizer will be used.
-    center_data : boolean (default = True)
-        If True, then compute the weighted mean of the input data and subtract
-        before fitting the model.
-    fit_offset : boolean (default = True)
-        If True, then fit a floating-mean sinusoid model.
-    Nterms : int (default = 1)
-        Number of Fourier frequencies to fit in the model
-    regularization : float, vector or None (default = None)
-        If specified, then add this regularization penalty to the
-        least squares fit.
-    regularize_by_trace : boolean (default = True)
-        If True, multiply regularization by the trace of the matrix
-
-    Examples
-    --------
-    >>> rng = np.random.RandomState(0)
-    >>> t = 100 * rng.rand(100)
-    >>> dy = 0.1
-    >>> omega = 10
-    >>> y = np.sin(omega * t) + dy * rng.randn(100)
-    >>> ls = LombScargle().fit(t, y, dy)
-    >>> ls.best_period
-    0.62827393156409295
-    >>> ls.score(ls.best_period)
-    array(0.9815178000850804)
-    >>> ls.predict([0, 0.5])
-    array([-0.01213809, -0.92700951])
-
-    See Also
-    --------
-    LombScargleAstroML
-    LombScargleMultiband
-    LombScargleMultibandFast
-    """
-    def __init__(self, optimizer=None, center_data=True, fit_offset=True,
-                 Nterms=1, regularization=None, regularize_by_trace=True):
-        self.center_data = center_data
-        self.fit_offset = fit_offset
-        self.Nterms = int(Nterms)
-        self.regularization = regularization
-        self.regularize_by_trace = regularize_by_trace
-        PeriodicModeler.__init__(self, optimizer)
-
-        if not self.center_data and not self.fit_offset:
[…]

### 2015-01-12 periodogram/supersmoother.py

@@ -1,88 +1,0 @@
-from __future__ import print_function, division
-
-"""
-Supersmoother code for periodic modeling
-"""
-import numpy as np
-
-
-try:
-    import supersmoother as ssm
-except ImportError:
-    raise ImportError("Package supersmoother is required. "
-                      "Use ``pip install supersmoother`` to install")
-
-
-from .modeler import PeriodicModeler
-
-
-class SuperSmoother(PeriodicModeler):
-    """Periodogram based on Friedman's SuperSmoother.
-
-    Parameters
-    ----------
-    optimizer : PeriodicOptimizer instance
-        Optimizer to use to find the best period. If not specified, the
-        LinearScanOptimizer will be used.
-
-    Examples
-    --------
-    >>> rng = np.random.RandomState(0)
-    >>> t = 100 * rng.rand(100)
-    >>> dy = 0.1
-    >>> omega = 10
-    >>> y = np.sin(omega * t) + dy * rng.randn(100)
-    >>> ssm = SuperSmoother().fit(t, y, dy)
-    >>> ssm.best_period
-    0.62826749832108475
-    >>> ssm.score(ls.best_period)
-    array(0.9951882158877049)
-    >>> ssm.predict([0, 0.5])
-    array([ 0.06759746, -0.90006247])
-
-    See Also
-    --------
-    LombScargle
-    """
-    def __init__(self, optimizer=None):
-        PeriodicModeler.__init__(self, optimizer)
-
-    def _fit(self, t, y, dy, filts):
-        if filts is not None:
-            raise NotImplementedError("``filts`` keyword is not supported")
-
-        # TODO: this should actually be a weighted median, probably...
-        mu = np.sum(y / dy ** 2) / np.sum(1 / dy ** 2)
-        self.baseline_err = np.mean(abs((y - mu) / dy))
-
-    def _predict(self, t, filts, period):
-        model = ssm.SuperSmoother().fit(self.t % period, self.y, self.dy)
-        return model.predict(t % period)
-
-    def _score(self, periods):
-        # double-up the data to allow periodicity on the fits
-        N = len(self.t)
-        N4 = N // 4
-        t = np.concatenate([self.t, self.t])
-        y = np.concatenate([self.y, self.y])
-        dy = np.concatenate([self.dy, self.dy])
-
-        results = []
-        for p in periods:
-            # compute doubled phase and sort
-            phase = t % p
-            phase[N:] += p
[…]

### 2015-01-12 multiband_LS/interactive.py
finish refactor of multiband_LS code

@@ -4,7 +4,7 @@ Utilities for creating IPython interactive plots to explore data & fits
 import numpy as np
 from IPython.html.widgets import interact
 from .data import fetch_light_curves
-from .lomb_scargle import LombScargleMultiband
+from . import LombScargleMultiband
 
 
 def interact_data():
@@ -60,7 +60,7 @@ def interact_multifit():
                                      Nterms_band=Nterms_band)
         model.fit(t, y, dy, filts)
         tfit = np.linspace(0, period, 500)[:, None]
-        yfits = model.predict(tfit, ['u', 'g', 'r', 'i', 'z'], omega)
+        yfits = model.predict(tfit, ['u', 'g', 'r', 'i', 'z'], period=period)
         colors = plt.rcParams['axes.color_cycle']
         
         ax0 = fig.add_subplot(gs[:2, :2])

### 2015-01-12 multiband_LS/lomb_scargle.py

@@ -195,6 +195,8 @@ class LombScargleAstroML(LombScargle):
     optimizer : PeriodicOptimizer instance
         Optimizer to use to find the best period. If not specified, the
         LinearScanOptimizer will be used.
+    Nterms : int (default = 1)
+        Number of terms for the fit. Only Nterms=1 is currently supported.
     center_data : boolean (default = True)
         If True, then compute the weighted mean of the input data and subtract
         before fitting the model.
@@ -221,8 +223,11 @@ class LombScargleAstroML(LombScargle):
     LombScargleMultiband
     LombScargleMultibandFast
     """
-    def __init__(self, optimizer=None, fit_offset=True, center_data=True,
-                 slow_version=False):
+    def __init__(self, optimizer=None, Nterms=1, fit_offset=True,
+                 center_data=True, slow_version=False):
+        if Nterms != 1:
+            raise ValueError("Only Nterms=1 is supported")
+
         LombScargle.__init__(self, optimizer=optimizer, Nterms=1,
                              center_data=center_data, fit_offset=fit_offset)
         if slow_version:

### 2015-01-12 multiband_LS/lomb_scargle_multiband.py

@@ -3,7 +3,7 @@ from __future__ import division, print_function
 import numpy as np
 
 from .modeler import PeriodicModeler
-from.lomb_scargle import LombScargle, LombScargleAstroML
+from.lomb_scargle import LombScargle
 
 
 class LombScargleMultiband(LombScargle):
@@ -135,17 +135,19 @@ class LombScargleMultiband(LombScargle):
         else:
             return np.transpose(np.vstack(cols))
 
-    def _predict(self, t, omega, filts=None):
+    def _predict(self, t, filts, period):
         vals = set(np.unique(filts))
         if not vals.issubset(self.unique_filts_):
             raise ValueError("filts does not match training data: "
-                             "{0}".format(set(self.unique_filts_) - vals))
+                             "input: {0} output: {1}"
+                             "".format(set(self.unique_filts_), set(vals)))
 
         t, filts = np.broadcast_arrays(t, filts)
         output_shape = t.shape
 
         t = t.ravel()
         filts = filts.ravel()
+        omega = 2 * np.pi / period
 
         # TODO: broadcast this
         ymeans = np.zeros(len(filts))
@@ -172,8 +174,8 @@ class LombScargleMultibandFast(PeriodicModeler):
         LinearScanOptimizer will be used.
     Nterms : integer (default = 1)
         Number of fourier terms to use in themodel
-    BaseModel : class type (default = LombScargleAstroML)
-        The base model to use for each individual band
+    BaseModel : class type (default = LombScargle)
+        The base model to use for each individual band.
 
     See Also
     --------
@@ -181,11 +183,10 @@ class LombScargleMultibandFast(PeriodicModeler):
     LombScargleAstroML
     LombScargleMultiband
     """
-    def __init__(self, optimizer=None, Nterms=1, BaseModel=LombScargleAstroML):
+    def __init__(self, optimizer=None, Nterms=1, BaseModel=LombScargle):
         # Note: center_data must be True, or else the chi^2 weighting will fail
         self.Nterms = Nterms
         self.BaseModel = BaseModel
-        self.period_range = period_range
         PeriodicModeler.__init__(self, optimizer)
 
     def _fit(self, t, y, dy, filts):
@@ -206,16 +207,17 @@ class LombScargleMultibandFast(PeriodicModeler):
         return np.asarray([model._best_params(omega)
                            for model in self.models_])
 
-    def _predict(self, t, period, filts):
+    def _predict(self, t, filts, period):
         vals = set(np.unique(filts))
[…]

### 2015-01-12 multiband_LS/modeler.py

@@ -38,7 +38,7 @@ class PeriodicModeler(object):
         else:
             self.t, self.y, self.dy, self.filts = np.broadcast_arrays(t, y, dy,
                                                                       filts)
-        self._fit(t, y, dy, filts)
+        self._fit(self.t, self.y, self.dy, filts=self.filts)
         return self
 
     def predict(self, t, filts=None, period=None):
@@ -67,12 +67,13 @@ class PeriodicModeler(object):
             t = np.asarray(t)
             if self.filts is not None:
                 raise ValueError("filts must be passed")
-            result = self._predict(t.ravel(), filts, period)
+            result = self._predict(t.ravel(), filts=filts, period=period)
         else:
             t, filts = np.broadcast_arrays(t, filts)
             if self.filts is None:
                 raise ValueError("filts passed to predict(), but not to fit()")
-            result = self._predict(t.ravel(), filts.ravel(), period)
+            result = self._predict(t.ravel(), filts=filts.ravel(),
+                                   period=period)
         return result.reshape(t.shape)
 
     def score(self, periods):
@@ -92,6 +93,8 @@ class PeriodicModeler(object):
         """
         periods = np.asarray(periods)
         return self._score(periods.ravel()).reshape(periods.shape)
+
+    periodogram = score
 
     @property
     def best_period(self):

### 2015-01-12 writeup/figures/fig_basic_example.py
update figure code for refactor

@@ -31,10 +31,8 @@ nights = nights[:Nobs]
 t = 57000 + nights + 0.05 * rng.randn(Nobs)
 dmag = 0.06 + 0.01 * rng.randn(Nobs)
 mag = rrlyrae.generated('r', t, err=dmag, corrected=False)
-omega_best = 2 * np.pi / rrlyrae.period
 
 periods = np.linspace(0.2, 1.4, 1000)
-omegas = 2 * np.pi / periods
 
 phase = (t / rrlyrae.period) % 1
 phasefit = np.linspace(0, 1, 1000)
@@ -58,10 +56,9 @@ model = LombScargle().fit(t, mag, dmag)
 phase = (t / model.best_period) % 1
 phasefit = np.linspace(0, 1, 1000)
 tfit = model.best_period * phasefit
-P = model.periodogram(omegas)
         
-lines = ax[2].plot(periods, model.periodogram(omegas), lw=1)
-ax[1].plot(phasefit, model.predict(tfit, omega=omega_best),
+lines = ax[2].plot(periods, model.periodogram(periods), lw=1)
+ax[1].plot(phasefit, model.predict(tfit),
            c=lines[0].get_color())
 ax[1].invert_yaxis()
 

### 2015-01-12 writeup/figures/fig_floating_mean.py

@@ -32,7 +32,6 @@ t = 57000 + nights + 0.1 * rng.randn(Nobs)
 dmag = 0.06 + 0.01 * rng.randn(Nobs)
 mag = (16 + 0.5 * np.sin(2 * np.pi * t / rrlyrae.period - 0.4)
        + dmag * rng.randn(Nobs))
-omega_best = 2 * np.pi / rrlyrae.period
 phase = (t / rrlyrae.period) % 1
 
 mask = mag < 16.05
@@ -41,7 +40,6 @@ phasefit = np.linspace(0, 1, 1000)
 tfit = rrlyrae.period * phasefit
 
 periods = np.linspace(0.2, 1.4, 1000)
-omegas = 2 * np.pi / periods
 
 fig = plt.figure(figsize=(10, 4))
 gs = plt.GridSpec(2, 2, left=0.07, right=0.95, wspace=0.15, bottom=0.15)
@@ -59,13 +57,13 @@ for fit_offset in [False, True]:
     i = int(fit_offset)
     model = LombScargle(fit_offset=fit_offset).fit(t[mask],
                                                    mag[mask], dmag[mask])
-    P = model.periodogram(omegas)
+    P = model.periodogram(periods)
     if fit_offset:
         label = 'floating mean'
     else:
         label = 'standard'
     lines = ax[0].plot(phasefit,
-                       model.predict(tfit, omega=2 * np.pi / rrlyrae.period),
+                       model.predict(tfit, period=rrlyrae.period),
                        label=label)
     ax[1 + i].plot(periods, P, lw=1, c=lines[0].get_color())
     ax[1 + i].set_title('{0} Periodogram'.format(label.title()))

### 2015-01-12 writeup/figures/fig_multiband_sim.py

@@ -14,7 +14,7 @@ import matplotlib.pyplot as plt
 import seaborn; seaborn.set()
 
 from multiband_LS.generated import RRLyraeObject
-from multiband_LS.lomb_scargle import LombScargleAstroML, LombScargleMultiband
+from multiband_LS import LombScargleAstroML, LombScargleMultiband
 
 
 # Choose a Sesar 2010 object to base our fits on
@@ -40,8 +40,7 @@ mags = np.array([rrlyrae.generated(band, t, err=dy, corrected=False)
 # Compute the lomb-scargle periodogram in each band
 
 periods = np.linspace(0.2, 0.9, 1000)
-omegas = 2 * np.pi / periods
-P = [LombScargleAstroML().fit(t, m, dy).periodogram(omegas) for m in mags]
+P = [LombScargleAstroML().fit(t, m, dy).periodogram(periods) for m in mags]
 
 fig, ax = plt.subplots(1, 2, figsize=(10, 4))
 fig.subplots_adjust(left=0.07, right=0.95, wspace=0.1, bottom=0.15)
@@ -79,7 +78,7 @@ mags = mags[np.arange(Nobs) % 5, np.arange(Nobs)]
 masks = [(filts == band) for band in 'ugriz']
 
 P = [LombScargleAstroML().fit(t[mask], mags[mask],
-                              dy[mask]).periodogram(omegas)
+                              dy[mask]).periodogram(periods)
      for mask in masks]
 
 fig = plt.figure(figsize=(10, 4))
@@ -112,7 +111,7 @@ ax[1].set_ylabel('power + offset' + 30 * ' ')
 for (i, Nbase, Nband) in [(0, 1, 0), (1, 0, 1)]:
     LS_multi = LombScargleMultiband(Nterms_base=Nbase, Nterms_band=Nband)
     LS_multi.fit(t, mags, dy, filts)
-    P_multi = LS_multi.periodogram(omegas)
+    P_multi = LS_multi.periodogram(periods)
     ax[2].plot(periods, P_multi + i, lw=1, color='gray')
     ax[2].text(0.89, 1.0 + i,
                "Nbase={0}, Nband={1}".format(Nbase, Nband),

### 2015-01-12 writeup/figures/fig_multiterm_example.py

@@ -32,10 +32,8 @@ nights = nights[:Nobs]
 t = 57000 + nights + 0.05 * rng.randn(Nobs)
 dmag = 0.06 + 0.01 * rng.randn(Nobs)
 mag = rrlyrae.generated('r', t, err=dmag, corrected=False)
-omega_best = 2 * np.pi / rrlyrae.period
 
 periods = np.linspace(0.2, 1.4, 1000)
-omegas = 2 * np.pi / periods
 
 phase = (t / rrlyrae.period) % 1
 phasefit = np.linspace(0, 1, 1000)
@@ -58,15 +56,15 @@ models = [1, 2, 6]
 
 for i, Nterms in enumerate(models):
     model = LombScargle(Nterms=Nterms).fit(t, mag, dmag)
-    P = model.periodogram(omegas)
+    P = model.periodogram(periods)
 
     label = "{0} terms".format(Nterms)
     if Nterms == 1:
         label = label[:-1]
         
-    lines = ax[0].plot(phasefit, model.predict(tfit, omega=omega_best),
+    lines = ax[0].plot(phasefit, model.predict(tfit, period=rrlyrae.period),
                        label=label)
-    ax[1 + i].plot(periods, model.periodogram(omegas),
+    ax[1 + i].plot(periods, model.periodogram(periods),
                    c=lines[0].get_color(), lw=1)
     ax[1 + i].set_title("{0}-term Periodogram".format(Nterms))
     ax[1 + i].set_xlim(0.2, 1.4)

### 2015-01-12 writeup/figures/fig_regularization_example.py

@@ -31,10 +31,8 @@ nights = nights[:Nobs]
 t = 57000 + nights + 0.05 * rng.randn(Nobs)
 dmag = 0.06 + 0.01 * rng.randn(Nobs)
 mag = rrlyrae.generated('r', t, err=dmag, corrected=False)
-omega_best = 2 * np.pi / rrlyrae.period
 
 periods = np.linspace(0.2, 1.4, 1000)
-omegas = 2 * np.pi / periods
 
 phase = (t / rrlyrae.period) % 1
 phasefit = np.linspace(0, 1, 1000)
@@ -59,17 +57,16 @@ models = [0.5 * sigma_r_inv ** 2, None]
 for i, reg in enumerate(models):
     model = LombScargle(Nterms=Nterms, regularization=reg,
                         regularize_by_trace=False).fit(t, mag, dmag)
-    P = model.periodogram(omegas)
 
     if reg is None:
         label = "unregularized"
     else:
         label = "regularized"
         
-    lines = ax[0].plot(phasefit, model.predict(tfit, omega=omega_best),
+    lines = ax[0].plot(phasefit, model.predict(tfit, period=rrlyrae.period),
                        label=label)
 
-    ax[1 + i].plot(periods, model.periodogram(omegas), lw=1,
+    ax[1 + i].plot(periods, model.periodogram(periods), lw=1,
                    c=lines[0].get_color())
     ax[1 + i].set_title("{0} Periodogram ({1} terms)".format(label.title(),
                                                              Nterms))

### 2015-01-21 writeup/paper.tex
rework paper and figures

@@
 
-The most popular parametric method is the Lomb-Scargle periodogram (discussed in detail in \S~\ref{sec:periodograms}.
+The most popular parametric method is the Lomb-Scargle periodogram, which is discussed in detail in \S~\ref{sec:periodograms}.
 The Lomb-Scargle periodogram is related to the $\chi^2$ for a least-square fit of a single sinusoid to data
@@
 multi-band case. The light curves in  each band are modeled as arbitrary truncated Fourier series, 
-with the period and phase shared across all bands. The key aspect is the use of Tikhonov regularization 
-(discussed in detail in \S \ref{regularization}) which drives most of the variability into the so-called base 
-model common to all bands, while fits for individual bands describe residuals relative to the base model 
+with the period, and optionally the phase, shared across all bands. The key aspect enabling this approach is the use of Tikhonov regularization 
+(discussed in detail in \S \ref{regularization}) which drives most of the variability into the so-called {\it base 
+model} common to all bands, while fits for individual bands describe residuals relative to the base model 
 and typically require lower-order Fourier series. This decrease in the required number of fit parameters is the
@@
 
-In \S 2, we provide a brief overview of periodograms, including the matrix-based formulation used in this work. 
-XXX Finish this paragraph after the paper is nearly completed...
+\todo{In \S 2, we provide a brief overview of periodograms, including the matrix-based formulation used in this work. In \S...}
 
@@
 where $\bar{y}$ is the mean and $V$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
-\citet{Lomb76} showed that this offset has a deeper effect: namely, it makes $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
+\citet{Lomb76} showed that this time-offset has a deeper effect: namely, it makes $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
 
@@
 Building on this result, \citet{Bretthorst88} explored the extension of these methods to more complicated models with multiple frequency terms, nonstationary frequencies, and other more sophisticated models within a Bayesian framework.
+
+While the important features of least squares frequency estimation via Lomb-Scargle periodograms has been discussed elsewhere, we will prevent a brief introduction to the subject in the following section.
+The matrix-based formalism we develop here will make clear how the method can be extended to more sophisticated models, including regularized multi-band methods.
 
@@
 \end{equation}
-With these definitions, the model in \eq{simplemodel} can be expressed as a simple linear product: $y(t|\omega,\theta) = X_\omega\theta$, and the model and reference $\chi^2$ can be written
+With these definitions, the model in \eq{simplemodel} can be expressed as a simple linear product, $y(t|\omega,\theta) = X_\omega\theta$, and the model and reference $\chi^2$ can be written
 
@@
 \end{eqnarray}
-The expression for the normalized periodogram can be computed by finding the value of $\theta$ which minizes $\chi^2(\omega)$ via standard methods and plugging the result into \eq{chi2PN} to find
+The expression for the normalized periodogram can be computed by finding via standard methods the value of $\theta$ which minizes $\chi^2(\omega)$, and plugging the result into \eq{chi2PN}.
+This yields
 \begin{equation}
@@
 This additional peak is due to beat frequency between the true period $P$ and the observing cadence of $\sim 1$ day.
-This beat frequency is the first in a large sequence: for nightly observations, we'd expect to find excess power at periods That is, for nightly observations, we'd expect excess power to be found at frequencies $P_n = P / (1 + nP)$ days, for any positive or negative integer $n$.
+This beat frequency is the first in a large sequence: for nightly observations, we'd expect to find excess power at periods $P_n = P / (1 + nP)$ days, for any integer $n$.
 The strong alias in \fig{basic_example} corresponds to the $n=1$ beat period $P_n=0.383$.
-Though it is possible to carefully correct for such aliasing via the estimated window function \citep[e.g.][]{Roberts87}, we'll ignore this detail in the current work.
+Though it is possible to carefully correct for such aliasing by iteratively removing contributions from the estimated window function \citep[e.g.][]{Roberts87}, we'll ignore this detail in the current work.
 
@@
 Here we visualize the data from the left panel, but folded as a function of phase with the best-fit single-term model over-plot.
-Here its apparent that the single-term model is biased: RR Lyrae light curves are, in general, much more complicated than a simple sinusoid.
-Nevertheless, the simplistic sinusoidal model does well is able to recover the correct frequency to a high degree of accuracy (roughly related to the width of the peak) and significance (roughly related to the height of the peak).
+Here it is apparent that the single-term model is highly biased: RR Lyrae light curves are, in general, much more complicated than a simple sinusoid.
+Nevertheless, the simplistic sinusoidal model is able to recover the correct frequency to a high degree of accuracy (roughly related to the width of the peak) and significance (roughly related to the height of the peak).
 For a more complete introduction to and discussion of the single-term normalized periodogram, refer to, e.g.
@@
 \begin{enumerate}
-  \item It is trivially extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the noise covariance matrix $\Sigma$
-  \item It is trivially extended to more sophisticated linear models by appropriately modifying the design matrix $X_\omega$.
+  \item It is trivially extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the {\it noise covariance matrix} $\Sigma$
+  \item It is trivially extended to more sophisticated linear models by appropriately modifying the {\it design matrix} $X_\omega$.
   \item It is trivially extended to include Tikhonov/L2-regularization terms (see \S \ref{regularization} for more details)  by adding an appropriate diagonal term to the {\it normal matrix} $X_\omega^T\Sigma^{-1}X_\omega$.
@@
 To mimick a potentially damaging selection effect, all observations with magnitude fainter than 16 are removed (indicated by the light-gray points).
-The standard and floating-mean periodograms are computed from the remaining data; these fits are shown over the data in the right panel.
-Because of this biased observing pattern, the mean of the observed data is not a good predictor of the true mean, and the standard model fails to recover the true period of 0.622 days, while the floating mean model still finds the correct period.
+The standard and floating-mean periodograms are computed from the remaining data; these fits are shown over the data in the left panel.
+Because of this biased observing pattern, the mean of the observed data is not a good predictor of the true mean, and the standard fixed-mean model fails to recover the true period of 0.622 days, while the floating mean model still finds the correct period.
   }
@@
 Where this becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases of the light curve: a toy example demonstrating this situation is shown in \fig{floating_mean}.
-The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias (right panel).
-Because of this observational bias, the mean of the observed data are a poor predictor of the true mean, causing the standard fixed-mean method to poorly fit the data.
+The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias (left panel).
+Because of this observational bias, the mean of the observed data are a poor predictor of the true mean, causing the standard fixed-mean method to poorly fit the data (upper-right panel).
 The floating-mean approach is able to automatically adjust for this bias, resulting in a periodogram which readily detects the input period of 0.622 days (lower-right panel).
@@
   \caption{
-    An illustration of a periodogram based on the truncated Fourier model.
+    An illustration of a periodogram based on several truncated Fourier models.
     The data are the same as those in \fig{basic_example}. Note that the
-    higher-order models will show periodogram peaks at multiples of the true
-    fundamental frequency $P_0$: this is because for integer $n$ less than
-    the number of Fourier terms in the model, $P_0$ is a higher harmonic of
-    the model at $P=nP_0$.
+    higher-order models will generally show periodogram peaks at multiples
+    of the true fundamental frequency $P_0$: this is because for integer $n$
+    less than the number of Fourier terms in the model, $P_0$ is a higher
+    harmonic of the model at $P=nP_0$. Additionally, the increased degrees of
+    freedom in the higher-order models let them fit better at any frequency,
+    which drives up the ``background'' level in the periodogram.
   }
@@
 \end{equation}
-Because this remains a linear model, it can be easily accommodated into the matrix formalism above.
+Because this model remains linear in the parameters $\theta$, it can be easily accommodated into the matrix formalism above.
 For example, an $M = 2$-term floating-mean model can be constructed by building a design matrix $X_\omega$ with $2M + 1 = 5$ columns:
@@
 
-First, we see in the right panel that all three models show a significant spike at the true period of $P_0 = 0.622$.
+First, we see in the right panel that all three models show a clear spike at the true period of $P_0 = 0.622$.
 The higher-order models, however, also show a a spike in power at $P_1 = 2 P_0$: the reason for this is that for and $M>1$-term model, the period $P_0$ is the first harmonic of a model with fundamental frequency $2P_0$, and the higher-order models contain the single-period result.
@@
     regularization term effectively damps the higher-order Fourier modes and
-    removes this oscillating behavior, leading to a model which is much more
-    robust to noise in the data.
+    removes this oscillating behavior, leading to a more robust model with
+    stronger periodogram peaks.
   }
@@
 The previous sections raise the question: how complicated a model should we use?
-It is easily shown that as we add more terms to the fit, the model will come closer and closer to the observed data.
+We have seen that as we add more terms to the fit, the model will come closer and closer to the observed data.
 When the number of model parameters equals the number of data points, we can expect to fit the data {\it perfectly} (though in most cases numerical inaccuracies in the fit prevent a truly perfect fit).
-For very high-order models, the result become very sensitive to the noise in the data, and we say we have {\it over-fit} the data.
+For very high-order models, such a fit becomes very sensitive to the noise in the data, and we say we have {\it over-fit} the data.
 This can be addressed by explicitly truncating the series, but we can also use a {\it regularization} term to mathematically ensure a less complicated model.
@@
 A regularization term is an explicit penalty on the magnitude of the model parameters $\theta$, and can take a number of forms.
-For computational simplicity here we'll use an {\it L2 regularization} (also known as Tikhonov Regularization), which is a quadratic penalty term in the model parameters added to the $\chi^2$.
-Mathematically, this is equivalent to the Bayesian approach of using a Gaussian prior on the model parameters.
+For computational simplicity here we'll use an {\it L2 regularization} -- also known as Tikhonov Regularization \citep{Tikhonov1963} or Ridge Regression \citep{Hoerl1970} -- which is a quadratic penalty term in the model parameters added to the $\chi^2$.
+Mathematically, this is equivalent to the Bayesian approach of using a zero-mean Gaussian prior on the model parameters.
 
@@
 \end{equation}
-Notice that the effect of this regularization term is to add a diagonal penalty to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$, which has the additional feature that it can correct ill-posed models where the normal matrix is not invertible.
-This feature of the regularization will become important below.
+Notice that the effect of this regularization term is to add a diagonal penalty to the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$, which has the additional feature that it can correct ill-posed models where the normal matrix is non-invertible.
+This feature of the regularization will become important for the multiband models discussed below.
 
@@
 \section{A Multiple-Band Model}
-To compute a periodogram for multi-band data, we'll take advantage of many of the nice features of the matrix form of the normalized periodogram, which allows us to specify an arbitrary linear model covering our data.
-We will construct a multi-term Fourier model with the following features:
+\sectlabel{multiband}
+To compute a periodogram for multi-band data, we'll take advantage of many of the nice features of the matrix form of the normalized periodogram, which easily accommodates arbitrary linear models covering our data.
+To account for data in multible bands, we will construct a multi-term Fourier model with the following features:
 \begin{enumerate}
@@
 1 & \sin(\omega t_3) & \cos(\omega t_3) & 0 & 0 & 0 & 1 & 0\\
-1 & \sin(\omega t_4) & \cos(\omega t_4) & 0 & 0 & 1 & 0 & 0\\
 \vdots & \vdots & \vdots & & & \vdots & &\\
-1 & \sin(\omega t_N) & \cos(\omega t_N) & 0 & 1 & 0 & 0 & 0\\
+1 & \sin(\omega t_N) & \cos(\omega t_N) & 0 & 0 & 1 & 0 & 0\\
 \end{array}
@@
 \end{equation}
-Here the nonzero entries of the final five columns indicate the $ugriz$ band of the given observation: the first row is a $u$-band measurement, the second is a $g$-band, the third is a $i$-band, etc.
+Here the nonzero entries of the final five columns are binary flags indicating the $ugriz$ band of the given observation: in this case, the first row is a $u$-band measurement, the second is a $g$-band, the third is a $i$-band, etc.
 
@@
 where $\epsilon$ is some small fraction of the trace of the normal matrix $[X_\omega^T\Sigma^{-1}X_\omega]$.
-The logic of this choice of regularization is that any component of the model which is common to all bands will be accounted for in the base terms: only components of the fit that differ in each band will affect the band terms.
-Setting $\epsilon$ to some small fraction of the trace ensures that the regularization effect will be small.
+The logic of this choice of regularization is that any component of the model which is common to all bands will be preferentially reflected in the base terms, with independent behavior reflected in the individual band terms.
+Setting $\epsilon$ to some small fraction of the trace ensures that the regularization effect on the remaining model will be small.
 With this regularization in place, the model is well-posed and \eq{LombScargleReg} can be used to straightforwardly compute the power.
@@
 
+This multiband approach, then, actually comprises a set of models indexed by their value of $N_{base}$ and $N_{band}$.
+The most fundamental models have $(N_{base}, N_{band}) = (1,0)$ and $(0,1)$, which we'll call the {\it shared-phase} and {\it multi-phase} models respectively.
+In the shared-phase model, all variability is assumed to be shared between the bands, with only the fixed offset between them allowed to float.
+In the multi-phase model, each band has independent variability around a shared fixed offset.
+
 \subsection{Relationship of Multiband and Single-band approaches}
 \sectlabel{relationship}
-There is a special case of the multi-band approach when $N_{base}=0$ and $N_{band}=1$.
-In this case, the base offset can be removed completely and the $X$ matrix can be arranged to be block-diagonal.
-A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here the independent components are a standard single-band generalized periodogram within each of the $K$ bands.
+The multi-phase $N_{base}=0$, $N_{band}=1$ model turns out to be a particularly special case.
+In it, the base offset can be removed completely and the design matrix $X_\omega$ can be arranged to be block-diagonal.
+A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here these independent components are a standard single-band generalized periodogram fit within each of the $K$ bands.
 
-For band $k$, we'll denote the single-band normalized periodogram as
+For band $k$, we'll denote the standard single-band normalized periodogram as
 \begin{equation}
@@
 \begin{equation}
-  P_N^{(0,1)}(\omega) = \frac{\sum_{k=1}^K\left[\chi^2_{0,k}P_N^{(k)}\right]}{\sum_{k=1}^K\chi^2_{0,k}}.
+  P_N^{(0,1)}(\omega) = \frac{\sum_{k=1}^K\chi^2_{0,k}P_N^{(k)}}{\sum_{k=1}^K\chi^2_{0,k}}.
 \end{equation}
-This tells us how we can construct a particular case of the multi-term periodogram from a weighted sum of periodograms in each band.
-Here the weights reflect both how many measurements there are in each band, as well as how much those measurements deviate from a simple constant reference model.
+This tells us how we can construct a particular case of the multi-term periodogram from a weighted sum of periodograms in each band, where the weights $\chi^2_{0,k}$ reflect both the number of measurements in each band and how much those measurements deviate from a simple constant reference model.
 
-\section{Examples}
-Here we'll show several examples of the multiband periodogram method in action.
-
-\subsection{Simulated Example}
+\subsection{Simulated Example of the Multiband Approach}
 \sectlabel{Simulated}
@@
   \centering
-  \includegraphics[width=\textwidth]{fig05.pdf}
+  \includegraphics[width=\textwidth]{fig05a.pdf}
+  \includegraphics[width=\textwidth]{fig05b.pdf}
   \caption{
-    An illustration of the typical approach to multiband periodograms,
-    in which each band is fit individually. The data consists of 60 coeval
-    {\it ugriz} observations spread over 180 nights, and is based on an
-    RR Lyrae template from \citet{Sesar2010}. With this much data in each
-    band, individual periodograms can be constructed and compared to find the
-    true period $P=0.622$ days
-  }
-  \figlabel{adhoc_example}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{fig06.pdf}
-  \caption{
-    Another realization of the data from \fig{adhoc_example}, with only a single
-    band observed each night (i.e. 12 observations per band, spread over 180
-    days). In this case, single-band data is not sufficient to recover the
-    periodogram peaks, and the classic band-by-band approach fails.
-    By utilizing all the data at once with the multiband approach
-    (with $N_{base}=1,~N_{band}=0$),
-    we faithfully reconstruct the periodogram and recover
-    the dominant power at period $P=0.622$ days.
+    An illustration of the performance of the multiband periodogram. The
+    upper panels show simulated $ugriz$ observations of an RR Lyrae light
+    curve in which all 5 bands are observed each night. With 60 observations
+    in each band, a periodogram computed from any single band is sufficient to
+    determine the true period of 0.622 days. The lower panels show the same
+    data, except with only a single $ugriz$ band observed each night (i.e.
+    12 observations per band). In this case, no single band has enough
+    information to detect the period. The shared-phase multiband approach
+    of \sect{multiband} (lower-right panel) combines the information from
+    all five bands, and results in a significant detection of the true period.
   }
@@
 
-\todo{We need the bottom-right panel to be bigger! It's our money result! How do we arrange the plot to do this? Also: label it new style and old style.}
+The upper panels of \fig{multiband_sim} show a multiband version of the simulated RR Lyrae light curve from \fig{basic_example}.
+The upper-right panel shows 60 nights of observations spread over a 6-month period, and for each night all five bands ({\it u,g,r,i,z}) are recorded.
+Using the typical approach from the literature prior to this paper, we individually compute the standard normalized periodogram within each band: the results are shown in the upper-right panel.
+Here the data is well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band, up to the aliasing effect discussed in \sect{simple_period}.
+Previous studies have made use of the information in multiple bands to choose between aliases and estimate uncertainties in determined periods \citep[e.g.][]{Oluseyi12,Sesar2010}.
+While this approach is sufficient for well-sampled data, it becomes problematic when the multi-band data is more sparsely sampled.
 
-\fig{adhoc_example} shows a simulated multiband version of the RR Lyrae light curve from \fig{basic_example}.
-The observations take place on 60 random nights over a 6-month period, and each night all five bands ({\it u,g,r,i,z}) are recorded.
-The left panel of \fig{adhoc_example} shows the observations folded as a function of phase.
-Using the typical approach from the literature prior to this paper, we individually compute the standard normalized periodogram within each band: the results are shown in the right panel.
-Here the data is well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band, up to the aliasing effect discussed in \sect{simple_period}.
+The lower panels of \fig{multiband_sim} show the same 60 nights of data, except with only a {\it single} band observation recorded each night.
+The lower-left panel again shows the observations as a function of phase, and the lower-right panels show the periodograms derived from the data.
+With only 12 observations for each indiviudal band, it is clear that there is not enough data to accurately determine the period within each single band.
+The shared-phase multiband approach (i.e.
+$(N_{base},N_{band})=(1,0)$), shown in the lower-right panel, makes use of all the information in a single model, and clearly recovers the true frequency of 0.622 days.
 
-\fig{multiband_sim} shows the same 60 nights of data, except that each night only a {\it single} band observation is recorded.
-The left panel again shows the observations as a function of phase, and the right panel shows the periodograms derived from the data.
-Now, with only 12 observations for each indiviudal band, there is not enough data to accurately determine the period within each single band.
-With the multiband approach (here computed for $(N_{base},N_{band})=(1,0)$ and $(0,1)$), seen in the lower-right panel of \fig{multiband_sim}, the period is cleanly recovered.
+\todo{Add a figure comparing models for different $(N_{base},N_{band})$ pairs.}
 
-Here we can clearly see the difference between the $(0,1)$ model, which we'll call {\it old-style}, and the $(1,0)$ model, which we'll call {\it new-style}.
+%Here we can clearly see the difference between the $(0,1)$ model, which we'll call an {\it multi-phase} model, and the $(1,0)$ model, which we'll call {\it single-phase} model.
 The old-style model, as discussed in \sect{relationship}, is equivalent to a weighted sum of single-band periodograms for each individual band.
@@
 
-\subsection{Stripe 82}
+\section{Application to Stripe 82 RR Lyrae}
 
-\begin{itemize}
-  \item Compute periods using majority method
-  \item Compute periods using unified method
-  \item Compare the results
-\end{itemize}
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig07a.pdf}
+  \includegraphics[width=\textwidth]{fig07b.pdf}
+  \caption{
+    Comparison of the Multiband algorithm and single-band supersmoother algorithm on 483 well-sampled RR Lyrae light curves from Stripe 82.
+    The upper panels show a representative lightcurve and periodogram fits, while the bottom panels compare the derived periods to the true periods \citep[Based on a combined SuperSmoother and template-fitting approach; see][]{Sesar2010}.
+    Shown for reference are the beat aliases (dotted lines) and the multiplicative alias (dashed line), along with counts of points aligned with each trend.
+    The supersmoother algorithm tends to err toward multiplicative aliases, while the multiband algorithm tends to err toward beat frequency aliases.
+    Both methods find the correct period among the top 5 significant peaks nearly 100\% of the time.
+  } 
+  \figlabel{compare_periods}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig08a.pdf}
+  \includegraphics[width=\textwidth]{fig08b.pdf}
+  \caption{
+    This figure shows a repeat of the results from \fig{compare_periods}, but the data is artificially reduced to only a single-band observation on each evening, a situation reflective of the observing cadence of future large-scale surveys (compare to \fig{compare_periods}).
+    In this case, the single-band SuperSmoother strategy used by \citet{Sesar2010} fails: there is simply not enough data in each band to recover an accurate period estimate, and the correct period is among the top 5 candidates in fewer than 60\% of cases.
+    The shared-phase multiband approach utilizes information from all five bands, and returns much more robust results: even with the reduced data, the true period is among the top 5 candidates in 94\% of cases.
+  } 
+  \figlabel{compare_periods_partial}
+\end{figure}
+
+\todo{Discuss the results in \fig{compare_periods} and \fig{compare_periods_partial}}
 
@@
 
-\subsection{Further Study}
-\begin{itemize}
-  \item issues with aliasing \& window function corrections
-  \item issues with physicality of model
-\end{itemize}
+\todo{Discuss issues with aliasing \& window function corrections}
+\todo{Discuss issues with physicality of the model?}
 

### 2015-01-21 writeup/figures/fig_compare_periods.py

@@ -14,7 +14,7 @@ from scipy.stats import mode
 import matplotlib.pyplot as plt
 import seaborn; seaborn.set()
 
-from multiband_LS import LombScargleMultiband
+from multiband_LS import LombScargleMultiband, SuperSmootherMultiband
 from compute_results import SuperSmoother1Band
 from multiband_LS.data import fetch_light_curves
 
@@ -24,10 +24,10 @@ from compute_results import get_period_results
 
 def plot_period_comparison(ax, Px_all, Py,
                            beats=[-3, -2, -1, 0, 1, 2, 3],
-                           aliases=[]):
+                           aliases=[], color=None):
     Px = Px_all[:, 0]
 
-    ax.plot(Px, Py, 'o', alpha=0.5, ms=5)
+    ax.plot(Px, Py, 'o', alpha=0.5, ms=5, color=color)
     P1 = np.linspace(0.1, 1.2)
 
     matches = lambda x, y, tol=0.01: np.sum(abs(x - y) < tol)
@@ -50,7 +50,7 @@ def plot_period_comparison(ax, Px_all, Py,
 
     for n in aliases:
         fn = lambda P, n=n: n * P
-        ax.plot(P1, fn(P1), ':k', alpha=0.7, lw=1, zorder=1)
+        ax.plot(P1, fn(P1), '--', color='gray', alpha=0.7, lw=1, zorder=1)
         if n < 1:
             ax.text(1.21, fn(1.2), str(matches(fn(Px), Py)),
                     size=10, va='center', ha='left', color='gray')
@@ -100,9 +100,11 @@ def plot_example_lightcurve(rrlyrae, lcid):
     models = [SuperSmoother1Band(band='g'),
               LombScargleMultiband(Nterms_base=1, Nterms_band=0)]
 
-    for axi, model in zip(ax[1:], models):
+    colors = seaborn.color_palette()
+
+    for axi, model, color in zip(ax[1:], models, colors):
         model.fit(t, y, dy, filts)
-        axi.plot(periods, model.score(periods), lw=1)
+        axi.plot(periods, model.score(periods), lw=1, color=color)
         axi.set_ylim(0, 1)
 
     ax[1].xaxis.set_major_formatter(plt.NullFormatter())
@@ -124,15 +126,19 @@ def plot_periods(ssm_file, mbls_file, rrlyrae):
     fig.subplots_adjust(left=0.07, right=0.95, wspace=0.1,
[…]

### 2015-01-21 writeup/figures/fig_multiband_sim.py

@@ -49,7 +49,7 @@ for i, band in enumerate('ugriz'):
                    fmt='.', label=band)
 ax[0].set_ylim(18, 14.5)
 ax[0].legend(loc='upper left', fontsize=12, ncol=3)
-ax[0].set_title('Folded Data (P={0:.3f} days)'.format(rrlyrae.period))
+ax[0].set_title('Folded Data, 5 bands per night (P={0:.3f} days)'.format(rrlyrae.period))
 ax[0].set_xlabel('phase')
 ax[0].set_ylabel('magnitude')
 
@@ -57,13 +57,13 @@ for i, band in enumerate(filts):
     offset = 4 - i
     ax[1].plot(periods, P[band] + offset, lw=1)
     ax[1].text(0.89, 0.7 + offset, band, fontsize=14, ha='right', va='top')
-ax[1].set_title('Periodogram in Each Band')
+ax[1].set_title('Standard Periodogram in Each Band')
 ax[1].set_ylim(0, 5)
 ax[1].yaxis.set_major_formatter(plt.NullFormatter())
 ax[1].set_xlabel('Period (days)')
 ax[1].set_ylabel('power + offset')
 
-fig.savefig('fig05.pdf')
+fig.savefig('fig05a.pdf')
 
 #----------------------------------------------------------------------
 # Second figure:
@@ -92,7 +92,7 @@ for band, mask in zip('ugriz', masks):
                    fmt='.', label=band)
 ax[0].set_ylim(18, 14.5)
 ax[0].legend(loc='upper left', fontsize=12, ncol=3)
-ax[0].set_title('Folded Data (P={0:.3f} days)'.format(rrlyrae.period))
+ax[0].set_title('Folded Data, 1 band per night (P={0:.3f} days)'.format(rrlyrae.period))
 ax[0].set_xlabel('phase')
 ax[0].set_ylabel('magnitude')
 
@@ -104,7 +104,7 @@ for i, band in enumerate('ugriz'):
 ax[1].set_title('Standard Periodogram in Each Band')
 ax[1].yaxis.set_major_formatter(plt.NullFormatter())
 ax[1].xaxis.set_major_formatter(plt.NullFormatter())
-ax[1].set_ylabel('power + offset' + 30 * ' ')
+ax[1].set_ylabel('power + offset')
 
 for (i, Nbase, Nband) in [(0, 1, 0)]:#, (1, 0, 1)]:
[…]

### 2015-01-21 multiband_LS/__init__.py
add imports to __init__

@@ -5,5 +5,5 @@ __version__ = "0.1"
 from .lomb_scargle import LombScargle, LombScargleAstroML
 from .lomb_scargle_multiband import (LombScargleMultiband,
                                      LombScargleMultibandFast)
-from .supersmoother import SuperSmoother
+from .supersmoother import SuperSmoother, SuperSmootherMultiband
 from .naive_multiband import NaiveMultiband

### 2015-01-21 multiband_LS/lomb_scargle_multiband.py

@@ -198,8 +198,7 @@ class LombScargleMultibandFast(PeriodicModeler):
 
     def _score(self, periods):
         # Total score is the sum of powers weighted by chi2-normalization
-        powers = np.array([model.score(periods)
-                           for model in self.models_])
+        powers = np.array([model.score(periods) for model in self.models_])
         chi2_0 = np.array([np.sum(model.yw_ ** 2) for model in self.models_])
         return np.dot(chi2_0 / chi2_0.sum(), powers)
 

### 2015-01-23 writeup/paper.tex
update paper and figures

@@
 
-%\usepackage{listings}
-%\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
-%\lstset{language=Python,
-%        basicstyle=\footnotesize\ttfamily,
-%        showspaces=false,
-%        showstringspaces=false,
-%        tabsize=2,
-%        breaklines=false,
-%        breakatwhitespace=true,
-%        identifierstyle=\ttfamily,
-%        keywordstyle=\bfseries\color[rgb]{0.133,0.545,0.133},
-%        commentstyle=\color[rgb]{0.133,0.545,0.133},
-%        stringstyle=\color[rgb]{0.627,0.126,0.941},
-%    }
+% Draft watermark:
+%\usepackage{draftwatermark}
+%\SetWatermarkLightness{0.9}
+%\SetWatermarkScale{4}
 
+% Some macros
 \newcommand{\todo}[1]{{\color{red} [TODO: #1]}}
@@
 
-\title{Periodograms for Multi-band Astronomical Time Series}
+\title{Periodograms for Multiband Astronomical Time Series}
 
@@
 heteroscedastic errors, the multiband periodogram significantly improves period finding for randomly 
-sampled multi-band light curves (e.g., Pan-STARRS, DES and LSST). The light curves in 
+sampled multiband light curves (e.g., Pan-STARRS, DES and LSST). The light curves in 
 each band are modeled as arbitrary truncated Fourier series, with the period and phase 
@@
 and randomly subsampled SDSS Stripe 82 data to demonstrate the superiority of this 
-method compared to other methods from the literature. The python implementation of 
+method compared to other methods from the literature. The Python implementation of 
 this method is made publicly available as part of the astroML distribution.
@@
 \citep [e.g. LINEAR,][]{LINEAR1}, or nearly-simultaneously in all bands at each observation \citep [e.g. SDSS,][]{Sesar2010}.
-For the case of simultaneously taken multi-band measurements, \cite{Suveges12} utilized the principal component
-method to optimaly extract the best period. Their method is essentially a multi-band generalization of the well-known
+For the case of simultaneously taken multiband measurements, \cite{Suveges12} utilized the principal component
+method to optimaly extract the best period. Their method is essentially a multiband generalization of the well-known
 two-band Welch-Stetson variability index \citep{Stetson1996}. Unfortunately, when data in each band are taken at
@@
 
-For surveys that obtain multi-band data one band at a time, such as Pan-STARRS \citep{Kaiser2010} and DES \citep{Flaugher08},
+For surveys that obtain multiband data one band at a time, such as Pan-STARRS \citep{Kaiser2010} and DES \citep{Flaugher08},
 and for future multicolor surveys such as LSST \citep{Ivezic08LSST}, this \adhoc{} approach is not optimal. In order to take 
@@
 The proposed method is essentially a generalization of the Lomb-Scargle method to 
-multi-band case. The light curves in  each band are modeled as arbitrary truncated Fourier series, 
+multiband case. The light curves in  each band are modeled as arbitrary truncated Fourier series, 
 with the period, and optionally the phase, shared across all bands. The key aspect enabling this approach is the use of Tikhonov regularization 
@@
 While the important features of least squares frequency estimation via Lomb-Scargle periodograms has been discussed elsewhere, we will prevent a brief introduction to the subject in the following section.
-The matrix-based formalism we develop here will make clear how the method can be extended to more sophisticated models, including regularized multi-band methods.
+The matrix-based formalism we develop here will make clear how the method can be extended to more sophisticated models, including regularized multiband methods.
 
@@
 \sectlabel{multiband}
-To compute a periodogram for multi-band data, we'll take advantage of many of the nice features of the matrix form of the normalized periodogram, which easily accommodates arbitrary linear models covering our data.
+To compute a periodogram for multiband data, we'll take advantage of many of the nice features of the matrix form of the normalized periodogram, which easily accommodates arbitrary linear models covering our data.
 To account for data in multible bands, we will construct a multi-term Fourier model with the following features:
 \begin{enumerate}
-  \item An $M_{base}$-term truncated Fourier fit which models a latent parameter, which we'll call the ``overall variability''.
-  \item A set of $M_{band}$-term truncated Fourier fits, each of which models the residual of a single band from this overall variability.
+  \item An $N_{base}$-term truncated Fourier fit which models a latent parameter, which we'll call the ``overall variability''.
+  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual of a single band from this overall variability.
 \end{enumerate}
-The total number of parameters for $K$ filters is then $M_K = (2M_{base} + 1) + K(2M_{band} + 1)$. As a result, for each band $b$ we have the following model of the observed magnitudes:
+The total number of parameters for $K$ filters is then $M_K = (2N_{base} + 1) + K(2N_{band} + 1)$. As a result, for each band $k$ we have the following model of the observed magnitudes:
 \begin{eqnarray}
-  y_b(t|\omega,\theta) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\\ 
-  &\theta^{(b)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(b)}_{2n - 1}\sin(n\omega t) + \theta^{(b)}_{2n}\cos(n\omega t)\right].&
+  y_k(t|\omega,\theta) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\\ 
+  &\theta^{(k)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(k)}_{2n - 1}\sin(n\omega t) + \theta^{(k)}_{2n}\cos(n\omega t)\right].&
 \end{eqnarray}
-The important feature of this model is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(b)}$ are determined individually.
+The important feature of this model is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(k)}$ are determined individually.
 
@@
 With this regularization in place, the model is well-posed and \eq{LombScargleReg} can be used to straightforwardly compute the power.
-We'll see an example of this in action in \sect{Simulated}.
+The effective number of free parameters for such a regularized $(N_{base}, N_{band})$ model with $K$ filters is
+$M_K^{eff} = 2N_{base}^{eff} + K(2N_{band} + 1)$ where $N_{base}^{eff} = \max(0, N_{base} - N_{band})$ is the effective number of base terms.
 
@@
 \sectlabel{relationship}
-The multi-phase $N_{base}=0$, $N_{band}=1$ model turns out to be a particularly special case.
-In it, the base offset can be removed completely and the design matrix $X_\omega$ can be arranged to be block-diagonal.
-A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here these independent components are a standard single-band generalized periodogram fit within each of the $K$ bands.
+The {\it multi-phase} $N_{base}=0$, $N_{band}=1$ model turns out to be a particularly special case.
+The base model is a simple global offset which is degenerate with the offsets in each band, so that the design matrix $X_\omega$ can be straightforwardly rearranged as block-diagonal.
+A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here these independent components amount to the single-band floating-mean model from \sect{floating_mean}, fit independently within each of the $K$ bands.
+This particular multiband model can give us insight into the relationship between single-band and multiband approaches.
 
-For band $k$, we'll denote the standard single-band normalized periodogram as
+For band $k$, we'll denote the single-band floating-mean periodogram as
 \begin{equation}
@@
 
-\subsection{Simulated Example of the Multiband Approach}
+\subsection{Multiband Periodogram for Simulated Data}
 \sectlabel{Simulated}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{fig06.pdf}
+  \caption{
+    Comparison of the periodograms produced by various multiband models.
+    The data is the same as that used in \fig{multiband_sim}. $N_{base}$ gives
+    the number of Fourier terms in the base model, and $N_{band}$ gives the
+    number of Fourier terms used to fit the residuals around this model within
+    each band. The characteristics discussed with previous figures are also
+    seen here: in particular, the level of ``background noise'' in the
+    periodogram grows with the model complexity $M$,
+  } 
+  \figlabel{multiband_models}
+\end{figure}
+
+Before applying the multiband method to real data, let's explore its effectiveness on a simulated RR Lyrae lightcurve.
 The upper panels of \fig{multiband_sim} show a multiband version of the simulated RR Lyrae light curve from \fig{basic_example}.
@@
 Previous studies have made use of the information in multiple bands to choose between aliases and estimate uncertainties in determined periods \citep[e.g.][]{Oluseyi12,Sesar2010}.
-While this approach is sufficient for well-sampled data, it becomes problematic when the multi-band data is more sparsely sampled.
+While this approach is sufficient for well-sampled data, it becomes problematic when the multiband data is more sparsely sampled.
 
 The lower panels of \fig{multiband_sim} show the same 60 nights of data, except with only a {\it single} band observation recorded each night.
-The lower-left panel again shows the observations as a function of phase, and the lower-right panels show the periodograms derived from the data.
+The lower-left panel shows the observations as a function of phase, and the lower-right panels show the periodograms derived from the data.
 With only 12 observations for each indiviudal band, it is clear that there is not enough data to accurately determine the period within each single band.
@@
 
-\todo{Add a figure comparing models for different $(N_{base},N_{band})$ pairs.}
+This shared-phase $(1,0)$ model is only one of the possible multiband options, however: \fig{multiband_models} shows multiband fits to this data for models with various combinations of $(N_{base},N_{band})$.
+We see here many of the characteristics noted above for single-band models: as discussed in \sect{multiterm}, increasing the number of Fourier terms leads to power at multiples of the fundamental period.
+Additionally, increasing the model complexity (roughly indexed by the effective number of free parameters $M^{eff}$) tends to increase the background level of the periodogram, potentially obscuring significant peaks.
+For this reason, models with $N_{base} > N_{band}$ are the most promising: they allow a flexible fit with minimal model complexity.
 
-%Here we can clearly see the difference between the $(0,1)$ model, which we'll call an {\it multi-phase} model, and the $(1,0)$ model, which we'll call {\it single-phase} model.
-The old-style model, as discussed in \sect{relationship}, is equivalent to a weighted sum of single-band periodograms for each individual band.
-The model thus has effectively 15 free parameters, and the phase and amplitude of each band varies individually.
-As with the truncated Fourier series, this added complexity leads to higher power at off-frequencies.
-The new-style model, on the other hand, is effectively a 7-parameter model where a single oscillation phase and amplitude is used to describe all the variability in the data.
-The only flexibility allowed in individual bands is the constant offset.
-This lower-parameter model produces a periodogram which nicely recovers the input frequency, and has a background level qualitatively similar to that of the single-band observations in \fig{adhoc_example}.
+Below, we'll apply the simplest of this class of models, the $(1, 0)$ shared-phase model, to data from the SDSS-2 survey.
 
@@
     The supersmoother algorithm tends to err toward multiplicative aliases, while the multiband algorithm tends to err toward beat frequency aliases.
-    Both methods find the correct period among the top 5 significant peaks nearly 100\% of the time.
+    Both methods find the correct period among the top 5 significant peaks around 99\% of the time.
   } 
@@
 \todo{Discuss issues with aliasing \& window function corrections}
+
 \todo{Discuss issues with physicality of the model?}

### 2015-01-23 writeup/figures/fig_compare_periods.py

@@ -65,6 +65,18 @@ def plot_period_comparison(ax, Px_all, Py,
     top_matches = np.any((abs(Px_all.T - Py) < 0.01), 0).sum()
     ax.text(1.17, 0.11, "Top 5 Matches: {0}/{1}".format(top_matches, len(Px)),
             size=10, ha='right', va='bottom')
+    
+
+def plot_filter(t, y, verbose=True):
+    """Simple filter which removes unnecessary points for plotting"""
+    dy = y[1:] - y[:-1]
+    mask = (dy[1:] * dy[:-1] <= 0)
+    mask = np.concatenate([[True], mask, [True]])
+
+    if verbose:
+        print("Reducing from {0} to {1} points for plotting"
+              "".format(len(t), mask.sum()))
+    return t[mask], y[mask]
 
 
 def plot_example_lightcurve(rrlyrae, lcid):
@@ -104,7 +116,8 @@ def plot_example_lightcurve(rrlyrae, lcid):
 
     for axi, model, color in zip(ax[1:], models, colors):
         model.fit(t, y, dy, filts)
-        axi.plot(periods, model.score(periods), lw=1, color=color)
+        per, scr = plot_filter(periods, model.score(periods))
+        axi.plot(per, scr, lw=1, color=color)
         axi.set_ylim(0, 1)
 
     ax[1].xaxis.set_major_formatter(plt.NullFormatter())

### 2015-01-23 writeup/figures/fig_multiband_models.py

@@ -1,0 +1,75 @@
+"""
+Here we plot a typical approach to the multi-band periodogram: treating each
+band separately, and taking a majority vote between the bands.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Use seaborn settings for plot styles
+import seaborn; seaborn.set()
+
+from multiband_LS.generated import RRLyraeObject
+from multiband_LS import (LombScargleAstroML, LombScargleMultiband,
+                          NaiveMultiband)
+
+
+# Choose a Sesar 2010 object to base our fits on
+lcid = 1019544
+rrlyrae = RRLyraeObject(lcid, random_state=0)
+print("Extinction A_r = {0:.4f}".format(rrlyrae.obsmeta['rExt']))
+
+# Generate data in a 6-month observing season
+Nobs = 60
+rng = np.random.RandomState(0)
+
+nights = np.arange(180)
+rng.shuffle(nights)
+nights = nights[:Nobs]
+
+# Find a subset of the simulated data. This is the same procedure as in
+# fig_multiband_sim
+t = 57000 + nights + 0.05 * rng.randn(Nobs)
+dy = 0.06 + 0.01 * rng.randn(Nobs)
+mags = np.array([rrlyrae.generated(band, t, err=dy, corrected=False)
+                 for band in 'ugriz'])
+filts = np.array([f for f in 'ugriz'])
+
+# Here's our subset
+filts = np.take(list('ugriz'), np.arange(Nobs), mode='wrap')
+mags = mags[np.arange(Nobs) % 5, np.arange(Nobs)]
+masks = [(filts == band) for band in 'ugriz']
+
+fig, ax = plt.subplots(5, sharex=True, sharey=True)
+fig.subplots_adjust(left=0.1, right=0.93, hspace=0.1)
+
+periods = np.linspace(0.2, 1.4, 1000)
+
+combos = [(1, 0), (0, 1), (2, 0), (2, 1), (2, 2)]
+
+for axi, (Nbase, Nband) in zip(ax, combos):
+    LS_multi = LombScargleMultiband(Nterms_base=Nbase,
+                                    Nterms_band=Nband)
+    LS_multi.fit(t, mags, dy, filts)
+    P_multi = LS_multi.periodogram(periods)
+    axi.plot(periods, P_multi, lw=1, color='gray')
+    
+    text = ('$N_{{base}}={0},\ N_{{band}}={1}\ \ (M^{{eff}}={2})$'
[…]

### 2015-01-23 writeup/figures/fig_multiband_sim.py

@@ -124,15 +124,14 @@ ax[2].set_ylabel('power')
 
 fig.savefig('fig05b.pdf')
 
-#----------------------------------------------------------------------
-# Write the results to file
-import pandas as pd
-df = pd.DataFrame({'t': t})
-df['band'] = filts
-df['mag'] = mags
-df['dmag'] = dy
-df = df.sort('t')
-
-df.to_csv('fig_data.dat', sep=' ', float_format='%.4f', index=False)
-
 plt.show()
+
+#----------------------------------------------------------------------
+# Uncomment this to Write the results to file
+#import pandas as pd
+#df = pd.DataFrame({'t': t})
+#df['band'] = filts
+#df['mag'] = mags
+#df['dmag'] = dy
+#df = df.sort('t')
+#df.to_csv('fig_data.dat', sep=' ', float_format='%.4f', index=False)

### 2015-01-24 writeup/paper.tex
more paper work; add appendix

@@
 \graphicspath{{figures/}}
+
+% For Python code
+\usepackage{listings}
+\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
+\lstset{language=Python,
+        basicstyle=\footnotesize\ttfamily,
+        showspaces=false,
+        showstringspaces=false,
+        tabsize=2,
+        breaklines=false,
+        breakatwhitespace=true,
+        identifierstyle=\ttfamily,
+        keywordstyle=\bfseries\color[rgb]{0.133,0.545,0.133},
+        commentstyle=\color[rgb]{0.133,0.545,0.133},
+        stringstyle=\color[rgb]{0.627,0.126,0.941},
+    }
 
@@
 
-\todo{In \S 2, we provide a brief overview of periodograms, including the matrix-based formulation used in this work. In \S...}
+\todo{In \S 2, we provide a brief overview of periodograms, including the matrix-based formulation used in this work. In \S... \citet[][hereafter S10]{Sesar2010}...}
 
@@
 
-Below, we'll apply the simplest of this class of models, the $(1, 0)$ shared-phase model, to data from the SDSS-2 survey.
+Below, we'll apply the simplest of this class of models, the $(1, 0)$ shared-phase model, to data from the Stripe 82 of the Sloan Digital Sky Survey.
 
 \section{Application to Stripe 82 RR Lyrae}
+
+Stripe 82 is a three hundred square degree equatorial region of the sky which was repeatedly imaged through multiple band-passes during the second phase of the Sloan Digital Sky Survey (SDSS II).
+As a real-world test of the multiband periodogram method, we consider the SDSS-2 observations of 483 RR Lyrae stars compiled and studied by S10.
+S10 determined periods for these stars based on an emperically-derived light curve templates.
+Because the template-fitting method is extremely computationally intensive, S10 first determined candidate periods by taking the top 5 results of the supersmoother \citep{Reimann94} algorithm applied to the $g$-band; template fits were then performed at each candidate period and the period with the best template fit was reported as the true period.
+In this section, we make use of this dataset to quantitatively evaluate the effectiveness of the multiband periodogram approach in the case of this real-world data.
+
+\subsection{Densely-sampled Multiband Data}
 
@@
 
+The full S10 RR Lyrae dataset consists of 483 objects with an average of 55 observations in each of the five SDSS $ugriz$ bands spread over just under ten years.
+In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother algorithm and the shared-phase $(0, 1)$-multiband algorithm\footnote{The supersmoother periodogram $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with the Lomb-Scargle periodogram's $\chi^2$ interpretation from \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$}.
+Here we have a case which is analogous to that shown for simulated data in the top panels of \fig{multiband_sim}: each band has enough data to easily locate candidate peaks, the best of which is selected via the S10 template-fitting procedure.
+
+The lower panels of \fig{compare_periods} compare the S10 period with the best periods obtained from the 1-band supersmoother (lower-left) and from the shared-phase multiband model (lower-right).
+To guide the eye, the figure includes indicators of the locations of beat aliases (dotted lines) and multiplicative aliases (dashed lines) of the S10 period.
+
+The best-fit supersmoother period matches the S10 period in 89\% of cases (421/483), while the best-fit multiband period matches the S10 period in 79\% of cases (382/483).
+The modes of failure are instructive: when the supersmoother model misses the S10 period, it tends to land on a multiplicative alias.
+This is due to the flexibility of the supersmoother model: a doubled period spreads the points out further, leading to fewer constraints in each neighborhood and thus a smaller average residual around the over-fit model.
+When the multiband model misses the S10 period, it tends to land on a beat alias between the S10 period and the 1-day observing cadence.
+This is due to the fact that the single-frequency periodic model significantly under-fits the data, and thus cannot distinguish deviations due to underfitting from deviations due to window function effects.
+
+In both models, the S10 period appears among the top 5 periods 99\% of the time: $477/483$ for supersmoother, and $480/483$ for multiband.\footnote{We would expect this correspondence to be 100\% in the case of the $g$-band supersmoother, which was the model used in the first pass of the S10 computation.
+This discrepancy can be chalked-up to the different supersmoother implementations used in S10 and in this work.
+The discrepant curves are those with very low signal-to-noise.} This suggests that had S10 used the multiband Lomb-Scargle rather than the supersmoother in the first pass for that study, the final results presented there would be for the most part unchanged.
+
+The results of this subsection show that the shared-phase multiband approach is comparable to the supersmoother approach for densely-sampled multiband data, although it has a tendency to get fooled by structure in the survey window function.
+
+\subsection{Sparsely-sampled Multiband Data}
+
 \begin{figure}
@@
   } 
-  \figlabel{compare_periods_partial}
+  \figlabel{compare_periods_reduced}
 \end{figure}
 
-\todo{Discuss the results in \fig{compare_periods} and \fig{compare_periods_partial}}
+Where we expect the multiband approach to have an advantage is when the data are sparsely sampled, with observations through only a single band each night.
+To simulate this, we reduce the Stripe 82 RR Lyrae dataset by a factor of 5, keeping only a single band of imaging each night, for an average of 11 observations of each object per band.
+This is much closer to the type of observation available in future multiband surveys.
+
+The upper panels of \fig{compare_periods_reduced} show an example light curve from this reduced dataset, along with the supersmoother and multiband periodograms derived from this data.
+Analogously to the lower panels of \fig{multiband_sim}, the single-band supersmoother model loses the true period within the noise, while the shared-phase multiband model still shows a noticeable spike at this true period.
+
+The lower panels of \fig{compare_periods_reduced} show the relationship between the S10 periods (based on the full dataset) and the periods derived with each model from this reduced dataset.
+It is clear that the supersmoother model is simply over-fitting noise with this few data points: the top period matches S10 in only 32\% of cases (compared to 87\% with the full dataset), and the top 5 periods contain the S10 period only 57\% of the time.
+The failure mode is much less regular as well: rather than being clustered near aliases, most of the points are scattered seemingly randomly around the space.
+
+The multiband method does much better on the reduced dataset.
+Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 66\% of the time (compared to 79\% with the full dataset), and the top 5 periods contain the S10 period 94\% of the time.
+This performance is due to the fact that the multiband algorithm has relatively few parameters, but is yet able to accommodate data from multiple observing bands.
+In particular, this suggests that with the multiterm periodogram, the S10 analysis could have been done effectively with only a small fraction of the available data.
+This bodes well for future surveys, where data on variable stars will be much more sparsely sampled.
 
@@
 
-\todo{Discuss issues with aliasing \& window function corrections}
+\todo{Zeljko, can you take a stab at this?}
 
-\todo{Discuss issues with physicality of the model?}
+\begin{enumerate}
+  \item Our method kicks ass, therefore we kick ass. QED.
+  \item Future work: account for beat-aliases for ground-based data
+  \item Future work: physically-motivated priors to further simplify the model
+  \item Opportunity to apply this to PanSTARRS data release
+\end{enumerate}
 
@@
 
-%\appendix
+\appendix
+\section{{\tt multiband\_LS}: Python Implementation of Multiband Periodogram}
+The algorithm outlined in this paper is available as an open-source Python package\footnote{\url{http://github.com/jakevdp/multiband_LS/}}, along with code to download the data and recreate all the figures in this paper.
+{\tt multiband\_LS} is a pure-Python package written to be compatible with both Python 2 and Python 3, and performs fast numerical computation via the {\tt numpy} package\footnote{http://www.numpy.org}.
+
+The API for the module is largely influenced by that of the {\tt scikit-learn} package\footnote{\url{http://scikit-learn.org}} \citep{scikit-learn}, in which models are class objects which can be ``fit()`` to data.
+Here is a basic example of how you can use the code to download data, fit a multiband model to the data, and compute the power at a few periods:
+
+\begin{lstlisting}
+from multiband_LS import LombScargleMultiband
+
+# Fetch the Sesar 2010 RR Lyrae data
+from multiband_LS.data import fetch_light_curves
+data = fetch_light_curves()
+t, mag, dmag, filts = data.get_lightcurve(data.ids[0],
+                                          return_1d=True)
+
+# Construct the multiband model
+model = LombScargleMultiband(Nterms_base=0, Nterms_band=1)
+model.fit(t, mag, dmag, filts)
+
+# Compute power at the following periods
+model.periodogram([0.2, 0.3, 0.4])
+\end{lstlisting}
+
+Other models are available as well. For example, here is how you can compute the periodogram under the supersmoother model:
+
+\begin{lstlisting}
+from multiband_LS import SuperSmoother
+
+# Construct the supersmoother model
+model = SuperSmoother()
+gband = (filts == 'g')
+model.fit(t[gband], mag[gband], dmag[gband])
+
+# Compute power at the following periods
+model.periodogram([0.2, 0.3, 0.4])
+\end{lstlisting}
+
+The code contains many more methods, and much more functionality that what is shown here. For updates and more information, visit \url{http://github.com/jakevdp/multiband_LS}.
 

### 2015-01-24 multiband_LS/data.py

@@ -49,6 +49,10 @@ class RRLyraeLC(object):
 
     @property
     def ids(self):
+        return list(self.ids_gen)
+
+    @property
+    def ids_gen(self):
         for f in self.filenames:
             if '/' not in f:
                 continue

### 2015-01-24 writeup/figures/fig_compare_periods.py

@@ -24,7 +24,7 @@ from compute_results import get_period_results
 
 def plot_period_comparison(ax, Px_all, Py,
                            beats=[-3, -2, -1, 0, 1, 2, 3],
-                           aliases=[], color=None):
+                           aliases=[2], color=None):
     Px = Px_all[:, 0]
 
     ax.plot(Px, Py, 'o', alpha=0.5, ms=5, color=color)
@@ -49,7 +49,7 @@ def plot_period_comparison(ax, Px_all, Py,
                     size=10, va='bottom', ha='center', color='gray')
 
     for n in aliases:
-        fn = lambda P, n=n: n * P
+        fn = lambda P, n=n: P / n
         ax.plot(P1, fn(P1), '--', color='gray', alpha=0.7, lw=1, zorder=1)
         if n < 1:
             ax.text(1.21, fn(1.2), str(matches(fn(Px), Py)),
@@ -141,7 +141,7 @@ def plot_periods(ssm_file, mbls_file, rrlyrae):
 
     colors = seaborn.color_palette()
 
-    plot_period_comparison(ax[0], ssm_periods, sesar_periods, aliases=[1/2],
+    plot_period_comparison(ax[0], ssm_periods, sesar_periods,
                            color=colors[0])
     plot_period_comparison(ax[1], mbls_periods, sesar_periods,
                            color=colors[1])

### 2015-01-24 writeup/figures/fig_compare_periods_reduced.py

@@ -12,7 +12,8 @@ lcid = list(rrlyrae.ids)[4]
 fig, ax = plot_example_lightcurve(rrlyrae, lcid)
 fig.savefig('fig08a.pdf')
 
-fig, ax = plot_periods(ssm_file='results/partial_supersmoother_g.npy',
+# XXX: use new results here when available
+fig, ax = plot_periods(ssm_file='results_old/partial_supersmoother_g.npy',
                        mbls_file='results/partial_multiband_1_0.npy',
                        rrlyrae=rrlyrae)
 fig.savefig('fig08b.pdf')

### 2015-01-24 writeup/figures/fig_example.py

@@ -1,0 +1,22 @@
+from multiband_LS import LombScargleMultiband, SuperSmoother
+
+# Fetch the Sesar 2010 RR Lyrae data
+from multiband_LS.data import fetch_light_curves
+data = fetch_light_curves()
+t, mag, dmag, filts = data.get_lightcurve(data.ids[0],
+                                          return_1d=True)
+
+# Construct the multiband model
+model = LombScargleMultiband(Nterms_base=0, Nterms_band=1)
+model.fit(t, mag, dmag, filts)
+
+# Compute power at the following periods
+model.periodogram([0.2, 0.3, 0.4])
+
+# Construct the supersmoother model
+model = SuperSmoother()
+gband = (filts == 'g')
+model.fit(t[gband], mag[gband], dmag[gband])
+
+# Compute power at the following periods
+model.periodogram([0.2, 0.3, 0.4])

### 2015-01-24 writeup/figures/fig_multiband_models.py

@@ -51,7 +51,7 @@ for axi, (Nbase, Nband) in zip(ax, combos):
                                     Nterms_band=Nband)
     LS_multi.fit(t, mags, dy, filts)
     P_multi = LS_multi.periodogram(periods)
-    axi.plot(periods, P_multi, lw=1, color='gray')
+    axi.plot(periods, P_multi, lw=1)
     
     text = ('$N_{{base}}={0},\ N_{{band}}={1}\ \ (M^{{eff}}={2})$'
             ''.format(Nbase, Nband, (2 * max(0, Nbase - Nband)

### 2015-01-24 writeup/figures/fig_num_observations.py

@@ -1,0 +1,35 @@
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+from multiband_LS.data import fetch_light_curves
+
+
+rrlyrae = fetch_light_curves()
+rrlyrae_partial = fetch_light_curves(partial=True)
+
+data = [rrlyrae.get_lightcurve(lcid, True)
+        for lcid in rrlyrae.ids]
+data_partial = [rrlyrae_partial.get_lightcurve(lcid, True)
+                for lcid in rrlyrae.ids]
+
+tminmax = np.array([[t.min(), t.max()] for t,_,_,_ in data])
+print(tminmax.min(0), tminmax.max(0))
+
+counts = np.array([[np.sum(filts == f) for f in 'ugriz']
+                   for t, y, dy, filts in data])
+counts_partial = np.array([[np.sum(filts == f) for f in 'ugriz']
+                           for t, y, dy, filts in data_partial])
+
+print(np.mean(counts, 0), np.median(counts, 0))
+print(np.mean(counts_partial, 0), np.median(counts_partial, 0))
+
+
+fig, ax = plt.subplots(2)
+sns.violinplot(counts, ax=ax[0], names='ugriz')
+sns.violinplot(counts_partial, ax=ax[1], names='ugriz')
+
+ax[0].set_ylim(20, 100)
+ax[1].set_ylim(0, 20)
+
+plt.show()

### 2015-01-24 gh-publisher-scripts/build.sh
add gh-publisher stuff

@@ -1,0 +1,15 @@
+#
+# Use this file to configure your build.
+#
+# You have the following variables available:
+#
+# GH_PUBLISHER_PROJECT_DIR - The root of your project repository.
+# GH_PUBLISHER_SCRIPTS_DIR - The gh-publisher-scripts directory.
+#
+# The current working directory is $GH_PUBLISHER_PROJECT_DIR.
+#
+echo "$PWD"
+
+cd "$GH_PUBLISHER_PROJECT_DIR"
+cd writeup; make; cd ..
+cp writeup/paper.pdf ./

### 2015-01-24 gh-publisher-scripts/copy.sh

@@ -1,0 +1,23 @@
+#
+# Use this file to configure whichever artifacts you wish to publish.
+#
+# You have the following variables available:
+#
+# GH_PUBLISHER_PROJECT_DIR - The root of your project repository.
+# GH_PUBLISHER_PUBLISH_DIR - The destination for files to be published.
+# GH_PUBLISHER_SCRIPTS_DIR - The gh-publisher-scripts directory.
+#
+# The current working directory is $GH_PUBLISHER_PROJECT_DIR.
+#
+
+find . \( \
+     -name \*.html -o \
+     -name \*.css -o \
+     -name \*.js -o \
+     -name \*.gif -o \
+     -name \*.jpeg -o \
+     -name \*.jpg -o \
+     -name \*.png -o \
+     -name \*.pdf \
+     \) -print0 |
+    rsync -av --files-from=- --from0 ./ "$GH_PUBLISHER_PUBLISH_DIR"

### 2015-01-24 gh-publisher-scripts/gh-publisher.sh

@@ -1,0 +1,101 @@
+#!/bin/bash
+
+set -eu
+
+function readlink_f() {
+  if [ $(uname) = 'Darwin' ]
+  then
+    DIR=$(echo "${1%/*}")
+    FILE=$(basename "$1")
+    (cd "$DIR" && echo "$(pwd -P)/$FILE")
+  else
+    readlink -f "$1"
+  fi
+}
+
+gh_token="${GH_TOKEN-}"
+git_name="${GIT_NAME-Automated build by gh-publisher}"
+git_email="${GIT_EMAIL-not.a.real.person@example.com}"
+git_remote="${GIT_REMOTE-origin}"
+git_publish_branch="${GIT_PUBLISH_BRANCH-gh-pages}"
+
+if [ -z "$gh_token" ]
+then
+  echo "GH_TOKEN is not set.  Cannot proceed." >&2
+  echo "Please see the installation instructions." >&2
+  exit 1
+fi
+
+scripts_dir=$(dirname $(readlink_f "$0"))
+scripts_dir_name=$(basename "$scripts_dir")
+project_dir=$(dirname "$scripts_dir")
+
+cd "$scripts_dir"
+git_url=$(git config "remote.${git_remote}.url" |
+          sed -e 's,git://*,https://,' -e "s,https://,https://${gh_token}@,")
+git_rev=$(git rev-parse HEAD)
+github_username=$(echo "$git_url" | sed -e 's,^https://[^/]*/\([^/]*\)/.*$,\1,')
+github_repo=$(echo "$git_url" | sed -e 's,^https://[^/]*/[^/]*/\([^/.]*\).*$,\1,')
+
+tmpdir=$(mktemp -t publish.XXXXXX -d)
+(cd "$tmpdir"
+ git clone "$git_url" "publish"
+ cd publish
+ git config user.name "$git_name"
+ git config user.email "$git_email"
+ if git branch -av | grep -q "remotes/origin/$git_publish_branch"
+ then
+   git checkout "$git_publish_branch"
+ else
+   git checkout --orphan "$git_publish_branch"
+   git reset .
+   git clean -dfx
+ fi
+)
+publish_dir="$tmpdir/publish"
+
+export GH_PUBLISHER_PROJECT_DIR="$project_dir"
+export GH_PUBLISHER_PUBLISH_DIR="$publish_dir"
+export GH_PUBLISHER_SCRIPTS_DIR="$scripts_dir"
+
+(cd "$project_dir" && /bin/bash "$scripts_dir/build.sh")
[…]

### 2015-01-24 writeup/paper.tex
enable draft watermark
paper: update reduced dataset & associated figures

@@
 % Draft watermark:
-%\usepackage{draftwatermark}
-%\SetWatermarkLightness{0.9}
-%\SetWatermarkScale{4}
+\usepackage{draftwatermark}
+\SetWatermarkLightness{0.9}
+\SetWatermarkScale{4}
 
@@
     Comparison of the Multiband algorithm and single-band supersmoother algorithm on 483 well-sampled RR Lyrae light curves from Stripe 82.
-    The upper panels show a representative lightcurve and periodogram fits, while the bottom panels compare the derived periods to the true periods \citep[Based on a combined SuperSmoother and template-fitting approach; see][]{Sesar2010}.
-    Shown for reference are the beat aliases (dotted lines) and the multiplicative alias (dashed line), along with counts of points aligned with each trend.
-    The supersmoother algorithm tends to err toward multiplicative aliases, while the multiband algorithm tends to err toward beat frequency aliases.
+    The upper panels show a representative lightcurve and periodogram fits, while the bottom panels compare the derived periods to the template-based periods reported in S10.
+    Shown for reference are the beat aliases (dotted lines) and the multiplicative alias (dashed lines), along with counts of points aligned with each trend.
+    The single-band supersmoother algorithm tends to err toward multiplicative aliases, while the multiband algorithm tends to err toward beat frequency aliases.
     Both methods find the correct period among the top 5 significant peaks around 99\% of the time.
@@
 The best-fit supersmoother period matches the S10 period in 89\% of cases (421/483), while the best-fit multiband period matches the S10 period in 79\% of cases (382/483).
-The modes of failure are instructive: when the supersmoother model misses the S10 period, it tends to land on a multiplicative alias.
-This is due to the flexibility of the supersmoother model: a doubled period spreads the points out further, leading to fewer constraints in each neighborhood and thus a smaller average residual around the over-fit model.
-When the multiband model misses the S10 period, it tends to land on a beat alias between the S10 period and the 1-day observing cadence.
-This is due to the fact that the single-frequency periodic model significantly under-fits the data, and thus cannot distinguish deviations due to underfitting from deviations due to window function effects.
+The modes of failure are instructive: when the supersmoother model misses the S10 period, it tends to land on a multiplicative alias (i.e.
+the dashed line).
+This is due to the flexibility of the supersmoother model: a doubled period spreads the points out, leading to fewer constraints in each neighborhood and thus a smaller average residual around model.
+In other words, the SuperSmoother tends to over-fit sparsely-sampled data.
+On the other hand, when the multiband model misses the S10 period, it tends to land on a beat alias between the S10 period and the 1-day observing cadence (i.e.
+the dotted lines).
+This is due to the fact that the single-frequency periodic model is biased, and significantly under-fits the data: it cannot distinguish residuals due to underfitting from residuals due to window function effects.
 
 In both models, the S10 period appears among the top 5 periods 99\% of the time: $477/483$ for supersmoother, and $480/483$ for multiband.\footnote{We would expect this correspondence to be 100\% in the case of the $g$-band supersmoother, which was the model used in the first pass of the S10 computation.
-This discrepancy can be chalked-up to the different supersmoother implementations used in S10 and in this work.
+This discrepancy here is likely due to the slightly different supersmoother implementations used in S10 and in this work.
 The discrepant curves are those with very low signal-to-noise.} This suggests that had S10 used the multiband Lomb-Scargle rather than the supersmoother in the first pass for that study, the final results presented there would be for the most part unchanged.
 
-The results of this subsection show that the shared-phase multiband approach is comparable to the supersmoother approach for densely-sampled multiband data, although it has a tendency to get fooled by structure in the survey window function.
+The results of this subsection show that the shared-phase multiband approach is comparable to the supersmoother approach for densely-sampled multiband data, although it has a tendency to get fooled by structure in the survey window.
+Correction for this based on the estimated window power may alleviate this; see \citet{Roberts87} for an example of such an approach.
 
@@
   \caption{
-    This figure shows a repeat of the results from \fig{compare_periods}, but the data is artificially reduced to only a single-band observation on each evening, a situation reflective of the observing cadence of future large-scale surveys (compare to \fig{compare_periods}).
-    In this case, the single-band SuperSmoother strategy used by \citet{Sesar2010} fails: there is simply not enough data in each band to recover an accurate period estimate, and the correct period is among the top 5 candidates in fewer than 60\% of cases.
-    The shared-phase multiband approach utilizes information from all five bands, and returns much more robust results: even with the reduced data, the true period is among the top 5 candidates in 94\% of cases.
+    This figure repeats the experiment shown in \fig{compare_periods}, but the data is artificially reduced to only a single-band observation on each evening, a situation reflective of the observing cadence of future large-scale surveys.
+    In this case, the single-band SuperSmoother strategy used as a first pass in S10 fails: there is simply not enough data in each band to recover an accurate period estimate, and the correct period is among the top 5 candidates in fewer than 50\% of cases.
+    The shared-phase multiband approach utilizes information from all five bands, and returns much more robust results: even with the greatly-reduced data, the true period is among the top 5 candidates in 93\% of cases.
   } 
@@
 
-Where we expect the multiband approach to have an advantage is when the data are sparsely sampled, with observations through only a single band each night.
-To simulate this, we reduce the Stripe 82 RR Lyrae dataset by a factor of 5, keeping only a single band of imaging each night, for an average of 11 observations of each object per band.
-This is much closer to the type of observation available in future multiband surveys.
+Above we saw that the multiband model is comparable to methods from the literature for densely-sampled data.
+Where we expect the multiband approach to gain an advantage is when the data are sparsely sampled, with observations through only a single band at each observation.
+To simulate this, we reduce the size of the Stripe 82 RR Lyrae dataset by a factor of 5, keeping only a single band of imaging each night, for an average of 11 observations of each object per band.
+This is much closer to the type of data which will be available in future multiband time-domain surveys.
 
 The upper panels of \fig{compare_periods_reduced} show an example light curve from this reduced dataset, along with the supersmoother and multiband periodograms derived from this data.
-Analogously to the lower panels of \fig{multiband_sim}, the single-band supersmoother model loses the true period within the noise, while the shared-phase multiband model still shows a noticeable spike at this true period.
+Analogously to the lower panels of \fig{multiband_sim}, the single-band supersmoother model loses the true period within the noise, while the shared-phase multiband model still shows a prominent spike at this true period.
 
 The lower panels of \fig{compare_periods_reduced} show the relationship between the S10 periods (based on the full dataset) and the periods derived with each model from this reduced dataset.
-It is clear that the supersmoother model is simply over-fitting noise with this few data points: the top period matches S10 in only 32\% of cases (compared to 87\% with the full dataset), and the top 5 periods contain the S10 period only 57\% of the time.
-The failure mode is much less regular as well: rather than being clustered near aliases, most of the points are scattered seemingly randomly around the space.
+It is clear that the supersmoother model is simply over-fitting noise with this few data points: the top period matches S10 in only 32\% of cases (compared to 87\% with the full dataset), and the top 5 periods contain the S10 period only 45\% of the time.
+The failure mode is much less predictable as well: rather than being clustered near aliases, most of the period determinations are scattered seemingly randomly around the space.
 
 The multiband method does much better on the reduced dataset.
-Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 66\% of the time (compared to 79\% with the full dataset), and the top 5 periods contain the S10 period 94\% of the time.
-This performance is due to the fact that the multiband algorithm has relatively few parameters, but is yet able to accommodate data from multiple observing bands.
+Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 64\% of the time (compared to 79\% with the full dataset), and the top 5 peaks contain the S10 period 94\% of the time.
+This performance is due to the fact that the multiband algorithm has relatively few parameters, but is yet able to flexibly accommodate data from multiple observing bands.
 In particular, this suggests that with the multiterm periodogram, the S10 analysis could have been done effectively with only a small fraction of the available data.
@@
 \begin{enumerate}
-  \item Our method kicks ass, therefore we kick ass. QED.
+  \item Our method rocks, therefore we rock. QED.
   \item Future work: account for beat-aliases for ground-based data
@@
 # Fetch the Sesar 2010 RR Lyrae data
-from multiband_LS.data import fetch_light_curves
-data = fetch_light_curves()
+from multiband_LS.data import fetch_rrlyrae
+data = fetch_rrlyrae()
 t, mag, dmag, filts = data.get_lightcurve(data.ids[0],

### 2015-01-24 multiband_LS/data.py

@@ -2,8 +2,8 @@
 Data downloaders for the multiband_LS stuff
 """
 
-__all__ = ['fetch_rrlyrae_templates', 'fetch_light_curves',
-           'fetch_lc_params', 'fetch_lc_fit_params']
+__all__ = ['fetch_rrlyrae_templates', 'fetch_rrlyrae',
+           'fetch_rrlyrae_lc_params', 'fetch_rrlyrae_fitdata']
 
 import os
 import tarfile
@@ -114,7 +114,7 @@ class RRLyraeLC(object):
         This is table 2 of Sesar 2010
         """
         if self._metadata is None:
-            self._metadata = fetch_lc_params()
+            self._metadata = fetch_rrlyrae_lc_params()
         i = np.where(self._metadata['id'] == lcid)[0]
         if len(i) == 0:
             raise ValueError("invalid lcid: {0}".format(lcid))
@@ -125,7 +125,7 @@ class RRLyraeLC(object):
         This is table 3 of Sesar 2010
         """
         if self._obsdata is None:
-            self._obsdata = fetch_lc_fit_params()
+            self._obsdata = fetch_rrlyrae_fitdata()
         i = np.where(self._obsdata['id'] == lcid)[0]
         if len(i) == 0:
             raise ValueError("invalid lcid: {0}".format(lcid))
@@ -135,17 +135,17 @@ class RRLyraeLC(object):
 class PartialRRLyraeLC(RRLyraeLC):
     """Class to get a partial Stripe 82 light curve: one band per night"""
     @classmethod
-    def from_rrlyrae(cls, rrlyrae, rseed=0):
+    def from_rrlyrae(cls, rrlyrae, offset=0):
         return cls(filename=rrlyrae.filename,
                    dirname=rrlyrae.dirname,
-                   rseed=rseed)
+                   offset=offset)
 
-    def __init__(self, filename, dirname='table1', rseed=0):
-        self.rseed = rseed
+    def __init__(self, filename, dirname='table1', offset=0):
+        self.offset = offset
         RRLyraeLC.__init__(self, filename, dirname)
 
     def __getstate__(self):
-        return (self.filename, self.dirname, self.rseed)
+        return (self.filename, self.dirname, self.offset)
 
     def __setstate__(self, args):
         self.__init__(*args)
[…]

### 2015-01-24 multiband_LS/generated.py

@@ -2,7 +2,7 @@
 
 import numpy as np
 from scipy.interpolate import interp1d
-from .data import fetch_rrlyrae_templates, fetch_light_curves
+from .data import fetch_rrlyrae_templates, fetch_rrlyrae
 
 
 class RRLyraeObject(object):
@@ -24,7 +24,7 @@ class RRLyraeObject(object):
     period : float
         Period of the RR Lyrae object
     """
-    lcdata = fetch_light_curves()
+    lcdata = fetch_rrlyrae()
     templates = fetch_rrlyrae_templates()
 
     # Extinction corrections: Table 1 from Berry et al. (2012, ApJ, 757, 166).

### 2015-01-24 multiband_LS/interactive.py

@@ -3,14 +3,14 @@ Utilities for creating IPython interactive plots to explore data & fits
 """
 import numpy as np
 from IPython.html.widgets import interact
-from .data import fetch_light_curves
+from .data import fetch_rrlyrae
 from . import LombScargleMultiband
 
 
 def interact_data():
     import matplotlib.pyplot as plt  # import here for backend safety
 
-    rrlyrae = fetch_light_curves()
+    rrlyrae = fetch_rrlyrae()
     lcids = np.fromiter(rrlyrae.ids, dtype=int)
     
     def plot_data(object_index=0, fold=False):
@@ -43,7 +43,7 @@ def interact_data():
 def interact_multifit():
     import matplotlib.pyplot as plt  # import here for backend safety
 
-    rrlyrae = fetch_light_curves()
+    rrlyrae = fetch_rrlyrae()
     lcids = np.fromiter(rrlyrae.ids, dtype=int)
     
     def plot_multifit(object_index=0, Nterms_base=4, Nterms_band=1):

### 2015-01-24 writeup/figures/compute_results.py

@@ -5,7 +5,7 @@ import numpy as np
 import os
 from datetime import datetime
 
-from multiband_LS.data import fetch_light_curves
+from multiband_LS.data import fetch_rrlyrae
 from multiband_LS import LombScargleMultiband, SuperSmoother, SuperSmootherMultiband
 
 
@@ -109,7 +109,7 @@ def get_period_results(outfile, lcids=None):
                                 "".format(outfile))
 
     if lcids is None:
-        lcids = fetch_light_curves().ids
+        lcids = fetch_rrlyrae().ids
 
     # turn generator into list
     lcids = list(lcids)
@@ -124,8 +124,8 @@ def get_period_results(outfile, lcids=None):
         
             
 if __name__ == '__main__':
-    rrlyrae = fetch_light_curves()
-    rrlyrae_partial = fetch_light_curves(partial=True)
+    rrlyrae = fetch_rrlyrae()
+    rrlyrae_partial = fetch_rrlyrae(partial=True)
 
     if not os.path.exists('results'):
         os.makedirs('results')

### 2015-01-24 writeup/figures/fig_compare_periods.py

@@ -16,7 +16,7 @@ import seaborn; seaborn.set()
 
 from multiband_LS import LombScargleMultiband, SuperSmootherMultiband
 from compute_results import SuperSmoother1Band
-from multiband_LS.data import fetch_light_curves
+from multiband_LS.data import fetch_rrlyrae
 
 from compute_results import get_period_results
 
@@ -157,8 +157,8 @@ def plot_periods(ssm_file, mbls_file, rrlyrae):
 
 
 if __name__ == '__main__':
-    rrlyrae = fetch_light_curves()
-    lcid = list(rrlyrae.ids)[4]
+    rrlyrae = fetch_rrlyrae()
+    lcid = list(rrlyrae.ids)[482]
 
     fig, ax = plot_example_lightcurve(rrlyrae, lcid)
     fig.savefig('fig07a.pdf')

### 2015-01-24 writeup/figures/fig_compare_periods_reduced.py

@@ -1,19 +1,18 @@
 """
 Plot figures comparing periods between multiband lomb scargle,
-supersmoother, and Sesar 2010 for partial dataset.
+supersmoother, and Sesar 2010 for reduced dataset.
 """
 import matplotlib.pyplot as plt
 from fig_compare_periods import plot_periods, plot_example_lightcurve
-from multiband_LS.data import fetch_light_curves
+from multiband_LS.data import fetch_rrlyrae
 
-rrlyrae = fetch_light_curves(partial=True)
-lcid = list(rrlyrae.ids)[4]
+rrlyrae = fetch_rrlyrae(partial=True)
+lcid = list(rrlyrae.ids)[482]
 
 fig, ax = plot_example_lightcurve(rrlyrae, lcid)
 fig.savefig('fig08a.pdf')
 
-# XXX: use new results here when available
-fig, ax = plot_periods(ssm_file='results_old/partial_supersmoother_g.npy',
+fig, ax = plot_periods(ssm_file='results/partial_supersmoother_g.npy',
                        mbls_file='results/partial_multiband_1_0.npy',
                        rrlyrae=rrlyrae)
 fig.savefig('fig08b.pdf')

### 2015-01-24 writeup/figures/fig_example.py

@@ -1,8 +1,8 @@
 from multiband_LS import LombScargleMultiband, SuperSmoother
 
 # Fetch the Sesar 2010 RR Lyrae data
-from multiband_LS.data import fetch_light_curves
-data = fetch_light_curves()
+from multiband_LS.data import fetch_rrlyrae
+data = fetch_rrlyrae()
 t, mag, dmag, filts = data.get_lightcurve(data.ids[0],
                                           return_1d=True)
 

### 2015-01-24 writeup/figures/fig_num_observations.py

@@ -2,11 +2,11 @@ import numpy as np
 import seaborn as sns
 import matplotlib.pyplot as plt
 
-from multiband_LS.data import fetch_light_curves
+from multiband_LS.data import fetch_rrlyrae
 
 
-rrlyrae = fetch_light_curves()
-rrlyrae_partial = fetch_light_curves(partial=True)
+rrlyrae = fetch_rrlyrae()
+rrlyrae_partial = fetch_rrlyrae(partial=True)
 
 data = [rrlyrae.get_lightcurve(lcid, True)
         for lcid in rrlyrae.ids]

### 2015-01-24 writeup/paper.tex
add acknowledgments

@@
 
-As an example of the standard periodogram in action, we perform a simple single-band harmonic analysis of simulated $r$-band observations of an RR Lyrae light curve, based on empirical templates derived in \citet{Sesar2010} (\fig{basic_example}).
+As an example of the standard periodogram in action, we perform a simple single-band harmonic analysis of simulated $r$-band observations of an RR Lyrae light curve, based on empirical templates derived in S10 (\fig{basic_example}).
 The observations are of a star with a period of 0.622 days, and take place on 60 random nights over a 6-month period, as seen in the left panel.
@@
 The full S10 RR Lyrae dataset consists of 483 objects with an average of 55 observations in each of the five SDSS $ugriz$ bands spread over just under ten years.
-In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother algorithm and the shared-phase $(0, 1)$-multiband algorithm\footnote{The supersmoother periodogram $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with the Lomb-Scargle periodogram's $\chi^2$ interpretation from \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$}.
+In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother algorithm and the shared-phase $(0, 1)$-multiband algorithm\footnote{The supersmoother ``periodogram'' $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with the Lomb-Scargle periodogram's $\chi^2$ interpretation from \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$}.
 Here we have a case which is analogous to that shown for simulated data in the top panels of \fig{multiband_sim}: each band has enough data to easily locate candidate peaks, the best of which is selected via the S10 template-fitting procedure.
@@
 
+{\it Acknowledgements:} JTV is supported by data science grants from the Alfred P.
+Sloan Foundation, the Gordon and Betty Moore Foundation, and the Washington Research Foundation.
+\todo{ZI is supported by...}  This work also made use of, and greatly benefitted from, free academic accounts provided by GitHub.
+
 \bibliographystyle{apj}

### 2015-01-25 writeup/paper.tex
add some tweaks and some references

@@
 
-The API for the module is largely influenced by that of the {\tt scikit-learn} package\footnote{\url{http://scikit-learn.org}} \citep{scikit-learn}, in which models are class objects which can be ``fit()`` to data.
-Here is a basic example of how you can use the code to download data, fit a multiband model to the data, and compute the power at a few periods:
+The API for the module is largely influenced by that of the {\tt scikit-learn} package\footnote{\url{http://scikit-learn.org}} \citep{scikit-learn}, in which models are Python class objects which can be fit to data with the \texttt{fit()} method.
+Here is a basic example of how you can use the package to download the data used in this paper, fit a multiband model to the data, and compute the power at a few periods:
 
@@
 # Compute power at the following periods
-model.periodogram([0.2, 0.3, 0.4])
+periods = [0.2, 0.3, 0.4, 0.5] # periods in days
+power = model.periodogram(periods)
 \end{lstlisting}
 
-Other models are available as well. For example, here is how you can compute the periodogram under the supersmoother model:
+Other models are available as well.
+For example, here is how you can compute the periodogram under the supersmoother model; this implementation of the supersmoother periodogram makes use of the \texttt{supersmoother} Python package \citep{Vanderplas2015}.
 
@@
 # Compute power at the following periods
-model.periodogram([0.2, 0.3, 0.4])
+power = model.periodogram(periods)
 \end{lstlisting}
 
-The code contains many more methods, and much more functionality that what is shown here. For updates and more information, visit \url{http://github.com/jakevdp/multiband_LS}.
+The models in the \texttt{multiband\_LS} package contain many more methods, and much more functionality that what is shown here.
+For updates, more examples, and more information, visit \url{http://github.com/jakevdp/multiband_LS}.
 

### 2015-01-28 writeup/paper.tex
JTV pass on paper; add conclusions

@@
 method compared to other methods from the literature. The Python implementation of 
-this method is made publicly available as part of the astroML distribution.
+this method, along with code to fully reproduce the results reported here, is available on GitHub.
 \end{abstract}
@@
 \section{Introduction}
+\sectlabel{introduction}
 
@@
 
-The most popular parametric method is the Lomb-Scargle periodogram, which is discussed in detail in \S~\ref{sec:periodograms}.
+The most popular parametric method is the Lomb-Scargle periodogram, which is discussed in detail in \sect{brief_overview}.
 The Lomb-Scargle periodogram is related to the $\chi^2$ for a least-square fit of a single sinusoid to data
@@
 A more general parameteric method based on the use of continuous-time autoregressive moving average (CARMA) model
-was recently introduced by \citep{Kelly14}. CARMA models can also treat non-uniformly sampled time series with 
+was recently introduced by \citet{Kelly14}. CARMA models can also treat non-uniformly sampled time series with 
 heteroscedastic measurement uncertainties, and can handle complex variability patterns.
@@
 advantage of the full information content in available data, it would be desirable to have a single estimate of the periodogram 
-which accounts for all observed data in a manner which is not dependent on the underlying spectrum of the object. 
+which accounts for all observed data in a manner which does not depend on the underlying spectrum of the object. 
 We propose such a method in this paper.
@@
 
-\todo{In \S 2, we provide a brief overview of periodograms, including the matrix-based formulation used in this work. In \S... \citet[][hereafter S10]{Sesar2010}...}
+In \sect{brief_overview} we provide a brief overview of least-squares periodic fitting, and in \sect{matrix_formalism} derive the matrix-based formalism for single-band periodic analysis used through the rest of this work.
+\sect{extending_periodogram} introduces several extensions and generalizations to the single-band model that the matrix formalism makes possible, including floating mean models, truncated Fourier models, and regularized models.
+In \sect{multiband}, we use the ideas behind these extensions to motivate the {\it multiband periodogram}, and show some examples of its use on simulated data.
+In \sect{stripe82} we apply this method to measurements of 483 RR Lyrae stars first explored by \citet[][hereafter S10]{Sesar2010}, and conclude in \sect{discussion}.
 
-\section{Brief Overview of Periodograms \label{sec:periodograms}}
+\section{Brief Overview of Periodograms}
+\sectlabel{brief_overview}
 
@@
   \eqlabel{LombScargle}
-  P_N(\omega) = \frac{1}{2\,V}\left[
+  P_N(\omega) = \frac{1}{2\,V_y}\left[
     \frac{\left[\sum_k(y_k - \bar{y})\cos\omega(t_k - \tau)\right]^2}
@@
 \end{equation}
-where $\bar{y}$ is the mean and $V$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
+where $\bar{y}$ is the mean and $V_y$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
 \citet{Lomb76} showed that this time-offset has a deeper effect: namely, it makes $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
-
 \begin{equation}
@@
 \end{equation}
-
 This long-recognized connection between spectral power and least squares fitting methods was solidified by \citet{Jaynes87}, who demonstrated that the normalized periodogram of Lomb and Scargle is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
@@
 
-While the important features of least squares frequency estimation via Lomb-Scargle periodograms has been discussed elsewhere, we will prevent a brief introduction to the subject in the following section.
-The matrix-based formalism we develop here will make clear how the method can be extended to more sophisticated models, including regularized multiband methods.
+While the important features of least squares frequency estimation via Lomb-Scargle periodograms has been discussed elsewhere, we will present a brief introduction to the subject in the following section.
+The matrix-based formalism we develop here will make clear how the method can be extended to more sophisticated models, including the multiband method.
 
-\section{Background: Matrix Formalism for Periodograms}
+\section{Standard Least Squares Spectral Fitting}
+\sectlabel{matrix_formalism}
 
@@
 \end{equation}
-For a given $\omega$, we can find the maximum likelihood estimate of the parameters $\theta$ by minimizing the $\chi^2$ of the model, which is given by
+For a given $\omega$, the maximum likelihood estimate of the parameters $\theta$ can be found by minimizing the $\chi^2$ of the model, which is given by
 \begin{equation}
@@
 
-\subsubsection{Matrix Formalism}
+\subsection{Matrix Formalism}
 The expressions related to the stationary sinusoid model can be expressed more compactly by defining the following matrices:
@@
 \end{equation}
-We note that this expression is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto V I$.
+We note that this expression is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto V_y I$.
 
-\subsubsection{Simple Period Finding}
+\subsection{Simple Single-band Period Finding}
 \sectlabel{simple_period}
@@
 The left panel shows the input data, while the right panels show the fit derived from the data.
-The top-right panel shows the periodogram, showing a distinct peak at the true period of 0.622 days, and the bottom-right panel shows the data as a function of phase as defined by this period.
+The upper-right panel shows the periodogram with a clear peak at the true period of 0.622 days, and the bottom-right panel shows the data as a function of the phase associated with this period.
 Note in the periodogram the presence of the typical aliasing effect, with power located at beat frequencies between the true period and the 1-day observing cadence (see \sect{simple_period} for further discussion).
@@
 The lower-right panel of \fig{basic_example} shows the maximum likelihood interpretation of this periodogram: it is a measure of the normalized $\chi^2$ for a single-term sinusoidal model.
-Here we visualize the data from the left panel, but folded as a function of phase with the best-fit single-term model over-plot.
+Here we visualize the data from the left panel, but folded as a function of phase, and overplotted with the best-fit single-term model.
 Here it is apparent that the single-term model is highly biased: RR Lyrae light curves are, in general, much more complicated than a simple sinusoid.
@@
 
-\subsection{Extending the Periodogram}
+\section{Extending the Periodogram}
+\sectlabel{extending_periodogram}
 We have shown two forms of the classic normalized periodogram: \eq{LombScargle} and \eq{LombScargle2}.
-Though the two expressions are equivalent, they each have distinct advantages.
-The expression in \eq{LombScargle} avoids the explicit construction of a matrix, and thus can be computed very efficiently.
-Furthermore, through some computational tricks taking advantage of the Fast Fourier Transform, expressions of the form of \eq{LombScargle} can be evaluated exactly for $N$ frequencies in $\mathcal{O}[\log{N}]$ time \citep{Press89}.
+Though the two expressions are equivalent, they differ in their utility.
+Because the expression in \eq{LombScargle} avoids the explicit construction of a matrix, it can be computed very efficiently.
+Furthermore, through clever use of the Fast Fourier Transform, expressions of the form of \eq{LombScargle} can be evaluated exactly for $N$ frequencies in $\mathcal{O}[\log{N}]$ time \citep{Press89}.
 
@@
 \end{enumerate}
-In the remainder of this section, we'll explore a few of these modifications and how they affect the periodogram and resulting model fits.
+In the remainder of this section, we will explore a few of these modifications and how they affect the periodogram and resulting model fits.
 
-\subsubsection{Stationary Sinusoid with Floating Mean}
+\subsection{Stationary Sinusoid with Floating Mean}
 \sectlabel{floating_mean}
@@
   \caption{
-    An illustration of the effect of the floating mean model for biased data.
+    An illustration of the effect of the floating mean model for censored data.
     The data consist of 80 observations drawn from a sinusoidal model.
@@
 The standard and floating-mean periodograms are computed from the remaining data; these fits are shown over the data in the left panel.
-Because of this biased observing pattern, the mean of the observed data is not a good predictor of the true mean, and the standard fixed-mean model fails to recover the true period of 0.622 days, while the floating mean model still finds the correct period.
+Because of this biased observing pattern, the mean of the observed data is a biased estimator of the true mean.
+The standard fixed-mean model in this case fails to recover the true period of 0.622 days, while the floating mean model still finds the correct period.
   }
@@
 
-As an example of one of these generalizations, consider the {\it generalized Lomb-Scargle} method of \citet{Zechmeister09}.
-This adjusts the classic normalized periodogram by allowing the mean of the model to be fit alongside the amplitudes:
+As an example of one of these generalizations, we'll consider what \citet{Zechmeister09} call the ``generalized Lomb-Scargle method'', and which we'll call the {\it floating-mean periodogram}.
+This method adjusts the classic normalized periodogram by fitting the mean of the model alongside the amplitudes:
 \begin{equation}
@@
 This model can be more accurate than the standard periodogram for certain observing cadences and selection functions.
-\citet{Zechmeister09} detail the modifications required to the harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
+\citet{Zechmeister09} detail the required modifications to the harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
 In the matrix formalism, the modification is much more straightforward: all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
+This column of ones corresponds to a third entry in the parameter vector $\theta$, and acts as a uniform constant offset for all datapoints.
 
@@
 The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias (left panel).
-Because of this observational bias, the mean of the observed data are a poor predictor of the true mean, causing the standard fixed-mean method to poorly fit the data (upper-right panel).
+Because of this observational bias, the mean of the observed data are a poor predictor of the true mean, causing the standard fixed-mean method to poorly fit the data and miss the input period (upper-right panel).
 The floating-mean approach is able to automatically adjust for this bias, resulting in a periodogram which readily detects the input period of 0.622 days (lower-right panel).
 
-\subsubsection{Truncated Fourier Models}
+\subsection{Truncated Fourier Models}
 \sectlabel{multiterm}
@@
   \caption{
-    An illustration of a periodogram based on several truncated Fourier models.
+    The model fits and periodograms for several truncated Fourier models.
     The data are the same as those in \fig{basic_example}. Note that the
@@
 A natural extension is to instead use a multiple-term sinusoidal model, with frequencies at integer multiples of the fundamental frequency.
-With $M$ Fourier terms, there are $2M + 1$ free parameters, and the model is given by
+With $N$ Fourier terms, there are $2N + 1$ free parameters, and the model is given by
 \begin{equation}
-  y(t|\omega,\theta) = \theta_0 + \sum_{n=1}^M \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right].
+  y(t|\omega,\theta) = \theta_0 + \sum_{n=1}^N \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right].
 \end{equation}
 Because this model remains linear in the parameters $\theta$, it can be easily accommodated into the matrix formalism above.
-For example, an $M = 2$-term floating-mean model can be constructed by building a design matrix $X_\omega$ with $2M + 1 = 5$ columns:
+For example, an $N = 2$-term floating-mean model can be constructed by building a design matrix $X_\omega$ with $2N + 1 = 5$ columns:
 \begin{equation}
@@
 Computing the power via \eq{LombScargle2} using $X_\omega^{(2)}$ will give the two-term periodogram.
-As $M$ grows, the size of the design matrix grows, but the periodogram can be computed in the same manner.
-\fig{multiterm_example} shows a few examples of this multiterm Fourier approach as applied to the simulated RR Lyrae light curve from \fig{basic_example}.
-There are several important features of this figure that give us insight into the subtleties of this type of multiterm fit.
+For larger $N$, more columns are added, but the periodogram can be computed in the same manner.
+\fig{multiterm_example} shows a few examples of this multiterm Fourier approach as applied to the simulated RR Lyrae light curve from \fig{basic_example}, and illustrates several important insights into the subtleties of this type of multiterm fit.
 
-First, we see in the right panel that all three models show a clear spike at the true period of $P_0 = 0.622$.
-The higher-order models, however, also show a a spike in power at $P_1 = 2 P_0$: the reason for this is that for and $M>1$-term model, the period $P_0$ is the first harmonic of a model with fundamental frequency $2P_0$, and the higher-order models contain the single-period result.
+First, we see in the right panel that all three models show a clear signal at the true period of $P_0 = 0.622$ days.
+The higher-order models, however, also show a a spike in power at $P_1 = 2 P_0$: the reason for this is that for and $N>1$-term model, the period $P_0$ is the first harmonic of a model with fundamental frequency $2P_0$, and the higher-order models contain the single-period result.
 
@@
 A more flexible higher-order model can better fit the data at all periods, not just the true period.
-Thus the observed power of a higher-order Fourier model will be everywhere higher than the power of a lower-order Fourier model.
+Thus in general the observed power of a higher-order Fourier model will be everywhere higher than the power of a lower-order Fourier model.
 
-\subsubsection{Regularized Models \label{regularization}}
+\subsection{Regularized Models \label{regularization}}
 
@@
 We have seen that as we add more terms to the fit, the model will come closer and closer to the observed data.
-When the number of model parameters equals the number of data points, we can expect to fit the data {\it perfectly} (though in most cases numerical inaccuracies in the fit prevent a truly perfect fit).
-For very high-order models, such a fit becomes very sensitive to the noise in the data, and we say we have {\it over-fit} the data.
-This can be addressed by explicitly truncating the series, but we can also use a {\it regularization} term to mathematically ensure a less complicated model.
+In the extreme case, when the number of model parameters equals the number of data points, the model can fit the data {\it exactly} regardless of frequency and the periodogram will be everywhere unity (though in most cases, numerical inaccuracies in the fit prevent a truly perfect fit).
+For very high-order models, the fit becomes very sensitive to the noise in the data, and we say we have {\it over-fit} the data.
+This can be addressed by explicitly truncating the series, but we can also use a {\it regularization} term to mathematically enforce a less complicated model.
 
@@
 For computational simplicity here we'll use an {\it L2 regularization} -- also known as Tikhonov Regularization \citep{Tikhonov1963} or Ridge Regression \citep{Hoerl1970} -- which is a quadratic penalty term in the model parameters added to the $\chi^2$.
-Mathematically, this is equivalent to the Bayesian approach of using a zero-mean Gaussian prior on the model parameters.
+Mathematically, this is equivalent in the Bayesian framework to using a zero-mean Gaussian prior on the model parameters.
 
-We encode our regularization in the matrix $\Lambda = {\rm diag}([\lambda_1, \lambda_2 \cdots \lambda_M])$ for a model with $M$ parameters, and our expression for ``regularized'' $\chi^2$ becomes
+We encode our regularization in the matrix $\Lambda = {\rm diag}([\lambda_1, \lambda_2 \cdots \lambda_M])$ for a model with $M$ parameters, and construct a ``regularized'' $\chi^2$:
 \begin{equation}
@@
   \eqlabel{LombScargleReg}
-  P_N(\omega) = \frac{y^T\Sigma^{-1}X_\omega~[X_\omega^T\Sigma^{-1}X_\omega + \Lambda]^{-1}~X_\omega^T\Sigma^{-1}y}{y^T\Sigma^{-1}y}.
+  P_{N,\Lambda}(\omega) = \frac{y^T\Sigma^{-1}X_\omega~[X_\omega^T\Sigma^{-1}X_\omega + \Lambda]^{-1}~X_\omega^T\Sigma^{-1}y}{y^T\Sigma^{-1}y}.
 \end{equation}
@@
 We use $\lambda = 0$ on the offset term, and make the penalty $\lambda_j$ progressively larger for each harmonic component.
-The result of the regularized model is less over-fitting to the input data, and a more distinct peaks in the associated periodogram.
+The result of the regularized model is less over-fitting to the input data, and stronger periodogram peaks.
 
@@
 \sectlabel{multiband}
-To compute a periodogram for multiband data, we'll take advantage of many of the nice features of the matrix form of the normalized periodogram, which easily accommodates arbitrary linear models covering our data.
-To account for data in multible bands, we will construct a multi-term Fourier model with the following features:
+In this section we will apply what was learned in the previous sections to construct the {\it multiband periodogram} which can flexibly account for heterogeneous sources of data for a single object.
+To compute such a periodogram, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
+The model contains the following features:
 \begin{enumerate}
@@
 For example, the $N_{base}=1$ and $N_{band}=0$ model corresponds to one with a simple single-term periodic base frequency, and an independent constant offset term in each band.
-The associated design matrix will look as follows:
+The associated design matrix depends on the particular data, but will look similar to this:
 \begin{equation}
@@
 1 & \sin(\omega t_1) & \cos(\omega t_1) & 1 & 0 & 0 & 0 & 0\\
-1 & \sin(\omega t_2) & \cos(\omega t_2) & 0 & 1 & 0 & 0 & 0\\
+1 & \sin(\omega t_2) & \cos(\omega t_2) & 0 & 0 & 0 & 0 & 1\\
 1 & \sin(\omega t_3) & \cos(\omega t_3) & 0 & 0 & 0 & 1 & 0\\
@@
 \end{equation}
-Here the nonzero entries of the final five columns are binary flags indicating the $ugriz$ band of the given observation: in this case, the first row is a $u$-band measurement, the second is a $g$-band, the third is a $i$-band, etc.
+Here the nonzero entries of the final five columns are binary flags indicating the $(u, g, r, i, z)$-band of the given observation: for this example, the first row is a $u$-band measurement, the second is a $z$-band, the third is a $i$-band, etc.
 
@@
 $X_\omega$ is low-rank), and thus the parameters of the best-fit model will be degenerate.
-Intuitively, this is due to the fact that if we add an overall offset to the base model, this can be perfectly accounted for by subtracting that same offset from each of the band columns.
+Intuitively, this is due to the fact that if we add an overall offset to the base model, this can be perfectly accounted for by subtracting that same offset from each residual.
 Mathematically, the result of this is that the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$ will be non-invertible, and thus the periodogram is ill-defined.
@@
 \sectlabel{relationship}
-The {\it multi-phase} $N_{base}=0$, $N_{band}=1$ model turns out to be a particularly special case.
+The {\it multi-phase} $(N_{base}=0, N_{band}=1)$ model turns out to be a particularly special case.
 The base model is a simple global offset which is degenerate with the offsets in each band, so that the design matrix $X_\omega$ can be straightforwardly rearranged as block-diagonal.
-A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here these independent components amount to the single-band floating-mean model from \sect{floating_mean}, fit independently within each of the $K$ bands.
+A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here these independent components amount to the single-band floating-mean model from \sect{floating_mean}, fit independently for each of the $K$ bands.
 This particular multiband model can give us insight into the relationship between single-band and multiband approaches.
@@
 \end{equation}
-This tells us how we can construct a particular case of the multi-term periodogram from a weighted sum of periodograms in each band, where the weights $\chi^2_{0,k}$ reflect both the number of measurements in each band and how much those measurements deviate from a simple constant reference model.
+We see that this particular multiband periodogram is identical to a weighted sum of standard periodograms in each band, where the weights $\chi^2_{0,k}$ are a reflection of both the number of measurements in each band, and how much those measurements deviate from a simple constant reference model.
 
@@
 
-Before applying the multiband method to real data, let's explore its effectiveness on a simulated RR Lyrae lightcurve.
+Before applying the multiband method to real data, we will here explore its effectiveness on a simulated RR Lyrae lightcurve.
 The upper panels of \fig{multiband_sim} show a multiband version of the simulated RR Lyrae light curve from \fig{basic_example}.
-The upper-right panel shows 60 nights of observations spread over a 6-month period, and for each night all five bands ({\it u,g,r,i,z}) are recorded.
-Using the typical approach from the literature prior to this paper, we individually compute the standard normalized periodogram within each band: the results are shown in the upper-right panel.
-Here the data is well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band, up to the aliasing effect discussed in \sect{simple_period}.
+The upper-left panel shows 60 nights of observations spread over a 6-month period, and for each night all five bands ({\it u,g,r,i,z}) are recorded.
+Using the typical approach from the literature, we individually compute the standard normalized periodogram within each band: the results are shown in the upper-right panel.
+The data are well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band, up to the aliasing effect discussed in \sect{simple_period}.
 Previous studies have made use of the information in multiple bands to choose between aliases and estimate uncertainties in determined periods \citep[e.g.][]{Oluseyi12,Sesar2010}.
-While this approach is sufficient for well-sampled data, it becomes problematic when the multiband data is more sparsely sampled.
+While this approach is sufficient for well-sampled data, it becomes problematic when the multiband data are sparsely sampled.
 
@@
 The shared-phase multiband approach (i.e.
-$(N_{base},N_{band})=(1,0)$), shown in the lower-right panel, makes use of all the information in a single model, and clearly recovers the true frequency of 0.622 days.
+$(N_{base}=1,N_{band}=0)$), shown in the lower-right panel, makes use of all the information in a single model, and clearly recovers the true frequency of 0.622 days.
 
-This shared-phase $(1,0)$ model is only one of the possible multiband options, however: \fig{multiband_models} shows multiband fits to this data for models with various combinations of $(N_{base},N_{band})$.
+This shared-phase $(1,0)$ model is only one of the possible multiband options, however: \fig{multiband_models} shows multiband fits to this data for models with various choices of $(N_{base},N_{band})$.
 We see here many of the characteristics noted above for single-band models: as discussed in \sect{multiterm}, increasing the number of Fourier terms leads to power at multiples of the fundamental period.
@@
 For this reason, models with $N_{base} > N_{band}$ are the most promising: they allow a flexible fit with minimal model complexity.
-
-Below, we'll apply the simplest of this class of models, the $(1, 0)$ shared-phase model, to data from the Stripe 82 of the Sloan Digital Sky Survey.
+Motivated by this, in the next section we'll apply the simplest of this class of models, the $(1, 0)$ shared-phase model, to data from the Stripe 82 of the Sloan Digital Sky Survey.
 
 \section{Application to Stripe 82 RR Lyrae}
-
-Stripe 82 is a three hundred square degree equatorial region of the sky which was repeatedly imaged through multiple band-passes during the second phase of the Sloan Digital Sky Survey (SDSS II).
-As a real-world test of the multiband periodogram method, we consider the SDSS-2 observations of 483 RR Lyrae stars compiled and studied by S10.
-S10 determined periods for these stars based on an emperically-derived light curve templates.
-Because the template-fitting method is extremely computationally intensive, S10 first determined candidate periods by taking the top 5 results of the supersmoother \citep{Reimann94} algorithm applied to the $g$-band; template fits were then performed at each candidate period and the period with the best template fit was reported as the true period.
-In this section, we make use of this dataset to quantitatively evaluate the effectiveness of the multiband periodogram approach in the case of this real-world data.
+\sectlabel{stripe82}
+Stripe 82 is a three hundred square degree equatorial region of the sky which was repeatedly imaged through multiple band-passes during phase II of the Sloan Digital Sky Survey \citep[SDSS II, see][]{Sesar2007}.
+Here we consider the SDSS II observations of 483 RR Lyrae stars compiled and studied by S10.
+S10 determined periods for these stars based on emperically-derived light curve templates.
+Because the template-fitting method is extremely computationally intensive, S10 first determined candidate periods by taking the top 5 results of the Supersmoother \citep{Reimann94} algorithm applied to the $g$-band; template fits were then performed at each candidate period and the period with the best template fit was reported as the true period.
+In this section, we make use of this dataset to quantitatively evaluate the effectiveness of the multiband periodogram approach.
 
@@
     The upper panels show a representative lightcurve and periodogram fits, while the bottom panels compare the derived periods to the template-based periods reported in S10.
-    Shown for reference are the beat aliases (dotted lines) and the multiplicative alias (dashed lines), along with counts of points aligned with each trend.
+    Shown for reference are the beat aliases (dotted lines) and the multiplicative alias (dashed lines): numbers along the top and right edges of the panels indicate the number of points aligned with each trend.
     The single-band supersmoother algorithm tends to err toward multiplicative aliases, while the multiband algorithm tends to err toward beat frequency aliases.
@@
 The full S10 RR Lyrae dataset consists of 483 objects with an average of 55 observations in each of the five SDSS $ugriz$ bands spread over just under ten years.
-In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother algorithm and the shared-phase $(0, 1)$-multiband algorithm\footnote{The supersmoother ``periodogram'' $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with the Lomb-Scargle periodogram's $\chi^2$ interpretation from \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$}.
+In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother algorithm and the shared-phase $(0, 1)$-multiband algorithm\footnote{The supersmoother ``periodogram'' $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$, where $\bar{r}_0$ is the residual around a constant model.}.
 Here we have a case which is analogous to that shown for simulated data in the top panels of \fig{multiband_sim}: each band has enough data to easily locate candidate peaks, the best of which is selected via the S10 template-fitting procedure.
@@
 This is due to the flexibility of the supersmoother model: a doubled period spreads the points out, leading to fewer constraints in each neighborhood and thus a smaller average residual around model.
-In other words, the SuperSmoother tends to over-fit sparsely-sampled data.
+In other words, the SuperSmoother tends to over-fit data which is sparsely-sampled.
 On the other hand, when the multiband model misses the S10 period, it tends to land on a beat alias between the S10 period and the 1-day observing cadence (i.e.
@@
 
-In both models, the S10 period appears among the top 5 periods 99\% of the time: $477/483$ for supersmoother, and $480/483$ for multiband.\footnote{We would expect this correspondence to be 100\% in the case of the $g$-band supersmoother, which was the model used in the first pass of the S10 computation.
+In both models, the S10 period appears among the top 5 periods 99\% of the time: $477/483$ for supersmoother, and $480/483$ for multiband.\footnote{We might expect this correspondence to be 100\% in the case of the $g$-band supersmoother, which was the model used in the first pass of the S10 computation.
 This discrepancy here is likely due to the slightly different supersmoother implementations used in S10 and in this work.
-The discrepant curves are those with very low signal-to-noise.} This suggests that had S10 used the multiband Lomb-Scargle rather than the supersmoother in the first pass for that study, the final results presented there would be for the most part unchanged.
+Objects showing this discrepancy are those with very low signal-to-noise.} This suggests that had S10 used the multiband Lomb-Scargle rather than the supersmoother in the first pass for that study, the final results presented there would be for the most part unchanged.
 
 The results of this subsection show that the shared-phase multiband approach is comparable to the supersmoother approach for densely-sampled multiband data, although it has a tendency to get fooled by structure in the survey window.
-Correction for this based on the estimated window power may alleviate this; see \citet{Roberts87} for an example of such an approach.
+Correction for this based on the estimated window power may alleviate this (see \citet{Roberts87} for an example of such an approach) though in practice selecting from among the top 5 peaks appears to be sufficient.
 
@@
   \caption{
-    This figure repeats the experiment shown in \fig{compare_periods}, but the data is artificially reduced to only a single-band observation on each evening, a situation reflective of the observing cadence of future large-scale surveys.
-    In this case, the single-band SuperSmoother strategy used as a first pass in S10 fails: there is simply not enough data in each band to recover an accurate period estimate, and the correct period is among the top 5 candidates in fewer than 50\% of cases.
+    This figure repeats the experiment shown in \fig{compare_periods} (see caption there for description), but the data is artificially reduced to only a single-band observation on each evening, a situation reflective of the observing cadence of future large-scale surveys.
+    In this case, the single-band SuperSmoother strategy used as a first pass in S10 fails: there is simply not enough data in each band to recover an accurate period estimate.
+The correct period is among the top 5 candidates in fewer than 50\% of cases.
     The shared-phase multiband approach utilizes information from all five bands, and returns much more robust results: even with the greatly-reduced data, the true period is among the top 5 candidates in 93\% of cases.
@@
 Where we expect the multiband approach to gain an advantage is when the data are sparsely sampled, with observations through only a single band at each observation.
-To simulate this, we reduce the size of the Stripe 82 RR Lyrae dataset by a factor of 5, keeping only a single band of imaging each night, for an average of 11 observations of each object per band.
+To simulate this, we reduce the size of the Stripe 82 RR Lyrae dataset by a factor of 5, keeping only a single band of imaging each night: an average of 11 observations of each object per band.
 This is much closer to the type of data which will be available in future multiband time-domain surveys.
@@
 The upper panels of \fig{compare_periods_reduced} show an example light curve from this reduced dataset, along with the supersmoother and multiband periodograms derived from this data.
-Analogously to the lower panels of \fig{multiband_sim}, the single-band supersmoother model loses the true period within the noise, while the shared-phase multiband model still shows a prominent spike at this true period.
+Analogously to the lower panels of \fig{multiband_sim}, the single-band supersmoother model loses the true period within the noise, while the shared-phase multiband model still shows prominent signal near the S10 period.
 
@@
 The multiband method does much better on the reduced dataset.
-Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 64\% of the time (compared to 79\% with the full dataset), and the top 5 peaks contain the S10 period 94\% of the time.
+Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 64\% of the time (compared to 79\% with the full dataset), and the top 5 peaks contain the S10 period 94\% of the time (compared to 99\% with the full dataset).
 This performance is due to the fact that the multiband algorithm has relatively few parameters, but is yet able to flexibly accommodate data from multiple observing bands.
@@
 \section{Discussion and Conclusion}
+\sectlabel{discussion}
 
-\todo{Zeljko, can you take a stab at this?}
+We have motivated and derived a multiband version of the classic Lomb-Scargle method for detecting periodicity in astronomical time-series.
+ Experiments on several hundred RR Lyrae stars from the Stripe 82 dataset indicate that this method outperforms methods used previously in the literature, especially for sparsely-sampled light curves with only single bands observed each night.
 
-\begin{enumerate}
-  \item Our method rocks, therefore we rock. QED.
-  \item Future work: account for beat-aliases for ground-based data
-  \item Future work: physically-motivated priors to further simplify the model
-  \item Opportunity to apply this to PanSTARRS data release
-\end{enumerate}
+The main deficiency of the model is its tendency to be tricked by the survey window function, a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle model bias from window bias.
+This could potentially be accounted for by estimating and filtering the effect of the window function; one potential method for this involves estimating the deconvolution of the window power and the observed power.
+It may also be possible to propose a multiband extension of CARMA or another forward-modeling approaches to detecting periodicity.
 
-{\it Acknowledgements:} JTV is supported by data science grants from the Alfred P.
+Another potentially fruitful way forward in the multiband case is to make use of physically-motivated priors in the light-curve shape.
+For example, the paths of RR Lyrae stars through color-color and color-magnitude space are constrained by known astrophysical processes in the structure of the stars, and making use of such a physically-motivated prior may help break degeneracies in period determination.
+
+Looking forward to future variable star catalogs from PanSTARRS, DES, and LSST, there are two important constraints that any analysis method must meet: the methods must be able to cope with heterogeneous and noisy observations through multiple band-passes, and the methods must be fast enough to be computable on millions or even billions of objects.
+The multiband method, through its combination of flexibility and model simplicity, meets the first: as shown above, it performs well in the case of sparsely-sampled multiband data.
+It also meets the second constraint: it consists of a fast linear model at each frequency, compared to a sliding-window model in the case of supersmoother, a nonlinear optimization in the case of template-fitting, and a markov chain monte carlo in the case of CARMA models.
+In our own benchmarks, we found the multiband method to be several times faster than the single-band supersmoother approach, and several orders of magnitude faster than the template fitting approach.
+This suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys.
+
+{\it Acknowledgements:} JTV is supported by the University of Washington eScience institute, including grants from the Alfred P.
 Sloan Foundation, the Gordon and Betty Moore Foundation, and the Washington Research Foundation.
-\todo{ZI is supported by...}  This work also made use of, and greatly benefitted from, free academic accounts provided by GitHub.
+\todo{ZI is supported by...}  The authors thank GitHub for providing free academic accounts which were essential in the development of this work.
 
@@
 \section{{\tt multiband\_LS}: Python Implementation of Multiband Periodogram}
+\sectlabel{multiband_LS}
 The algorithm outlined in this paper is available as an open-source Python package\footnote{\url{http://github.com/jakevdp/multiband_LS/}}, along with code to download the data and recreate all the figures in this paper.
-{\tt multiband\_LS} is a pure-Python package written to be compatible with both Python 2 and Python 3, and performs fast numerical computation via the {\tt numpy} package\footnote{http://www.numpy.org}.
+{\tt multiband\_LS} is a pure-Python package written to be compatible with both Python 2 and Python 3, and performs fast numerical computation through dependencies on {\tt numpy} \citep{numpy}\footnote{\url{http://www.numpy.org}} and {\tt astroML} \citep{astroML}\footnote{\url{http://www.astroml.org}}, which offer optimized implementatinos of numerical methods in Python.
 
-The API for the module is largely influenced by that of the {\tt scikit-learn} package\footnote{\url{http://scikit-learn.org}} \citep{scikit-learn}, in which models are Python class objects which can be fit to data with the \texttt{fit()} method.
-Here is a basic example of how you can use the package to download the data used in this paper, fit a multiband model to the data, and compute the power at a few periods:
+The API for the module is largely influenced by that of the {\tt scikit-learn} package \citep{scikit-learn, sklearn_API}\footnote{\url{http://scikit-learn.org}}, in which models are Python class objects which can be fit to data with the \texttt{fit()} method.
+Here is a basic example of how you can use {\tt multiband\_LS} to download the data used in this paper, fit a multiband model to the data, and compute the power at a few periods:
 
@@
 from multiband_LS import LombScargleMultiband
+import numpy as np
 
@@
 # Compute power at the following periods
-periods = [0.2, 0.3, 0.4, 0.5] # periods in days
+periods = np.linspace(0.2, 1.4, 1000) # periods in days
 power = model.periodogram(periods)

### 2015-01-28 multiband_LS/__init__.py
change fig code to use gatspy

@@ -1,9 +1,0 @@
-"""General Periodic Modeling for Astronomical Time Series"""
-
-__version__ = "0.1"
-
-from .lomb_scargle import LombScargle, LombScargleAstroML
-from .lomb_scargle_multiband import (LombScargleMultiband,
-                                     LombScargleMultibandFast)
-from .supersmoother import SuperSmoother, SuperSmootherMultiband
-from .naive_multiband import NaiveMultiband

### 2015-01-28 multiband_LS/data.py

@@ -1,264 +1,0 @@
-"""
-Data downloaders for the multiband_LS stuff
-"""
-
-__all__ = ['fetch_rrlyrae_templates', 'fetch_rrlyrae',
-           'fetch_rrlyrae_lc_params', 'fetch_rrlyrae_fitdata']
-
-import os
-import tarfile
-import gzip
-from io import BytesIO
-
-import numpy as np
-from astroML.datasets.tools import download_with_progress_bar
-
-
-def set_data_directory(data_directory):
-    global DATA_DIRECTORY
-    DATA_DIRECTORY = data_directory
-
-    if not os.path.exists(DATA_DIRECTORY):
-        os.makedirs(DATA_DIRECTORY)
-
-DATA_DIRECTORY = ''
-set_data_directory(os.path.abspath(os.path.join(os.path.dirname(__file__),
-                                                '_data_downloads')))
-
-SESAR_RRLYRAE_URL = 'http://www.astro.washington.edu/users/bsesar/S82_RRLyr/'
-
-
-class RRLyraeLC(object):
-    """Container for accessing RR Lyrae Light Curve data."""
-    def __init__(self, filename, dirname='table1'):
-        self.filename = filename
-        self.dirname = dirname
-        self.data = tarfile.open(filename)
-        self._metadata = None
-        self._obsdata = None
-
-    def __getstate__(self):
-        return (self.filename, self.dirname)
-
-    def __setstate__(self, args):
-        self.__init__(*args)
-
-    @property
-    def filenames(self):
-        return self.data.getnames()
-
-    @property
-    def ids(self):
-        return list(self.ids_gen)
-
-    @property
-    def ids_gen(self):
-        for f in self.filenames:
-            if '/' not in f:
-                continue
-            f = f.split('/')[1].split('.')
-            if len(f) == 1:
-                continue
-            else:
-                yield int(f[0])
-
-    def get_lightcurve(self, star_id, return_1d=False):
-        """Get the light curves for the given ID
-
-        Parameters
-        ----------
-        star_id : int
-            A valid integer star id representing an object in the dataset
-
-        Returns
-        -------
-        t, y, dy : np.ndarrays (if return_1d == False)
-            Times, magnitudes, and magnitude errors.
[…]

### 2015-01-28 multiband_LS/generated.py

@@ -1,88 +1,0 @@
-"""Tools to generate light curves"""
-
-import numpy as np
-from scipy.interpolate import interp1d
-from .data import fetch_rrlyrae_templates, fetch_rrlyrae
-
-
-class RRLyraeObject(object):
-    """Generate RR Lyrae light curves from Sesar 2010 lightcurves
-
-    Parameters
-    ----------
-    lcid : int
-        Valid light curve ID from the Sesar 2010 RR Lyrae dataset
-    random_state : int
-        Random number generator seed
-
-    Attributes
-    ----------
-    lcdata : RRLyraeLC object
-        Container for the RR Lyrae light curve dataset
-    templates : RRLyraeTemplates object
-        Container for the RR Lyrae template dataset
-    period : float
-        Period of the RR Lyrae object
-    """
-    lcdata = fetch_rrlyrae()
-    templates = fetch_rrlyrae_templates()
-
-    # Extinction corrections: Table 1 from Berry et al. (2012, ApJ, 757, 166).
-    ext_correction = {'u': 1.810,
-                      'g': 1.400,
-                      'r': 1.0,
-                      'i': 0.759,
-                      'z': 0.561 }
-    
-    @classmethod
-    def _template_func(cls, num, band, mu=0, A=1):
-        template_id = "{0:.0f}{1}".format(num, band)
-    
-        phase, amp = cls.templates.get_template(template_id)
-        phase = np.concatenate([phase, [1]])
-        amp = np.concatenate([amp, amp[-1:]])
-
-        return interp1d(phase, mu + A * amp)
-    
-    def __init__(self, lcid, random_state=None):
-        self.lcid = lcid
-        self.meta = self.lcdata.get_metadata(lcid)
-        self.obsmeta = self.lcdata.get_obsmeta(lcid)
-        self.rng = np.random.RandomState(random_state)
-        
-    @property
-    def period(self):
-        return self.meta['P']
-    
-    def observed(self, band, corrected=True):
-        if band not in 'ugriz':
-            raise ValueError("band='{0}' not recognized".format(band))
-        i = 'ugriz'.find(band)
-        t, y, dy = self.lcdata.get_lightcurve(self.lcid)
-
-        if corrected:
-            ext = self.obsmeta['rExt'] * self.ext_correction[band]
-        else:
-            ext = 0
-
-        return t[:, i], y[:, i] - ext, dy[:, i]
-
-    def generated(self, band, t, err=None, corrected=True):
[…]

### 2015-01-28 multiband_LS/lomb_scargle.py

@@ -1,242 +1,0 @@
-from __future__ import division, print_function
-
-import numpy as np
-
-from .modeler import PeriodicModeler
-
-
-class LombScargle(PeriodicModeler):
-    """Lomb-Scargle Periodogram Implementation
-
-    This is a generalized periodogram implementation using the matrix formalism
-    outlined in VanderPlas & Ivezic 2015. It allows computation of periodograms
-    and best-fit models for both the classic normalized periodogram and
-    truncated Fourier series generalizations.
-
-    Parameters
-    ----------
-    optimizer : PeriodicOptimizer instance
-        Optimizer to use to find the best period. If not specified, the
-        LinearScanOptimizer will be used.
-    center_data : boolean (default = True)
-        If True, then compute the weighted mean of the input data and subtract
-        before fitting the model.
-    fit_offset : boolean (default = True)
-        If True, then fit a floating-mean sinusoid model.
-    Nterms : int (default = 1)
-        Number of Fourier frequencies to fit in the model
-    regularization : float, vector or None (default = None)
-        If specified, then add this regularization penalty to the
-        least squares fit.
-    regularize_by_trace : boolean (default = True)
-        If True, multiply regularization by the trace of the matrix
-
-    Examples
-    --------
-    >>> rng = np.random.RandomState(0)
-    >>> t = 100 * rng.rand(100)
-    >>> dy = 0.1
-    >>> omega = 10
-    >>> y = np.sin(omega * t) + dy * rng.randn(100)
-    >>> ls = LombScargle().fit(t, y, dy)
-    >>> ls.best_period
-    0.62827393156409295
-    >>> ls.score(ls.best_period)
-    array(0.9815178000850804)
-    >>> ls.predict([0, 0.5])
-    array([-0.01213809, -0.92700951])
-
-    See Also
-    --------
-    LombScargleAstroML
-    LombScargleMultiband
-    LombScargleMultibandFast
-    """
-    def __init__(self, optimizer=None, center_data=True, fit_offset=True,
-                 Nterms=1, regularization=None, regularize_by_trace=True):
-        self.center_data = center_data
-        self.fit_offset = fit_offset
-        self.Nterms = int(Nterms)
-        self.regularization = regularization
-        self.regularize_by_trace = regularize_by_trace
-        PeriodicModeler.__init__(self, optimizer)
-
-        if not self.center_data and not self.fit_offset:
[…]

### 2015-01-28 multiband_LS/lomb_scargle_multiband.py

@@ -1,222 +1,0 @@
-from __future__ import division, print_function
-
-import numpy as np
-
-from .modeler import PeriodicModeler
-from.lomb_scargle import LombScargle
-
-
-class LombScargleMultiband(LombScargle):
-    """Multiband Periodogram Implementation
-
-    This implements the generalized multi-band periodogram described in
-    VanderPlas & Ivezic 2015.
-
-    Parameters
-    ----------
-    optimizer : PeriodicOptimizer instance
-        Optimizer to use to find the best period. If not specified, the
-        LinearScanOptimizer will be used.
-    Nterms_base : integer (default = 1)
-        number of frequency terms to use for the base model common to all bands
-    Nterms_band : integer (default = 1)
-        number of frequency terms to use for the residuals between the base
-        model and each individual band
-    reg_base : float or None (default = None)
-        amount of regularization to use on the base model parameters
-    reg_band : float or None (default = 1E-6)
-        amount of regularization to use on the band model parameters
-    regularize_by_trace : bool (default = True)
-        if True, then regularization is expressed in units of the trace of
-        the normal matrix
-    center_data : boolean (default = True)
-        if True, then center the y data prior to the fit
-
-    See Also
-    --------
-    LombScargle
-    LombScargleAstroML
-    LombScargleMultibandFast
-    """
-    def __init__(self, optimizer=None, Nterms_base=1, Nterms_band=1,
-                 reg_base=None, reg_band=1E-6, regularize_by_trace=True,
-                 center_data=True):
-        self.Nterms_base = Nterms_base
-        self.Nterms_band = Nterms_band
-        self.reg_base = reg_base
-        self.reg_band = reg_band
-        self.regularize_by_trace = regularize_by_trace
-        self.center_data = center_data
-        PeriodicModeler.__init__(self, optimizer)
-
-    def _fit(self, t, y, dy, filts):
-        self.unique_filts_ = np.unique(filts)
-        self.ymean_ = self._compute_ymean()
-
-        masks = [(filts == filt) for filt in self.unique_filts_]
-        self.ymean_by_filt_ = [LombScargle._compute_ymean(self,
-                                                          y=y[mask],
-                                                          dy=dy[mask])
-                               for mask in masks]
-
-        self.yw_ = self._construct_y(weighted=True)
[…]

### 2015-01-28 multiband_LS/modeler.py

@@ -1,126 +1,0 @@
-from __future__ import division, print_function
-
-import numpy as np
-
-from .optimizer import LinearScanOptimizer
-
-
-class PeriodicModeler(object):
-    """Base class for periodic modeling"""
-    def __init__(self, optimizer=None, *args, **kwargs):
-        if optimizer is None:
-            optimizer = LinearScanOptimizer()
-        if not hasattr(optimizer, 'best_period'):
-            raise ValueError("optimizer must be an PeriodicOptimizer instance. "
-                             "{0} not valid".format(optimizer))
-        self.optimizer = optimizer
-        self.args = args
-        self.kwargs = kwargs
-        self._best_period = None
-
-    def fit(self, t, y, dy=None, filts=None):
-        """Fit the multiterm Periodogram model to the data.
-
-        Parameters
-        ----------
-        t : array_like, one-dimensional
-            sequence of observation times
-        y : array_like, one-dimensional
-            sequence of observed values
-        dy : float or array_like (optional)
-            errors on observed values
-        filts : array_like (optional)
-            The array specifying the filter/bandpass for each observation.
-        """
-        if filts is None:
-            self.t, self.y, self.dy = np.broadcast_arrays(t, y, dy)
-            self.filts = None
-        else:
-            self.t, self.y, self.dy, self.filts = np.broadcast_arrays(t, y, dy,
-                                                                      filts)
-        self._fit(self.t, self.y, self.dy, filts=self.filts)
-        return self
-
-    def predict(self, t, filts=None, period=None):
-        """Compute the best-fit model at ``t`` for a given frequency omega
-
-        Parameters
-        ----------
-        t : float or array_like
-            times at which to predict
-        filts : array_like (optional)
-            the array specifying the filter/bandpass for each observation. This
-            is used only in multiband periodograms.
-        period : float (optional)
-            The period at which to compute the model. If not specified, it
-            will be computed via the optimizer provided at initialization.
-
-        Returns
-        -------
-        y : np.ndarray
-            predicted model values at times t
-        """
-        if period is None:
-            period = self.best_period
-
-        if filts is None:
-            t = np.asarray(t)
-            if self.filts is not None:
-                raise ValueError("filts must be passed")
-            result = self._predict(t.ravel(), filts=filts, period=period)
-        else:
-            t, filts = np.broadcast_arrays(t, filts)
-            if self.filts is None:
-                raise ValueError("filts passed to predict(), but not to fit()")
[…]

### 2015-01-28 multiband_LS/naive_multiband.py

@@ -1,101 +1,0 @@
-"""
-Naive Multiband Methods
-
-This basically amounts to a band-by-band single band approach, followed by
-some sort of majority vote among the peaks of the individual periodograms.
-"""
-from __future__ import division, print_function
-
-import numpy as np
-
-from .modeler import PeriodicModeler
-from .lomb_scargle import LombScargle
-from .utils import mode_in_range
-
-
-class NaiveMultiband(PeriodicModeler):
-    """Naive version of multiband fitting
-
-    Parameters
-    ----------
-    BaseModel : PeriodicModeler instance
-        Single-band model to use on data from each band.
-    *args, **kwargs :
-        additional arguments are passed to BaseModel on construction.
-    """
-    def __init__(self, optimizer=None, BaseModel=LombScargle,
-                 *args, **kwargs):
-        self.BaseModel = BaseModel
-        self.args = args
-        self.kwargs = kwargs
-        PeriodicModeler.__init__(self, optimizer)
-
-    def _fit(self, t, y, dy, filts):
-        t, y, dy, filts = np.broadcast_arrays(t, y, dy, filts)
-        unique_filts = np.unique(filts)
-
-        masks = [(filts == filt) for filt in unique_filts]
-        self.models_ = dict([(filt,
-                              self.BaseModel(*self.args,
-                                             **self.kwargs).fit(t[mask],
-                                                                y[mask],
-                                                                dy[mask]))
-                             for filt, mask in zip(unique_filts, masks)])
-        
-    def _predict(self, t, filts, period):
-        fset = set(np.unique(filts))
-        if not fset.issubset(self.models_.keys()):
-            raise ValueError("filts does not match training data: "
-                             "input: {0} output: {1}"
-                             "".format(set(self.model_.keys()), fset))
-        
-        result = np.zeros_like(t)
-        for filt, model in self.models_.items():
-            mask = (filts == filt)
-            result[mask] = model.predict(t[mask], period=period)
-        return result
-
-    def _score(self, periods):
-        raise NotImplementedError("score is not implmented for NaiveMultiband")
-
-    def scores(self, periods):
-        """Compute the scores under the various models
-
-        Parameters
-        ----------
-        periods : array_like
-            array of periods at which to compute scores
-
-        Returns
-        -------
-        scores : dict
-            Dictionary of scores. Dictionary keys are the unique filter names
-            passed to fit()
-        """
-        return dict([(filt, model.score(periods))
-                     for (filt, model) in self.models_.items()])
-
[…]

### 2015-01-28 multiband_LS/optimizer.py

@@ -1,111 +1,0 @@
-from __future__ import division, print_function
-
-import sys
-
-import numpy as np
-
-
-class PeriodicOptimizer(object):
-    def __init__(self, *args, **kwargs):
-        self.args = args
-        self.kwargs = kwargs
-
-    def find_best_periods(self, model, n_periods=5, return_scores=False):
-        raise NotImplementedError()
-
-    def best_period(self, model):
-        periods = self.find_best_periods(model, n_periods=1,
-                                         return_scores=False)
-        return periods[0]
-
-
-class LinearScanOptimizer(PeriodicOptimizer):
-    """Optimizer based on a linear scan of candidate frequencies.
-
-    Parameters / Attributes
-    -----------------------
-    period_range : tuple
-        (min_period, max_period) for the linear scan
-    verbose : int (default = 1)
-        Level of verbosity for the optimization process. If greater than zero,
-        information about the optimization will be printed to stdout.
-    first_pass_coverage : float (default = 5.0)
-        estimated number of points across the width of a typical peak for the
-        initial scan.
-    final_pass_coverage : float (default = 500.0)
-        estimated number of points across the width of a typical peak within
-        the final scan.
-    """
-    def __init__(self, period_range=(0.2, 1.2), verbose=1,
-                 first_pass_coverage=5, final_pass_coverage=500):
-        self.period_range = period_range
-        self.verbose = verbose
-        self.first_pass_coverage = first_pass_coverage
-        self.final_pass_coverage = final_pass_coverage
-        
-    def find_best_periods(self, model, n_periods=5, return_scores=False):
-        """Find the `n_periods` best periods in the model"""
-
-        # compute the estimated peak width from the data range
-        tmin, tmax = np.min(model.t), np.max(model.t)
-        width = 2 * np.pi / (tmax - tmin)
-
-        # our candidate steps in omega is controlled by period_range & coverage
-        omega_step = width / self.first_pass_coverage
-        omega_min = 2 * np.pi / np.max(self.period_range)
-        omega_max = 2 * np.pi / np.min(self.period_range)
-        omegas = np.arange(omega_min, omega_max + omega_step, omega_step)
-        periods = 2 * np.pi / omegas
-
-        # print some updates if desired
-        if self.verbose:
-            print("Finding optimal frequency:")
-            print(" - Using omega_step = {0:.5f}".format(omega_step))
-            print(" - Computing periods at {0:.0f} steps "
[…]

### 2015-01-28 multiband_LS/supersmoother.py

@@ -1,109 +1,0 @@
-"""
-Supersmoother code for periodic modeling
-"""
-from __future__ import print_function, division
-
-import numpy as np
-
-try:
-    import supersmoother as ssm
-except ImportError:
-    raise ImportError("Package supersmoother is required. "
-                      "Use ``pip install supersmoother`` to install")
-
-from .modeler import PeriodicModeler
-
-
-class SuperSmoother(PeriodicModeler):
-    """Periodogram based on Friedman's SuperSmoother.
-
-    Parameters
-    ----------
-    optimizer : PeriodicOptimizer instance
-        Optimizer to use to find the best period. If not specified, the
-        LinearScanOptimizer will be used.
-
-    Examples
-    --------
-    >>> rng = np.random.RandomState(0)
-    >>> t = 100 * rng.rand(100)
-    >>> dy = 0.1
-    >>> omega = 10
-    >>> y = np.sin(omega * t) + dy * rng.randn(100)
-    >>> ssm = SuperSmoother().fit(t, y, dy)
-    >>> ssm.best_period
-    0.62826749832108475
-    >>> ssm.score(ls.best_period)
-    array(0.9951882158877049)
-    >>> ssm.predict([0, 0.5])
-    array([ 0.06759746, -0.90006247])
-
-    See Also
-    --------
-    LombScargle
-    """
-    def __init__(self, optimizer=None):
-        PeriodicModeler.__init__(self, optimizer)
-
-    def _fit(self, t, y, dy, filts):
-        if filts is not None:
-            raise NotImplementedError("``filts`` keyword is not supported")
-
-        # TODO: this should actually be a weighted median, probably...
-        mu = np.sum(y / dy ** 2) / np.sum(1 / dy ** 2)
-        self.baseline_err = np.mean(abs((y - mu) / dy))
-
-    def _predict(self, t, filts, period):
-        model = ssm.SuperSmoother(period=period).fit(self.t, self.y, self.dy)
-        return model.predict(t)
-
-    def _score(self, periods):
-        return np.asarray([1 - (ssm.SuperSmoother(period=p)
-                                            .fit(self.t, self.y, self.dy)
-                                            .cv_error(skip_endpoints=False)
-                                / self.baseline_err)
-                           for p in periods])
-        
-
-class SuperSmootherMultiband(PeriodicModeler):
-    """
-    Simple multi-band SuperSmoother, with each band smoothed independently
-    
-    Parameters
-    ----------
-    optimizer : PeriodicOptimizer instance
[…]

### 2015-01-28 multiband_LS/tests/test_base_modeler.py

@@ -1,94 +1,0 @@
-import numpy as np
-from numpy.testing import assert_allclose
-
-from ..modeler import PeriodicModeler
-from .. import LombScargle, LombScargleAstroML, SuperSmoother
-from .. import LombScargleMultiband, LombScargleMultibandFast, NaiveMultiband
-
-
-class FakeModeler(PeriodicModeler):
-    """Fake periodic modeler for testing PeriodicModeler base"""
-    def _fit(self, t, y, dy, filts):
-        pass
-
-    def _predict(self, t, filts, period):
-        return np.ones(len(t))
-
-    def _score(self, periods):
-        return np.exp(-np.abs(periods - np.round(periods)))
-
-
-def test_modeler_base():
-    """Smoke-test of PeriodicModeler base class"""
-    t = np.linspace(0, 10, 100)
-    y = np.random.rand(len(t))
-    dy = 0.1
-    filts = None
-
-    model = FakeModeler()
-
-    # test setting the period range for the optimizer
-    model.optimizer.period_range = (0.8, 1.2)
-
-    # test fitting the model
-    model.fit(t, y, dy, filts)
-
-    # test the score function
-    assert_allclose(model.score(1.0), 1.0)
-
-    # test the best_period property
-    assert_allclose(model.best_period, 1, rtol=1E-4)
-
-    # test the predict function
-    assert_allclose(model.predict(t, filts), 1)
-
-
-def _generate_data(N=100, omega=10, theta=[10, 2, 3], dy=1, rseed=0):
-    """Generate some data for testing"""
-    rng = np.random.RandomState(rseed)
-    t = 20 * (2 * np.pi / omega) * rng.rand(N)
-    y = theta[0] + theta[1] * np.sin(omega * t) + theta[2] * np.cos(omega * t)
-    dy = dy * (0.5 + rng.rand(N))
-    y += dy * rng.randn(N)
-    return t, y, dy
-
-
-def test_modelers_smoketest():
-    t, y, dy = _generate_data()
-    periods = np.linspace(0.2, 1.0, 5)
-
-    def check_model(Model):
-        model = Model()
-        model.fit(t, y, dy)
-
-        # Make optimization fast
-        model.optimizer.period_range = (0.5, 0.52)
-        model.optimizer.final_pass_coverage = 0
-        model.best_period
-
-        model.score(periods)
-        model.predict(t)
-
-    for Model in (LombScargle, LombScargleAstroML, SuperSmoother):
-        yield check_model, Model
-
[…]

### 2015-01-28 multiband_LS/tests/test_lomb_scargle.py

@@ -1,113 +1,0 @@
-import numpy as np
-from numpy.testing import assert_allclose, assert_, assert_equal
-from nose import SkipTest
-
-from .. import LombScargle, LombScargleAstroML
-
-
-def _generate_data(N=100, omega=10, theta=[10, 2, 3], dy=1, rseed=0):
-    """Generate some data for testing"""
-    rng = np.random.RandomState(rseed)
-    t = 20 * (2 * np.pi / omega) * rng.rand(N)
-    y = theta[0] + theta[1] * np.sin(omega * t) + theta[2] * np.cos(omega * t)
-    dy = dy * (0.5 + rng.rand(N))
-    y += dy * rng.randn(N)
-    return t, y, dy
-
-
-def test_lomb_scargle(N=100, omega=10):
-    """Test whether the standard and generalized lomb-scargle
-    give close to the same results for non-centered data"""
-    t, y, dy = _generate_data(N, omega)
-    omegas = np.linspace(1, omega + 1, 100)
-    periods = 2 * np.pi / omegas
-
-    def check_model(Model):
-        P1 = Model(fit_offset=True).fit(t, y, dy).score(periods)
-        P2 = Model(fit_offset=False).fit(t, y, dy).score(periods)
-
-        rms = np.sqrt(np.mean((P1 - P2) ** 2))
-        assert_(rms < 0.005)
-
-    for Model in [LombScargle, LombScargleAstroML]:
-        yield check_model, Model
-
-
-def test_dy_scalar(N=100, omega=10):
-    """Test whether the standard and generalized lomb-scargle
-    give close to the same results for non-centered data"""
-    t, y, dy = _generate_data(N, omega)
-
-    # Make dy array all the same
-    dy[:] = dy.mean()
-    omegas = np.linspace(1, omega + 1, 100)
-    periods = 2 * np.pi / omegas
-
-    def check_model(Model):
-        assert_equal(Model().fit(t, y, dy).score(periods),
-                     Model().fit(t, y, dy[0]).score(periods))
-
-    for Model in [LombScargle, LombScargleAstroML]:
-        yield check_model, Model
-
-
-def test_vs_astroML(N=100, omega=10):
-    t, y, dy = _generate_data(N, omega)
-    omegas = np.linspace(omega - 4, omega + 4, 100)
-    periods = 2 * np.pi / omegas
-
-    def compare_models(model1, model2):
-        P = [model.fit(t, y, dy).score(periods)
-             for model in (model1, model2)]
[…]

### 2015-01-28 multiband_LS/tests/test_lomb_scargle_multiband.py

@@ -1,32 +1,0 @@
-import numpy as np
-from numpy.testing import assert_allclose, assert_
-from nose import SkipTest
-
-from .. import (LombScargle, LombScargleMultiband, LombScargleMultibandFast)
-
-
-def _generate_data(N=100, omega=10, theta=[10, 2, 3], dy=1, rseed=0):
-    """Generate some data for testing"""
-    rng = np.random.RandomState(rseed)
-    t = 20 * (2 * np.pi / omega) * rng.rand(N)
-    y = theta[0] + theta[1] * np.sin(omega * t) + theta[2] * np.cos(omega * t)
-    dy = dy * (0.5 + rng.rand(N))
-    y += dy * rng.randn(N)
-    return t, y, dy
-
-
-def test_lomb_scargle_multiband(N=100, omega=10):
-    """Test that results are the same with/without filter labels"""
-    t, y, dy = _generate_data(N, omega)
-    omegas = np.linspace(1, omega + 1, 100)
-    periods = 2 * np.pi / omegas
-
-    model = LombScargle(center_data=False, fit_offset=True)
-    P_singleband = model.fit(t, y, dy).score(periods)
-
-    filts = np.ones(N)
-    model_mb = LombScargleMultiband(center_data=False)
-    P_multiband = model_mb.fit(t, y, dy, filts).score(periods)
-    assert_allclose(P_multiband, P_singleband)
-    assert_allclose(model._best_params(omega),
-                    model_mb._best_params(omega)[:3])

### 2015-01-28 multiband_LS/tests/test_supersmoother.py

@@ -1,43 +1,0 @@
-import numpy as np
-from numpy.testing import assert_allclose, assert_, assert_equal
-from nose import SkipTest
-
-from .. import SuperSmoother
-
-
-def _generate_data(N=100, omega=10, theta=[10, 2, 3], dy=1, rseed=0):
-    """Generate some data for testing"""
-    rng = np.random.RandomState(rseed)
-    t = 10 * (2 * np.pi / omega) * rng.rand(N)
-    y = theta[0] + theta[1] * np.sin(omega * t) + theta[2] * np.cos(omega * t)
-    dy = dy * (0.5 + rng.rand(N))
-    y += dy * rng.randn(N)
-    return t, y, dy
-
-
-def test_supersmoother(N=100, omega=10):
-    """Test whether the standard and generalized lomb-scargle
-    give close to the same results for non-centered data"""
-    t, y, dy = _generate_data(N, omega)
-    omegas = np.linspace(1, omega + 1, 100)
-    periods = 2 * np.pi / omegas
-
-    model = SuperSmoother().fit(t, y, dy)
-    model.optimizer.period_range = (2 * np.pi / (omega + 1),
-                                    2 * np.pi / (omega - 1))
-    model.optimizer.final_pass_coverage = 0
-    assert_allclose(omega, 2 * np.pi / model.best_period, rtol=0.02)
-
-
-def test_dy_scalar(N=100, omega=10):
-    """Test whether the standard and generalized lomb-scargle
-    give close to the same results for non-centered data"""
-    t, y, dy = _generate_data(N, omega)
-
-    # Make dy array all the same
-    dy[:] = dy.mean()
-    omegas = np.linspace(1, omega + 1, 10)
-    periods = 2 * np.pi / omegas
-
-    assert_equal(SuperSmoother().fit(t, y, dy).score(periods),
-                 SuperSmoother().fit(t, y, dy[0]).score(periods))

### 2015-01-28 setup.py

@@ -1,41 +1,0 @@
-from distutils.core import setup
-
-DESCRIPTION = "Methods for periodic fitting of multiband data"
-LONG_DESCRIPTION = open('README.md').read()
-NAME = "multiband_LS"
-AUTHOR = "Jake VanderPlas"
-AUTHOR_EMAIL = "jakevdp@uw.edu"
-MAINTAINER = "Jake VanderPlas"
-MAINTAINER_EMAIL = "jakevdp@uw.edu"
-URL = 'http://github.com/jakevdp/multiband_LS'
-DOWNLOAD_URL = 'http://github.com/jakevdp/multiband_LS'
-LICENSE = 'BSD 3-clause'
-
-import supersmoother
-VERSION = supersmoother.__version__
-
-setup(name=NAME,
-      version=VERSION,
-      description=DESCRIPTION,
-      long_description=LONG_DESCRIPTION,
-      author=AUTHOR,
-      author_email=AUTHOR_EMAIL,
-      maintainer=MAINTAINER,
-      maintainer_email=MAINTAINER_EMAIL,
-      url=URL,
-      download_url=DOWNLOAD_URL,
-      license=LICENSE,
-      packages=['multiband_LS',
-                'multiband_LS.tests',
-            ],
-      classifiers=[
-        'Development Status :: 4 - Beta',
-        'Environment :: Console',
-        'Intended Audience :: Science/Research',
-        'License :: OSI Approved :: BSD License',
-        'Natural Language :: English',
-        'Programming Language :: Python :: 2.6',
-        'Programming Language :: Python :: 2.7',
-        'Programming Language :: Python :: 3.3',
-        'Programming Language :: Python :: 3.4'],
-     )

### 2015-01-28 writeup/figures/compute_results.py

@@ -5,8 +5,9 @@ import numpy as np
 import os
 from datetime import datetime
 
-from multiband_LS.data import fetch_rrlyrae
-from multiband_LS import LombScargleMultiband, SuperSmoother, SuperSmootherMultiband
+from gatspy.datasets import fetch_rrlyrae
+from gatspy.periodic import (LombScargleMultiband, SuperSmoother,
+                             SuperSmootherMultiband)
 
 
 class SuperSmoother1Band(SuperSmoother):
@@ -135,8 +136,8 @@ if __name__ == '__main__':
     client = Client()
     dview = client.direct_view()
     with dview.sync_imports():
-        from multiband_LS import (LombScargleMultiband, SuperSmoother,
-                                  SuperSmootherMultiband)
+        from gatspy.periodic import (LombScargleMultiband, SuperSmoother,
+                                     SuperSmootherMultiband)
 
     # Now time to compute the results. Here are the keywords used throughout:
     kwargs = dict(Nperiods=5, save_every=1, parallel=True, client=client)

### 2015-01-28 writeup/figures/fig_compare_periods.py

@@ -14,9 +14,9 @@ from scipy.stats import mode
 import matplotlib.pyplot as plt
 import seaborn; seaborn.set()
 
-from multiband_LS import LombScargleMultiband, SuperSmootherMultiband
+from gatspy.periodic import LombScargleMultiband, SuperSmootherMultiband
 from compute_results import SuperSmoother1Band
-from multiband_LS.data import fetch_rrlyrae
+from gatspy.datasets import fetch_rrlyrae
 
 from compute_results import get_period_results
 

### 2015-01-28 writeup/figures/fig_compare_periods_reduced.py

@@ -4,7 +4,7 @@ supersmoother, and Sesar 2010 for reduced dataset.
 """
 import matplotlib.pyplot as plt
 from fig_compare_periods import plot_periods, plot_example_lightcurve
-from multiband_LS.data import fetch_rrlyrae
+from gatspy.datasets import fetch_rrlyrae
 
 rrlyrae = fetch_rrlyrae(partial=True)
 lcid = list(rrlyrae.ids)[482]

### 2015-01-28 writeup/figures/fig_example.py

@@ -1,22 +1,0 @@
-from multiband_LS import LombScargleMultiband, SuperSmoother
-
-# Fetch the Sesar 2010 RR Lyrae data
-from multiband_LS.data import fetch_rrlyrae
-data = fetch_rrlyrae()
-t, mag, dmag, filts = data.get_lightcurve(data.ids[0],
-                                          return_1d=True)
-
-# Construct the multiband model
-model = LombScargleMultiband(Nterms_base=0, Nterms_band=1)
-model.fit(t, mag, dmag, filts)
-
-# Compute power at the following periods
-model.periodogram([0.2, 0.3, 0.4])
-
-# Construct the supersmoother model
-model = SuperSmoother()
-gband = (filts == 'g')
-model.fit(t[gband], mag[gband], dmag[gband])
-
-# Compute power at the following periods
-model.periodogram([0.2, 0.3, 0.4])

### 2015-01-28 writeup/figures/fig_multiband_sim.py

@@ -8,14 +8,14 @@ import matplotlib.pyplot as plt
 # Use seaborn settings for plot styles
 import seaborn; seaborn.set()
 
-from multiband_LS.generated import RRLyraeObject
-from multiband_LS import (LombScargleAstroML, LombScargleMultiband,
-                          NaiveMultiband)
+from gatspy.datasets import RRLyraeGenerated
+from gatspy.periodic import (LombScargleAstroML, LombScargleMultiband,
+                             NaiveMultiband)
 
 
 # Choose a Sesar 2010 object to base our fits on
 lcid = 1019544
-rrlyrae = RRLyraeObject(lcid, random_state=0)
+rrlyrae = RRLyraeGenerated(lcid, random_state=0)
 print("Extinction A_r = {0:.4f}".format(rrlyrae.obsmeta['rExt']))
 
 # Generate data in a 6-month observing season

### 2015-01-28 writeup/figures/fig_multiband_supersmoother.py

@@ -13,13 +13,13 @@ import matplotlib.pyplot as plt
 # Use seaborn settings for plot styles
 import seaborn; seaborn.set()
 
-from multiband_LS.generated import RRLyraeObject
-from multiband_LS import SuperSmoother
+from gatspy.datasets import RRLyraeGenerated
+from gatspy.periodic import SuperSmoother
 
 
 # Choose a Sesar 2010 object to base our fits on
 lcid = 1019544
-rrlyrae = RRLyraeObject(lcid, random_state=0)
+rrlyrae = RRLyraeGenerated(lcid, random_state=0)
 print("Extinction A_r = {0:.4f}".format(rrlyrae.obsmeta['rExt']))
 
 # Generate data in a 6-month observing season
@@ -39,7 +39,7 @@ periods = np.linspace(0.2, 0.9, 1000)
 
 #----------------------------------------------------------------------
 # First figure:
-# Compute the lomb-scargle periodogram in each band
+# Compute the supersmoother periodogram in each band
 
 P = [SuperSmoother().fit(t, m, dy).periodogram(periods) for m in mags]
 

### 2015-01-28 writeup/figures/fig_multiterm_example.py

@@ -13,13 +13,13 @@ import matplotlib.pyplot as plt
 # Use seaborn settings for plot styles
 import seaborn; seaborn.set()
 
-from multiband_LS.generated import RRLyraeObject
-from multiband_LS.lomb_scargle import LombScargle
+from gatspy.datasets import RRLyraeGenerated
+from gatspy.periodic import LombScargle
 
 
 # Choose a Sesar 2010 object to base our fits on
 lcid = 1019544
-rrlyrae = RRLyraeObject(lcid, random_state=0)
+rrlyrae = RRLyraeGenerated(lcid, random_state=0)
 
 # Generate data in a 6-month observing season
 Nobs = 60

### 2015-01-28 writeup/figures/fig_num_observations.py

@@ -2,7 +2,7 @@ import numpy as np
 import seaborn as sns
 import matplotlib.pyplot as plt
 
-from multiband_LS.data import fetch_rrlyrae
+from gatspy.datasets import fetch_rrlyrae
 
 
 rrlyrae = fetch_rrlyrae()

### 2015-01-28 writeup/figures/fig_regularization_example.py

@@ -12,13 +12,13 @@ import matplotlib.pyplot as plt
 # Use seaborn settings for plot styles
 import seaborn; seaborn.set()
 
-from multiband_LS.generated import RRLyraeObject
-from multiband_LS.lomb_scargle import LombScargle
+from gatspy.datasets import RRLyraeGenerated
+from gatspy.periodic import LombScargle
 
 
 # Choose a Sesar 2010 object to base our fits on
 lcid = 1019544
-rrlyrae = RRLyraeObject(lcid, random_state=0)
+rrlyrae = RRLyraeGenerated(lcid, random_state=0)
 
 # Generate data in a 6-month observing season
 Nobs = 60

### 2015-01-28 writeup/figures/compute_results.py
update for API change in gatspy

@@ -62,7 +62,7 @@ def compute_and_save_periods(rrlyrae, Model, outfile,
     def find_periods(lcid, Nperiods=Nperiods,
                      rrlyrae=rrlyrae, Model=Model,
                      model_args=model_args, model_kwds=model_kwds):
-        t, y, dy, filts = rrlyrae.get_lightcurve(lcid, return_1d=True)
+        t, y, dy, filts = rrlyrae.get_lightcurve(lcid)
         model = Model(*(model_args or ()), **(model_kwds or {}))
         model.optimizer.period_range = (0.2, 1.2)
         model.optimizer.verbose = 0

### 2015-01-28 writeup/figures/fig_compare_periods.py

@@ -89,7 +89,7 @@ def plot_example_lightcurve(rrlyrae, lcid):
           fig.add_subplot(gs[0, 1]),
           fig.add_subplot(gs[1, 1])]
 
-    t, y, dy, filts = rrlyrae.get_lightcurve(lcid, True)
+    t, y, dy, filts = rrlyrae.get_lightcurve(lcid)
 
     # don't plot data with huge errorbars
     mask = (dy < 1)

### 2015-01-28 writeup/figures/fig_num_observations.py

@@ -8,9 +8,9 @@ from gatspy.datasets import fetch_rrlyrae
 rrlyrae = fetch_rrlyrae()
 rrlyrae_partial = fetch_rrlyrae(partial=True)
 
-data = [rrlyrae.get_lightcurve(lcid, True)
+data = [rrlyrae.get_lightcurve(lcid)
         for lcid in rrlyrae.ids]
-data_partial = [rrlyrae_partial.get_lightcurve(lcid, True)
+data_partial = [rrlyrae_partial.get_lightcurve(lcid)
                 for lcid in rrlyrae.ids]
 
 tminmax = np.array([[t.min(), t.max()] for t,_,_,_ in data])

### 2015-01-28 writeup/paper.tex
update paper: multiband_LS -> gatspy

@@
 \appendix
-\section{{\tt multiband\_LS}: Python Implementation of Multiband Periodogram}
-\sectlabel{multiband_LS}
-The algorithm outlined in this paper is available as an open-source Python package\footnote{\url{http://github.com/jakevdp/multiband_LS/}}, along with code to download the data and recreate all the figures in this paper.
-{\tt multiband\_LS} is a pure-Python package written to be compatible with both Python 2 and Python 3, and performs fast numerical computation through dependencies on {\tt numpy} \citep{numpy}\footnote{\url{http://www.numpy.org}} and {\tt astroML} \citep{astroML}\footnote{\url{http://www.astroml.org}}, which offer optimized implementatinos of numerical methods in Python.
+\section{Python Implementation of Multiband Periodogram}
+\sectlabel{gatspy}
+The algorithm outlined in this paper is available in {\tt gatspy}, an open-source Python package for general astronomical time-series analysis\footnote{\url{http://github.com/jakevdp/gatspy/}}.
+Along with the periodogram implementation, it also contains code to download all the data used in this work.
+Code to reproduce this paper, including all figures, is available as well\footnote{\url{http://github.com/jakevdp/multiband_LS/}}.
+
+{\tt gatspy} is a pure-Python package written to be compatible with both Python 2 and Python 3, and performs fast numerical computation through dependencies on {\tt numpy} \citep{numpy}\footnote{\url{http://www.numpy.org}} and {\tt astroML} \citep{astroML}\footnote{\url{http://www.astroml.org}}, which offer optimized implementatinos of numerical methods in Python.
 
@@
 \begin{lstlisting}
-from multiband_LS import LombScargleMultiband
+from gatspy.periodic import LombScargleMultiband
 import numpy as np
@@
 # Fetch the Sesar 2010 RR Lyrae data
-from multiband_LS.data import fetch_rrlyrae
+from gatspy.datasets import fetch_rrlyrae
 data = fetch_rrlyrae()
-t, mag, dmag, filts = data.get_lightcurve(data.ids[0],
-                                          return_1d=True)
+t, mag, dmag, filts = data.get_lightcurve(data.ids[0])
 
@@
 \begin{lstlisting}
-from multiband_LS import SuperSmoother
+from gatspy.periodic import SuperSmoother
 
@@
 
-# Compute power at the following periods
+# Compute power at the given periods
 power = model.periodogram(periods)
@@
 
-The models in the \texttt{multiband\_LS} package contain many more methods, and much more functionality that what is shown here.
-For updates, more examples, and more information, visit \url{http://github.com/jakevdp/multiband_LS}.
+The models in the \texttt{gatspy} package contain many more methods, and much more functionality that what is shown here.
+For updates, more examples, and more information, visit \url{http://github.com/jakevdp/gatspy/}.
 

### 2015-01-28 writeup/paper.tex
minor edits in Conclusions

@@
 We have motivated and derived a multiband version of the classic Lomb-Scargle method for detecting periodicity in astronomical time-series.
- Experiments on several hundred RR Lyrae stars from the Stripe 82 dataset indicate that this method outperforms methods used previously in the literature, especially for sparsely-sampled light curves with only single bands observed each night.
+ Experiments on several hundred RR Lyrae stars from the SDSS Stripe 82 dataset indicate that this method outperforms methods used previously in the literature, especially for sparsely-sampled light curves with only single bands observed each night.
 
@@
 This could potentially be accounted for by estimating and filtering the effect of the window function; one potential method for this involves estimating the deconvolution of the window power and the observed power.
-It may also be possible to propose a multiband extension of CARMA or another forward-modeling approaches to detecting periodicity.
+It may also be possible to propose a multiband extension of CARMA \citep{Kelly14},
+or another forward-modeling approaches to detecting periodicity.
 
 Another potentially fruitful way forward in the multiband case is to make use of physically-motivated priors in the light-curve shape.
-For example, the paths of RR Lyrae stars through color-color and color-magnitude space are constrained by known astrophysical processes in the structure of the stars, and making use of such a physically-motivated prior may help break degeneracies in period determination.
+For example, the paths of RR Lyrae stars through color-color and color-magnitude space are constrained by known astrophysical processes in the structure of the stars \citep[e.g., see Fig.
+5 in][]{Szabo2014}, and making use of such a physically-motivated prior may help break degeneracies in period determination.
 
@@
 The multiband method, through its combination of flexibility and model simplicity, meets the first: as shown above, it performs well in the case of sparsely-sampled multiband data.
-It also meets the second constraint: it consists of a fast linear model at each frequency, compared to a sliding-window model in the case of supersmoother, a nonlinear optimization in the case of template-fitting, and a markov chain monte carlo in the case of CARMA models.
+It also meets the second constraint: it consists of a fast linear model at each frequency, compared to a sliding-window model in the case of supersmoother, a nonlinear optimization in the case of template-fitting, and a Markov Chain Monte Carlo in the case of CARMA models.
 In our own benchmarks, we found the multiband method to be several times faster than the single-band supersmoother approach, and several orders of magnitude faster than the template fitting approach.
@@
 Sloan Foundation, the Gordon and Betty Moore Foundation, and the Washington Research Foundation.
-\todo{ZI is supported by...}  The authors thank GitHub for providing free academic accounts which were essential in the development of this work.
+The authors thank GitHub for providing free academic accounts which were essential in the development of this work.
 
@@
 
-{\tt gatspy} is a pure-Python package written to be compatible with both Python 2 and Python 3, and performs fast numerical computation through dependencies on {\tt numpy} \citep{numpy}\footnote{\url{http://www.numpy.org}} and {\tt astroML} \citep{astroML}\footnote{\url{http://www.astroml.org}}, which offer optimized implementatinos of numerical methods in Python.
+{\tt gatspy} is a pure-Python package written to be compatible with both Python 2 and Python 3, and performs fast numerical computation through dependencies on {\tt numpy} \citep{numpy}\footnote{\url{http://www.numpy.org}} and {\tt astroML} \citep{astroML}\footnote{\url{http://www.astroml.org}}, which offer optimized implementations of numerical methods in Python.
 

### 2015-01-28 writeup/paper.tex
minor edits in Conclusions

@@
 In our own benchmarks, we found the multiband method to be several times faster than the single-band supersmoother approach, and several orders of magnitude faster than the template fitting approach.
-This suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys.
+This suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys
+and we make our tools publicly available on GitHub.
 

### 2015-01-30 LSST_sims/compute_results.py
update LSST sim code

@@ -2,8 +2,33 @@ import os
 from datetime import datetime
 
 import numpy as np
+import numpy
 
 from LSSTsims import LSSTsims
+from gatspy.periodic import (LombScargleMultiband, SuperSmoother,
+                             SuperSmootherMultiband, LombScargleMultibandFast)
+
+
+class SuperSmoother1Band(SuperSmootherMultiband):
+    """
+    Convenience class to fit a single band of data with supersmoother
+
+    This class ignores all data not associated with the given band.
+
+    The main reason for this is that it can then be used as a stand-in for
+    any multiband class.
+    """
+    def __init__(self, optimizer=None, band='g'):
+        self.band = band
+        SuperSmootherMultiband.__init__(self, optimizer)
+
+    def _fit(self, t, y, dy, filts):
+        mask = (filts == self.band)
+        self.t, self.y, self.dy, self.filts = (t[mask], y[mask],
+                                               dy[mask], filts[mask])
+        self.unique_filts_ = numpy.unique(self.filts)
+        return SuperSmootherMultiband._fit(self, self.t, self.y,
+                                           self.dy, self.filts)
 
 
 def find_res(results, **kwargs):
@@ -55,13 +80,21 @@ def compute_and_save_periods(Model, outfile,
     # Define a function which, given a key, computes the desired periods.
     def find_periods(key, Nperiods=Nperiods, Model=Model, LSSTsims=LSSTsims,
                      model_args=model_args, model_kwds=model_kwds):
+        import numpy as np
         lsstsim = LSSTsims()
         t, y, dy, filts = lsstsim.generate_lc(*key, random_state=0)
+
         model = Model(*(model_args or ()), **(model_kwds or {}))
         model.optimizer.period_range = (0.2, 1.2)
         model.optimizer.verbose = 0
         model.fit(t, y, dy, filts)
-        return key + (model.find_best_periods(Nperiods),)
+        try:
+            periods = model.find_best_periods(Nperiods)
+        except np.linalg.LinAlgError:
+            periods = np.nan + np.zeros(Nperiods)
+        except ValueError:
+            periods = np.nan + np.zeros(Nperiods)
+        return key + (periods,)
 
     # Set up the iterator over results
     results = np.zeros(len(keys), dtype=dtype)
@@ -100,24 +133,60 @@ def gather_results(outfile, pointing_indices, ndays, rmags, template_indices):
     results = np.array([find_res(results, **dict(zip(dtype.names, key)))
[…]

### 2015-01-30 LSST_sims/plot_results.py

@@ -4,33 +4,56 @@ import matplotlib.pyplot as plt
 from compute_results import gather_results
 from gatspy.datasets import fetch_rrlyrae
 
-rng = np.random.RandomState(0)
-template_indices = rng.randint(0, 483, 5)
+def plot_results(outfile, pointing_indices, ndays, rmags, template_indices):
+    results = gather_results(outfile,
+                             pointing_indices=pointing_indices,
+                             ndays=ndays,
+                             rmags=rmags,
+                             template_indices=template_indices)
+
+    rrlyrae = fetch_rrlyrae()
+    periods = np.reshape([rrlyrae.get_metadata(rrlyrae.ids[i])['P']
+                          for i in template_indices.ravel()],
+                         template_indices.shape)
+
+    nans = np.isnan(results)
+    results[nans] = 0
+
+    matches = np.any(abs(results - periods[:, :, None]) < 0.01, -1)
+    #matches[np.any(nans, -1)] = np.nan
+    matches = matches.reshape(matches.shape[:2] + (-1,))
+
+    frac = np.nanmean(matches, -1)
+
+    rmags = rmags.ravel()
+    ndays = ndays.ravel()
+
+    for i in range(len(rmags)):
+        plt.plot(ndays / 365., frac[i], label='r = {0:.2f}'.format(rmags[i]))
+    plt.grid(True)
+    plt.legend(loc='lower right')
+
+
+
+
+template_indices = np.arange(5 * 23).reshape(5, 23).T
 pointing_indices = np.arange(1, 24)[:, None]
-ndays = np.array([365, 3 * 365, 5 * 365, 10 * 365])[:, None, None]
+ndays = np.array([90, 180, 365, 2*365])[:, None, None]
 rmags = np.array([20, 22, 24.5])[:, None, None, None]
 
-results = gather_results('results.npy',
-                         pointing_indices=pointing_indices,
-                         ndays=ndays,
-                         rmags=rmags,
-                         template_indices=template_indices)
+template_indices = template_indices[:, :1]
 
-rrlyrae = fetch_rrlyrae()
-periods = np.asarray([rrlyrae.get_metadata(rrlyrae.ids[0])['P']
-                      for i in template_indices.ravel()])
+plot_results('resultsLSST.npy',
+             pointing_indices=pointing_indices,
+             ndays=ndays,
+             rmags=rmags,
[…]

### 2015-01-30 writeup/paper.tex
Add LCB14 discussion

@@
 \begin{eqnarray}
-  y_k(t|\omega,\theta) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\\ 
+  \eqlabel{multiband_model}
+  y_k(t|\omega,\theta) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\nonumber\\ 
   &\theta^{(k)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(k)}_{2n - 1}\sin(n\omega t) + \theta^{(k)}_{2n}\cos(n\omega t)\right].&
@@
 
+\subsection{Comparison with \citet{Long2014}}
+As we were finishing this study, it was pointed out to us that another group had released a preprint independently addressing the multiband periodogram case \citep[][hereafter LCB14]{Long2014}, and come up with a solutions very similar to the ones presented here.
+They present two methods, the ``Multiband Generalized Lomb-Scargle'' (MGLS) which is effectively identical to the $(1, 0)$ multi-phase model, and the ``Penalized Generalized Lomb-Scargle'' (PGLS), which is similar in spirit to our $(0, 1)$ shared-phase model.
+
+In the PGLS model, they start with a multi-phase model, fitting independent $N=1$ term fits to each band, and apply a nonlinear regularization term which penalizes differences in the amplitude and phase.
+In terms of the formalism used in this work, the PGLS model minimizes a regularized $\chi^2$ of the form
+\begin{equation}
+  \chi^2_{PGLS} = \sum_{k=1}^K \bigg[~\chi^2_{GLS}(D^{(K)}) + J_A(A^{(k)}) + J_\phi(\phi^{(k)})~\bigg].
+\end{equation}
+for $K$ bands, where $\chi^2_{GLS}(D^{(K)})$ is the $\chi^2$ of the standard floating mean model on the single-band data $D^{(K)}$, and $J_A$ and $J_\phi$ are regularization/penalty terms which are a function of the amplitude $A^{k}$ and phase $\phi^{(k)}$ of each model.
+In terms of our linear model parameters $\theta^{(k)}$, this amplitude and phase can be expressed:
+\begin{eqnarray}
+ A^{(k)} &=& \sqrt{(\theta_1^{(k)})^2 + (\theta_1^{(k)})^2}\nonumber\\
+ \phi^{(k)} &=& \arctan(\theta_2 / \theta_1)
+\end{eqnarray}
+The selected form of these regularization terms penalizes deviations of the amplitude and phase from a common mean between the bands; in this sense the PGLS model can be considered a conceptual mid-point between our shared-phase and multi-phase models.
+Within the formalism proposed here, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in the shared-phase model, though the nature of the resulting regularization will differ.
+
+Computationally, the PGLS model requires a nonlinear optimization at each frequency $\omega$, and is thus much more expensive than the straightforward linear optimization of our shared-phase model.
+For this reason, LCB14 propose a clever method by which nested models are used to reduce the number of nonlinear optimizations used: essentially, by showing that the (linear) MGLS $\chi^2$ is a lower-bound of the (non-linear) PGLS $\chi^2$, it is possible to iteratively reduce the number of PGLS computations required to minimize the $\chi^2$ among a grid of frequencies.
+Such an optimization could also be applied in the case of our shared-phase model, but is not necessary here due to its relative speed.
+When applying the method to a very large number of light curves in, e.g. LSST however, such a computational trick may prove very useful.
+
+Given these important distinctions between the models proposed here and in LCB14, in future work we plan to do a detailed comparison of the two means of model regularization.
+
 {\it Acknowledgements:} JTV is supported by the University of Washington eScience institute, including grants from the Alfred P.

### 2015-01-30 LSST_sims/compute_results.py
last of LSST sim code

@@ -138,19 +138,25 @@ def gather_results(outfile, pointing_indices, ndays, rmags, template_indices):
 if __name__ == '__main__':
     from gatspy.periodic import LombScargleMultiband
 
-    # Need some imports on the engine
-    from IPython.parallel import Client
-    client = Client()
-    dview = client.direct_view()
-    with dview.sync_imports():
-        import numpy
-        from gatspy.periodic import (LombScargleMultiband, SuperSmoother,
-                                     SuperSmootherMultiband)
+    parallel = False
+
+    if parallel:
+        # Need some imports on the engine
+        from IPython.parallel import Client
+        client = Client()
+        dview = client.direct_view()
+        with dview.sync_imports():
+            import numpy
+            from gatspy.periodic import (LombScargleMultiband, SuperSmoother,
+                                         SuperSmootherMultiband)
+    else:
+        client = None
 
     template_indices = np.arange(5 * 23).reshape(5, 23).T
     pointing_indices = np.arange(1, 24)[:, None]
     ndays = np.array([90, 180, 365, 2*365])[:, None, None]
     rmags = np.array([20, 22, 24.5])[:, None, None, None]
+
 
     compute_and_save_periods(LombScargleMultiband, 'resultsLSST.npy',
                              model_kwds=dict(Nterms_base=1, Nterms_band=0),
@@ -158,7 +164,7 @@ if __name__ == '__main__':
                              ndays=ndays,
                              rmags=rmags,
                              template_indices=template_indices,
-                             parallel=True, client=client,
+                             parallel=parallel, client=client,
                              save_every=1)
 
     template_indices = template_indices[:, :1]
@@ -169,7 +175,7 @@ if __name__ == '__main__':
                              ndays=ndays,
                              rmags=rmags,
                              template_indices=template_indices,
-                             parallel=True, client=client,
+                             parallel=parallel, client=client,
                              save_every=1)
 
     for i, band in enumerate('ugriz'):
@@ -180,7 +186,7 @@ if __name__ == '__main__':
                                  ndays=ndays,
                                  rmags=rmags,
                                  template_indices=template_indices,
-                                 parallel=True, client=client,
+                                 parallel=parallel, client=client,
                                  save_every=1)
 
     compute_and_save_periods(SuperSmootherMultiband, 'resultsLSST_ssmulti.npy',
@@ -188,5 +194,5 @@ if __name__ == '__main__':
                              ndays=ndays,
                              rmags=rmags,
                              template_indices=template_indices,
[…]

### 2015-01-30 LSST_sims/plot_results.py

@@ -4,6 +4,7 @@ import matplotlib.pyplot as plt
 from compute_results import gather_results
 from gatspy.datasets import fetch_rrlyrae
 
+
 def plot_results(outfile, pointing_indices, ndays, rmags, template_indices):
     results = gather_results(outfile,
                              pointing_indices=pointing_indices,
@@ -29,9 +30,40 @@ def plot_results(outfile, pointing_indices, ndays, rmags, template_indices):
     ndays = ndays.ravel()
 
     for i in range(len(rmags)):
-        plt.plot(ndays / 365., frac[i], label='r = {0:.2f}'.format(rmags[i]))
+        plt.plot(ndays / 365, frac[i], label='r = {0:.2f}'.format(rmags[i]))
     plt.grid(True)
     plt.legend(loc='lower right')
+
+
+def plot_results_multi(outfiles, pointing_indices, ndays, rmags, template_indices):
+    results = [gather_results(outfile,
+                              pointing_indices=pointing_indices,
+                              ndays=ndays,
+                              rmags=rmags,
+                              template_indices=template_indices)[:, :, :, :, :1]
+               for outfile in outfiles[:4]]
+    results = np.concatenate(results, axis=-1)
+
+    rrlyrae = fetch_rrlyrae()
+    periods = np.reshape([rrlyrae.get_metadata(rrlyrae.ids[i])['P']
+                          for i in template_indices.ravel()],
+                         template_indices.shape)
+
+    nans = np.isnan(results)
+    results[nans] = 0
+
+    matches = np.any(abs(results - periods[:, :, None]) < 0.01, -1)
+    #matches[np.any(nans, -1)] = np.nan
+    matches = matches.reshape(matches.shape[:2] + (-1,))
+
+    frac = np.nanmean(matches, -1)
+
+    rmags = rmags.ravel()
+    ndays = ndays.ravel()
+
+    for i in range(len(rmags)):
+        plt.plot(ndays / 365, frac[i], label='r = {0:.2f}'.format(rmags[i]))
+    plt.grid(True)
 
 
 
@@ -41,19 +73,24 @@ pointing_indices = np.arange(1, 24)[:, None]
 ndays = np.array([90, 180, 365, 2*365])[:, None, None]
 rmags = np.array([20, 22, 24.5])[:, None, None, None]
 
-template_indices = template_indices[:, :1]
-
 plot_results('resultsLSST.npy',
              pointing_indices=pointing_indices,
              ndays=ndays,
              rmags=rmags,
[…]

### 2015-01-30 writeup/paper.tex
fixed two typos in sec 7.1

@@
 \subsection{Comparison with \citet{Long2014}}
-As we were finishing this study, it was pointed out to us that another group had released a preprint independently addressing the multiband periodogram case \citep[][hereafter LCB14]{Long2014}, and come up with a solutions very similar to the ones presented here.
-They present two methods, the ``Multiband Generalized Lomb-Scargle'' (MGLS) which is effectively identical to the $(1, 0)$ multi-phase model, and the ``Penalized Generalized Lomb-Scargle'' (PGLS), which is similar in spirit to our $(0, 1)$ shared-phase model.
+As we were finishing this study, it was pointed out to us that another group had released a preprint independently addressing the multiband periodogram case \citep[][hereafter LCB14]{Long2014}, and come up with solutions very similar to the ones presented here.
+They present two methods, the ``Multiband Generalized Lomb-Scargle'' (MGLS) which is effectively identical to the $(1, 0)$ multi-phase model
+here, and the ``Penalized Generalized Lomb-Scargle'' (PGLS), which is similar in spirit to our $(0, 1)$ shared-phase model.
 

### 2015-01-30 writeup/paper.tex
fixed two typos in sec 7.1

@@
 The selected form of these regularization terms penalizes deviations of the amplitude and phase from a common mean between the bands; in this sense the PGLS model can be considered a conceptual mid-point between our shared-phase and multi-phase models.
-Within the formalism proposed here, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in the shared-phase model, though the nature of the resulting regularization will differ.
+Within the formalism proposed here, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in our shared-phase model, though the nature of the resulting regularization will differ.
 

### 2015-01-30 writeup/paper.tex
fixed two typos in sec 7.1

@@
 For this reason, LCB14 propose a clever method by which nested models are used to reduce the number of nonlinear optimizations used: essentially, by showing that the (linear) MGLS $\chi^2$ is a lower-bound of the (non-linear) PGLS $\chi^2$, it is possible to iteratively reduce the number of PGLS computations required to minimize the $\chi^2$ among a grid of frequencies.
-Such an optimization could also be applied in the case of our shared-phase model, but is not necessary here due to its relative speed.
-When applying the method to a very large number of light curves in, e.g. LSST however, such a computational trick may prove very useful.
+Such an optimization could also be applied in the case of our shared-phase model, but is not necessary here due to its already high speed.
+However, when applying the method to a very large number of light curves, as in e.g. LSST, such a computational trick may prove very useful.
 

### 2015-01-30 LSST_sims/LSSTsims.py
fix LSST sims

@@ -20,7 +20,7 @@ class LSSTsims(object):
         return [np.loadtxt(files.extractfile(member), dtype=dtype)
                 for member in files]
 
-    def generate_lc(self, pointing_index, n_days, rmag, S82index,
+    def generate_lc(self, pointing_index, n_days, gmag, S82index,
                     random_state=None):
         gen = RRLyraeGenerated(self.S82data.ids[S82index],
                                random_state=random_state)
@@ -30,20 +30,18 @@ class LSSTsims(object):
         t = pointing['mjd']
         filts = pointing['filter']
         m5 = pointing['m5']
-
-        # HACK: assume y-band and z-band are the same,
-        # as we don't have y-band RR-Lyrae templates
-        filts[filts == 5] = 4
         
         # generate magnitudes; no errors
         mag = np.zeros_like(t)
-        for i, filt in enumerate('ugriz'):
+        for i, filt in enumerate('ugrizy'):
             mask = (filts == i)
+            # HACK: for y-band, use z-band template
+            if filt == 'y': filt = 'z'
             mag[mask] = gen.generated(filt, t[mask])
 
         # adjust mags to desired r-band mean
-        rmag_mean = mag[filts == 2].mean()
-        mag += (rmag - rmag_mean)
+        gmag_mean = mag[filts == 1].mean()
+        mag += (gmag - gmag_mean)
 
         # compute magnitude error from m5 (eq 5 of Ivezic 2014 LSST paper)
         gamma = np.array([0.037, 0.038, 0.039, 0.039, 0.040, 0.040])
@@ -57,17 +55,15 @@ class LSSTsims(object):
 
 if __name__ == '__main__':
     F = LSSTsims()
-    for p in F.pointings:
-        print(len(p))
-    exit()
-    t, mag, dmag, filts = F.generate_lc(0, n_days=3650, rmag=24.0)
+    t, mag, dmag, filts = F.generate_lc(0, n_days=3650, gmag=24.0, S82index=0)
     print(len(t))
 
     import matplotlib.pyplot as plt
     import seaborn; seaborn.set()
 
     # Note: each night has multiple obs.
-    for f in filts:
+    for f in np.unique(filts):
         mask = (filts == f)
-        plt.errorbar(t[mask], mag[mask], dmag[mask], fmt='.')
+        plt.errorbar(t[mask], mag[mask], dmag[mask], fmt='.', label='ugrizy'[f])
+    plt.legend()
     plt.show()

### 2015-01-30 LSST_sims/cachedb.py

@@ -1,0 +1,208 @@
+import pickle
+import os
+import numpy as np
+
+from numpy.testing import assert_equal, assert_allclose
+
+
+class CacheDB(object):
+    def __init__(self, filename):
+        raise NotImplementedError()
+
+    def key_exists(self, key):
+        raise NotImplementedError()
+
+    def add_row(self, key, val, save=False):
+        raise NotImplementedError()
+
+    def get_row(self, key):
+        raise NotImplementedError()
+
+    def save(self):
+        raise NotImplementedError()
+
+    def load(self):
+        raise NotImplementedError()
+
+    def compute(self, func, key, *args, **kwargs):
+        if self.key_exists(key):
+            return self.get_row(key)
+        else:
+            return self.add_row(key, func(key, *args, **kwargs))
+
+    def compute_iter(self, func, keys, *args, **kwargs):
+        return [self.compute(func, key, *args, **kwargs)
+                for key in keys]
+
+
+class PickleCache(CacheDB):
+    """Pickle-backed Persistency Helper"""
+    def __init__(self, filename):
+        self.filename = filename
+        self.db = self.load() if os.path.exists(filename) else {}
+
+    def key_exists(self, key):
+        return key in self.db
+
+    def add_row(self, key, val, save=False):
+        try:
+            self.db[key] = val
+            return val
+        finally:
+            if save:
+                self.save()
+
+    def get_row(self, key):
+        return self.db[key]
+
+    def save(self):
+        with open(self.filename, 'wb') as f:
+            pickle.dump(self.db, f)
+
+    def load(self):
+        with open(self.filename, 'rb') as f:
+            return pickle.load(f)
+
+
+class NumpyCache(PickleCache):
+    """Numpy Binary-backed Persistency Helper"""
+    def __init__(self, filename):
+        if not filename.endswith('.npy'):
+            filename += '.npy'
+        PickleCache.__init__(self, filename)
+
+    def save(self):
+        np.save(self.filename, self.dict_to_array(self.db))
+
+    def load(self):
+        return self.array_to_dict(np.load(self.filename))
+
+    @staticmethod
+    def array_to_dict(arr):
+        keys = arr['key']
+        vals = arr['val']
+
+        if keys.dtype.names is not None:
+            names = keys.dtype.names
[…]

### 2015-01-30 LSST_sims/plot_results.py

@@ -41,7 +41,7 @@ def plot_results_multi(outfiles, pointing_indices, ndays, rmags, template_indice
                               ndays=ndays,
                               rmags=rmags,
                               template_indices=template_indices)[:, :, :, :, :1]
-               for outfile in outfiles[:4]]
+               for outfile in outfiles]
     results = np.concatenate(results, axis=-1)
 
     rrlyrae = fetch_rrlyrae()
@@ -82,7 +82,7 @@ plot_results('resultsLSST.npy',
 template_indices = template_indices[:, :1]
 
 plot_results_multi(['resultsLSST_ssm_{0}.npy'.format(band)
-                    for band in 'ugriz'],
+                    for band in 'g'],
                    pointing_indices=pointing_indices,
                    ndays=ndays,
                    rmags=rmags,

### 2015-01-31 LSST_sims/cachedb.py
mapcache

@@ -1,208 +1,0 @@
-import pickle
-import os
-import numpy as np
-
-from numpy.testing import assert_equal, assert_allclose
-
-
-class CacheDB(object):
-    def __init__(self, filename):
-        raise NotImplementedError()
-
-    def key_exists(self, key):
-        raise NotImplementedError()
-
-    def add_row(self, key, val, save=False):
-        raise NotImplementedError()
-
-    def get_row(self, key):
-        raise NotImplementedError()
-
-    def save(self):
-        raise NotImplementedError()
-
-    def load(self):
-        raise NotImplementedError()
-
-    def compute(self, func, key, *args, **kwargs):
-        if self.key_exists(key):
-            return self.get_row(key)
-        else:
-            return self.add_row(key, func(key, *args, **kwargs))
-
-    def compute_iter(self, func, keys, *args, **kwargs):
-        return [self.compute(func, key, *args, **kwargs)
-                for key in keys]
-
-
-class PickleCache(CacheDB):
-    """Pickle-backed Persistency Helper"""
-    def __init__(self, filename):
-        self.filename = filename
-        self.db = self.load() if os.path.exists(filename) else {}
-
-    def key_exists(self, key):
-        return key in self.db
-
-    def add_row(self, key, val, save=False):
-        try:
-            self.db[key] = val
-            return val
-        finally:
-            if save:
-                self.save()
-
-    def get_row(self, key):
-        return self.db[key]
-
-    def save(self):
-        with open(self.filename, 'wb') as f:
-            pickle.dump(self.db, f)
-
-    def load(self):
-        with open(self.filename, 'rb') as f:
-            return pickle.load(f)
-
-
-class NumpyCache(PickleCache):
-    """Numpy Binary-backed Persistency Helper"""
-    def __init__(self, filename):
-        if not filename.endswith('.npy'):
-            filename += '.npy'
-        PickleCache.__init__(self, filename)
-
-    def save(self):
-        np.save(self.filename, self.dict_to_array(self.db))
-
-    def load(self):
-        return self.array_to_dict(np.load(self.filename))
-
-    @staticmethod
-    def array_to_dict(arr):
-        keys = arr['key']
-        vals = arr['val']
-
-        if keys.dtype.names is not None:
-            names = keys.dtype.names
[…]

### 2015-01-31 LSST_sims/mapcache.py

@@ -1,0 +1,204 @@
+"""
+mapcache.py: a simple mapped object persistency model, with various backends.
+"""
+
+import pickle
+import os
+import numpy as np
+
+
+class MapCache(object):
+    """Base Class for mapped object persistency"""
+    def __init__(self, filename):
+        raise NotImplementedError()
+
+    def key_exists(self, key):
+        raise NotImplementedError()
+
+    def add_row(self, key, val, save=True):
+        raise NotImplementedError()
+
+    def get_row(self, key):
+        raise NotImplementedError()
+
+    def save(self):
+        raise NotImplementedError()
+
+    def load(self):
+        raise NotImplementedError()
+
+    def compute(self, func, key, *args, **kwargs):
+        if self.key_exists(key):
+            return self.get_row(key)
+        else:
+            return self.add_row(key, func(key, *args, **kwargs))
+
+    def compute_iter(self, func, keys, *args, **kwargs):
+        return [self.compute(func, key, *args, **kwargs)
+                for key in keys]
+
+
+class PickleCache(MapCache):
+    """Pickle-backed Persistency Helper"""
+    def __init__(self, filename):
+        self.filename = filename
+        self.dct = self.load() if os.path.exists(filename) else {}
+
+    def key_exists(self, key):
+        return key in self.dct
+
+    def add_row(self, key, val, save=True):
+        try:
+            self.dct[key] = val
+            return val
+        finally:
+            if save: self.save()
+
+    def get_row(self, key):
+        return self.dct[key]
+
+    def save(self):
+        with open(self.filename, 'wb') as f:
+            pickle.dump(self.dct, f)
+
+    def load(self):
+        with open(self.filename, 'rb') as f:
+            return pickle.load(f)
+
+
+class NumpyCache(PickleCache):
+    """Numpy storage-backed Persistency Helper"""
+    def __init__(self, filename):
+        if not filename.endswith('.npy'):
+            filename += '.npy'
+        PickleCache.__init__(self, filename)
+
+    def save(self):
+        np.save(self.filename, _dict_to_array(self.dct))
+
+    def load(self):
+        return _array_to_dict(np.load(self.filename))
+
+
+#----------------------------------------------------------------------
+# Utilities:
+
+def _array_to_dict(arr):
[…]

### 2015-02-03 writeup/paper.tex
fixed two typos in sec 7.1
add LSST discussion

@@
 
+\section{Prospects for LSST}
+\sectlabel{LSST}
+Previously, \citet{Oluseyi12} evaluated the prospects of period finding in early LSST data, and found results which were not encouraging.
+Using the conservative criterion of a 2/3 majority among the top single-band supersmoother fits in the $g$, $r$, and $i$ bands, they showed that, depending on spectral type, finding reliable periods for the brightest ($g \sim 20$) RR Lyrae stars will require several years of LSST data, while periods for some of the faintest ($g \sim 25$) stars will not be reliable with even ten years of data!
+Here we show that a strategy involving the multiband periodogram along with S10-style template fitting greatly improves this outlook.
+
+The results of the previous section show that even in the case of sparsely-sampled multiband data, the shared-phase periodogram finds the true period among its top candidates.
+Furthermore, even with such sparse data, the S10 template-fitting performs well, though computationally it is extremely expensive if a small set of candidate periods cannot be provided (B.
+Sesar, private communication).
+This suggests the following procedure for determining periods in future multiband datasets:
+
+\begin{enumerate}
+\item As a first pass, find a set of candidate frequencies using the multiband periodogram. This is a fast linear optimization that can be trivially parallelized.
+\item Within these candidate frequencies, use the much more costly template-fitting procedure to choose the optimal period from among the handful of candidates.
+\item Compute a goodness-of-fit statistic of this template-fitting result to determine whether the fit is suitable; if not, then apply the template-fitting procedure across the full period range.
+\end{enumerate}
+
+Here we briefly explore simulated LSST observations of RR Lyrae stars in order to gauge the effectiveness of this strategy.
+
+To accomplish this, we \todo{Zeljko fill in details about bands \& cadences}
+and simulated a set of 50 RR Lyrae observations with the S10 templates, at each of a range of apparent magnitudes between $g=20$ and $g=24.5$.
+Given the capability of template-fitting to choose among candidate periods, we use a more relaxed period-matching criterion than in \citet{Oluseyi12}: when evaluating the single-band supersmoother, we require that the true period is among the five periods determined independently in the $u, g, r, i, z$ bands; in the multiband case we require that the true period is among the top five peaks in the multiband periodogram.
+
+\fig{LSST_sims} shows the fraction of stars where this period matching criterion is met as a function of $g$-band magnitude and subset of LSST data.
+The solid lines show the multiband results; the dashed lines show the single-band supersmoother results; and the shading helps guide the eye for the sake of comparison.
+In all cases, the multiband result is superior to the single-band result.
+This is especially so for faint stars, where the multiband results greatly improve over the supersmoother results, which due to model flexibility tend to over-fit the noisy data.
+
+In particular, these results point to much rosier prospects for science with variable stars than previously reported.
+In particular, even with only six months of LSST data, we can expect to correctly identify the periods for over 60\% of stars brighter than $g=22$; with the first two years of LSST observations, this increases to nearly 100\%.
+Part of this improvement is due to the performance of the shared-phase multiband model with noisy data, and part of this improvement is due to the relaxed period-matching constraints enabled by the hybrid approach of non-parametric and template-fitting proposed above.
+
+\todo{Say more about Brani's template fitting?}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{fig09.pdf}
+  \caption{
+  } 
+  \figlabel{LSST_sims}
+\end{figure}
+
 \section{Discussion and Conclusion}
@@
 In our own benchmarks, we found the multiband method to be several times faster than the single-band supersmoother approach, and several orders of magnitude faster than the template fitting approach.
-This suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys
-and we make our tools publicly available on GitHub.
 
-\subsection{Comparison with \citet{Long2014}}
-As we were finishing this study, it was pointed out to us that another group had released a preprint independently addressing the multiband periodogram case \citep[][hereafter LCB14]{Long2014}, and come up with solutions very similar to the ones presented here.
-They present two methods, the ``Multiband Generalized Lomb-Scargle'' (MGLS) which is effectively identical to the $(1, 0)$ multi-phase model
-here, and the ``Penalized Generalized Lomb-Scargle'' (PGLS), which is similar in spirit to our $(0, 1)$ shared-phase model.
+The strengths and weaknesses of the method suggest a hybrid approach to finding periodicity in sparsely-sampled multiband data: a first pass with the fast multiband method, followed by a second pass using the more computationally intensive template-fitting method to select among these candidate periods.
+Despite pessimism in previous studies, our experiments with simulated LSST data lead us to expect such a hybrid approach to successfully identify periods in the vast majority of RR Lyrae stars brighter than $g\sim 22.5$
+This suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys.
 
-In the PGLS model, they start with a multi-phase model, fitting independent $N=1$ term fits to each band, and apply a nonlinear regularization term which penalizes differences in the amplitude and phase.
-In terms of the formalism used in this work, the PGLS model minimizes a regularized $\chi^2$ of the form
-\begin{equation}
-  \chi^2_{PGLS} = \sum_{k=1}^K \bigg[~\chi^2_{GLS}(D^{(K)}) + J_A(A^{(k)}) + J_\phi(\phi^{(k)})~\bigg].
-\end{equation}
-for $K$ bands, where $\chi^2_{GLS}(D^{(K)})$ is the $\chi^2$ of the standard floating mean model on the single-band data $D^{(K)}$, and $J_A$ and $J_\phi$ are regularization/penalty terms which are a function of the amplitude $A^{k}$ and phase $\phi^{(k)}$ of each model.
-In terms of our linear model parameters $\theta^{(k)}$, this amplitude and phase can be expressed:
-\begin{eqnarray}
- A^{(k)} &=& \sqrt{(\theta_1^{(k)})^2 + (\theta_1^{(k)})^2}\nonumber\\
- \phi^{(k)} &=& \arctan(\theta_2 / \theta_1)
-\end{eqnarray}
-The selected form of these regularization terms penalizes deviations of the amplitude and phase from a common mean between the bands; in this sense the PGLS model can be considered a conceptual mid-point between our shared-phase and multi-phase models.
-Within the formalism proposed here, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in our shared-phase model, though the nature of the resulting regularization will differ.
-
-Computationally, the PGLS model requires a nonlinear optimization at each frequency $\omega$, and is thus much more expensive than the straightforward linear optimization of our shared-phase model.
-For this reason, LCB14 propose a clever method by which nested models are used to reduce the number of nonlinear optimizations used: essentially, by showing that the (linear) MGLS $\chi^2$ is a lower-bound of the (non-linear) PGLS $\chi^2$, it is possible to iteratively reduce the number of PGLS computations required to minimize the $\chi^2$ among a grid of frequencies.
-Such an optimization could also be applied in the case of our shared-phase model, but is not necessary here due to its already high speed.
-However, when applying the method to a very large number of light curves, as in e.g. LSST, such a computational trick may prove very useful.
-
-Given these important distinctions between the models proposed here and in LCB14, in future work we plan to do a detailed comparison of the two means of model regularization.
+We have released a Python implementation of the multiband periodogram on GitHub, along with Python code to reproduce all results and figures in this work; see \sect{gatspy}.
+As we were finalizing this manuscript, we were made aware of a preprint of an independent exploration of a similar approach to multiband light curves \citep{Long14}; we compare the two approaches in \sect{long_comparison}.
 
@@
 
+\section{Comparison with \citet{Long2014}}
+\sectlabel{long_comparison}
+As we were finishing this study, it was pointed out to us that another group had released a preprint independently addressing the multiband periodogram case \citep[][hereafter LCB14]{Long2014}, and come up with solutions very similar to the ones presented here.
+They present two methods, the ``Multiband Generalized Lomb-Scargle'' (MGLS) which is effectively identical to the $(1, 0)$ multi-phase model
+here, and the ``Penalized Generalized Lomb-Scargle'' (PGLS), which is similar in spirit to our $(0, 1)$ shared-phase model.
+
+In the PGLS model, they start with a multi-phase model, fitting independent $N=1$ term fits to each band, and apply a nonlinear regularization term which penalizes differences in the amplitude and phase.
+In terms of the formalism used in this work, the PGLS model minimizes a regularized $\chi^2$ of the form
+\begin{equation}
+  \chi^2_{PGLS} = \sum_{k=1}^K \bigg[~\chi^2_{GLS}(D^{(K)}) + J_A(A^{(k)}) + J_\phi(\phi^{(k)})~\bigg].
+\end{equation}
+for $K$ bands, where $\chi^2_{GLS}(D^{(K)})$ is the $\chi^2$ of the standard floating mean model on the single-band data $D^{(K)}$, and $J_A$ and $J_\phi$ are regularization/penalty terms which are a function of the amplitude $A^{k}$ and phase $\phi^{(k)}$ of each model.
+In terms of our linear model parameters $\theta^{(k)}$, this amplitude and phase can be expressed:
+\begin{eqnarray}
+ A^{(k)} &=& \sqrt{(\theta_1^{(k)})^2 + (\theta_1^{(k)})^2}\nonumber\\
+ \phi^{(k)} &=& \arctan(\theta_2 / \theta_1)
+\end{eqnarray}
+The selected form of these regularization terms penalizes deviations of the amplitude and phase from a common mean between the bands; in this sense the PGLS model can be considered a conceptual mid-point between our shared-phase and multi-phase models.
+Within the formalism proposed here, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in our shared-phase model, though the nature of the resulting regularization will differ.
+
+Computationally, the PGLS model requires a nonlinear optimization at each frequency $\omega$, and is thus much more expensive than the straightforward linear optimization of our shared-phase model.
+For this reason, LCB14 propose a clever method by which nested models are used to reduce the number of nonlinear optimizations used: essentially, by showing that the (linear) MGLS $\chi^2$ is a lower-bound of the (non-linear) PGLS $\chi^2$, it is possible to iteratively reduce the number of PGLS computations required to minimize the $\chi^2$ among a grid of frequencies.
+Such an optimization could also be applied in the case of our shared-phase model, but is not necessary here due to its already high speed.
+Nevertheless, when applying the method to a very large number of light curves, as in e.g. LSST, such a computational trick may prove very useful.
+
+Given these important distinctions between the models proposed here and in LCB14, in future work we plan to do a detailed comparison of the two means of model regularization.
+
 \end{document}

### 2015-02-03 writeup/paper.tex
add LSST results to abstract

@@
 and randomly subsampled SDSS Stripe 82 data to demonstrate the superiority of this 
-method compared to other methods from the literature. The Python implementation of 
-this method, along with code to fully reproduce the results reported here, is available on GitHub.
+method compared to other methods from the literature, and find that this method will be able to
+efficiently determine the correct period in the majority of LSST's bright RR Lyrae stars with
+as little as six months of LSST data.
+The Python implementation of this method, along with code to fully reproduce the results
+reported here, is available on GitHub.
 \end{abstract}
@@
   \caption{
+    Fraction of periods correctly determined for LSST RR Lyrae as a function
+    of observing time and $g$-band magnitude, for the multiband (solid lines)
+    and single-band supersmoother (dashed lines) approaches.
   } 

### 2015-02-03 writeup/paper.tex
ZI edits

@@
 Using the conservative criterion of a 2/3 majority among the top single-band supersmoother fits in the $g$, $r$, and $i$ bands, they showed that, depending on spectral type, finding reliable periods for the brightest ($g \sim 20$) RR Lyrae stars will require several years of LSST data, while periods for some of the faintest ($g \sim 25$) stars will not be reliable with even ten years of data!
-Here we show that a strategy involving the multiband periodogram along with S10-style template fitting greatly improves this outlook.
+Here we show that a strategy involving the multiband periodogram along with S10-style template fitting greatly improves this outlook
+(note that the template fitting method is too slow to apply to such a large dataset as expected from LSST; the compute time budget
+to estimate periods for a billion stars on a 1000-core machine in a year is only 30 seconds per light curve).
 
-The results of the previous section show that even in the case of sparsely-sampled multiband data, the shared-phase periodogram finds the true period among its top candidates.
-Furthermore, even with such sparse data, the S10 template-fitting performs well, though computationally it is extremely expensive if a small set of candidate periods cannot be provided (B.
-Sesar, private communication).
+The results of the previous section show that even in the case of sparsely-sampled multiband data, the shared-phase periodogram 
+finds the true period among its top candidates. Furthermore, even with such sparse data, the S10 template-fitting performs well, 
+though computationally it is extremely expensive (hours of computing time for a well-sampled period grid with at least 10$^4$ points; B.
+Sesar, private communication) if a small set of candidate periods cannot be provided.
 This suggests the following procedure for determining periods in future multiband datasets:
@@
 
-To accomplish this, we \todo{Zeljko fill in details about bands \& cadences}
-and simulated a set of 50 RR Lyrae observations with the S10 templates, at each of a range of apparent magnitudes between $g=20$ and $g=24.5$.
+To accomplish this, we use simulated LSST cadence \citep{opsim1, opsim2, opsim3} in 20 arbitrarily chosen fields
+that are representative of the anticipated main survey temporal coverage.
+We simulated a set of 50 RR Lyrae observations with the S10 templates, with a range of apparent magnitudes between 
+$g=20$ and $g=24.5$, corresponding to bright-to-faint range of LSST main-survey observations. We compute expected 
+photometric errors using eqs.~4--6  from \citet{Ivezic08LSST}. 
 Given the capability of template-fitting to choose among candidate periods, we use a more relaxed period-matching criterion than in \citet{Oluseyi12}: when evaluating the single-band supersmoother, we require that the true period is among the five periods determined independently in the $u, g, r, i, z$ bands; in the multiband case we require that the true period is among the top five peaks in the multiband periodogram.
@@
 
-In particular, these results point to much rosier prospects for science with variable stars than previously reported.
+In particular, these results point to much more promising prospects for science with variable stars than previously reported.
 In particular, even with only six months of LSST data, we can expect to correctly identify the periods for over 60\% of stars brighter than $g=22$; with the first two years of LSST observations, this increases to nearly 100\%.
@@
 
-\todo{Say more about Brani's template fitting?}
+\todo{Say more about Brani's template fitting? Zeljko: I don't know. I would guess that fitting templates to
+these periods will uncover the true period. However, the referee could ask how would you know in practice
+what template to use. Thus, we'd have to try all $\sim$400 templates from Branimir's library and hope
+that the true period would be uncovered. Branimir has code to do that so we could ask him. Or we could 
+add this to Discussion. I am slightly more in favor of closing this question by demonstrating that it
+can be done. What do you think?}
 
@@
 The strengths and weaknesses of the method suggest a hybrid approach to finding periodicity in sparsely-sampled multiband data: a first pass with the fast multiband method, followed by a second pass using the more computationally intensive template-fitting method to select among these candidate periods.
-Despite pessimism in previous studies, our experiments with simulated LSST data lead us to expect such a hybrid approach to successfully identify periods in the vast majority of RR Lyrae stars brighter than $g\sim 22.5$
-This suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys.
+Despite pessimism in previous studies, our experiments with simulated LSST data lead us to expect such a hybrid approach to successfully identify periods in the vast majority of RR Lyrae stars brighter than $g\sim 22.5$.
+This finding suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys.
 
 We have released a Python implementation of the multiband periodogram on GitHub, along with Python code to reproduce all results and figures in this work; see \sect{gatspy}.
-As we were finalizing this manuscript, we were made aware of a preprint of an independent exploration of a similar approach to multiband light curves \citep{Long14}; we compare the two approaches in \sect{long_comparison}.
+As we were finalizing this manuscript, we were made aware of a preprint of an independent exploration of a similar approach to multiband light curves \citep{Long2014}; we compare the two approaches in \sect{long_comparison}.
 

### 2015-02-03 writeup/paper.tex
JTV: another pass at conclusions

--- 2015-02-03 writeup/paper.tex
@@
 
+\subsection{Potential Improvements to the Multiband Method}
+The fact that the multiband model does not select the top frequency each time points to its weaknesses: first, as a Fourier-like model, it tends to be tricked by a structured survey window function common in astronomical datasets.
+This is a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle model bias from window bias.
+This could potentially be accounted for by correcting for the effect of the estimated window function; one potential method for this involves estimating the deconvolution of the window power and the observed power \citep{Roberts87}.
+It may also be possible to propose a multiband extension of CARMA \citep{Kelly14}, or another forward-modeling approach to detecting periodicity.
+
+Another potentially fruitful way forward in the multiband case would be to make use of physically-motivated priors in the light-curve shape.
+This approach could allow the model bias to be decreased without a commensurate increase in model complexity, which is what causes poor performance in the case of sparsely-sampled noisy data.
+As an example of such a physically-motivated prior, consider that the paths of RR Lyrae stars through color-color and color-magnitude space are constrained by known astrophysical processes in the structure of the stars \citep[e.g., see Fig.
+5 in][]{Szabo2014}.
+Making use of this information could help break degeneracies in period determination with higher-order models.
+
 \section{Prospects for LSST}
@@
 Previously, \citet{Oluseyi12} evaluated the prospects of period finding in early LSST data, and found results which were not encouraging.
-Using the conservative criterion of a 2/3 majority among the top single-band supersmoother fits in the $g$, $r$, and $i$ bands, they showed that, depending on spectral type, finding reliable periods for the brightest ($g \sim 20$) RR Lyrae stars will require several years of LSST data, while periods for some of the faintest ($g \sim 25$) stars will not be reliable with even ten years of data!
-Here we show that a strategy involving the multiband periodogram along with S10-style template fitting greatly improves this outlook.
+Using the conservative criterion of a 2/3 majority among the top single-band supersmoother periods in the $g$, $r$, and $i$ bands, they showed that, depending on spectral type, finding reliable periods for the brightest ($g \sim 20$) RR Lyrae stars will require several years of LSST data, while periods for some of the faintest ($g \sim 25$) stars will not be reliable with even ten years of data!
 
-The results of the previous section show that even in the case of sparsely-sampled multiband data, the shared-phase periodogram finds the true period among its top candidates.
-Furthermore, even with such sparse data, the S10 template-fitting performs well, though computationally it is extremely expensive if a small set of candidate periods cannot be provided (B.
-Sesar, private communication).
-This suggests the following procedure for determining periods in future multiband datasets:
+One potential way forward is to move away from nonparametric models like supersmoother and lomb-scargle to template-fitting methods such as those used in S10.
+Indeed, such methods perform well even for sparsely-sampled multiband data such as those from the PanSTARRS survey (B.
+Sesar, private communication); the primary drawback is that such blind template fits are computationally extremely expensive: they involve nonlinear optimizations over each of several hundred candidate templates at each of tens of thousands of candidate frequencies.
+Thus the template-fitting method, though extremely accurate, in practice requires several hours of computing time for a well-sampled period grid for a single source, compared to several seconds for the multiband periodogram proposed here.
+Note that several hours per object is orders-of-magnitude too slow in the case of LSST; to estimate periods for a billion stars on a 1000-core machine in a year requires a compute-time budget of only 30 seconds per light curve.
 
+Because of the computational expense of the pure template-fitting method, when working with SDSS II data S10 performed a first-pass with a single-band supersmoother to establish candidate periods, which were in turn evaluated with template-fitting approach.
+Here we show that such a hybrid strategy combining the multiband periodogram and the S10 template fits will be useful for determining periodicity of variables in early LSST data releases, greatly improving on the outlook presented in \citet{Oluseyi12}.
+Rather than doing the full analysis including the final template fits, we will focus on the ability of the multiband periodogram to quickly provide suitable candidate periods, and here assume that the S10 template algorithm will then select or reject the optimial period from this set.
+
+We suggest the following procedure for determining periods in future multiband datasets:
 \begin{enumerate}
@@
    \item Within these candidate frequencies, use the much more costly template-fitting procedure to choose the optimal period from among the handful of candidates.
-\item Compute a goodness-of-fit statistic of this template-fitting result to determine whether the fit is suitable; if not, then apply the template-fitting procedure across the full period range.
+   \item Compute a goodness-of-fit statistic for the best-fit template to determine whether the fit is suitable; if not, then apply the template-fitting procedure across the full period range.
 \end{enumerate}
+Here we briefly explore simulated LSST observations of RR Lyrae stars in order to gauge the effectiveness of the first step in this strategy; the effectiveness of the template-fitting step will be explored further in future work.
 
-Here we briefly explore simulated LSST observations of RR Lyrae stars in order to gauge the effectiveness of this strategy.
-
-To accomplish this, we \todo{Zeljko fill in details about bands \& cadences}
-and simulated a set of 50 RR Lyrae observations with the S10 templates, at each of a range of apparent magnitudes between $g=20$ and $g=24.5$.
+We use a simulated LSST cadence \citep{opsim1, opsim2, opsim3} in 25 arbitrarily chosen fields
+that are representative of the anticipated main survey temporal coverage.
+We simulate a set of 50 RR Lyrae observations with the S10 templates, with a range of apparent magnitudes between 
+$g=20$ and $g=24.5$, corresponding to bright-to-faint range of LSST main-survey observations, and with expected 
+photometric errors computed using eqs.~4--6  from \citet{Ivezic08LSST}. 
 Given the capability of template-fitting to choose among candidate periods, we use a more relaxed period-matching criterion than in \citet{Oluseyi12}: when evaluating the single-band supersmoother, we require that the true period is among the five periods determined independently in the $u, g, r, i, z$ bands; in the multiband case we require that the true period is among the top five peaks in the multiband periodogram.
@@
 The solid lines show the multiband results; the dashed lines show the single-band supersmoother results; and the shading helps guide the eye for the sake of comparison.
-In all cases, the multiband result is superior to the single-band result.
-This is especially so for faint stars, where the multiband results greatly improve over the supersmoother results, which due to model flexibility tend to over-fit the noisy data.
+Because of our relaxed matching criteria, the dashed line here is everywhere higher than the \citet{Oluseyi12} results (compare to Figure 15 in that work); thus the supersmoother result here can be considered a best-case scenario for single-band fits.
+Nevertheless, without fail the multiband result is superior to the single-band result.
+The improvement is most apparent for faint stars, where the greater model flexibility of the supersmoother causes it to over-fit the noisy data.
 
-In particular, these results point to much rosier prospects for science with variable stars than previously reported.
+These results point to much more promising prospects for science with variable stars than previously reported.
 In particular, even with only six months of LSST data, we can expect to correctly identify the periods for over 60\% of stars brighter than $g=22$; with the first two years of LSST observations, this increases to nearly 100\%.
-Part of this improvement is due to the performance of the shared-phase multiband model with noisy data, and part of this improvement is due to the relaxed period-matching constraints enabled by the hybrid approach of non-parametric and template-fitting proposed above.
+Part of this improvement is due to the performance of the shared-phase multiband model with noisy data, and part of this improvement is due to the relaxed period-matching constraints enabled by the hybrid approach of non-parametric and template-based period determination.
 
-\todo{Say more about Brani's template fitting?}
+%\todo{Say more about Brani's template fitting? Zeljko: I don't know. I would guess that fitting templates to
+%these periods will uncover the true period. However, the referee could ask how would you know in practice
+%what template to use. Thus, we'd have to try all $\sim$400 templates from Branimir's library and hope
+%that the true period would be uncovered. Branimir has code to do that so we could ask him. Or we could 
+%add this to Discussion. I am slightly more in favor of closing this question by demonstrating that it
+%can be done. What do you think?}
+
+% Shout-out to @OverheardOnAph from @jakevdp. I love your work.
 
@@
 Experiments on several hundred RR Lyrae stars from the SDSS Stripe 82 dataset indicate that this method outperforms methods used previously in the literature, especially for sparsely-sampled light curves with only single bands observed each night.
-
-The main deficiency of the model is its tendency to be tricked by the survey window function, a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle model bias from window bias.
-This could potentially be accounted for by estimating and filtering the effect of the window function; one potential method for this involves estimating the deconvolution of the window power and the observed power.
-It may also be possible to propose a multiband extension of CARMA \citep{Kelly14},
-or another forward-modeling approaches to detecting periodicity.
-
-Another potentially fruitful way forward in the multiband case is to make use of physically-motivated priors in the light-curve shape.
-For example, the paths of RR Lyrae stars through color-color and color-magnitude space are constrained by known astrophysical processes in the structure of the stars \citep[e.g., see Fig.
-5 in][]{Szabo2014}, and making use of such a physically-motivated prior may help break degeneracies in period determination.
+While there are potential areas of improvement involving corrections to window function artifacts and accounting for physically-motivated priors, the straightforward multiband model outperforms previous \adhoc{} approaches to multiband data.
 
@@
 
-The strengths and weaknesses of the method suggest a hybrid approach to finding periodicity in sparsely-sampled multiband data: a first pass with the fast multiband method, followed by a second pass using the more computationally intensive template-fitting method to select among these candidate periods.
-Despite pessimism in previous studies, our experiments with simulated LSST data lead us to expect such a hybrid approach to successfully identify periods in the vast majority of RR Lyrae stars brighter than $g\sim 22.5$
-This suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys.
+The strengths and weaknesses of the multiband method suggest a hybrid approach to finding periodicity in sparsely-sampled multiband data: a first pass with the fast multiband method, followed by a second pass using the more computationally intensive template-fitting method to select among these candidate periods.
+Despite pessimism in previous studies, our experiments with simulated LSST data lead us to expect such a hybrid approach to successfully identify periods in the vast majority of RR Lyrae stars brighter than $g\sim 22.5$.
+This finding suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys.
 
-We have released a Python implementation of the multiband periodogram on GitHub, along with Python code to reproduce all results and figures in this work; see \sect{gatspy}.
-As we were finalizing this manuscript, we were made aware of a preprint of an independent exploration of a similar approach to multiband light curves \citep{Long14}; we compare the two approaches in \sect{long_comparison}.
+We have released a Python implementation of the multiband periodogram on GitHub, along with Python code to reproduce all results and figures in this work; this is described in \app{gatspy}.
+As we were finalizing this manuscript, we were made aware of a preprint of an independent exploration of a similar approach to multiband light curves \citep{Long2014}; we compare the two approaches in \app{long_comparison}.
 
@@
 Along with the periodogram implementation, it also contains code to download all the data used in this work.
-Code to reproduce this paper, including all figures, is available as well\footnote{\url{http://github.com/jakevdp/multiband_LS/}}.
+Code to reproduce this paper, including all figures, is available in a separate repository\footnote{\url{http://github.com/jakevdp/multiband_LS/}}.
 
@@
 \sectlabel{long_comparison}
-As we were finishing this study, it was pointed out to us that another group had released a preprint independently addressing the multiband periodogram case \citep[][hereafter LCB14]{Long2014}, and come up with solutions very similar to the ones presented here.
+As we were finishing this study, we learned that another group had released a preprint independently addressing the multiband periodogram case, and come up with a solution very similar to the one presented here \citep[][hereafter LCB14]{Long2014}.
 They present two methods, the ``Multiband Generalized Lomb-Scargle'' (MGLS) which is effectively identical to the $(1, 0)$ multi-phase model
@@
 \begin{eqnarray}
- A^{(k)} &=& \sqrt{(\theta_1^{(k)})^2 + (\theta_1^{(k)})^2}\nonumber\\
+ A^{(k)} &=& \sqrt{(\theta_1^{(k)})^2 + (\theta_2^{(k)})^2}\nonumber\\
  \phi^{(k)} &=& \arctan(\theta_2 / \theta_1)
@@
 The selected form of these regularization terms penalizes deviations of the amplitude and phase from a common mean between the bands; in this sense the PGLS model can be considered a conceptual mid-point between our shared-phase and multi-phase models.
-Within the formalism proposed here, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in our shared-phase model, though the nature of the resulting regularization will differ.
+Within the formalism proposed in the current work, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in our shared-phase model, though the nature of the resulting regularization will differ.
 
 Computationally, the PGLS model requires a nonlinear optimization at each frequency $\omega$, and is thus much more expensive than the straightforward linear optimization of our shared-phase model.
-For this reason, LCB14 propose a clever method by which nested models are used to reduce the number of nonlinear optimizations used: essentially, by showing that the (linear) MGLS $\chi^2$ is a lower-bound of the (non-linear) PGLS $\chi^2$, it is possible to iteratively reduce the number of PGLS computations required to minimize the $\chi^2$ among a grid of frequencies.
+For this reason, LCB14 proposes a clever method by which nested models are used to reduce the number of nonlinear optimizations used: essentially, by showing that the (linear) MGLS $\chi^2$ is a lower-bound of the (non-linear) PGLS $\chi^2$, it is possible to iteratively reduce the number of PGLS computations required to minimize the $\chi^2$ among a grid of frequencies.
 Such an optimization could also be applied in the case of our shared-phase model, but is not necessary here due to its already high speed.
-Nevertheless, when applying the method to a very large number of light curves, as in e.g. LSST, such a computational trick may prove very useful.
+Nevertheless, when applying the method to a very large number of light curves, as in e.g.~LSST, such a computational trick may prove very useful.
 
-Given these important distinctions between the models proposed here and in LCB14, in future work we plan to do a detailed comparison of the two means of model regularization.
+Given these important distinctions between the models proposed here and in LCB14, in future work we plan to do a detailed comparison of the two means of multiband model regularization.
 

### 2015-02-03 writeup/paper.tex
JTV final pass

@@
 \begin{abstract}
-This paper introduces the {\it multiband periodogram}, a general extension of the standard 
+This paper introduces the {\it multiband periodogram}, a general extension of the well-known 
 Lomb-Scargle approach for detecting periodic signals in time-domain data. In addition to 
@@
 as little as six months of LSST data.
-The Python implementation of this method, along with code to fully reproduce the results
+A Python implementation of this method, along with code to fully reproduce the results
 reported here, is available on GitHub.
@@
 with the period, and optionally the phase, shared across all bands. The key aspect enabling this approach is the use of Tikhonov regularization 
-(discussed in detail in \S \ref{regularization}) which drives most of the variability into the so-called {\it base 
+(discussed in detail in \sect{regularization}) which drives most of the variability into the so-called {\it base 
 model} common to all bands, while fits for individual bands describe residuals relative to the base model 
@@
 
-In \sect{brief_overview} we provide a brief overview of least-squares periodic fitting, and in \sect{matrix_formalism} derive the matrix-based formalism for single-band periodic analysis used through the rest of this work.
+The remainder of the paper is organized as follows:
+in \sect{brief_overview} we provide a brief review of least-squares periodic fitting, and in \sect{matrix_formalism} derive the matrix-based formalism for single-band periodic analysis used through the rest of this work.
 \sect{extending_periodogram} introduces several extensions and generalizations to the single-band model that the matrix formalism makes possible, including floating mean models, truncated Fourier models, and regularized models.
 In \sect{multiband}, we use the ideas behind these extensions to motivate the {\it multiband periodogram}, and show some examples of its use on simulated data.
-In \sect{stripe82} we apply this method to measurements of 483 RR Lyrae stars first explored by \citet[][hereafter S10]{Sesar2010}, and conclude in \sect{discussion}.
+In \sect{stripe82} we apply this method to measurements of 483 RR Lyrae stars first explored by \citet[][hereafter S10]{Sesar2010}, and in \sect{LSST} explore the performance of the method for simulated observations from the LSST survey.
+We conclude in \sect{discussion}.
 
-\section{Brief Overview of Periodograms}
+\section{Brief Overview of Periodic Analysis}
 \sectlabel{brief_overview}
@@
 The detection and quantification of periodicity in time-varying signals is an important area of data analysis within modern time-domain astronomical surveys.
-For evenly-spaced data, the {\it Schuster periodogram}, introduced in 1905, gives a quantitative measure of the periodicity of data as a function of the angular frequency $\omega$.
-For data $\{y_k\}_{k=1}^N$ measured at equal intervals $t_k = t_0 + k\Delta t$, the Schuster periodogram, which measures the spectral power as a function of the angular frequency, is given by
+For evenly-spaced data, the {\it periodogram}, a term coined by \citet{Schuster98}, gives a quantitative measure of the periodicity of data as a function of the angular frequency $\omega$.
+For data $\{y_k\}_{k=1}^N$ measured at equal intervals $t_k = t_0 + k\Delta t$, Schuster's periodogram, which measures the spectral power as a function of the angular frequency, is given by
 \begin{equation}
@@
 
-While the important features of least squares frequency estimation via Lomb-Scargle periodograms has been discussed elsewhere, we will present a brief introduction to the subject in the following section.
-The matrix-based formalism we develop here will make clear how the method can be extended to more sophisticated models, including the multiband method.
+While the important features of least squares frequency estimation via Lomb-Scargle periodograms have been discussed elsewhere, we will present a brief introduction to the subject in the following section.
+The matrix-based formalism we develop here will make clear how the method can be extended to more sophisticated models, including the multiband periodogram proposed in this work.
 
@@
 
-In this section we present a quick quantitative introduction to the least squares fitting formulation of the normalized periodogram of \eq{LombScargle}. We denote $N$ observed data points as
+In this section we present a brief quantitative introduction to the least squares fitting formulation of the normalized periodogram of \eq{LombScargle}. We denote $N$ observed data points as
 \begin{equation}
@@
 \begin{enumerate}
-  \item It is trivially extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the {\it noise covariance matrix} $\Sigma$
+  \item It is trivially extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the {\it noise covariance matrix} $\Sigma$.
   \item It is trivially extended to more sophisticated linear models by appropriately modifying the {\it design matrix} $X_\omega$.
-  \item It is trivially extended to include Tikhonov/L2-regularization terms (see \S \ref{regularization} for more details)  by adding an appropriate diagonal term to the {\it normal matrix} $X_\omega^T\Sigma^{-1}X_\omega$.
+  \item It is trivially extended to include Tikhonov/L2-regularization terms (see \sect{regularization} for more details)  by adding an appropriate diagonal term to the {\it normal matrix} $X_\omega^T\Sigma^{-1}X_\omega$.
 \end{enumerate}
@@
 
-\subsection{Regularized Models \label{regularization}}
+\subsection{Regularized Models}
+\sectlabel{regularization}
 
@@
 We have seen that as we add more terms to the fit, the model will come closer and closer to the observed data.
-In the extreme case, when the number of model parameters equals the number of data points, the model can fit the data {\it exactly} regardless of frequency and the periodogram will be everywhere unity (though in most cases, numerical inaccuracies in the fit prevent a truly perfect fit).
+In the extreme case, when the number of model parameters equals the number of data points, the model can fit the data {\it exactly} regardless of frequency and the periodogram will be everywhere unity (though in most cases, numerical inaccuracies prevent a truly perfect fit).
 For very high-order models, the fit becomes very sensitive to the noise in the data, and we say we have {\it over-fit} the data.
@@
 To compute such a periodogram, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
-The model contains the following features:
+The multiband model contains the following features:
 \begin{enumerate}
@@
 \end{equation}
-Here the nonzero entries of the final five columns are binary flags indicating the $(u, g, r, i, z)$-band of the given observation: for this example, the first row is a $u$-band measurement, the second is a $z$-band, the third is a $i$-band, etc.
+Here the nonzero entries of the final five columns are binary flags indicating the $(u, g, r, i, z)$-band of the given observation: for this example, the first row is a $u$-band measurement, the second is a $z$-band, the third is a $i$-band, etc., as indicated by the position of the nonzero matrix element within the row.
 
@@
 The {\it multi-phase} $(N_{base}=0, N_{band}=1)$ model turns out to be a particularly special case.
-The base model is a simple global offset which is degenerate with the offsets in each band, so that the design matrix $X_\omega$ can be straightforwardly rearranged as block-diagonal.
+Here the base model is a simple global offset which is degenerate with the offsets in each band, so that the design matrix $X_\omega$ can be straightforwardly rearranged as block-diagonal.
 A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here these independent components amount to the single-band floating-mean model from \sect{floating_mean}, fit independently for each of the $K$ bands.
@@
 The shared-phase multiband approach (i.e.
-$(N_{base}=1,N_{band}=0)$), shown in the lower-right panel, makes use of all the information in a single model, and clearly recovers the true frequency of 0.622 days.
+$N_{base}=1,N_{band}=0$), shown in the lower-right panel, fits a single model to the full data and clearly recovers the true frequency of 0.622 days.
 
 This shared-phase $(1,0)$ model is only one of the possible multiband options, however: \fig{multiband_models} shows multiband fits to this data for models with various choices of $(N_{base},N_{band})$.
-We see here many of the characteristics noted above for single-band models: as discussed in \sect{multiterm}, increasing the number of Fourier terms leads to power at multiples of the fundamental period.
-Additionally, increasing the model complexity (roughly indexed by the effective number of free parameters $M^{eff}$) tends to increase the background level of the periodogram, potentially obscuring significant peaks.
+We see here many of the characteristics noted above for single-band models: as discussed in \sect{multiterm}, increasing the number of Fourier terms leads to power at multiples of the fundamental period, and increased model complexity (roughly indexed by the effective number of free parameters $M^{eff}$) tends to increase the background level of the periodogram, obscuring significant peaks.
 For this reason, models with $N_{base} > N_{band}$ are the most promising: they allow a flexible fit with minimal model complexity.
@@
 Stripe 82 is a three hundred square degree equatorial region of the sky which was repeatedly imaged through multiple band-passes during phase II of the Sloan Digital Sky Survey \citep[SDSS II, see][]{Sesar2007}.
-Here we consider the SDSS II observations of 483 RR Lyrae stars compiled and studied by S10.
-S10 determined periods for these stars based on emperically-derived light curve templates.
+Here we consider the SDSS II observations of 483 RR Lyrae stars compiled and studied by S10, in which periods for these stars were determined based on emperically-derived light curve templates.
 Because the template-fitting method is extremely computationally intensive, S10 first determined candidate periods by taking the top 5 results of the Supersmoother \citep{Reimann94} algorithm applied to the $g$-band; template fits were then performed at each candidate period and the period with the best template fit was reported as the true period.
@@
     Shown for reference are the beat aliases (dotted lines) and the multiplicative alias (dashed lines): numbers along the top and right edges of the panels indicate the number of points aligned with each trend.
-    The single-band supersmoother algorithm tends to err toward multiplicative aliases, while the multiband algorithm tends to err toward beat frequency aliases.
+    The single-band supersmoother model tends to err toward multiplicative aliases, while the multiband model tends to err toward beat frequency aliases.
     Both methods find the correct period among the top 5 significant peaks around 99\% of the time.
@@
 The full S10 RR Lyrae dataset consists of 483 objects with an average of 55 observations in each of the five SDSS $ugriz$ bands spread over just under ten years.
-In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother algorithm and the shared-phase $(0, 1)$-multiband algorithm\footnote{The supersmoother ``periodogram'' $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$, where $\bar{r}_0$ is the residual around a constant model.}.
+In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother model and the shared-phase $(0, 1)$-multiband model\footnote{The supersmoother ``periodogram'' $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$, where $\bar{r}_0$ is the mean absolute residual around a constant model.}.
 Here we have a case which is analogous to that shown for simulated data in the top panels of \fig{multiband_sim}: each band has enough data to easily locate candidate peaks, the best of which is selected via the S10 template-fitting procedure.
@@
 the dashed line).
-This is due to the flexibility of the supersmoother model: a doubled period spreads the points out, leading to fewer constraints in each neighborhood and thus a smaller average residual around model.
+This is due to the flexibility of supersmoother: a doubled period spreads the points out, leading to fewer constraints in each neighborhood and thus a smaller average residual around model.
 In other words, the SuperSmoother tends to over-fit data which is sparsely-sampled.
@@
 Above we saw that the multiband model is comparable to methods from the literature for densely-sampled data.
-Where we expect the multiband approach to gain an advantage is when the data are sparsely sampled, with observations through only a single band at each observation.
+Where we expect the multiband approach to gain an advantage is when the data are sparsely sampled, with data through only a single band at each observation time.
 To simulate this, we reduce the size of the Stripe 82 RR Lyrae dataset by a factor of 5, keeping only a single band of imaging each night: an average of 11 observations of each object per band.
@@
 It is clear that the supersmoother model is simply over-fitting noise with this few data points: the top period matches S10 in only 32\% of cases (compared to 87\% with the full dataset), and the top 5 periods contain the S10 period only 45\% of the time.
-The failure mode is much less predictable as well: rather than being clustered near aliases, most of the period determinations are scattered seemingly randomly around the space.
+The failure mode is much less predictable as well: rather than being clustered near aliases, most of the period determinations are scattered seemingly randomly around the parameter space.
 
@@
 
-Another potentially fruitful way forward in the multiband case would be to make use of physically-motivated priors in the light-curve shape.
+One potentially fruitful avenue of research which we do not study here is the application of L1 regularization (also known as Lasso regression) to the higher-order periodogram.
+This is similar in spirit to the Tikhonov regularization discussed in \sect{regularization}, but tends toward sparsity in the model parameters \citep[see, e.g.][for a discussion]{ICVG2014}.
+Such an approach could provide a useful tradeoff between model complexity and bias in the case of higher-order truncated Fourier models.
+
+Another potentially fruitful way forward in the multiband case would be to define and make use of physically-motivated priors in the light-curve shape.
 This approach could allow the model bias to be decreased without a commensurate increase in model complexity, which is what causes poor performance in the case of sparsely-sampled noisy data.
@@
 
-One potential way forward is to move away from nonparametric models like supersmoother and lomb-scargle to template-fitting methods such as those used in S10.
-Indeed, such methods perform well even for sparsely-sampled multiband data such as those from the PanSTARRS survey (B.
-Sesar, private communication); the primary drawback is that such blind template fits are computationally extremely expensive: they involve nonlinear optimizations over each of several hundred candidate templates at each of tens of thousands of candidate frequencies.
-Thus the template-fitting method, though extremely accurate, in practice requires several hours of computing time for a well-sampled period grid for a single source, compared to several seconds for the multiband periodogram proposed here.
+One potential remedy is to move away from general models like supersmoother and lomb-scargle to specific template-fitting methods such as those used in S10.
+Indeed, such methods perform well even for sparsely-sampled multiband data such as those from the PanSTARRS survey; the primary drawback is that such blind template fits are computationally extremely expensive: they involve nonlinear optimizations over each of several hundred candidate templates at each of tens of thousands of candidate frequencies (B.
+Sesar, private communication).
+Thus the template-fitting method, though extremely accurate, in practice requires several hours of CPU time for a well-sampled period grid for a single source (compared to several seconds for the multiband periodogram proposed here).
 Note that several hours per object is orders-of-magnitude too slow in the case of LSST; to estimate periods for a billion stars on a 1000-core machine in a year requires a compute-time budget of only 30 seconds per light curve.
@@
 Here we show that such a hybrid strategy combining the multiband periodogram and the S10 template fits will be useful for determining periodicity of variables in early LSST data releases, greatly improving on the outlook presented in \citet{Oluseyi12}.
-Rather than doing the full analysis including the final template fits, we will focus on the ability of the multiband periodogram to quickly provide suitable candidate periods, and here assume that the S10 template algorithm will then select or reject the optimial period from this set.
 
@@
 Here we briefly explore simulated LSST observations of RR Lyrae stars in order to gauge the effectiveness of the first step in this strategy; the effectiveness of the template-fitting step will be explored further in future work.
+Rather than doing the full analysis including the final template fits, we will focus on the ability of the multiband periodogram to quickly provide suitable candidate periods under the assumption that the S10 template algorithm will then select or reject the optimial period from this set.
+
+\subsection{LSST Simulations}
 
@@
 The solid lines show the multiband results; the dashed lines show the single-band supersmoother results; and the shading helps guide the eye for the sake of comparison.
-Because of our relaxed matching criteria, the dashed line here is everywhere higher than the \citet{Oluseyi12} results (compare to Figure 15 in that work); thus the supersmoother result here can be considered a best-case scenario for single-band fits.
-Nevertheless, without fail the multiband result is superior to the single-band result.
-The improvement is most apparent for faint stars, where the greater model flexibility of the supersmoother causes it to over-fit the noisy data.
+Because of our relaxed matching criteria, even the single-band supersmoother results here are much more optimistic than the \citet{Oluseyi12} results (compare to Figure 15 in that work): the supersmoother result here can be considered a best-case scenario for \adhoc{} single-band fits.
+Without fail, the multiband result exceeds this best-case single-band result; the improvement is most apparent for faint stars, where the greater model flexibility of the supersmoother causes it to over-fit the noisy data.
 
@@
 In particular, even with only six months of LSST data, we can expect to correctly identify the periods for over 60\% of stars brighter than $g=22$; with the first two years of LSST observations, this increases to nearly 100\%.
-Part of this improvement is due to the performance of the shared-phase multiband model with noisy data, and part of this improvement is due to the relaxed period-matching constraints enabled by the hybrid approach of non-parametric and template-based period determination.
+Part of this improvement is due to the performance of the shared-phase multiband model with noisy data, and part of this improvement is due to the relaxed period-matching constraints enabled by the hybrid approach of periodogram-based and template-based period determination.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{fig09.pdf}
+  \includegraphics[width=0.7\textwidth]{fig09.pdf}
   \caption{
@@
 Looking forward to future variable star catalogs from PanSTARRS, DES, and LSST, there are two important constraints that any analysis method must meet: the methods must be able to cope with heterogeneous and noisy observations through multiple band-passes, and the methods must be fast enough to be computable on millions or even billions of objects.
-The multiband method, through its combination of flexibility and model simplicity, meets the first: as shown above, it performs well in the case of sparsely-sampled multiband data.
-It also meets the second constraint: it consists of a fast linear model at each frequency, compared to a sliding-window model in the case of supersmoother, a nonlinear optimization in the case of template-fitting, and a Markov Chain Monte Carlo in the case of CARMA models.
+The multiband method, through its combination of flexibility and model simplicity, meets the first constraint: as shown above, in the case of sparsely-sampled noisy multiband data, it out-performs previous approaches to period determination.
+It also meets the second constraint: it requires the solution of a simple linear model at each frequency, compared to a sliding-window model in the case of supersmoother, a nonlinear optimization in the case of template-fitting, and a Markov Chain Monte Carlo in the case of CARMA models.
 In our own benchmarks, we found the multiband method to be several times faster than the single-band supersmoother approach, and several orders of magnitude faster than the template fitting approach.
@@
 The strengths and weaknesses of the multiband method suggest a hybrid approach to finding periodicity in sparsely-sampled multiband data: a first pass with the fast multiband method, followed by a second pass using the more computationally intensive template-fitting method to select among these candidate periods.
-Despite pessimism in previous studies, our experiments with simulated LSST data lead us to expect such a hybrid approach to successfully identify periods in the vast majority of RR Lyrae stars brighter than $g\sim 22.5$.
-This finding suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future surveys.
+Despite pessimism in previous studies, our experiments with simulated LSST data lead us to expect such a hybrid approach to successfully identify periods in the vast majority of RR Lyrae stars brighter than $g\sim 22.5$ in the first months of the survey.
+This finding suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future multiband surveys.
 
 We have released a Python implementation of the multiband periodogram on GitHub, along with Python code to reproduce all results and figures in this work; this is described in \app{gatspy}.
-As we were finalizing this manuscript, we were made aware of a preprint of an independent exploration of a similar approach to multiband light curves \citep{Long2014}; we compare the two approaches in \app{long_comparison}.
+As we were finalizing this manuscript, we were made aware of a preprint of an independent exploration of a similar approach to multiband light curves \citep{Long2014}; we discuss the similarities and differences between these two approaches in \app{long_comparison}.
 

### 2015-02-04 writeup/paper.tex
ZI edits for caption fig. 9

@@
 Here we briefly explore simulated LSST observations of RR Lyrae stars in order to gauge the effectiveness of the first step in this strategy; the effectiveness of the template-fitting step will be explored further in future work.
-Rather than doing the full analysis including the final template fits, we will focus on the ability of the multiband periodogram to quickly provide suitable candidate periods under the assumption that the S10 template algorithm will then select or reject the optimial period from this set.
+Rather than doing the full analysis including the final template fits, we will focus on the ability of the multiband periodogram to quickly provide suitable candidate periods under the assumption that the S10 template algorithm will then select or reject the optimal period from this set.
 
@@
     Fraction of periods correctly determined for LSST RR Lyrae as a function
-    of observing time and $g$-band magnitude, for the multiband (solid lines)
+    of the length of the observing season and the mean $g$-band magnitude, for the multiband (solid lines)
     and single-band supersmoother (dashed lines) approaches.

### 2015-02-04 writeup/paper.tex
paper: spell check and add small changes

@@
 Lomb-Scargle approach for detecting periodic signals in time-domain data. In addition to 
-advantanges of the Lomb-Scargle method such as treatment of non-uniform sampling and
+advantages of the Lomb-Scargle method such as treatment of non-uniform sampling and
 heteroscedastic errors, the multiband periodogram significantly improves period finding for randomly 
@@
 maximum of the periodogram is found by grid search \citep[for details see, e.g.][]{ICVG2014}.
-A more general parameteric method based on the use of continuous-time autoregressive moving average (CARMA) model
+A more general parametric method based on the use of continuous-time autoregressive moving average (CARMA) model
 was recently introduced by \citet{Kelly14}. CARMA models can also treat non-uniformly sampled time series with 
@@
 For the case of simultaneously taken multiband measurements, \cite{Suveges12} utilized the principal component
-method to optimaly extract the best period. Their method is essentially a multiband generalization of the well-known
+method to optimally extract the best period. Their method is essentially a multiband generalization of the well-known
 two-band Welch-Stetson variability index \citep{Stetson1996}. Unfortunately, when data in each band are taken at
@@
 This long-recognized connection between spectral power and least squares fitting methods was solidified by \citet{Jaynes87}, who demonstrated that the normalized periodogram of Lomb and Scargle is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
-Building on this result, \citet{Bretthorst88} explored the extension of these methods to more complicated models with multiple frequency terms, nonstationary frequencies, and other more sophisticated models within a Bayesian framework.
+Building on this result, \citet{Bretthorst88} explored the extension of these methods to more complicated models with multiple frequency terms, non-stationary frequencies, and other more sophisticated models within a Bayesian framework.
 
@@
 \end{eqnarray}
-The expression for the normalized periodogram can be computed by finding via standard methods the value of $\theta$ which minizes $\chi^2(\omega)$, and plugging the result into \eq{chi2PN}.
+The expression for the normalized periodogram can be computed by finding via standard methods the value of $\theta$ which minimizes $\chi^2(\omega)$, and plugging the result into \eq{chi2PN}.
 This yields
@@
     The data consist of 80 observations drawn from a sinusoidal model.
-To mimick a potentially damaging selection effect, all observations with magnitude fainter than 16 are removed (indicated by the light-gray points).
+To mimic a potentially damaging selection effect, all observations with magnitude fainter than 16 are removed (indicated by the light-gray points).
 The standard and floating-mean periodograms are computed from the remaining data; these fits are shown over the data in the left panel.
@@
 In the matrix formalism, the modification is much more straightforward: all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
-This column of ones corresponds to a third entry in the parameter vector $\theta$, and acts as a uniform constant offset for all datapoints.
+This column of ones corresponds to a third entry in the parameter vector $\theta$, and acts as a uniform constant offset for all data points.
 
@@
 The lower-left panel shows the observations as a function of phase, and the lower-right panels show the periodograms derived from the data.
-With only 12 observations for each indiviudal band, it is clear that there is not enough data to accurately determine the period within each single band.
+With only 12 observations for each individual band, it is clear that there is not enough data to accurately determine the period within each single band.
 The shared-phase multiband approach (i.e.
@@
 Stripe 82 is a three hundred square degree equatorial region of the sky which was repeatedly imaged through multiple band-passes during phase II of the Sloan Digital Sky Survey \citep[SDSS II, see][]{Sesar2007}.
-Here we consider the SDSS II observations of 483 RR Lyrae stars compiled and studied by S10, in which periods for these stars were determined based on emperically-derived light curve templates.
+Here we consider the SDSS II observations of 483 RR Lyrae stars compiled and studied by S10, in which periods for these stars were determined based on empirically-derived light curve templates.
 Because the template-fitting method is extremely computationally intensive, S10 first determined candidate periods by taking the top 5 results of the Supersmoother \citep{Reimann94} algorithm applied to the $g$-band; template fits were then performed at each candidate period and the period with the best template fit was reported as the true period.
@@
 These results point to much more promising prospects for science with variable stars than previously reported.
-In particular, even with only six months of LSST data, we can expect to correctly identify the periods for over 60\% of stars brighter than $g=22$; with the first two years of LSST observations, this increases to nearly 100\%.
+In particular, even with only six months of LSST data, we can expect to correctly identify the periods for over 60\% of stars brighter than $g=22$; with the first two years of LSST observations, this increases to nearly 100\%; with five years of data, the multiband method identifies the correct period for 100\% of even the faintest stars.
 Part of this improvement is due to the performance of the shared-phase multiband model with noisy data, and part of this improvement is due to the relaxed period-matching constraints enabled by the hybrid approach of periodogram-based and template-based period determination.
@@
     and single-band supersmoother (dashed lines) approaches.
+    The multiband method is superior to the single-band supersmoother approach in all cases, and especially for the faintest objects.
   } 
@@
 
-{\it Acknowledgements:} JTV is supported by the University of Washington eScience institute, including grants from the Alfred P.
+{\it Acknowledgments:} JTV is supported by the University of Washington eScience institute, including grants from the Alfred P.
 Sloan Foundation, the Gordon and Betty Moore Foundation, and the Washington Research Foundation.
@@
 \sectlabel{gatspy}
-The algorithm outlined in this paper is available in {\tt gatspy}, an open-source Python package for general astronomical time-series analysis\footnote{\url{http://github.com/jakevdp/gatspy/}}.
+The algorithm outlined in this paper is available in {\tt gatspy}, an open-source Python package for general astronomical time-series analysis\footnote{\url{http://github.com/astroml/gatspy/}} \citep{gatspy}.
 Along with the periodogram implementation, it also contains code to download all the data used in this work.
@@
 The models in the \texttt{gatspy} package contain many more methods, and much more functionality that what is shown here.
-For updates, more examples, and more information, visit \url{http://github.com/jakevdp/gatspy/}.
+For updates, more examples, and more information, visit \url{http://github.com/astroml/gatspy/}.
 

### 2015-02-04 writeup/paper.tex
final read-through

@@
 % Draft watermark:
-\usepackage{draftwatermark}
-\SetWatermarkLightness{0.9}
-\SetWatermarkScale{4}
+%\usepackage{draftwatermark}
+%\SetWatermarkLightness{0.9}
+%\SetWatermarkScale{4}
 
@@
 fits for individual bands describe residuals relative to the base model and typically require
-lower-order Fourier series. This decrease in the required number of fit parameters is the
+lower-order Fourier series. This decrease in the effective model complexity is the
 main reason for improved performance. We use simulated light curves 
@@
 two-band Welch-Stetson variability index \citep{Stetson1996}. Unfortunately, when data in each band are taken at
-different times, the  principal component approach in not applicable. In such cases, past studies have generally relied 
+different times, such an approach in not applicable. In such cases, past studies have generally relied 
 on \adhoc{} methods such as a majority vote among multiple single-band estimates of the 
@@
 model} common to all bands, while fits for individual bands describe residuals relative to the base model 
-and typically require lower-order Fourier series. This decrease in the required number of fit parameters is the
+and typically require lower-order Fourier series. This decrease in effective model complexity is the
 main reason for improved performance.
@@
 \end{equation}
-For the single-term Fourier model, it can be shown \citep[See, e.g.][]{ICVG2014} that
+For the single-term Fourier model, it can be shown \citep[see, e.g.][]{ICVG2014} that
 \begin{equation}
@@
 Here we visualize the data from the left panel, but folded as a function of phase, and overplotted with the best-fit single-term model.
-Here it is apparent that the single-term model is highly biased: RR Lyrae light curves are, in general, much more complicated than a simple sinusoid.
+This visualization makes it apparent that the single-term model is highly biased: RR Lyrae light curves are, in general, much more complicated than a simple sinusoid.
 Nevertheless, the simplistic sinusoidal model is able to recover the correct frequency to a high degree of accuracy (roughly related to the width of the peak) and significance (roughly related to the height of the peak).
@@
 \subsection{Potential Improvements to the Multiband Method}
-The fact that the multiband model does not select the top frequency each time points to its weaknesses: first, as a Fourier-like model, it tends to be tricked by a structured survey window function common in astronomical datasets.
-This is a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle model bias from window bias.
+The fact that the multiband model does not select the top frequency each time points to its weaknesses: first, as a Fourier-like model, it tends to respond to frequency structure in the window function as well as frequency structure in the data.
+This is a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle bias in the model from bias due to the survey window.
 This could potentially be accounted for by correcting for the effect of the estimated window function; one potential method for this involves estimating the deconvolution of the window power and the observed power \citep{Roberts87}.
@@
 
-One potentially fruitful avenue of research which we do not study here is the application of L1 regularization (also known as Lasso regression) to the higher-order periodogram.
-This is similar in spirit to the Tikhonov regularization discussed in \sect{regularization}, but tends toward sparsity in the model parameters \citep[see, e.g.][for a discussion]{ICVG2014}.
+One potentially fruitful avenue of research which we do not study here is the application of other types of regularization to the higher-order periodogram.
+In particular, L1 regularization (also known as Lasso regression) could lead to interesting results: L1 regularization is similar in spirit to the Tikhonov regularization discussed in \sect{regularization}, but tends toward sparsity in the model parameters \citep[see, e.g.][for a discussion]{ICVG2014}.
 Such an approach could provide a useful tradeoff between model complexity and bias in the case of higher-order truncated Fourier models.
 
-Another potentially fruitful way forward in the multiband case would be to define and make use of physically-motivated priors in the light-curve shape.
+Another potentially interesting extension of the multiband case would be to define and make use of physically-motivated priors in the light-curve shape.
 This approach could allow the model bias to be decreased without a commensurate increase in model complexity, which is what causes poor performance in the case of sparsely-sampled noisy data.
@@
 
-\section{Prospects for LSST}
+\section{Prospects for Multiband Periodograms with LSST}
 \sectlabel{LSST}
@@
 Sesar, private communication).
-Thus the template-fitting method, though extremely accurate, in practice requires several hours of CPU time for a well-sampled period grid for a single source (compared to several seconds for the multiband periodogram proposed here).
+Thus the template-fitting method, though it produces accurate periods, in practice requires several hours of CPU time for a well-sampled period grid for a single source (compared to several seconds for the multiband periodogram proposed here).
 Note that several hours per object is orders-of-magnitude too slow in the case of LSST; to estimate periods for a billion stars on a 1000-core machine in a year requires a compute-time budget of only 30 seconds per light curve.
@@
    \item As a first pass, find a set of candidate frequencies using the multiband periodogram. This is a fast linear optimization that can be trivially parallelized.
-   \item Within these candidate frequencies, use the much more costly template-fitting procedure to choose the optimal period from among the handful of candidates.
+   \item Within these candidate frequencies, use the more costly template-fitting procedure to choose the optimal period from among the handful of candidates.
    \item Compute a goodness-of-fit statistic for the best-fit template to determine whether the fit is suitable; if not, then apply the template-fitting procedure across the full period range.
@@
 The solid lines show the multiband results; the dashed lines show the single-band supersmoother results; and the shading helps guide the eye for the sake of comparison.
-Because of our relaxed matching criteria, even the single-band supersmoother results here are much more optimistic than the \citet{Oluseyi12} results (compare to Figure 15 in that work): the supersmoother result here can be considered a best-case scenario for \adhoc{} single-band fits.
+Because of our relaxed matching criteria, even the single-band supersmoother results here are much more optimistic than the \citet{Oluseyi12} results (compare to Figure 15 in that work): the supersmoother result here can be considered representative of a best-case scenario for \adhoc{} single-band fits.
 Without fail, the multiband result exceeds this best-case single-band result; the improvement is most apparent for faint stars, where the greater model flexibility of the supersmoother causes it to over-fit the noisy data.
 
-These results point to much more promising prospects for science with variable stars than previously reported.
+The performance of the multiband periodogram points to much more promising prospects for science with variable stars than previously reported.
 In particular, even with only six months of LSST data, we can expect to correctly identify the periods for over 60\% of stars brighter than $g=22$; with the first two years of LSST observations, this increases to nearly 100\%; with five years of data, the multiband method identifies the correct period for 100\% of even the faintest stars.
@@
 The multiband method, through its combination of flexibility and model simplicity, meets the first constraint: as shown above, in the case of sparsely-sampled noisy multiband data, it out-performs previous approaches to period determination.
-It also meets the second constraint: it requires the solution of a simple linear model at each frequency, compared to a sliding-window model in the case of supersmoother, a nonlinear optimization in the case of template-fitting, and a Markov Chain Monte Carlo in the case of CARMA models.
+It also meets the second constraint: it requires the solution of a simple linear model at each frequency, compared to a rank-based sliding-window model in the case of supersmoother, a nonlinear optimization in the case of template-fitting, and a Markov Chain Monte Carlo analysis in the case of CARMA models.
 In our own benchmarks, we found the multiband method to be several times faster than the single-band supersmoother approach, and several orders of magnitude faster than the template fitting approach.
@@
 The strengths and weaknesses of the multiband method suggest a hybrid approach to finding periodicity in sparsely-sampled multiband data: a first pass with the fast multiband method, followed by a second pass using the more computationally intensive template-fitting method to select among these candidate periods.
-Despite pessimism in previous studies, our experiments with simulated LSST data lead us to expect such a hybrid approach to successfully identify periods in the vast majority of RR Lyrae stars brighter than $g\sim 22.5$ in the first months of the survey.
+Despite pessimism in previous studies, our experiments with simulated LSST data indicate that such a hybrid approach will successfully identify periods in the majority of RR Lyrae stars brighter than $g\sim 22.5$ in the first months of the survey, and the majority of the faintest detected stars with several years of data.
 This finding suggests that the multiband periodogram could have an important role to play in the analysis of variable stars in future multiband surveys.

### 2015-02-05 writeup/paper.tex
Change to 2 column w/ emulateapj

@@
-\documentclass[12pt,preprint]{aastex}
+\documentclass{emulateapj}
 
@@
 Most famously, \citet{Lomb76} and \citet{Scargle82} extended earlier work to define the {\it normalized periodogram}:
-\begin{equation}
+\begin{eqnarray}
   \eqlabel{LombScargle}
-  P_N(\omega) = \frac{1}{2\,V_y}\left[
-    \frac{\left[\sum_k(y_k - \bar{y})\cos\omega(t_k - \tau)\right]^2}
-    {\sum_k \cos^2\omega(t_k - \tau)}
-    +
-    \frac{\left[\sum_k(y_k - \bar{y})\sin\omega(t_k - \tau)\right]^2}
-    {\sum_k \sin^2\omega(t_k - \tau)}
-\right],
-\end{equation}
+  P_N(\omega) = \frac{1}{2\,V_y}
+  \Bigg[&\frac{\left[\sum_k(y_k - \bar{y})\cos\omega(t_k - \tau)\right]^2}
+    {\sum_k \cos^2\omega(t_k - \tau)} +\nonumber\\
+   & \frac{\left[\sum_k(y_k - \bar{y})\sin\omega(t_k - \tau)\right]^2}
+    {\sum_k \sin^2\omega(t_k - \tau)}\Bigg],
+\end{eqnarray}
 where $\bar{y}$ is the mean and $V_y$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
@@
 The expressions related to the stationary sinusoid model can be expressed more compactly by defining the following matrices:
-\begin{equation}
+\begin{eqnarray}
 X_\omega = \left[
@@
 \end{array}
-\right];~~
+\right]; \nonumber\\
 y = \left[
@@
 \right]
-\end{equation}
+\end{eqnarray}
 With these definitions, the model in \eq{simplemodel} can be expressed as a simple linear product, $y(t|\omega,\theta) = X_\omega\theta$, and the model and reference $\chi^2$ can be written
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
@@
   \figlabel{basic_example}
-\end{figure}
+\end{figure*}
 
@@
   \figlabel{floating_mean}
-\end{figure}
+\end{figure*}
 
@@
   \figlabel{multiterm_example}
-\end{figure}
+\end{figure*}
 
@@
   \figlabel{regularized_example}
-\end{figure}
+\end{figure*}
 
@@
   \eqlabel{multiband_model}
-  y_k(t|\omega,\theta) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\nonumber\\ 
+  &y_k(t|\omega,\theta) =
+  \theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right] +&\nonumber\\ 
   &\theta^{(k)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(k)}_{2n - 1}\sin(n\omega t) + \theta^{(k)}_{2n}\cos(n\omega t)\right].&
@@
   \figlabel{multiband_sim}
-\end{figure}
+\end{figure*}
 
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{fig06.pdf}
+  \includegraphics[width=0.5\textwidth]{fig06.pdf}
   \caption{
@@
   \figlabel{compare_periods}
-\end{figure}
+\end{figure*}
 
@@
   \figlabel{compare_periods_reduced}
-\end{figure}
+\end{figure*}
 
@@
   \centering
-  \includegraphics[width=0.7\textwidth]{fig09.pdf}
+  \includegraphics[width=0.5\textwidth]{fig09.pdf}
   \caption{
@@
 Along with the periodogram implementation, it also contains code to download all the data used in this work.
-Code to reproduce this paper, including all figures, is available in a separate repository\footnote{\url{http://github.com/jakevdp/multiband_LS/}}.
+Code to reproduce this paper, including all figures, is available in a separate repository\footnote{\url{http://github.com/jakevdp/multiband\_LS/}}.
 
@@
 
-\section{Comparison with \citet{Long2014}}
+\section{Comparison with Long (2014)}
 \sectlabel{long_comparison}

### 2015-05-12 writeup/paper.tex
Modify Welch-Stetson ref; closes Issue #2
Modify Jaynes (1987) reference; he doesn't mention Lomb-Scargle in particular

@@
 method to optimally extract the best period. Their method is essentially a multiband generalization of the well-known
-two-band Welch-Stetson variability index \citep{Stetson1996}. Unfortunately, when data in each band are taken at
+two-band Welch-Stetson variability index \citep{WelchStetson1993}. Unfortunately, when data in each band are taken at
 different times, such an approach in not applicable. In such cases, past studies have generally relied 
@@
 \end{equation}
-This long-recognized connection between spectral power and least squares fitting methods was solidified by \citet{Jaynes87}, who demonstrated that the normalized periodogram of Lomb and Scargle is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
+This long-recognized connection between spectral power and least squares fitting methods was solidified by \citet{Jaynes87}, who demonstrated that the least-squares periodogram method is a sufficient statistic for inferences about a stationary-frequency signal in the presence of Gaussian noise.
 Building on this result, \citet{Bretthorst88} explored the extension of these methods to more complicated models with multiple frequency terms, non-stationary frequencies, and other more sophisticated models within a Bayesian framework.

### 2015-05-12 writeup/paper.tex
respond to referee comments – red letters

@@
 \newcommand{\eq}[1]{\Eq{#1}}
+\newcommand{\eqs}[2]{Equations~(\ref{eq:#1})-(\ref{eq:#2})}
 \newcommand{\eqlabel}[1]{\label{eq:#1}}
@@
 \newcommand{\sect}[1]{\Sect{#1}}
+\newcommand{\sects}[1]{Sections~#1}
 \newcommand{\App}[1]{Appendix~\ref{sect:#1}}
@@
 \newcommand{\sectlabel}[1]{\label{sect:#1}}
+
+\usepackage[normalem]{ulem}
+\newcommand{\new}[1]{{\color{red} #1}}
+\newcommand{\old}[1]{{\sout{#1}}}
 
@@
 lower-order Fourier series. This decrease in the effective model complexity is the
-main reason for improved performance. We use simulated light curves 
+main reason for improved performance. \new{After a pedagogical development of the formalism
+of least-squares spectral analysis which motivates the essential features of the multiband model,}
+we use simulated light curves 
 and randomly subsampled SDSS Stripe 82 data to demonstrate the superiority of this 
@@
 efficiently determine the correct period in the majority of LSST's bright RR Lyrae stars with
-as little as six months of LSST data.
+as little as six months of LSST data, \new{a vast improvement over the 10+ years of data
+reported to be required by previous studies.}
 A Python implementation of this method, along with code to fully reproduce the results
@@
 advantage of the full information content in available data, it would be desirable to have a single estimate of the periodogram 
-which accounts for all observed data in a manner which does not depend on the underlying spectrum of the object. 
+which accounts for all observed data in a manner which does not depend on assumptions about the underlying spectrum of the object. 
 We propose such a method in this paper.
@@
 
-The remainder of the paper is organized as follows:
-in \sect{brief_overview} we provide a brief review of least-squares periodic fitting, and in \sect{matrix_formalism} derive the matrix-based formalism for single-band periodic analysis used through the rest of this work.
+The remainder of the paper is organized as follows.
+\new{\sects{2-4} offer a review of essential concepts in least squares modeling and least squares spectral analysis, as well as their relationship to common periodogram estimates:}
+in \sect{brief_overview} we provide a brief review of least-squares periodic fitting, and in \sect{standard_least_squares} derive the matrix-based formalism for single-band least squares spectral analysis used through the rest of this work.
 \sect{extending_periodogram} introduces several extensions and generalizations to the single-band model that the matrix formalism makes possible, including floating mean models, truncated Fourier models, and regularized models.
-In \sect{multiband}, we use the ideas behind these extensions to motivate the {\it multiband periodogram}, and show some examples of its use on simulated data.
+\new{\sects{5-7} present our new developments:}
+in \sect{multiband}, we use the ideas and formalism of \sects{2-4} to motivate the {\it multiband periodogram}, and show some examples of its use on simulated data.
 In \sect{stripe82} we apply this method to measurements of 483 RR Lyrae stars first explored by \citet[][hereafter S10]{Sesar2010}, and in \sect{LSST} explore the performance of the method for simulated observations from the LSST survey.
@@
 \end{eqnarray}
-where $\bar{y}$ is the mean and $V_y$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
-\citet{Lomb76} showed that this time-offset has a deeper effect: namely, it makes $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
+where $\bar{y}$ is the mean and $V_y$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which \new{orthogonalizes the model and} makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
+\citet{Lomb76} showed that this time-offset has a deeper effect: namely, it \new{gives $P_N$ a similar form to previous extensions of $C(\omega)$, while leaving} $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
 \begin{equation}
@@
 While the important features of least squares frequency estimation via Lomb-Scargle periodograms have been discussed elsewhere, we will present a brief introduction to the subject in the following section.
-The matrix-based formalism we develop here will make clear how the method can be extended to more sophisticated models, including the multiband periodogram proposed in this work.
+\old{The matrix-based formalism we develop here will make clear how the method} \new{In particular, expressing the problem in a matrix-based formalism makes clear how the basic approach motivated by \citet{Lomb76}, \citet{Scargle82}, and others} can be extended to more sophisticated models, including the multiband periodogram proposed in this work.
 
 \section{Standard Least Squares Spectral Fitting}
-\sectlabel{matrix_formalism}
+\sectlabel{standard_least_squares}
 
@@
 where $t_k$ is the time of observation, $y_k$ is the observed value (typically a magnitude), and $\sigma_k$ describes the Gaussian errors on each value.
-Without loss of generality we will assume that the data $y_k$ are centered such that the measurements within each band satisfy
+\old{Without loss of generality} \new{For notational simplicity} we will assume that the data $y_k$ are centered such that the measurements within each band satisfy
 \begin{equation}
@@
 where the weights are $w_k = \sigma_k^{-2}$.
+\new{Though this assumption is important for the models presented in this section, it will become superfluous in later sections.}
 
@@
 \end{equation}
-where $P_N(\omega)$ is the normalized periodogram given in \eq{LombScargle} and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$.
+where $P_N(\omega)$ is the normalized periodogram given in \eq{LombScargle}\new{\footnote{\new{An important feature of the Lomb-Scargle approach is the modification of the model with the time-offset $\tau$ tuned to orthogonalize the harmonic basis across the irregular times $\{t_i\}$.
+This orthogonalization cancels cross-terms in the expression of $\chi^2$, greatly reducing the complexity of computing $P_N$.
+As discussed in footnote \ref{footnote:ortho2}, however, this orthogonalization does not change the final form of the periodogram and so we can safely ignore it for the purposes of this work.}}\label{footnote:ortho1}}
+and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$.
 
 \subsection{Matrix Formalism}
+\sectlabel{matrix_formalism}
 The expressions related to the stationary sinusoid model can be expressed more compactly by defining the following matrices:
@@
 \begin{eqnarray}
+  \eqlabel{LS_chi2}
   \chi^2(\omega) &=& (y - X_\omega\theta)^T\Sigma^{-1}(y - X_\omega\theta)\\
@@
 \end{equation}
-We note that this expression is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto V_y I$.
+We note that \eq{LombScargle2} is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto V_y I$.
+\footnote{\new{In order to directly match the Lomb-Scargle approach, we need the equivalent of the $\tau$ parameter which orthogonalizes the basis across the observed times $\{t_i\}$.
+Such an orthogonalization is accomplished via the transformations $X_\omega \to X_\omega V_\omega$ and $\theta \to V_\omega^T \theta$, where $V_\omega$ is the orthogonal matrix of eigenvectors of the covariance $X_\omega^T \Sigma^{-1} X_\omega$.
+The $V_\omega$ terms trivially cancel out of \eqs{LS_chi2}{LombScargle2} and the results of this section are unchanged.
+The matrix formalism makes clear that this result applies to all the extensions of the periodogram mentioned in this work.}\label{footnote:ortho2}}
 
@@
 This visualization makes it apparent that the single-term model is highly biased: RR Lyrae light curves are, in general, much more complicated than a simple sinusoid.
-Nevertheless, the simplistic sinusoidal model is able to recover the correct frequency to a high degree of accuracy (roughly related to the width of the peak) and significance (roughly related to the height of the peak).
+Nevertheless, the simplistic sinusoidal model is able to recover the correct frequency to a high degree of accuracy (roughly related to the width of the peak) and significance (roughly related to the height of the peak; see \citet{Scargle82} for details).
 For a more complete introduction to and discussion of the single-term normalized periodogram, refer to, e.g.
@@
 
-\section{Extending the Periodogram}
+\section{Generalizing the Periodogram Model}
 \sectlabel{extending_periodogram}
@@
 
-As an example of one of these generalizations, we'll consider what \citet{Zechmeister09} call the ``generalized Lomb-Scargle method'', and which we'll call the {\it floating-mean periodogram}.
-This method adjusts the classic normalized periodogram by fitting the mean of the model alongside the amplitudes:
+As an example of one of these generalizations, we'll consider \old{what Zechmeister \& Kurster (2009) call the ``generalized Lomb-Scargle method'', and which we'll call the floating-mean periodogram.}
+\new{what has variously been called the {\it Date-compensated periodogram} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle method} \citep{Zechmeister09}.
+Here we use the term {\it floating-mean periodogram}.} This method adjusts the classic normalized periodogram by fitting the mean of the model alongside the amplitudes:
 \begin{equation}
@@
 This model can be more accurate than the standard periodogram for certain observing cadences and selection functions.
-\citet{Zechmeister09} detail the required modifications to the harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
+\citet{Zechmeister09} detail the required modifications to the \new{orthogonalized} harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
 In the matrix formalism, the modification is much more straightforward: all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
@@
 
-For well-sampled data, there is usually very little difference between a standard periodogram and a floating-mean periodogram.
+For well-sampled data, there is usually very little difference between a \old{standard periodogram and a floating-mean periodogram} \new{floating-mean periodogram and a standard periodogram on pre-centered data}.
 Where this becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases of the light curve: a toy example demonstrating this situation is shown in \fig{floating_mean}.
 The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias (left panel).
-Because of this observational bias, the mean of the observed data are a poor predictor of the true mean, causing the standard fixed-mean method to poorly fit the data and miss the input period (upper-right panel).
+Because of this observational bias, the mean of the observed data is a poor predictor of the true mean, causing the standard method to poorly fit the data and miss the input period (upper-right panel).
 The floating-mean approach is able to automatically adjust for this bias, resulting in a periodogram which readily detects the input period of 0.622 days (lower-right panel).
@@
 As mentioned above, the standard periodogram is equivalent to fitting a single-term stationary sinusoidal model to the data.
-A natural extension is to instead use a multiple-term sinusoidal model, with frequencies at integer multiples of the fundamental frequency.
+A natural extension is to instead use a multiple-term sinusoidal model, with frequencies at integer multiples of the fundamental frequency \citep[See e.g.][]{Bretthorst88}.
 With $N$ Fourier terms, there are $2N + 1$ free parameters, and the model is given by
@@
 \end{equation}
-Because this model remains linear in the parameters $\theta$, it can be easily accommodated into the matrix formalism above.
+Because this model remains linear in the parameters $\theta$, it can be easily accommodated into the matrix formalism of \sect{matrix_formalism}.
 For example, an $N = 2$-term floating-mean model can be constructed by building a design matrix $X_\omega$ with $2N + 1 = 5$ columns:
@@
 \sectlabel{multiband}
-In this section we will apply what was learned in the previous sections to construct the {\it multiband periodogram} which can flexibly account for heterogeneous sources of data for a single object.
+In this section we will combine the ideas of the previous sections to construct the {\it multiband periodogram} which can flexibly account for heterogeneous sources of data for a single object.
 To compute such a periodogram, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
@@
   &y_k(t|\omega,\theta) =
-  \theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right] +&\nonumber\\ 
-  &\theta^{(k)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(k)}_{2n - 1}\sin(n\omega t) + \theta^{(k)}_{2n}\cos(n\omega t)\right].&
+  \theta_0 + \sum_{n=1}^{N_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right] +&\nonumber\\ 
+  &\theta^{(k)}_0 + \sum_{n=1}^{N_{band}} \left[\theta^{(k)}_{2n - 1}\sin(n\omega t) + \theta^{(k)}_{2n}\cos(n\omega t)\right].&
 \end{eqnarray}
@@
 \begin{equation}
-  \Lambda^{(1,0)} = {\rm diag}([0, 0, 0, \epsilon, \epsilon, \epsilon, \epsilon, \epsilon])
+  \Lambda^{(1,0)} = {\rm diag}([0, 0, 0, \lambda, \lambda, \lambda, \lambda, \lambda])
 \end{equation}
-where $\epsilon$ is some small fraction of the trace of the normal matrix $[X_\omega^T\Sigma^{-1}X_\omega]$.
-The logic of this choice of regularization is that any component of the model which is common to all bands will be preferentially reflected in the base terms, with independent behavior reflected in the individual band terms.
-Setting $\epsilon$ to some small fraction of the trace ensures that the regularization effect on the remaining model will be small.
+where $\lambda$ controls the degree of regularization.
+As $\lambda$ grows large, the model will preferentially push power into the base terms, while minimizing the deviations of the model for each individual band.
+
+Here we will choose $\lambda$ to be some small fraction of the trace of the normal matrix $[X_\omega^T\Sigma^{-1}X_\omega]$.
+This choice ensures the multiband periodogram is well-defined, while maintaining the flexibility of the model in accounting for independent band-to-band variation.
 With this regularization in place, the model is well-posed and \eq{LombScargleReg} can be used to straightforwardly compute the power.
@@
 \end{equation}
-We see that this particular multiband periodogram is identical to a weighted sum of standard periodograms in each band, where the weights $\chi^2_{0,k}$ are a reflection of both the number of measurements in each band, and how much those measurements deviate from a simple constant reference model.
+Thus the $(N_{base}=0, N_{band}=1)$ multi-phase  multiband periodogram is identical to a weighted sum of standard periodograms in each band, where the weights $\chi^2_{0,k}$ are a reflection of both the number of measurements in each band, and how much those measurements deviate from a simple constant reference model.
 
@@
 The data are well-enough sampled that a distinct period of 0.622 days can be recognized within each individual band, up to the aliasing effect discussed in \sect{simple_period}.
-Previous studies have made use of the information in multiple bands to choose between aliases and estimate uncertainties in determined periods \citep[e.g.][]{Oluseyi12,Sesar2010}.
+Previous studies have made use of the information in multiple bands to choose between aliases and estimate uncertainties in determined periods \citep[e.g.][]{Sesar2010,Oluseyi12}.
 While this approach is sufficient for well-sampled data, it becomes problematic when the multiband data are sparsely sampled.
@@
 With only 12 observations for each individual band, it is clear that there is not enough data to accurately determine the period within each single band.
-The shared-phase multiband approach (i.e.
-$N_{base}=1,N_{band}=0$), shown in the lower-right panel, fits a single model to the full data and clearly recovers the true frequency of 0.622 days.
+The shared-phase $(N_{base}=1,N_{band}=0)$ multiband approach, shown in the lower-right panel, fits a single model to the full data and clearly recovers the true frequency of 0.622 days.
+\new{The key observation is that while methods based on the standard periodogram are suitable for densely-sampled data, the multiband periodogram is superior for sparsely-sampled multiband observations.}
 
-This shared-phase $(1,0)$ model is only one of the possible multiband options, however: \fig{multiband_models} shows multiband fits to this data for models with various choices of $(N_{base},N_{band})$.
+This shared-phase $(1,0)$ model is only one of the possible multiband options, however: \fig{multiband_models} compares multiband fits to this data for models with various choices of $(N_{base},N_{band})$.
 We see here many of the characteristics noted above for single-band models: as discussed in \sect{multiterm}, increasing the number of Fourier terms leads to power at multiples of the fundamental period, and increased model complexity (roughly indexed by the effective number of free parameters $M^{eff}$) tends to increase the background level of the periodogram, obscuring significant peaks.
@@
 
-One potentially fruitful avenue of research which we do not study here is the application of other types of regularization to the higher-order periodogram.
+One potentially fruitful avenue of research which we do not study here is the adjustment of the regularization terms in the model, and the application of other types of regularization to the higher-order periodogram.
 In particular, L1 regularization (also known as Lasso regression) could lead to interesting results: L1 regularization is similar in spirit to the Tikhonov regularization discussed in \sect{regularization}, but tends toward sparsity in the model parameters \citep[see, e.g.][for a discussion]{ICVG2014}.
-Such an approach could provide a useful tradeoff between model complexity and bias in the case of higher-order truncated Fourier models.
+Such an approach could provide a useful tradeoff between model complexity and bias in the case of higher-order truncated Fourier models, though comes at a higher computational cost.
 
@@
 Sesar, private communication).
-Thus the template-fitting method, though it produces accurate periods, in practice requires several hours of CPU time for a well-sampled period grid for a single source (compared to several seconds for the multiband periodogram proposed here).
+Thus the template-fitting method, though it can produce accurate periods, in practice requires several hours of CPU time for a well-sampled period grid for a single source (compared to several seconds for the multiband periodogram proposed here).
 Note that several hours per object is orders-of-magnitude too slow in the case of LSST; to estimate periods for a billion stars on a 1000-core machine in a year requires a compute-time budget of only 30 seconds per light curve.
@@
 \begin{equation}
-  \chi^2_{PGLS} = \sum_{k=1}^K \bigg[~\chi^2_{GLS}(D^{(K)}) + J_A(A^{(k)}) + J_\phi(\phi^{(k)})~\bigg].
+  \chi^2_{PGLS} = \sum_{k=1}^K \bigg[~\chi^2_{GLS}(D^{(k)}) + J_A(A^{(k)}) + J_\phi(\phi^{(k)})~\bigg].
 \end{equation}
-for $K$ bands, where $\chi^2_{GLS}(D^{(K)})$ is the $\chi^2$ of the standard floating mean model on the single-band data $D^{(K)}$, and $J_A$ and $J_\phi$ are regularization/penalty terms which are a function of the amplitude $A^{k}$ and phase $\phi^{(k)}$ of each model.
+where $K$ is the number of bands, $\chi^2_{GLS}(D^{(k)})$ is the $\chi^2$ of the standard floating mean model on the single-band data $D^{(k)}$, and $J_A$ and $J_\phi$ are regularization/penalty terms which are a function of the amplitude $A^{k}$ and phase $\phi^{(k)}$ of each model.
 In terms of our linear model parameters $\theta^{(k)}$, this amplitude and phase can be expressed:
@@
  A^{(k)} &=& \sqrt{(\theta_1^{(k)})^2 + (\theta_2^{(k)})^2}\nonumber\\
- \phi^{(k)} &=& \arctan(\theta_2 / \theta_1)
+ \phi^{(k)} &=& \arctan(\theta_2^{(k)} / \theta_1^{(k)})
 \end{eqnarray}
 The selected form of these regularization terms penalizes deviations of the amplitude and phase from a common mean between the bands; in this sense the PGLS model can be considered a conceptual mid-point between our shared-phase and multi-phase models.
-Within the formalism proposed in the current work, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in our shared-phase model, though the nature of the resulting regularization will differ.
+Within the formalism proposed in the current work, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\lambda$ used in our shared-phase model, though the precise nature of the resulting regularization will differ.
 
@@
 
-Given these important distinctions between the models proposed here and in LCB14, in future work we plan to do a detailed comparison of the two means of multiband model regularization.
+Given these important distinctions between the models proposed here and in LCB14, in future work we plan to do a detailed comparison of the two approaches to multiband model regularization.
 

### 2015-05-12 writeup/paper.tex
typo in abstract

@@
 efficiently determine the correct period in the majority of LSST's bright RR Lyrae stars with
-as little as six months of LSST data, \new{a vast improvement over the 10+ years of data
+as little as six months of LSST data, \new{a vast improvement over the years of data
 reported to be required by previous studies.}

### 2015-05-12 writeup/paper.tex
add note on floating-mean model

@@
 \end{equation}
-This model can be more accurate than the standard periodogram for certain observing cadences and selection functions.
+\old{This model} \new{The periodogram derived from this model} can be more accurate than the standard periodogram for certain observing cadences and selection functions.
 \citet{Zechmeister09} detail the required modifications to the \new{orthogonalized} harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.

### 2015-05-12 writeup/paper.tex
further updates to comments

@@
 where $\bar{y}$ is the mean and $V_y$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which \new{orthogonalizes the model and} makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
-\citet{Lomb76} showed that this time-offset has a deeper effect: namely, it \new{gives $P_N$ a similar form to previous extensions of $C(\omega)$, while leaving} $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
+\citet{Lomb76} showed that this time-offset has a deeper effect: namely, it \old{leaves} \new{gives $P_N$ a similar form to previous extensions of $C(\omega)$, while leaving} $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
 \begin{equation}
@@
 While the important features of least squares frequency estimation via Lomb-Scargle periodograms have been discussed elsewhere, we will present a brief introduction to the subject in the following section.
-\old{The matrix-based formalism we develop here will make clear how the method} \new{In particular, expressing the problem in a matrix-based formalism makes clear how the basic approach motivated by \citet{Lomb76}, \citet{Scargle82}, and others} can be extended to more sophisticated models, including the multiband periodogram proposed in this work.
+\old{The matrix-based formalism we develop here will make clear how the method} \new{In particular, we re-express the problem in a matrix-based formalism that makes clear how the basic approach motivated by \citet{Lomb76}, \citet{Scargle82}, and others} can be extended to more sophisticated models, including the multiband periodogram proposed in this work.
 
@@
 \sectlabel{standard_least_squares}
-
 In this section we present a brief quantitative introduction to the least squares fitting formulation of the normalized periodogram of \eq{LombScargle}. We denote $N$ observed data points as
@@
 where the weights are $w_k = \sigma_k^{-2}$.
-\new{Though this assumption is important for the models presented in this section, it will become superfluous in later sections.}
+\new{Though this assumption is important for the models presented in this section, it will become superfluous with the floating-mean models described in later sections.}
 
@@
 This orthogonalization cancels cross-terms in the expression of $\chi^2$, greatly reducing the complexity of computing $P_N$.
-As discussed in footnote \ref{footnote:ortho2}, however, this orthogonalization does not change the final form of the periodogram and so we can safely ignore it for the purposes of this work.}}\label{footnote:ortho1}}
+As discussed in footnote \ref{footnote:ortho2}, however, this orthogonalization does not change the resulting periodogram and so it can savely be ignored for the purposes of this work.}}\label{footnote:ortho1}}
 and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$.
@@
 We note that \eq{LombScargle2} is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto V_y I$.
-\footnote{\new{In order to directly match the Lomb-Scargle approach, we need the equivalent of the $\tau$ parameter which orthogonalizes the basis across the observed times $\{t_i\}$.
+\footnote{\new{For direct comparison to the Lomb-Scargle approach, we need the equivalent of the $\tau$ parameter which orthogonalizes the basis across the observed times $\{t_i\}$.
 Such an orthogonalization is accomplished via the transformations $X_\omega \to X_\omega V_\omega$ and $\theta \to V_\omega^T \theta$, where $V_\omega$ is the orthogonal matrix of eigenvectors of the covariance $X_\omega^T \Sigma^{-1} X_\omega$.
 The $V_\omega$ terms trivially cancel out of \eqs{LS_chi2}{LombScargle2} and the results of this section are unchanged.
-The matrix formalism makes clear that this result applies to all the extensions of the periodogram mentioned in this work.}\label{footnote:ortho2}}
+The general matrix formalism used here makes clear that this result applies to all the periodogram extensions mentioned in this work.}\label{footnote:ortho2}}
 
@@
 As an example of one of these generalizations, we'll consider \old{what Zechmeister \& Kurster (2009) call the ``generalized Lomb-Scargle method'', and which we'll call the floating-mean periodogram.}
-\new{what has variously been called the {\it Date-compensated periodogram} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle method} \citep{Zechmeister09}.
+\new{what has variously been called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle method} \citep{Zechmeister09}.
 Here we use the term {\it floating-mean periodogram}.} This method adjusts the classic normalized periodogram by fitting the mean of the model alongside the amplitudes:
@@
 
-For well-sampled data, there is usually very little difference between a \old{standard periodogram and a floating-mean periodogram} \new{floating-mean periodogram and a standard periodogram on pre-centered data}.
+For well-sampled data, there is usually very little difference between a \old{standard periodogram and a floating-mean periodogram} \new{a standard periodogram on pre-centered data and a floating-mean periodogram}.
 Where this becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases of the light curve: a toy example demonstrating this situation is shown in \fig{floating_mean}.
@@
     all five bands, and results in a significant detection of the true period.
+    This indicates that while methods based on the standard periodogram are
+    suitable for densely-sampled multiband data, the multiband periodogram
+    is superior for sparsely-sampled multiband observations.
   }
@@
 The shared-phase $(N_{base}=1,N_{band}=0)$ multiband approach, shown in the lower-right panel, fits a single model to the full data and clearly recovers the true frequency of 0.622 days.
-\new{The key observation is that while methods based on the standard periodogram are suitable for densely-sampled data, the multiband periodogram is superior for sparsely-sampled multiband observations.}
+\new{The key result is that while methods based on the standard periodogram are suitable for densely-sampled data, the multiband periodogram is superior for sparsely-sampled multiband observations.}
 
@@
     Both methods find the correct period among the top 5 significant peaks around 99\% of the time.
+    \new{This suggests that for densely-sampled multiband surveys, the multiband periodogram will match the results of standard methods (but see \fig{compare_periods_reduced}).}
   } 
@@
     The shared-phase multiband approach utilizes information from all five bands, and returns much more robust results: even with the greatly-reduced data, the true period is among the top 5 candidates in 93\% of cases.
+    \new{This suggests that for sparsely-sampled multiband survey data (such as that expected from LSST) the multiband periodogram will produce superior results when compared to standard methods -- see \fig{LSST_sims}.}
   } 

### 2015-05-12 writeup/paper.tex
add Brandt matrix reference

@@
 \sectlabel{matrix_formalism}
-The expressions related to the stationary sinusoid model can be expressed more compactly by defining the following matrices:
+\new{A standard way of compactly expressing least squares models is via matrix expressions \citep[See e.g.][]{Brandt1970}. Likewise,}
+the expressions related to the stationary sinusoid model can be expressed more compactly by defining the following matrices:
 \begin{eqnarray}

### 2015-05-12 writeup/paper.tex
add note about centering

@@
 where $t_k$ is the time of observation, $y_k$ is the observed value (typically a magnitude), and $\sigma_k$ describes the Gaussian errors on each value.
-\old{Without loss of generality} \new{For notational simplicity} we will assume that the data $y_k$ are centered such that the measurements within each band satisfy
+\old{Without loss of generality} \new{For notational simplicity} we will assume \new{without loss of generality} that the data $y_k$ are centered such that the measurements within each band satisfy
 \begin{equation}
@@
 where the weights are $w_k = \sigma_k^{-2}$.
-\new{Though this assumption is important for the models presented in this section, it will become superfluous with the floating-mean models described in later sections.}
+\new{Though this assumption is essential to the models presented in this section, it will become superfluous with the floating-mean models described in later sections.}
 

### 2015-05-15 writeup/paper.tex
add references to figure 9 caption

@@
     Fraction of periods correctly determined for LSST RR Lyrae as a function
-    of the length of the observing season and the mean $g$-band magnitude, for the multiband (solid lines)
-    and single-band supersmoother (dashed lines) approaches.
+    of the length of the observing season and the mean $g$-band magnitude, for the multiband periodogram approach \new{(method of this work; solid lines)}
+    and single-band supersmoother approach \new{\citep[method of][dashed lines]{Oluseyi12}}.
     The multiband method is superior to the single-band supersmoother approach in all cases, and especially for the faintest objects.

### 2015-05-15 writeup/paper.tex
Add table summarizing results

@@
 To guide the eye, the figure includes indicators of the locations of beat aliases (dotted lines) and multiplicative aliases (dashed lines) of the S10 period.
+\new{Numerical results are summarized in the upper rows of Table \ref{table:results}.}
 
-The best-fit supersmoother period matches the S10 period in 89\% of cases (421/483), while the best-fit multiband period matches the S10 period in 79\% of cases (382/483).
+The best-fit supersmoother period matches the S10 period in 87\% of cases (421/483), while the best-fit multiband period matches the S10 period in 79\% of cases (382/483).
 The modes of failure are instructive: when the supersmoother model misses the S10 period, it tends to land on a multiplicative alias (i.e.
@@
 
-The results of this subsection show that the shared-phase multiband approach is comparable to the supersmoother approach for densely-sampled multiband data, although it has a tendency to get fooled by structure in the survey window.
+The results of this subsection show that the shared-phase multiband approach is comparable to the single-band supersmoother approach for densely-sampled multiband data, although it has a tendency to get fooled by structure in the survey window.
 Correction for this based on the estimated window power may alleviate this (see \citet{Roberts87} for an example of such an approach) though in practice selecting from among the top 5 peaks appears to be sufficient.
@@
 
-The lower panels of \fig{compare_periods_reduced} show the relationship between the S10 periods (based on the full dataset) and the periods derived with each model from this reduced dataset.
-It is clear that the supersmoother model is simply over-fitting noise with this few data points: the top period matches S10 in only 32\% of cases (compared to 87\% with the full dataset), and the top 5 periods contain the S10 period only 45\% of the time.
+The lower panels of \fig{compare_periods_reduced} show the relationship between the S10 periods (based on the full dataset) and the periods derived with each model from this reduced dataset, \new{and these results are summarized in the lower rows of Table \ref{table:results}}.
+It is clear that the supersmoother model is simply over-fitting noise with this few data points: the top period matches S10 in only 23\% of cases (compared to 87\% with the full dataset), and the top 5 periods contain the S10 period only 45\% of the time.
 The failure mode is much less predictable as well: rather than being clustered near aliases, most of the period determinations are scattered seemingly randomly around the parameter space.
 
-The multiband method does much better on the reduced dataset.
-Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 64\% of the time (compared to 79\% with the full dataset), and the top 5 peaks contain the S10 period 94\% of the time (compared to 99\% with the full dataset).
-This performance is due to the fact that the multiband algorithm has relatively few parameters, but is yet able to flexibly accommodate data from multiple observing bands.
+\old{The multiband method outperforms the single-band method on the sparse dataset.} \new{While the multiband method performed comparably to the S10 method on dense data, it far outperforms S10 on the sparse dataset.} Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 64\% of the time (compared to 79\% with the full dataset), and the top 5 peaks contain the S10 period 94\% of the time (compared to 99\% with the full dataset).
+This performance is due to the fact that the multiband algorithm has relatively few parameters, but is yet able to flexibly accommodate noisy data from multiple observing bands.
 In particular, this suggests that with the multiterm periodogram, the S10 analysis could have been done effectively with only a small fraction of the available data.
 This bodes well for future surveys, where data on variable stars will be much more sparsely sampled.
+
+\begin{table*}
+\centering
+  \caption{Period Determination from dense and sparse data (483 total)}
+  \begin{tabular}{|l|l|l|l|l|l|l|}
+  \hline
+   Data & Method & Match among top 5 & Top peak matches & Beat Aliases & Multiplicative Aliases  \\
+  \hline\hline
+  Dense data (\fig{compare_periods})
+  & g-band Supersmoother & 477 (98.8\%) & 421 (87.2\%)& 31 & 34 \\
+  & Multi-band Periodogram & 480 (99.4\%) & 382 (79.1\%) & 94 & 5 \\
+  \hline\hline
+  Sparse data (\fig{compare_periods_reduced})
+  & g-band Supersmoother & 219 (45.3\%) & 113 (23.4\%) & 101 & 4 \\
+  & Multi-band Periodogram & 449 (93.0\%) & 308 (63.8\%) & 136 & 7 \\
+  \hline
+  \end{tabular}
+  \label{table:results}
+\end{table*}
 

### 2015-05-15 writeup/paper.tex
multiplicative -> harmonic

@@
     The upper panels show a representative lightcurve and periodogram fits, while the bottom panels compare the derived periods to the template-based periods reported in S10.
-    Shown for reference are the beat aliases (dotted lines) and the multiplicative alias (dashed lines): numbers along the top and right edges of the panels indicate the number of points aligned with each trend.
-    The single-band supersmoother model tends to err toward multiplicative aliases, while the multiband model tends to err toward beat frequency aliases.
+    Shown for reference are the beat aliases (dotted lines) and the \old{multiplicative} \new{first harmonic} alias (dashed lines): numbers along the top and right edges of the panels indicate the number of points aligned with each trend.
+    The single-band supersmoother model tends to err toward \old{multiplicative} \new{harmonic} aliases, while the multiband model tends to err toward beat frequency aliases.
     Both methods find the correct period among the top 5 significant peaks around 99\% of the time.
@@
 The lower panels of \fig{compare_periods} compare the S10 period with the best periods obtained from the 1-band supersmoother (lower-left) and from the shared-phase multiband model (lower-right).
-To guide the eye, the figure includes indicators of the locations of beat aliases (dotted lines) and multiplicative aliases (dashed lines) of the S10 period.
+To guide the eye, the figure includes indicators of the locations of beat aliases (dotted lines) and \old{multiplicative} \new{first harmonic} aliases (dashed lines) of the S10 period.
 \new{Numerical results are summarized in the upper rows of Table \ref{table:results}.}
@@
 The best-fit supersmoother period matches the S10 period in 87\% of cases (421/483), while the best-fit multiband period matches the S10 period in 79\% of cases (382/483).
-The modes of failure are instructive: when the supersmoother model misses the S10 period, it tends to land on a multiplicative alias (i.e.
+The modes of failure are instructive: when the supersmoother model misses the S10 period, it tends to land on a \old{multiplicative} \new{harmonic} alias (i.e.
 the dashed line).
@@
   \hline
-   Data & Method & Match among top 5 & Top peak matches & Beat Aliases & Multiplicative Aliases  \\
+   Data & Method & Match among top 5 & Top peak matches & Beat Aliases & \old{Multiplicative} \new{Harmonic} Aliases  \\
   \hline\hline

### 2015-05-19 writeup/paper.tex
multiplicative -> harmonic
remove remove red text

--- 2015-05-15 writeup/paper.tex
@@
 lower-order Fourier series. This decrease in the effective model complexity is the
-main reason for improved performance. \new{After a pedagogical development of the formalism
-of least-squares spectral analysis which motivates the essential features of the multiband model,}
+main reason for improved performance. After a pedagogical development of the formalism
+of least-squares spectral analysis which motivates the essential features of the multiband model,
 we use simulated light curves 
@@
 efficiently determine the correct period in the majority of LSST's bright RR Lyrae stars with
-as little as six months of LSST data, \new{a vast improvement over the years of data
-reported to be required by previous studies.}
+as little as six months of LSST data, a vast improvement over the years of data
+reported to be required by previous studies.
 A Python implementation of this method, along with code to fully reproduce the results
@@
 The remainder of the paper is organized as follows.
-\new{\sects{2-4} offer a review of essential concepts in least squares modeling and least squares spectral analysis, as well as their relationship to common periodogram estimates:}
+\sects{2-4} offer a review of essential concepts in least squares modeling and least squares spectral analysis, as well as their relationship to common periodogram estimates:
 in \sect{brief_overview} we provide a brief review of least-squares periodic fitting, and in \sect{standard_least_squares} derive the matrix-based formalism for single-band least squares spectral analysis used through the rest of this work.
 \sect{extending_periodogram} introduces several extensions and generalizations to the single-band model that the matrix formalism makes possible, including floating mean models, truncated Fourier models, and regularized models.
-\new{\sects{5-7} present our new developments:}
+\sects{5-7} present our new developments:
 in \sect{multiband}, we use the ideas and formalism of \sects{2-4} to motivate the {\it multiband periodogram}, and show some examples of its use on simulated data.
@@
 \end{eqnarray}
-where $\bar{y}$ is the mean and $V_y$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which \new{orthogonalizes the model and} makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
-\citet{Lomb76} showed that this time-offset has a deeper effect: namely, it \old{leaves} \new{gives $P_N$ a similar form to previous extensions of $C(\omega)$, while leaving} $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
+where $\bar{y}$ is the mean and $V_y$ is the variance of the data $\{y_k\}$, and $\tau$ is the time-offset which orthogonalizes the model and makes $P_N(\omega)$ independent of a translation in $t$ \citep[see][for an in-depth discussion]{NumRec}.
+\citet{Lomb76} showed that this time-offset has a deeper effect: namely, it gives $P_N$ a similar form to previous extensions of $C(\omega)$, while leaving $P_N$ identical to the estimate of harmonic content given a least-squares fit to a single-component sinusoidal model,
 \begin{equation}
@@
 While the important features of least squares frequency estimation via Lomb-Scargle periodograms have been discussed elsewhere, we will present a brief introduction to the subject in the following section.
-\old{The matrix-based formalism we develop here will make clear how the method} \new{In particular, we re-express the problem in a matrix-based formalism that makes clear how the basic approach motivated by \citet{Lomb76}, \citet{Scargle82}, and others} can be extended to more sophisticated models, including the multiband periodogram proposed in this work.
+In particular, we re-express the problem in a matrix-based formalism that makes clear how the basic approach motivated by \citet{Lomb76}, \citet{Scargle82}, and others can be extended to more sophisticated models, including the multiband periodogram proposed in this work.
 
@@
 where $t_k$ is the time of observation, $y_k$ is the observed value (typically a magnitude), and $\sigma_k$ describes the Gaussian errors on each value.
-\old{Without loss of generality} \new{For notational simplicity} we will assume \new{without loss of generality} that the data $y_k$ are centered such that the measurements within each band satisfy
+For notational simplicity we will assume without loss of generality that the data $y_k$ are centered such that the measurements within each band satisfy
 \begin{equation}
@@
 where the weights are $w_k = \sigma_k^{-2}$.
-\new{Though this assumption is essential to the models presented in this section, it will become superfluous with the floating-mean models described in later sections.}
+Though this assumption is essential to the models presented in this section, it will become superfluous with the floating-mean models described in later sections.
 
@@
 \end{equation}
-where $P_N(\omega)$ is the normalized periodogram given in \eq{LombScargle}\new{\footnote{\new{An important feature of the Lomb-Scargle approach is the modification of the model with the time-offset $\tau$ tuned to orthogonalize the harmonic basis across the irregular times $\{t_i\}$.
+where $P_N(\omega)$ is the normalized periodogram given in \eq{LombScargle}\footnote{An important feature of the Lomb-Scargle approach is the modification of the model with the time-offset $\tau$ tuned to orthogonalize the harmonic basis across the irregular times $\{t_i\}$.
 This orthogonalization cancels cross-terms in the expression of $\chi^2$, greatly reducing the complexity of computing $P_N$.
-As discussed in footnote \ref{footnote:ortho2}, however, this orthogonalization does not change the resulting periodogram and so it can savely be ignored for the purposes of this work.}}\label{footnote:ortho1}}
+As discussed in footnote \ref{footnote:ortho2}, however, this orthogonalization does not change the resulting periodogram and so it can savely be ignored for the purposes of this work.}\label{footnote:ortho1}
 and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$.
@@
 \sectlabel{matrix_formalism}
-\new{A standard way of compactly expressing least squares models is via matrix expressions \citep[See e.g.][]{Brandt1970}. Likewise,}
+A standard way of compactly expressing least squares models is via matrix expressions \citep[See e.g.][]{Brandt1970}. Likewise,
 the expressions related to the stationary sinusoid model can be expressed more compactly by defining the following matrices:
@@
 We note that \eq{LombScargle2} is equivalent to \eq{LombScargle} in the homoscedastic case with $\Sigma \propto V_y I$.
-\footnote{\new{For direct comparison to the Lomb-Scargle approach, we need the equivalent of the $\tau$ parameter which orthogonalizes the basis across the observed times $\{t_i\}$.
+\footnote{For direct comparison to the Lomb-Scargle approach, we need the equivalent of the $\tau$ parameter which orthogonalizes the basis across the observed times $\{t_i\}$.
 Such an orthogonalization is accomplished via the transformations $X_\omega \to X_\omega V_\omega$ and $\theta \to V_\omega^T \theta$, where $V_\omega$ is the orthogonal matrix of eigenvectors of the covariance $X_\omega^T \Sigma^{-1} X_\omega$.
 The $V_\omega$ terms trivially cancel out of \eqs{LS_chi2}{LombScargle2} and the results of this section are unchanged.
-The general matrix formalism used here makes clear that this result applies to all the periodogram extensions mentioned in this work.}\label{footnote:ortho2}}
+The general matrix formalism used here makes clear that this result applies to all the periodogram extensions mentioned in this work.\label{footnote:ortho2}}
 
@@
 
-As an example of one of these generalizations, we'll consider \old{what Zechmeister \& Kurster (2009) call the ``generalized Lomb-Scargle method'', and which we'll call the floating-mean periodogram.}
-\new{what has variously been called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle method} \citep{Zechmeister09}.
-Here we use the term {\it floating-mean periodogram}.} This method adjusts the classic normalized periodogram by fitting the mean of the model alongside the amplitudes:
+As an example of one of these generalizations, we'll consider 
+what has variously been called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle method} \citep{Zechmeister09}.
+Here we use the term {\it floating-mean periodogram}.
+This method adjusts the classic normalized periodogram by fitting the mean of the model alongside the amplitudes:
 \begin{equation}
@@
 \end{equation}
-\old{This model} \new{The periodogram derived from this model} can be more accurate than the standard periodogram for certain observing cadences and selection functions.
-\citet{Zechmeister09} detail the required modifications to the \new{orthogonalized} harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
+The periodogram derived from this model can be more accurate than the standard periodogram for certain observing cadences and selection functions.
+\citet{Zechmeister09} detail the required modifications to the orthogonalized harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
 In the matrix formalism, the modification is much more straightforward: all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
@@
 
-For well-sampled data, there is usually very little difference between a \old{standard periodogram and a floating-mean periodogram} \new{a standard periodogram on pre-centered data and a floating-mean periodogram}.
+For well-sampled data, there is usually very little difference between a standard periodogram on pre-centered data and a floating-mean periodogram.
 Where this becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases of the light curve: a toy example demonstrating this situation is shown in \fig{floating_mean}.
@@
 The shared-phase $(N_{base}=1,N_{band}=0)$ multiband approach, shown in the lower-right panel, fits a single model to the full data and clearly recovers the true frequency of 0.622 days.
-\new{The key result is that while methods based on the standard periodogram are suitable for densely-sampled data, the multiband periodogram is superior for sparsely-sampled multiband observations.}
+The key result is that while methods based on the standard periodogram are suitable for densely-sampled data, the multiband periodogram is superior for sparsely-sampled multiband observations.
 
@@
     The upper panels show a representative lightcurve and periodogram fits, while the bottom panels compare the derived periods to the template-based periods reported in S10.
-    Shown for reference are the beat aliases (dotted lines) and the multiplicative alias (dashed lines): numbers along the top and right edges of the panels indicate the number of points aligned with each trend.
-    The single-band supersmoother model tends to err toward multiplicative aliases, while the multiband model tends to err toward beat frequency aliases.
+    Shown for reference are the beat aliases (dotted lines) and the first harmonic alias (dashed lines): numbers along the top and right edges of the panels indicate the number of points aligned with each trend.
+    The single-band supersmoother model tends to err toward harmonic aliases, while the multiband model tends to err toward beat frequency aliases.
     Both methods find the correct period among the top 5 significant peaks around 99\% of the time.
-    \new{This suggests that for densely-sampled multiband surveys, the multiband periodogram will match the results of standard methods (but see \fig{compare_periods_reduced}).}
+    This suggests that for densely-sampled multiband surveys, the multiband periodogram will match the results of standard methods (but see \fig{compare_periods_reduced}).
   } 
@@
 The lower panels of \fig{compare_periods} compare the S10 period with the best periods obtained from the 1-band supersmoother (lower-left) and from the shared-phase multiband model (lower-right).
-To guide the eye, the figure includes indicators of the locations of beat aliases (dotted lines) and multiplicative aliases (dashed lines) of the S10 period.
-\new{Numerical results are summarized in the upper rows of Table \ref{table:results}.}
+To guide the eye, the figure includes indicators of the locations of beat aliases (dotted lines) and first harmonic aliases (dashed lines) of the S10 period.
+Numerical results are summarized in the upper rows of Table \ref{table:results}.
 
 The best-fit supersmoother period matches the S10 period in 87\% of cases (421/483), while the best-fit multiband period matches the S10 period in 79\% of cases (382/483).
-The modes of failure are instructive: when the supersmoother model misses the S10 period, it tends to land on a multiplicative alias (i.e.
+The modes of failure are instructive: when the supersmoother model misses the S10 period, it tends to land on a harmonic alias (i.e.
 the dashed line).
@@
     The shared-phase multiband approach utilizes information from all five bands, and returns much more robust results: even with the greatly-reduced data, the true period is among the top 5 candidates in 93\% of cases.
-    \new{This suggests that for sparsely-sampled multiband survey data (such as that expected from LSST) the multiband periodogram will produce superior results when compared to standard methods -- see \fig{LSST_sims}.}
+    This suggests that for sparsely-sampled multiband survey data (such as that expected from LSST) the multiband periodogram will produce superior results when compared to standard methods -- see \fig{LSST_sims}.
   } 
@@
 
-The lower panels of \fig{compare_periods_reduced} show the relationship between the S10 periods (based on the full dataset) and the periods derived with each model from this reduced dataset, \new{and these results are summarized in the lower rows of Table \ref{table:results}}.
+The lower panels of \fig{compare_periods_reduced} show the relationship between the S10 periods (based on the full dataset) and the periods derived with each model from this reduced dataset, and these results are summarized in the lower rows of Table \ref{table:results}.
 It is clear that the supersmoother model is simply over-fitting noise with this few data points: the top period matches S10 in only 23\% of cases (compared to 87\% with the full dataset), and the top 5 periods contain the S10 period only 45\% of the time.
@@
 
-\old{The multiband method outperforms the single-band method on the sparse dataset.} \new{While the multiband method performed comparably to the S10 method on dense data, it far outperforms S10 on the sparse dataset.} Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 64\% of the time (compared to 79\% with the full dataset), and the top 5 peaks contain the S10 period 94\% of the time (compared to 99\% with the full dataset).
+While the multiband method performed comparably to the S10 method on dense data, it far outperforms S10 on the sparse dataset.
+Even with an 80\% reduction in the number of observations, the multiband method matches the S10 period 64\% of the time (compared to 79\% with the full dataset), and the top 5 peaks contain the S10 period 94\% of the time (compared to 99\% with the full dataset).
 This performance is due to the fact that the multiband algorithm has relatively few parameters, but is yet able to flexibly accommodate noisy data from multiple observing bands.
@@
   \hline
-   Data & Method & Match among top 5 & Top peak matches & Beat Aliases & Multiplicative Aliases  \\
+   Data & Method & Match among top 5 & Top peak matches & Beat Aliases & Harmonic Aliases  \\
   \hline\hline
@@
     Fraction of periods correctly determined for LSST RR Lyrae as a function
-    of the length of the observing season and the mean $g$-band magnitude, for the multiband periodogram approach \new{(method of this work; solid lines)}
-    and single-band supersmoother approach \new{\citep[method of][dashed lines]{Oluseyi12}}.
+    of the length of the observing season and the mean $g$-band magnitude, for the multiband periodogram approach (method of this work; solid lines)
+    and single-band supersmoother approach \citep[method of][dashed lines]{Oluseyi12}.
     The multiband method is superior to the single-band supersmoother approach in all cases, and especially for the faintest objects.

### 2015-08-18 writeup/paper.tex
respond to second ref report

@@
 and can treat non-uniformly sampled time series with heteroscedastic measurement uncertainties. 
-The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency and basis functions at different
-frequencies are not orthogonal. As a result, the periodogram has many local maxima and thus in practice the global 
-maximum of the periodogram is found by grid search \citep[for details see, e.g.][]{ICVG2014}.
+\new{The underlying model of the Lomb-Scargle periodogram is nonlinear in frequency, and thus the likelihood surface in frequency is non-convex.
+This non-convexity is readily apparent in the many local maxima of the periodogram, which makes it difficult to find the maximum via standard numerical optimization routines.
+Thus in practice the global 
+maximum of the periodogram is often found by a brute-force grid search \citep[for details see, e.g.][]{ICVG2014}.}
+\old{The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency and basis functions at different frequencies are not orthogonal. 
+As a result, the periodogram has many local maxima. Thus in practice the global 
+maximum of the periodogram is often found by a brute-force grid search (for details see, e.g. Ivezic 2014)}
 A more general parametric method based on the use of continuous-time autoregressive moving average (CARMA) model
@@
 This orthogonalization cancels cross-terms in the expression of $\chi^2$, greatly reducing the complexity of computing $P_N$.
-As discussed in footnote \ref{footnote:ortho2}, however, this orthogonalization does not change the resulting periodogram and so it can savely be ignored for the purposes of this work.}\label{footnote:ortho1}
+As discussed in footnote \ref{footnote:ortho2}, however, this orthogonalization does not change the resulting periodogram and so it can safely be ignored for the purposes of this work.}\label{footnote:ortho1}
 and $\chi^2_0$ is the reference $\chi^2$ for a constant model, which due to the assumption in \eq{ycentered} is simply $\chi^2_0 = \sum_k (y_k/\sigma_k)^2$.
@@
 Such an orthogonalization is accomplished via the transformations $X_\omega \to X_\omega V_\omega$ and $\theta \to V_\omega^T \theta$, where $V_\omega$ is the orthogonal matrix of eigenvectors of the covariance $X_\omega^T \Sigma^{-1} X_\omega$.
-The $V_\omega$ terms trivially cancel out of \eqs{LS_chi2}{LombScargle2} and the results of this section are unchanged.
+The $V_\omega$ terms straightforwardly cancel out of \eqs{LS_chi2}{LombScargle2} and the results of this section are unchanged.
 The general matrix formalism used here makes clear that this result applies to all the periodogram extensions mentioned in this work.\label{footnote:ortho2}}
@@
 \begin{enumerate}
-  \item It is trivially extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the {\it noise covariance matrix} $\Sigma$.
-  \item It is trivially extended to more sophisticated linear models by appropriately modifying the {\it design matrix} $X_\omega$.
-  \item It is trivially extended to include Tikhonov/L2-regularization terms (see \sect{regularization} for more details)  by adding an appropriate diagonal term to the {\it normal matrix} $X_\omega^T\Sigma^{-1}X_\omega$.
+  \item It is straightforwardly extended to heteroscedastic and/or correlated measurement noise in the data $y_k$ through appropriate modification of the {\it noise covariance matrix} $\Sigma$.
+  \item It is straightforwardly extended to more sophisticated linear models by appropriately modifying the {\it design matrix} $X_\omega$.
+  \item It is straightforwardly extended to include Tikhonov/L2-regularization terms (see \sect{regularization} for more details)  by adding an appropriate diagonal term to the {\it normal matrix} $X_\omega^T\Sigma^{-1}X_\omega$.
 \end{enumerate}
@@
 For well-sampled data, there is usually very little difference between a standard periodogram on pre-centered data and a floating-mean periodogram.
-Where this becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases of the light curve: a toy example demonstrating this situation is shown in \fig{floating_mean}.
+Where this difference becomes important is if selection effects or observing cadences cause there to be preferentially more observations at certain phases of the light curve: a toy example demonstrating this situation is shown in \fig{floating_mean}.
 The data are drawn from a sinusoid with Gaussian errors, and data with a magnitude fainter than 16 are removed to simulate an observational bias (left panel).
@@
     The model fits and periodograms for several truncated Fourier models.
-    The data are the same as those in \fig{basic_example}. Note that the
+    The data are the same as those in \fig{basic_example}. Note that
+    \new{in addition to the previously-seen 0.38-day alias,} the
     higher-order models will generally show periodogram peaks at multiples
@@
 Thus in general the observed power of a higher-order Fourier model will be everywhere higher than the power of a lower-order Fourier model.
+\new{One might hope that when adding terms, the correct-period model would show more of an improvement than the incorrect-period model (and thus the maximum would become more pronounced), but this does not generally hold.
+Consider that in the extreme limit in which the number model parameters is equal to the number of data points, the model has enough flexibility to fit the data perfectly at {\it every} frequency, and the resulting periodogram would be a flat line!
+Thus when adding more terms, the boost in power to the background must be larger on average than the boost in power to the signal.}
 
@@
 \sectlabel{multiband}
-In this section we will combine the ideas of the previous sections to construct the {\it multiband periodogram} which can flexibly account for heterogeneous sources of data for a single object.
-To compute such a periodogram, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
+In this section we will combine the ideas of the previous sections to construct the {\it multiband periodogram} which flexibly account for heterogeneous sources of data for a single object.
+\new{To start with, we might consider one of two na{\"i}ve approaches:
+
+First, we might consider ignoring band labels and simply compute a single standard Lomb-Scargle periodogram over the full dataset.
+This amounts to the assumption that a single sinusoid fits each band, and in practice will fail to account for astrophysical variability between bands.
+
+Second, we might consider computing a standard Lomb-Scargle periodogram in each band, and using the additivity of $\chi^2$ along with \eq{chi2PN} to construct a multiband periodogram.
+This amounts to the assumption that the bands have completely independent phases and amplitudes, and has too many free parameters to be useful in most cases of interest (see \sect{relationship} for further discussion).}
+
+To compute \old{such} a periodogram \new{which strikes a balance between these two extremes}, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
 The multiband model contains the following features:
 \begin{enumerate}
-  \item An $N_{base}$-term truncated Fourier fit which models a latent parameter, which we'll call the ``overall variability''.
-  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual of a single band from this overall variability.
+  \item An $N_{base}$-term truncated Fourier fit which models \new{the shared variability among all $K$ observed bands}. \old{a latent parameter, which we'll call the ``overall variability''.}
+  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual within a single band from the \old{overall variability} \new{shared variability accounted for above}.
 \end{enumerate}
-The total number of parameters for $K$ filters is then $M_K = (2N_{base} + 1) + K(2N_{band} + 1)$. As a result, for each band $k$ we have the following model of the observed magnitudes:
+The total number of parameters for $K$ bands is then $M_K = (2N_{base} + 1) + K(2N_{band} + 1)$. As a result, for each band $k$ we have the following model of the observed magnitudes:
 \begin{eqnarray}
@@
 The important feature of this model is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(k)}$ are determined individually.
+\new{Note the potential for confusion: $N_{band}$ here is not the number of observed bands, but the number of Fourier components fit to the data each of the $K$ observed bands.}
 
@@
 Columns corresponding to the base model and the matching observation band will have nonzero entries; all other columns will be filled with zeros.
-For example, the $N_{base}=1$ and $N_{band}=0$ model corresponds to one with a simple single-term periodic base frequency, and an independent constant offset term in each band.
+For example, the $(N_{base},N_{band})=(1,0)$ model corresponds to one with a simple single-term periodic base frequency, and an independent constant offset term in each band.
 The associated design matrix depends on the particular data, but will look similar to this:
@@
 \sectlabel{relationship}
-The {\it multi-phase} $(N_{base}=0, N_{band}=1)$ model turns out to be a particularly special case.
+The multi-phase model with $(N_{base},N_{band})=(0,1)$ turns out to be a particularly special case.
 Here the base model is a simple global offset which is degenerate with the offsets in each band, so that the design matrix $X_\omega$ can be straightforwardly rearranged as block-diagonal.
@@
 \end{equation}
-Thus the $(N_{base}=0, N_{band}=1)$ multi-phase  multiband periodogram is identical to a weighted sum of standard periodograms in each band, where the weights $\chi^2_{0,k}$ are a reflection of both the number of measurements in each band, and how much those measurements deviate from a simple constant reference model.
+Thus the $(N_{base},N_{band}=(0,1)$ multi-phase multiband periodogram is identical to a weighted sum of standard periodograms in each band, where the weights $\chi^2_{0,k}$ are a reflection of both the number of measurements in each band, and how much those measurements deviate from a simple constant reference model.
 
@@
 With only 12 observations for each individual band, it is clear that there is not enough data to accurately determine the period within each single band.
-The shared-phase $(N_{base}=1,N_{band}=0)$ multiband approach, shown in the lower-right panel, fits a single model to the full data and clearly recovers the true frequency of 0.622 days.
+The shared-phase $(N_{base},N_{band})=(1,0)$ multiband approach, shown in the lower-right panel, fits a single model to the full data and clearly recovers the true frequency of 0.622 days.
 The key result is that while methods based on the standard periodogram are suitable for densely-sampled data, the multiband periodogram is superior for sparsely-sampled multiband observations.
@@
 \subsection{Potential Improvements to the Multiband Method}
-The fact that the multiband model does not select the top frequency each time points to its weaknesses: first, as a Fourier-like model, it tends to respond to frequency structure in the window function as well as frequency structure in the data.
-This is a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle bias in the model from bias due to the survey window.
+\new{A well-known (though often unrecognized) deficiency of Lomb-Scargle-type periodograms on unevenly-sampled data is that they do not measure the power of the signal in question, but the power of the signal {\it convolved with the observing with the survey window function}.
+For regularly-sampled timeseries, this convolution is the source of the aliasing related to the Nyquist sampling limit; for non-regular sampling, this aliasing generally happens to some degree at {\it all} frequencies!
+Thus even if the signal has a single well-defined period, this aliasing generally causes multiple maxima in the measured periodogram.
+The multiband periodogram, as a generalization of Lomb-Scargle, shares this difficulty:}
+\old{The fact that the multiband model does not select the top frequency each time points to its weaknesses: first, as a Fourier-like model,} it tends to respond to frequency structure in the window function as well as frequency structure in the data.
+This can be viewed as a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle bias in the model from bias due to the survey window.
+
 This could potentially be accounted for by correcting for the effect of the estimated window function; one potential method for this involves estimating the deconvolution of the window power and the observed power \citep{Roberts87}.
@@
 
-One potentially fruitful avenue of research which we do not study here is the adjustment of the regularization terms in the model, and the application of other types of regularization to the higher-order periodogram.
+Another potentially fruitful avenue of research which we do not study here is the adjustment of the regularization terms in the model, and the application of other types of regularization to the higher-order periodogram.
 In particular, L1 regularization (also known as Lasso regression) could lead to interesting results: L1 regularization is similar in spirit to the Tikhonov regularization discussed in \sect{regularization}, but tends toward sparsity in the model parameters \citep[see, e.g.][for a discussion]{ICVG2014}.
@@
 \begin{enumerate}
-   \item As a first pass, find a set of candidate frequencies using the multiband periodogram. This is a fast linear optimization that can be trivially parallelized.
+   \item As a first pass, find a set of candidate frequencies using the multiband periodogram. This is a fast linear optimization that can be straightforwardly parallelized.
    \item Within these candidate frequencies, use the more costly template-fitting procedure to choose the optimal period from among the handful of candidates.

### 2015-08-18 writeup/paper.tex
more modifications for referee 2

@@
 and can treat non-uniformly sampled time series with heteroscedastic measurement uncertainties. 
+\old{The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency and basis functions at different frequencies are not orthogonal. 
+As a result, the periodogram has many local maxima.}
 \new{The underlying model of the Lomb-Scargle periodogram is nonlinear in frequency, and thus the likelihood surface in frequency is non-convex.
-This non-convexity is readily apparent in the many local maxima of the periodogram, which makes it difficult to find the maximum via standard numerical optimization routines.
+This non-convexity is readily apparent in the many local maxima of the periodogram, which makes it difficult to find the maximum via standard numerical optimization routines.}
 Thus in practice the global 
-maximum of the periodogram is often found by a brute-force grid search \citep[for details see, e.g.][]{ICVG2014}.}
-\old{The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency and basis functions at different frequencies are not orthogonal. 
-As a result, the periodogram has many local maxima. Thus in practice the global 
-maximum of the periodogram is often found by a brute-force grid search (for details see, e.g. Ivezic 2014)}
+maximum of the periodogram is often found by a brute-force grid search \citep[for details see, e.g.][]{ICVG2014}.
 A more general parametric method based on the use of continuous-time autoregressive moving average (CARMA) model
@@
 Thus in general the observed power of a higher-order Fourier model will be everywhere higher than the power of a lower-order Fourier model.
-\new{One might hope that when adding terms, the correct-period model would show more of an improvement than the incorrect-period model (and thus the maximum would become more pronounced), but this does not generally hold.
-Consider that in the extreme limit in which the number model parameters is equal to the number of data points, the model has enough flexibility to fit the data perfectly at {\it every} frequency, and the resulting periodogram would be a flat line!
-Thus when adding more terms, the boost in power to the background must be larger on average than the boost in power to the signal.}
+\new{One might hope that when adding terms, the correct-period model would show more of an improvement than the incorrect-period model (and thus the periodogram maximum would become more pronounced), but this does not generally hold.
+Consider that in the extreme limit in which the number model parameters is equal to the number of data points, the model has enough flexibility to fit the data perfectly at {\it every} frequency, and the resulting periodogram would be a flat line at maximum power!
+This can only be the case if, on average, addition of terms preferentially boosts the background level.}
 
@@
 
-First, we might consider ignoring band labels and simply compute a single standard Lomb-Scargle periodogram over the full dataset.
-This amounts to the assumption that a single sinusoid fits each band, and in practice will fail to account for astrophysical variability between bands.
+First, we might ignore band labels entirely and simply compute a single standard Lomb-Scargle periodogram over the full dataset.
+This amounts to the assumption that one global model suitably fits each band, and in practice will fail due to the astrophysical variability between bands.
 
-Second, we might consider computing a standard Lomb-Scargle periodogram in each band, and using the additivity of $\chi^2$ along with \eq{chi2PN} to construct a multiband periodogram.
+Second, we might treat each band entirely independently and compute a standard Lomb-Scargle periodogram on each, and use the additivity of $\chi^2$ along with \eq{chi2PN} to construct a multiband periodogram.
 This amounts to the assumption that the bands have completely independent phases and amplitudes, and has too many free parameters to be useful in most cases of interest (see \sect{relationship} for further discussion).}
@@
 To compute \old{such} a periodogram \new{which strikes a balance between these two extremes}, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
-The multiband model contains the following features:
+The multiband model presented here contains the following features:
 \begin{enumerate}
@@
 The important feature of this model is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(k)}$ are determined individually.
-\new{Note the potential for confusion: $N_{band}$ here is not the number of observed bands, but the number of Fourier components fit to the data each of the $K$ observed bands.}
+\new{Note the potential for confusion: $N_{band}$ here is not the number of observed bands, but the number of Fourier components fit to the residuals in each of the $K$ observed bands.}
 
@@
 For regularly-sampled timeseries, this convolution is the source of the aliasing related to the Nyquist sampling limit; for non-regular sampling, this aliasing generally happens to some degree at {\it all} frequencies!
-Thus even if the signal has a single well-defined period, this aliasing generally causes multiple maxima in the measured periodogram.
+Because of this, even a signal with a single well-defined period will result in a Lomb-Scargle periodogram with multiple maxima at locations which depend on both the underlying signal and the precise observing window.
+
 The multiband periodogram, as a generalization of Lomb-Scargle, shares this difficulty:}
@@
 This could potentially be accounted for by correcting for the effect of the estimated window function; one potential method for this involves estimating the deconvolution of the window power and the observed power \citep{Roberts87}.
-It may also be possible to propose a multiband extension of CARMA \citep{Kelly14}, or another forward-modeling approach to detecting periodicity.
+It may also be possible to propose a multiband extension of, e.g., CARMA \citep{Kelly14} or another forward-modeling approach to detecting periodicity.
 

### 2015-08-20 writeup/paper.tex
more tweaks in response to second referee

@@
 \new{The underlying model of the Lomb-Scargle periodogram is nonlinear in frequency, and thus the likelihood surface in frequency is non-convex.
-This non-convexity is readily apparent in the many local maxima of the periodogram, which makes it difficult to find the maximum via standard numerical optimization routines.}
+This non-convexity is readily apparent in the many local maxima of the typical periodogram, which makes it difficult to find the maximum via standard numerical optimization routines.}
 Thus in practice the global 
@@
 advantage of the full information content in available data, it would be desirable to have a single estimate of the periodogram 
-which accounts for all observed data in a manner which does not depend on assumptions about the underlying spectrum of the object. 
+which accounts for all observed data in a manner independent of assumptions about the underlying spectrum of the object.
 We propose such a method in this paper.
@@
 model} common to all bands, while fits for individual bands describe residuals relative to the base model 
-and typically require lower-order Fourier series. This decrease in effective model complexity is the
+and typically require lower-order Fourier series. This regularization-driven decrease in effective model complexity is the
 main reason for improved performance.
@@
 Thus in general the observed power of a higher-order Fourier model will be everywhere higher than the power of a lower-order Fourier model.
+
 \new{One might hope that when adding terms, the correct-period model would show more of an improvement than the incorrect-period model (and thus the periodogram maximum would become more pronounced), but this does not generally hold.
@@
 We use $\lambda = 0$ on the offset term, and make the penalty $\lambda_j$ progressively larger for each harmonic component.
-The result of the regularized model is less over-fitting to the input data, and stronger periodogram peaks.
+The regularization prevents overfitting (left panel), and results in more prominent periodogram peaks (right panel).
 
@@
 \sectlabel{multiband}
-In this section we will combine the ideas of the previous sections to construct the {\it multiband periodogram} which flexibly account for heterogeneous sources of data for a single object.
-\new{To start with, we might consider one of two na{\"i}ve approaches:
+In this section we will combine the ideas of the previous sections to construct the {\it multiband periodogram} which flexibly accounts for heterogeneous sources of data for a single object.
+\new{To start with, we might consider one of two na{\"i}ve approaches to the multi-band problem:
 
 First, we might ignore band labels entirely and simply compute a single standard Lomb-Scargle periodogram over the full dataset.
-This amounts to the assumption that one global model suitably fits each band, and in practice will fail due to the astrophysical variability between bands.
+This amounts to the assumption that one global model suitably fits each band, and in practice will perform poorly due to the astrophysical variability between bands: in other words, the model is too simple and under-fits the data.
 
 Second, we might treat each band entirely independently and compute a standard Lomb-Scargle periodogram on each, and use the additivity of $\chi^2$ along with \eq{chi2PN} to construct a multiband periodogram.
-This amounts to the assumption that the bands have completely independent phases and amplitudes, and has too many free parameters to be useful in most cases of interest (see \sect{relationship} for further discussion).}
+This amounts to the assumption that the bands have completely independent phases and amplitudes, and has too many free parameters to be useful in most cases of interest.
+In other words, the model is too complex and over-fits the data (see \sect{relationship} for further discussion).}
 
@@
 \begin{enumerate}
-  \item An $N_{base}$-term truncated Fourier fit which models \new{the shared variability among all $K$ observed bands}. \old{a latent parameter, which we'll call the ``overall variability''.}
-  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual within a single band from the \old{overall variability} \new{shared variability accounted for above}.
+\item An $N_{base}$-term truncated Fourier \new{``base model'' which models the shared variability among all $K$ observed bands}.
+\old{a latent parameter, which we'll call the ``overall variability''.}
+  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual within a single band from the \new{shared variability accounted for in the base model.} \old{overall variability}.
 \end{enumerate}
@@
 
-On examination of the above matrix, it's clear that the columns are not linearly independent (i.e.
-$X_\omega$ is low-rank), and thus the parameters of the best-fit model will be degenerate.
-Intuitively, this is due to the fact that if we add an overall offset to the base model, this can be perfectly accounted for by subtracting that same offset from each residual.
+On examination of the above matrix, it's clear that the columns are not linearly independent (i.e. $X_\omega$ is low-rank), and thus the parameters of the best-fit model will be degenerate.
+Intuitively, this is due to the fact that if we add an overall offset to the base model, this can be perfectly accounted for by subtracting that same offset from each residual model.
 Mathematically, the result of this is that the normal matrix $X_\omega^T\Sigma^{-1}X_\omega$ will be non-invertible, and thus the periodogram is ill-defined.
@@
 \sectlabel{relationship}
+\new{
+With this formalism in place, we can return to the na{\"i}ve models discussed at the beginning of \sect{multiband}.
+The first is simply a standard Lomb-Scargle, ignoring the band information.
+The second, in which each band is fit independently, turns out to be equivalent to the $(N_{base},N_{band})=(0,1)$ model defined above.
+}
+\old{
 The multi-phase model with $(N_{base},N_{band})=(0,1)$ turns out to be a particularly special case.
+}
 Here the base model is a simple global offset which is degenerate with the offsets in each band, so that the design matrix $X_\omega$ can be straightforwardly rearranged as block-diagonal.
 A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here these independent components amount to the single-band floating-mean model from \sect{floating_mean}, fit independently for each of the $K$ bands.
-This particular multiband model can give us insight into the relationship between single-band and multiband approaches.
+\old{This particular multiband model can give us insight into the relationship between single-band and multiband approaches.}
 
@@
 \end{equation}
-Thus the $(N_{base},N_{band}=(0,1)$ multi-phase multiband periodogram is identical to a weighted sum of standard periodograms in each band, where the weights $\chi^2_{0,k}$ are a reflection of both the number of measurements in each band, and how much those measurements deviate from a simple constant reference model.
+Thus the $(N_{base},N_{band})=(0,1)$  multiband periodogram is identical to a weighted sum of standard periodograms in each band, where the weights $\chi^2_{0,k}$ are a reflection of both the number of measurements in each band and how much those measurements deviate from a simple constant reference model.
 
@@
 \subsection{Potential Improvements to the Multiband Method}
-\new{A well-known (though often unrecognized) deficiency of Lomb-Scargle-type periodograms on unevenly-sampled data is that they do not measure the power of the signal in question, but the power of the signal {\it convolved with the observing with the survey window function}.
-For regularly-sampled timeseries, this convolution is the source of the aliasing related to the Nyquist sampling limit; for non-regular sampling, this aliasing generally happens to some degree at {\it all} frequencies!
+\new{A well-known (though often unrecognized) difficulty of Lomb-Scargle-type periodograms on unevenly-sampled data is that they do not measure the power of the signal in question, but the power of the signal {\it convolved with the observing with the survey window function}.
+For regularly-sampled timeseries, this convolution is the source of the perfect aliasing beyond the Nyquist sampling limit; for non-regular sampling, this aliasing generally happens to some degree at {\it all} frequencies!
 Because of this, even a signal with a single well-defined period will result in a Lomb-Scargle periodogram with multiple maxima at locations which depend on both the underlying signal and the precise observing window.
@@
 \old{The fact that the multiband model does not select the top frequency each time points to its weaknesses: first, as a Fourier-like model,} it tends to respond to frequency structure in the window function as well as frequency structure in the data.
-This can be viewed as a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle bias in the model from bias due to the survey window.
+This can be viewed as a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle bias in the model from bias due to features in the survey window.
 

### 2015-08-21 writeup/paper.tex
more tweaks in response to second referee

@@
 Neither the Supersmoother algorithm nor the PDM method require \apriori{} knowledge of the light curve shape.
-We note that neither method produces posterior probability distribution for the period but rather a single point 
-estimate.
 
@@
 and can treat non-uniformly sampled time series with heteroscedastic measurement uncertainties. 
-\old{The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency and basis functions at different frequencies are not orthogonal. 
+The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency \old{and basis functions at different frequencies are not orthogonal. 
 As a result, the periodogram has many local maxima.}
-\new{The underlying model of the Lomb-Scargle periodogram is nonlinear in frequency, and thus the likelihood surface in frequency is non-convex.
+\new{and so the likelihood surface in frequency is non-convex.
 This non-convexity is readily apparent in the many local maxima of the typical periodogram, which makes it difficult to find the maximum via standard numerical optimization routines.}
@@
 maximum of the periodogram is often found by a brute-force grid search \citep[for details see, e.g.][]{ICVG2014}.
+
 A more general parametric method based on the use of continuous-time autoregressive moving average (CARMA) model
@@
 
-A weakness of the above methods is that they require homogeneous measurements -- for astronomy data, this means 
+A weakness of all these standard methods is that they require homogeneous measurements -- for astronomy data, this means 
 that successive measurements must be taken through a single photometric bandpass (filter). This has not been a major
@@
 where the weights are $w_k = \sigma_k^{-2}$.
-Though this assumption is essential to the models presented in this section, it will become superfluous with the floating-mean models described in later sections.
+Though this assumption is essential to the simpler models presented in this section, it will become superfluous with the floating-mean models described in later sections.
 
@@
 \end{equation}
-The periodogram derived from this model can be more accurate than the standard periodogram for certain observing cadences and selection functions.
-\citet{Zechmeister09} detail the required modifications to the orthogonalized harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
+The periodogram derived from this model can be more accurate than the standard \new{pre-centered} periodogram for certain observing cadences and selection functions, \new{and espeically when searching for long-period varaibility or working with very few samples \citep{Cumming99}.} \citet{Zechmeister09} detail the required modifications to the orthogonalized harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
 In the matrix formalism, the modification is much more straightforward: all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
@@
 
-\new{One might hope that when adding terms, the correct-period model would show more of an improvement than the incorrect-period model (and thus the periodogram maximum would become more pronounced), but this does not generally hold.
-Consider that in the extreme limit in which the number model parameters is equal to the number of data points, the model has enough flexibility to fit the data perfectly at {\it every} frequency, and the resulting periodogram would be a flat line at maximum power!
+\new{One might hope that when adding terms, the correct-period model would show more of an improvement than the incorrect-period model (and thus the periodogram maximum would become more pronounced in comparison to the background), but this does not generally hold.
+Consider that in the extreme limit in which the number model parameters is equal to the number of data points, the model has enough flexibility to fit the data perfectly at {\it every} frequency, and the resulting periodogram would be everywhere unity!
 This can only be the case if, on average, addition of terms preferentially boosts the background level.}
@@
 
-The previous sections raise the question: how complicated a model should we use?
-We have seen that as we add more terms to the fit, the model will come closer and closer to the observed data.
-In the extreme case, when the number of model parameters equals the number of data points, the model can fit the data {\it exactly} regardless of frequency and the periodogram will be everywhere unity (though in most cases, numerical inaccuracies prevent a truly perfect fit).
-For very high-order models, the fit becomes very sensitive to the noise in the data, and we say we have {\it over-fit} the data.
-This can be addressed by explicitly truncating the series, but we can also use a {\it regularization} term to mathematically enforce a less complicated model.
+The previous sections raise the question: how complicated a model should we use? We have seen that as we add more terms to the fit, the model will more closely describe the observed data.
+\old{In the extreme case, when the number of model parameters equals the number of data points, the model can fit the data {\it exactly} regardless of frequency and the periodogram will be everywhere unity (though in most cases, numerical inaccuracies prevent a truly perfect fit).}
+For very high-order models, however, such a close fit {\it over-fits} the data: that is, the fit is more responsive to statistical noise in the observations than to the underlying signal.
+This can be addressed by explicitly truncating the series at some number of terms, but we can also use a {\it regularization} term to mathematically enforce model simplicity.
 
@@
 \new{
-With this formalism in place, we can return to the na{\"i}ve models discussed at the beginning of \sect{multiband}.
-The first is simply a standard Lomb-Scargle, ignoring the band information.
+With this formalism in place, we can return briefly to the na{\"i}ve models discussed at the beginning of \sect{multiband}.
+The first, which ignores band information, is simply a standard Lomb-Scargle over the heterogeneous data.
 The second, in which each band is fit independently, turns out to be equivalent to the $(N_{base},N_{band})=(0,1)$ model defined above.
@@
 \end{equation}
-and it can easily be shown that the $P_N$ can be constructed as a weighted sum of $P_N^{(k)}$:
+and it can be shown straightforwaredly that $P_N^{(0,1)}$ can be constructed as a weighted sum of $P_N^{(k)}$:
 \begin{equation}
@@
 The full S10 RR Lyrae dataset consists of 483 objects with an average of 55 observations in each of the five SDSS $ugriz$ bands spread over just under ten years.
-In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother model and the shared-phase $(0, 1)$-multiband model\footnote{The supersmoother ``periodogram'' $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$, where $\bar{r}_0$ is the mean absolute residual around a constant model.}.
+In the upper panels of \fig{compare_periods} we show the observed data for one of these objects, along with the periodogram derived with the single-band supersmoother model\footnote{The supersmoother ``periodogram'' $P_{SS}$ is constructed from the minimum sum of weighted model residuals $\bar{r}_{min}$ in analogy with \eq{chi2PN}: $P_{SS}(\omega) = 1 - \bar{r}_{min}(\omega) / \bar{r}_0$, where $\bar{r}_0$ is the mean absolute residual around a constant model.} and the shared-phase $(0, 1)$-multiband model.
 Here we have a case which is analogous to that shown for simulated data in the top panels of \fig{multiband_sim}: each band has enough data to easily locate candidate peaks, the best of which is selected via the S10 template-fitting procedure.

### 2015-08-26 writeup/paper.tex
remove new/old markers

@@
 and can treat non-uniformly sampled time series with heteroscedastic measurement uncertainties. 
-The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency \old{and basis functions at different frequencies are not orthogonal. 
-As a result, the periodogram has many local maxima.}
-\new{and so the likelihood surface in frequency is non-convex.
-This non-convexity is readily apparent in the many local maxima of the typical periodogram, which makes it difficult to find the maximum via standard numerical optimization routines.}
+The underlying model of the Lomb–Scargle periodogram is nonlinear in frequency and so the likelihood surface in frequency is non-convex.
+This non-convexity is readily apparent in the many local maxima of the typical periodogram, which makes it difficult to find the maximum via standard numerical optimization routines.
 Thus in practice the global 
@@
 \end{equation}
-The periodogram derived from this model can be more accurate than the standard \new{pre-centered} periodogram for certain observing cadences and selection functions, \new{and espeically when searching for long-period varaibility or working with very few samples \citep{Cumming99}.} \citet{Zechmeister09} detail the required modifications to the orthogonalized harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
+The periodogram derived from this model can be more accurate than the standard \pre-centered periodogram for certain observing cadences and selection functions, and espeically when searching for long-period varaibility or working with very few samples \citep{Cumming99}.
+\citet{Zechmeister09} detail the required modifications to the orthogonalized harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
 In the matrix formalism, the modification is much more straightforward: all that is required is to add a column of ones to the $X_\omega$ matrix before computing the power via \eq{LombScargle2}.
@@
     The data are the same as those in \fig{basic_example}. Note that
-    \new{in addition to the previously-seen 0.38-day alias,} the
+    in addition to the previously-seen 0.38-day alias, the
     higher-order models will generally show periodogram peaks at multiples
@@
 
-\new{One might hope that when adding terms, the correct-period model would show more of an improvement than the incorrect-period model (and thus the periodogram maximum would become more pronounced in comparison to the background), but this does not generally hold.
+One might hope that when adding terms, the correct-period model would show more of an improvement than the incorrect-period model (and thus the periodogram maximum would become more pronounced in comparison to the background), but this does not generally hold.
 Consider that in the extreme limit in which the number model parameters is equal to the number of data points, the model has enough flexibility to fit the data perfectly at {\it every} frequency, and the resulting periodogram would be everywhere unity!
-This can only be the case if, on average, addition of terms preferentially boosts the background level.}
+This can only be the case if, on average, addition of terms preferentially boosts the background level.
 
@@
 The previous sections raise the question: how complicated a model should we use? We have seen that as we add more terms to the fit, the model will more closely describe the observed data.
-\old{In the extreme case, when the number of model parameters equals the number of data points, the model can fit the data {\it exactly} regardless of frequency and the periodogram will be everywhere unity (though in most cases, numerical inaccuracies prevent a truly perfect fit).}
 For very high-order models, however, such a close fit {\it over-fits} the data: that is, the fit is more responsive to statistical noise in the observations than to the underlying signal.
@@
 In this section we will combine the ideas of the previous sections to construct the {\it multiband periodogram} which flexibly accounts for heterogeneous sources of data for a single object.
-\new{To start with, we might consider one of two na{\"i}ve approaches to the multi-band problem:
+To start with, we might consider one of two na{\"i}ve approaches to the multi-band problem:
 
@@
 This amounts to the assumption that the bands have completely independent phases and amplitudes, and has too many free parameters to be useful in most cases of interest.
-In other words, the model is too complex and over-fits the data (see \sect{relationship} for further discussion).}
+In other words, the model is too complex and over-fits the data (see \sect{relationship} for further discussion).
 
-To compute \old{such} a periodogram \new{which strikes a balance between these two extremes}, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
+To compute a periodogram which strikes a balance between these two extremes, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
 The multiband model presented here contains the following features:
 \begin{enumerate}
-\item An $N_{base}$-term truncated Fourier \new{``base model'' which models the shared variability among all $K$ observed bands}.
-\old{a latent parameter, which we'll call the ``overall variability''.}
-  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual within a single band from the \new{shared variability accounted for in the base model.} \old{overall variability}.
+\item An $N_{base}$-term truncated Fourier ``base model'' which models the shared variability among all $K$ observed bands. 
+  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual within a single band from the shared variability accounted for in the base model.
 \end{enumerate}
@@
 The important feature of this model is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(k)}$ are determined individually.
-\new{Note the potential for confusion: $N_{band}$ here is not the number of observed bands, but the number of Fourier components fit to the residuals in each of the $K$ observed bands.}
+Note the potential for confusion: $N_{band}$ here is not the number of observed bands, but the number of Fourier components fit to the residuals in each of the $K$ observed bands.
 
@@
 \sectlabel{relationship}
-\new{
 With this formalism in place, we can return briefly to the na{\"i}ve models discussed at the beginning of \sect{multiband}.
@@
 The second, in which each band is fit independently, turns out to be equivalent to the $(N_{base},N_{band})=(0,1)$ model defined above.
-}
-\old{
-The multi-phase model with $(N_{base},N_{band})=(0,1)$ turns out to be a particularly special case.
-}
 Here the base model is a simple global offset which is degenerate with the offsets in each band, so that the design matrix $X_\omega$ can be straightforwardly rearranged as block-diagonal.
 A block-diagonal design matrix in a linear model indicates that components of the model are being solved independently: here these independent components amount to the single-band floating-mean model from \sect{floating_mean}, fit independently for each of the $K$ bands.
-\old{This particular multiband model can give us insight into the relationship between single-band and multiband approaches.}
 
@@
 \subsection{Potential Improvements to the Multiband Method}
-\new{A well-known (though often unrecognized) difficulty of Lomb-Scargle-type periodograms on unevenly-sampled data is that they do not measure the power of the signal in question, but the power of the signal {\it convolved with the observing with the survey window function}.
+A well-known (though often unrecognized) difficulty of Lomb-Scargle-type periodograms on unevenly-sampled data is that they do not measure the power of the signal in question, but the power of the signal {\it convolved with the observing with the survey window function}.
 For regularly-sampled timeseries, this convolution is the source of the perfect aliasing beyond the Nyquist sampling limit; for non-regular sampling, this aliasing generally happens to some degree at {\it all} frequencies!
@@
 
-The multiband periodogram, as a generalization of Lomb-Scargle, shares this difficulty:}
-\old{The fact that the multiband model does not select the top frequency each time points to its weaknesses: first, as a Fourier-like model,} it tends to respond to frequency structure in the window function as well as frequency structure in the data.
+The multiband periodogram, as a generalization of Lomb-Scargle, shares this difficulty: it tends to respond to frequency structure in the window function as well as frequency structure in the data.
 This can be viewed as a result of the very model simplicity which causes its success in the case of sparse multiband data: it cannot disentangle bias in the model from bias due to features in the survey window.

### 2017-01-05 writeup/paper.tex
Typo: \pre->pre

@@
 \end{equation}
-The periodogram derived from this model can be more accurate than the standard \pre-centered periodogram for certain observing cadences and selection functions, and espeically when searching for long-period varaibility or working with very few samples \citep{Cumming99}.
+The periodogram derived from this model can be more accurate than the standard pre-centered periodogram for certain observing cadences and selection functions, and espeically when searching for long-period varaibility or working with very few samples \citep{Cumming99}.
 \citet{Zechmeister09} detail the required modifications to the orthogonalized harmonic formalism of \eq{LombScargle} to allow the mean to float in the model.
